### 2022-06-09 Manuscript/src/main.tex
feat: cleaned code and started writing paper

%! Author = sbbfti
%! Date = 8/06/2022

% Preamble
\documentclass[11pt]{article}

\title{PMV model comparison}
\author{Federico Tartarini, Stefano Schiavon}

\usepackage{mypreamble}

\usepackage[printonlyused]{acronym}

\begin{document}

    \maketitle

    \begin{abstract}
        The \ac{pmv} model developed by Fanger predicts thermal sensation.
        It is used in its original formulation in the ISO 7730:2005, while the ASHRAE 55:2020 uses a modified version \acs{pmv-ce} to better account for air speeds.
        We compared their results against XXX thermal sensation votes collected in buildings.
    \end{abstract}

    \section*{Nomenclature}
    \input{myacronyms.tex}

    \section{Introduction}\label{sec:introduction}
    In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the ISO 7730:2005 Standard~\cite{iso7730} in its original form.
    The \ac{pmv} is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
    It is used both by researchers and practitioners.
    However, Cheung et al. (2019) demonstrated that the accuracy of the \ac{pmv} in predicting self-reported thermal sensation of people is only \qty{34}{\percent}~\cite{Cheung2019}.

    ASHRAE-55 2020 Standard utilizes a modified version of the PMV model [4] that calculates a Cooling Effect (CE) using the Standard Effective Temperature (SET) equation for air speeds higher than \qty{0.1}{\m\per\s}.
    The value of CE is then subtracted from both the measured dry-bulb air temperature (tdb) and \ac{tr} and these results become the input into the PMV model.
    Since the CE already accounts for convective heat losses from the person to the environment, the air speed used to calculate the PMV value is set to 0.1 m/s.

    \printbibliography

\end{document}

### 2022-06-09 Manuscript/src/myacronyms.tex
feat: cleaned code and started writing paper

%! Author = federico
%! Date = 10/06/2020

%\section*{Nomenclature}
\renewcommand{\baselinestretch}{0.75}\normalsize
\renewcommand{\aclabelfont}[1]{\textsc{\acsfont{#1}}}
\begin{acronym}[longest]

    \acro{tdb}[$t_{db}$]{dry-bulb air temperature\acroextra{, \unit{\celsius}}}
    \acro{rh}[$RH$]{relative humidity\acroextra{, \%}}
    \acro{v}[$V$]{average air speed\acroextra{, m/s}}
    \acro{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, \unit{\celsius}}}
    \acro{clo}[$I_{cl}$]{total clothing insulation\acroextra{, clo}}
    \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}

    \acro{pmv}[PMV]{Predicted Mean Vote}
    \acro{pmv-ce}[PMV$_{CE}$]{PMV calculated in accordance with the ASHRAE 55:2020}
    \acro{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
    \acro{set}[SET]{Standard Effective Temperature\acroextra{, \unit{\celsius}}}
    \acro{ce}[CE]{Cooling Effect\acroextra{, \unit{\celsius}}}

\end{acronym}
\renewcommand{\baselinestretch}{1}\normalsize

### 2022-06-09 main.py

@@ -1,17 +1,12 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib import cm
-from matplotlib.colors import ListedColormap, LinearSegmentedColormap
-from pythermalcomfort.models import pmv_ppd, set_tmp, two_nodes
-from pythermalcomfort.utilities import v_relative, clo_dynamic
+import matplotlib as mpl
+mpl.use("Qt5Agg")  # or can use 'TkAgg', whatever you have/prefer
+
+from matplotlib.colors import LinearSegmentedColormap
 import matplotlib.pyplot as plt
 import pandas as pd
 import seaborn as sns
-import statsmodels.api as sm
 import numpy as np
 import warnings
-from scipy import stats
-import matplotlib.style
 import matplotlib as mpl
 from scipy import stats
 from sklearn.metrics import r2_score, mean_absolute_error
@@ -29,7 +24,7 @@ def save_var_latex(key, value):
 
     dict_var = {}
 
-    file_path = "Manuscript/Variables/results.csv"
+    file_path = "Manuscript/Variables/results.dat"
 
     try:
         with open(file_path, newline="") as file:
@@ -71,10 +66,6 @@ def calculate_new_indices(df_):
     df_["thermal_sensation_round"] = df_["thermal_sensation"].round()
     df_["diff_ts_pmv"] = df_[["thermal_sensation", "pmv"]].diff(axis=1)["pmv"]
     df_["diff_ts_pmv_ce"] = df_[["thermal_sensation", "pmv_ce"]].diff(axis=1)["pmv_ce"]
-
-    # check delta between the PMV I calculated and the one Toby did
-    df_["PMV - pmv"] = df_["PMV"] - df_["pmv"]
-    df_["PMV - pmv_ce"] = df_["PMV"] - df_["pmv_ce"]
     df_["thermal_sensation_round - pmv_ce_round"] = df_["thermal_sensation"] - df_["pmv_ce_round"]
 
     return df_
@@ -509,6 +500,7 @@ def plot_distribution_variable():
             axs[ix].set(yticks=(np.arange(0, 1.8, 0.3)))
     sns.despine(bottom=True, left=True)
     plt.savefig("./Manuscript/Figures/dist_input_data.png", dpi=300)
+    plt.show()
 
 
 if __name__ == "__main__":
@@ -555,7 +547,10 @@ if __name__ == "__main__":
     # import data
[…]

### 2022-06-09 Manuscript/src/delete.tex
feat: imported elsevier template

--- 2022-06-09 Manuscript/src/main.tex
@@
 
+\usepackage{glossaries}
+
+\newglossaryentry{iso}{
+      name = ISO 7730:2005 ,
+      description = ISO Standard to calculate PMV and PPD,
+}
+
+\newglossaryentry{ashrae}{
+      name = ASHRAE 55:2020 ,
+      description = ASHRAE Standard to calculate PMV and PPD,
+}
+
 \begin{document}
@@
         The \ac{pmv} model developed by Fanger predicts thermal sensation.
-        It is used in its original formulation in the ISO 7730:2005, while the ASHRAE 55:2020 uses a modified version \acs{pmv-ce} to better account for air speeds.
+        It is used in its original formulation in the \gls{iso}, while the \gls{ashrae} uses a modified version \acs{pmv-ce} to better account for air speeds.
         We compared their results against XXX thermal sensation votes collected in buildings.
@@
 
-    ASHRAE-55 2020 Standard utilizes a modified version of the PMV model [4] that calculates a Cooling Effect (CE) using the Standard Effective Temperature (SET) equation for air speeds higher than \qty{0.1}{\m\per\s}.
+    ASHRAE 55:2020 Standard utilizes a modified version of the PMV model [4] that calculates a Cooling Effect (CE) using the Standard Effective Temperature (SET) equation for air speeds higher than \qty{0.1}{\m\per\s}.
     The value of CE is then subtracted from both the measured dry-bulb air temperature (tdb) and \ac{tr} and these results become the input into the PMV model.

### 2022-06-09 Manuscript/src/main.tex
feat: imported elsevier template

@@
 %! Author = sbbfti
-%! Date = 8/06/2022
+%! Date = 10/06/2020
 
-% Preamble
-\documentclass[11pt]{article}
-
-\title{PMV model comparison}
-\author{Federico Tartarini, Stefano Schiavon}
+\documentclass[review, 3p]{elsarticle}
+%\documentclass[final]{elsarticle}
+%\documentclass[final,5p,times,twocolumn]{elsarticle}
 
@@
 
-\usepackage[printonlyused]{acronym}
-
 \begin{document}
 
-    \maketitle
+    \input{sections/titleAbstract}
 
-    \begin{abstract}
-        The \ac{pmv} model developed by Fanger predicts thermal sensation.
-        It is used in its original formulation in the ISO 7730:2005, while the ASHRAE 55:2020 uses a modified version \acs{pmv-ce} to better account for air speeds.
-        We compared their results against XXX thermal sensation votes collected in buildings.
-    \end{abstract}
-
-    \section*{Nomenclature}
     \input{myacronyms.tex}
 
-    \section{Introduction}\label{sec:introduction}
-    In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the ISO 7730:2005 Standard~\cite{iso7730} in its original form.
-    The \ac{pmv} is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
-    It is used both by researchers and practitioners.
-    However, Cheung et al. (2019) demonstrated that the accuracy of the \ac{pmv} in predicting self-reported thermal sensation of people is only \qty{34}{\percent}~\cite{Cheung2019}.
+    \linenumbers
 
-    ASHRAE-55 2020 Standard utilizes a modified version of the PMV model [4] that calculates a Cooling Effect (CE) using the Standard Effective Temperature (SET) equation for air speeds higher than \qty{0.1}{\m\per\s}.
-    The value of CE is then subtracted from both the measured dry-bulb air temperature (tdb) and \ac{tr} and these results become the input into the PMV model.
-    Since the CE already accounts for convective heat losses from the person to the environment, the air speed used to calculate the PMV value is set to 0.1 m/s.
+    \input{sections/introduction}
 
-    \printbibliography
+    \input{sections/methodology}
+
+    \input{sections/results}
+
+    \input{sections/discussion}
+
+    \input{sections/conclusions}
+
+    \bibliography{mybibfile}
 

### 2022-06-09 Manuscript/src/myacronyms.tex
feat: imported elsevier template

@@
 
-%\section*{Nomenclature}
+\section*{Nomenclature}
 \renewcommand{\baselinestretch}{0.75}\normalsize
@@
 
-    \acro{tdb}[$t_{db}$]{dry-bulb air temperature\acroextra{, \unit{\celsius}}}
+    \acro{t-db}[$t_{db}$]{dry-bulb air temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-wb}[$t_{wb}$]{wet-bulb air temperature\acroextra{, $^{\circ}$C}}
+    \acro{ti}[$t_{i}$]{indoor air temperature\acroextra{, $^{\circ}$C}}
+    \acro{tout}[$t_{out}$]{outdoor air temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-op}[$t_{o}$]{operative air temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-cl}[$t_{cl}$]{clothing temperature\acroextra{, $^{\circ}$C}}
+    \acro{tg}[$t_{g}$]{globe temperature\acroextra{, $^{\circ}$C}}
     \acro{rh}[$RH$]{relative humidity\acroextra{, \%}}
     \acro{v}[$V$]{average air speed\acroextra{, m/s}}
-    \acro{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, \unit{\celsius}}}
+    \acro{t-r}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, $^{\circ}$C}}
     \acro{clo}[$I_{cl}$]{total clothing insulation\acroextra{, clo}}
+    \acro{i-cl}[$i_{cl}$]{permeation efficiency of water vapor through the clothing layer}
     \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
@@
     \acro{pmv}[PMV]{Predicted Mean Vote}
-    \acro{pmv-ce}[PMV$_{CE}$]{PMV calculated in accordance with the ASHRAE 55:2020}
     \acro{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
-    \acro{set}[SET]{Standard Effective Temperature\acroextra{, \unit{\celsius}}}
-    \acro{ce}[CE]{Cooling Effect\acroextra{, \unit{\celsius}}}
+    \acro{set}[SET]{Standard Effective Temperature\acroextra{, $^{\circ}$C}}
+    \acro{ce}[CE]{Cooling Effect\acroextra{, $^{\circ}$C}}
+    \acro{phs}[PHS]{Predicted Heat Strain}
+
+    \acro{BMS}[BMS]{Building Management System}
+    \acro{HVAC}[HVAC]{Heating, Ventilation, and Air Conditioning}
+    \acro{VAV}[VAV]{Variable Air Volume}
+    \acro{AHU}[AHU]{Air Handling Unit}
+
+    \acro{p-sk}[$p_{sk,s}$]{water vapor pressure at skin\acroextra{, kPa}}
+    \acro{p-a}[$p_{a}$]{water vapor pressure in ambient air\acroextra{, kPa}}
+
+    \acro{wmo}[WMO]{World Meteorological Organization}
+    \acro{who}[WHO]{World Health Organization}
+    \acro{cdc}[CDC]{Centers for Disease Control and Prevention}
+    \acro{noaa}[NOAA]{National Oceanic and Atmospheric Administration}
+    \acro{epa}[EPA]{United States Environmental Protection Agency}
+    \acro{iea}[IEA]{International Energy Agency}
+    \acro{un}[UN]{United Nations}
 

### 2022-06-09 Manuscript/src/sections/introduction.tex
feat: imported elsevier template

--- 2022-06-09 Manuscript/src/delete.tex
@@
 %! Author = sbbfti
-%! Date = 8/06/2022
+%! Date = 10/06/2020
 
-% Preamble
-\documentclass[11pt]{article}
-
-\title{PMV model comparison}
-\author{Federico Tartarini, Stefano Schiavon}
-
-\usepackage{mypreamble}
-
-\usepackage[printonlyused]{acronym}
-
-\usepackage{glossaries}
-
-\newglossaryentry{iso}{
-      name = ISO 7730:2005 ,
-      description = ISO Standard to calculate PMV and PPD,
-}
-
-\newglossaryentry{ashrae}{
-      name = ASHRAE 55:2020 ,
-      description = ASHRAE Standard to calculate PMV and PPD,
-}
-
-\begin{document}
-
-    \maketitle
-
-    \begin{abstract}
-        The \ac{pmv} model developed by Fanger predicts thermal sensation.
-        It is used in its original formulation in the \gls{iso}, while the \gls{ashrae} uses a modified version \acs{pmv-ce} to better account for air speeds.
-        We compared their results against XXX thermal sensation votes collected in buildings.
-    \end{abstract}
-
-    \section*{Nomenclature}
-    \input{myacronyms.tex}
+% Document
 
 \section{Introduction}\label{sec:introduction}
+
     In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the ISO 7730:2005 Standard~\cite{iso7730} in its original form.
@@
     Since the CE already accounts for convective heat losses from the person to the environment, the air speed used to calculate the PMV value is set to 0.1 m/s.
-
-    \printbibliography
-
-\end{document}

### 2022-06-09 Manuscript/src/sections/titleAbstract.tex
feat: imported elsevier template

%! Author = sbbfti
%! Date = 10/06/2020

\begin{frontmatter}

\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
    The \ac{pmv} model developed by Fanger predicts thermal sensation.
%    It is used in its original formulation in the \gls{iso}, while the \gls{ashrae} uses a modified version \acs{pmv-ce} to better account for air speeds.
    We compared their results against XXX thermal sensation votes collected in buildings.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

### 2022-06-09 Manuscript/src/myglossary.tex
feat: added glossary

\usepackage[nonumberlist,nogroupskip]{glossaries}

\setglossarystyle{super}

\newglossaryentry{ipm}{
    name={Incremental PM},
    description={Incremental Personal Model}
}

\newglossaryentry{gpm}{
    name={Warm start PM},
    description={Warm start Personal Model}
}

\newglossaryentry{pm}{
    name={PM},
    description={Personal Model}
}
\newglossaryentry{ib}
{
name={iButton\textsuperscript{\tiny\textregistered}},
description={Wireless temperature sensors},
plural={iButtons\textsuperscript{\tiny\textregistered}},
firstplural={iButtons\textsuperscript{\tiny\textregistered}}
}

\newglossaryentry{ubibot}
{
name={UbiBot\textsuperscript{\tiny\textregistered}},
description={Wireless environmental sensor},
}

\newglossaryentry{7730}
{
name={ISO~7730:2005},
description={ISO~7730:2005 is a thermal comfort standard developed by ISO}
}

\newglossaryentry{55}
{
name={ASHRAE~55--2020},
description={ASHRAE~55--2020 is a thermal comfort standard developed by ANSI and ASHRAE}
}

\makenoidxglossaries

### 2022-06-09 Manuscript/src/sections/introduction.tex
feat: added glossary

@@
 
-    In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the ISO 7730:2005 Standard~\cite{iso7730} in its original form.
+    In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
     The \ac{pmv} is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.

### 2022-06-09 Manuscript/src/sections/titleAbstract.tex
paper: updated abstract

@@
 
-\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
+    \title{Title of the paper}
 
-%% Group authors per affiliation:
-\author{Elsevier\fnref{myfootnote}}
-\address{Radarweg 29, Amsterdam}
-\fntext[myfootnote]{Since 1880.}
+    \author[label1]{Federico Tartarini\corref{mycorrespondingauthor}}
+    \ead{federicotartarini@berkeley.edu}
+    \author[label2]{Stefano Schiavon}
 
-%% or include affiliations in footnotes:
-\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
-\ead[url]{www.elsevier.com}
+    \address[label1]{Berkeley Education Alliance for Research in Singapore, Singapore}
+    \address[label2]{Center for the Built Environment, University of California, Berkeley, CA, USA}
 
-\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
     \cortext[mycorrespondingauthor]{Corresponding author}
-\ead{support@elsevier.com}
-
-\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
-\address[mysecondaryaddress]{360 Park Avenue South, New York}
 
@@
     \begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
+        Personal thermal comfort model \sep Ecological momentary assessment \sep Skin temperature \sep Machine learning \sep Internet of Things (IoT)
     \end{keyword}

### 2022-06-09 Manuscript/src/myglossary.tex
paper: minor edits

@@
 \setglossarystyle{super}
-
-\newglossaryentry{ipm}{
-    name={Incremental PM},
-    description={Incremental Personal Model}
-}
-
-\newglossaryentry{gpm}{
-    name={Warm start PM},
-    description={Warm start Personal Model}
-}
-
-\newglossaryentry{pm}{
-    name={PM},
-    description={Personal Model}
-}
-\newglossaryentry{ib}
-{
-name={iButton\textsuperscript{\tiny\textregistered}},
-description={Wireless temperature sensors},
-plural={iButtons\textsuperscript{\tiny\textregistered}},
-firstplural={iButtons\textsuperscript{\tiny\textregistered}}
-}
-
-\newglossaryentry{ubibot}
-{
-name={UbiBot\textsuperscript{\tiny\textregistered}},
-description={Wireless environmental sensor},
-}
 
@@
 
+\newglossaryentry{rhrn}
+{
+name={right-here-right-now},
+description={Right here right now thermal comfort survey}
+}
+
+\newglossaryentry{pmv-ce}
+{
+name={PMV\textsubscript{CE}},
+description={PMV calculated in accordance with the ASHRAE~55--2020}
+}
+
+\newglossaryentry{db2}
+{
+name={Comfort DB},
+description={ASHRAE Global Thermal Comfort Database II}
+}
+
 \makenoidxglossaries

### 2022-06-09 Manuscript/src/sections/introduction.tex
paper: minor edits

@@
 %! Date = 10/06/2020
-
-% Document
 
@@
     In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
+    The \ac{pmv} model has been widely adopted worldwide both among researchers and practitioners.
     The \ac{pmv} is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
-    It is used both by researchers and practitioners.
     However, Cheung et al. (2019) demonstrated that the accuracy of the \ac{pmv} in predicting self-reported thermal sensation of people is only \qty{34}{\percent}~\cite{Cheung2019}.
 
-    ASHRAE 55:2020 Standard utilizes a modified version of the PMV model [4] that calculates a Cooling Effect (CE) using the Standard Effective Temperature (SET) equation for air speeds higher than \qty{0.1}{\m\per\s}.
-    The value of CE is then subtracted from both the measured dry-bulb air temperature (tdb) and \ac{tr} and these results become the input into the PMV model.
-    Since the CE already accounts for convective heat losses from the person to the environment, the air speed used to calculate the PMV value is set to 0.1 m/s.
+    To overcome this problem, the \gls{55} Standard introduced a modified version of the \ac{pmv}, the \gls{pmv-ce}~\cite{ashrae552020}.
+    For \ac{v} higher than \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
+    The value of \ac{ce}, calculated using the \ac{set} equation, is subtracted from both the value of \ac{tdb} and \ac{tr} and these results become the input into the \ac{pmv} model.
+    The \ac{ce} already accounts for convective heat losses from the person to the environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+    The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) are not modified.
+    Consequently, given the same thermal environment and same occupants, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+    This may be a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national building codes worldwide.
+    In addition, over the course of the last \num{50} years, other formulations of the \ac{pmv} model have been formulated by other researchers.
+    To our knowledge, no other study compared the accuracy all the aforementioned  two aforementioned PMV models.
+    One study compared the accuracy of the \ac{pmv-ce} and the \ac{pmv} model.
+
+    This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} formulations developed over the course of the last years with self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.

### 2022-06-09 Manuscript/src/sections/methodology.tex
paper: minor edits

%! Author = sbbfti
%! Date = 10/06/2020

\section{Methodology}\label{sec:methodology}

    The \gls{db2} has over 107,000 independent recordings from more than 50 field investigations throughout the world and it is the largest of its kind.
    \Ac{tsv} were collected using a \gls{rhrn} survey and environmental parameters were measured and logged while participants completed their survey.
    We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
%    All the results were calculated using the function pmv_ppd [6] included in pythermalcomfort v1.10.0 a Python package for thermal comfort calculations.

\subsection{Data Processing and Cleaning}\label{subsec:data-processing-and-cleaning}
The Comfort DB contains data collected by researchers in cross-sectional field studies.
However, not all researchers logged all the six parameters required to calculate the PMV.\@
We, therefore, decided to remove all entries that did not have any of the following variables: tdb, relative humidity (RH), total clothing insulation (Icl), metabolic rate (M), and airspeed (V).
These constraints filtered out approximately \qty{28}{\percent} of the data in the Comfort DB.\@
We decided to keep only entries that did not have \ac{tr} data.
However, since this is a required input, we used the following methodology to fill in missing data.
If the \ac{to} was available \ac{tr} was calculated using the following equation: 2to – tdb.
Alternatively, we assumed \ac{tr} equal to \ac{tdb}.
We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \qty{44}{\percent} drop in the total entries in the database and there is evidence that in many conditions is possible to assume assumed \ac{tr} equal to tdb [7].
We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: https://github.com/FedericoTartarini/paper-cobee-2022.git so other users can test different assumptions.

Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points which did not meet the inclusion criteria of both Standards.
The limits we used to filter the data are as follows: \qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius}, \qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius}, \qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s}, \qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo}, 0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa, and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
We used these inputs to calculate the \ac{pmv} values.

The \ac{pmv} model is only applicable when its absolute value is lower than 2 \cite{Fanger1970, iso7730}.
However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with |\ac{tsv}| or |\ac{pmv}| $\leq$ than 3.5.
We used Python to analyse and plot the results.

### 2022-06-09 Manuscript/src/sections/results.tex
paper: minor edits

%! Author = sbbfti
%! Date = 10/06/2020

\section{Results}\label{sec:results}

A total of \var{tot_surveys} entries met the inclusion criteria listed in the Methodology Section.
The distribution of the six input variables used to calculate the PMV indices is depicted in Figure~\ref{fig:dist_input_data}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/dist_input_data}
    \caption{Distribution of the input variables used to calculate the \ac{pmv} values.
    The data are shown using boxen-plots (letter-value plots).
    They depict the median as the centerline and each successive level outward contains half of the remaining data.}
    \label{fig:dist_input_data}
\end{figure}

### 2022-06-09 Manuscript/src/sections/titleAbstract.tex
paper: minor edits

@@
 
-    \title{Title of the paper}
+    \title{PMV formulations: which one is the most accurate?}
 
@@
     \begin{abstract}
-        The \ac{pmv} model developed by Fanger predicts thermal sensation.
-    %    It is used in its original formulation in the \gls{iso}, while the \gls{ashrae} uses a modified version \acs{pmv-ce} to better account for air speeds.
-        We compared their results against XXX thermal sensation votes collected in buildings.
+        The \ac{pmv} model was originally developed by Fanger to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
+        His model has seen widespread adoption and it is currently used in the \gls{7730}.
+        However, to compensate for the model low prediction accuracy several other formulations have been proposed, one of which is currently used in the \gls{55}.
+        This is a source of confusion for many practitioners and researchers who find it difficult to settle on which formulation they should use.
+        We aim to solve this problem by comparing the accuracy of the main \ac{pmv} formulations against \var{tot_surveys} thermal sensation votes collected in buildings.
+        % We found that for V>0.1m/s the \gls{pmv-ce} is significantly more accurate (48%) than the PMV (42%) in predicting thermal neutrality.
+        % In 95% of the cases, the error was less than one point.
+        % The PMV and PMVCE accuracy were 35% and 36% for |PMV|≤2 and 43% and 44% for |PMV|≤1, respectively.
+        % Neither PMV model could accurately predict ‘cold’, ‘cool’, ‘warm’, or ‘hot’ thermal sensations.
+        % Consequently, the PMV should only be used in conditions close to neutrality within the range |PMV|≤1.
+        % We open-sourced the source code.
     \end{abstract}

### 2022-06-09 main.py

@@ -24,7 +24,7 @@ def save_var_latex(key, value):
 
     dict_var = {}
 
-    file_path = "Manuscript/Variables/results.dat"
+    file_path = "Manuscript/src/mydata.dat"
 
     try:
         with open(file_path, newline="") as file:
@@ -216,7 +216,7 @@ def bar_chart(data, ind="tsv", show_per=True, figletter=False):
     if figletter:
         plt.gcf().text(0.05, 0.95, f"{figletter})", weight="bold")
 
-    plt.savefig(f"./Manuscript/Figures/bar_plot_{ind}_Vmin_{data.vel.min()}.png", dpi=300)
+    plt.savefig(f"./Manuscript/src/figures/bar_plot_{ind}_Vmin_{data.vel.min()}.png", dpi=300)
 
 
 def legend_pmv():
@@ -351,7 +351,7 @@ def scatter_plot(data, ind="tsv", x_jitter=0):
         axs[ix].grid(axis="x")
 
     plt.tight_layout()
-    plt.savefig("./Manuscript/Figures/scatter_tsv_pmv.png", dpi=300)
+    plt.savefig("./Manuscript/src/figures/scatter_tsv_pmv.png", dpi=300)
 
 
 def plot_error_prediction(data):
@@ -478,11 +478,11 @@ def plot_error_prediction(data):
             )
 
     plt.savefig(
-        f"./Manuscript/Figures/prediction_error_Vmin_{data.vel.min()}.png", dpi=300
+        f"./Manuscript/src/figures/prediction_error_Vmin_{data.vel.min()}.png", dpi=300
     )
 
 
-def plot_distribution_variable():
+def plot_distribution_variable(data=df):
     f, axs = plt.subplots(1, 6, constrained_layout=True, figsize=(8, 3))
 
     for ix, var in enumerate(["ta", "tr", "vel", "clo", "met", "rh"]):
@@ -499,7 +499,7 @@ def plot_distribution_variable():
         if var == "clo":
             axs[ix].set(yticks=(np.arange(0, 1.8, 0.3)))
     sns.despine(bottom=True, left=True)
-    plt.savefig("./Manuscript/Figures/dist_input_data.png", dpi=300)
+    plt.savefig("./Manuscript/src/figures/dist_input_data.png", dpi=300)
     plt.show()
 
 
@@ -553,8 +553,8 @@ if __name__ == "__main__":
     # calculate rounded values
     df = calculate_new_indices(df_=df)
 
[…]

### 2022-06-15 Manuscript/src/sections/introduction.tex
paper: working on first draft

@@
-%! Author = sbbfti
-%! Date = 10/06/2020
-
 \section{Introduction}\label{sec:introduction}
@@
 The \ac{pmv} model has been widely adopted worldwide both among researchers and practitioners.
-    The \ac{pmv} is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
-    However, Cheung et al. (2019) demonstrated that the accuracy of the \ac{pmv} in predicting self-reported thermal sensation of people is only \qty{34}{\percent}~\cite{Cheung2019}.
+It is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
+However, it has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019}.
 
-    To overcome this problem, the \gls{55} Standard introduced a modified version of the \ac{pmv}, the \gls{pmv-ce}~\cite{ashrae552020}.
+To overcome this problem, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \gls{pmv-ce}~\cite{ashrae552020}.
 For \ac{v} higher than \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-    The value of \ac{ce}, calculated using the \ac{set} equation, is subtracted from both the value of \ac{tdb} and \ac{tr} and these results become the input into the \ac{pmv} model.
-    The \ac{ce} already accounts for convective heat losses from the person to the environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
-    The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) are not modified.
-    Consequently, given the same thermal environment and same occupants, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
-    This may be a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national building codes worldwide.
-    In addition, over the course of the last \num{50} years, other formulations of the \ac{pmv} model have been formulated by other researchers.
+The value of \ac{ce}, which is calculated using the \ac{set} equation, is subtracted from both the value of \ac{tdb} and \ac{tr}.
+The resulting values become the new inputs into the \ac{pmv} model.
+The \ac{ce} accounts for convective heat losses from the person to the environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+While the other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
+Consequently, for a given thermal environment, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+This is a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national and international building codes worldwide.
+To further complicate this issue, over the course of the last \num{50} years, several other formulations of the \ac{pmv} model have been proposed.
+Among the most notable formulations there are the:
+\begin{enumerate}
+    \item \gls{epmv}
+    \item \gls{apmv}
+    \item SET \ac{pmv}
+    \item Gagge \ac{pmv}
+    \item Marcel \ac{pmv}
+\end{enumerate}
+
+In 2002, Fanger, P. O. and Toftum, J. developed the \gls{epmv} an extension of the \ac{pmv}.
+The \gls{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
+The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
+They, consequently, propose to reduce the metabolic rate used to calculate the \ac{pmv} by \qty{6.7}{\percent} for every scale unit of \ac{pmv} above neutral.
+The resulting \ac{pmv} value is multiplied by the \ac{ef}.
+The \ac{ef} for non-air-conditioned buildings can be determined by the length of warm weather throughout the year and whether such buildings may be compared to many others in the region that are air-conditioned.
+If the temperature is hot all year or most of the year and there are no or few other air-conditioned buildings, \ac{ef} may be \num{0.5}, but if there are many other air-conditioned buildings, \ac{ef} may be \num{0.7}.
+For non-air-conditioned buildings in locations where the temperature is warm solely during the summer and no or few buildings have air-conditioning \ac{ef} is \numrange{0.7}{0.8}, whereas it is \numrange{0.8}{.9} in areas where there are numerous air-conditioned buildings.
+In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
+To better account for occupants adaptation Yao, R. et al. (2009) developed the Adaptive Predicted Mean
+ Vote Model~\gls{apmv}.
+They used a ``Black Box'' model to determine an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
+The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
+
 To our knowledge, no other study compared the accuracy all the aforementioned two aforementioned PMV models.
-    One study compared the accuracy of the \ac{pmv-ce} and the \ac{pmv} model.
+One study compared the accuracy of the \gls{pmv-ce} and the \ac{pmv} model.
 

### 2022-06-15 Manuscript/src/sections/titleAbstract.tex
paper: working on first draft

@@
         The \ac{pmv} model was originally developed by Fanger to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
-        His model has seen widespread adoption and it is currently used in the \gls{7730}.
-        However, to compensate for the model low prediction accuracy several other formulations have been proposed, one of which is currently used in the \gls{55}.
-        This is a source of confusion for many practitioners and researchers who find it difficult to settle on which formulation they should use.
+        It has seen widespread adoption and it is currently used in the \gls{7730}.
+        However, to compensate for the model low prediction accuracy, several other formulations have been proposed, one of which is currently used in the \gls{55}.
+        This is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
         We aim to solve this problem by comparing the accuracy of the main \ac{pmv} formulations against \var{tot_surveys} thermal sensation votes collected in buildings.

### 2022-06-15 main.py

@@ -1,4 +1,5 @@
 import matplotlib as mpl
+
 mpl.use("Qt5Agg")  # or can use 'TkAgg', whatever you have/prefer
 
 from matplotlib.colors import LinearSegmentedColormap
@@ -12,6 +13,7 @@ from scipy import stats
 from sklearn.metrics import r2_score, mean_absolute_error
 import psychrolib
 import matplotlib.patches as patches
+from pythermalcomfort.models import two_nodes, athb
 
 warnings.filterwarnings("ignore")
 
@@ -19,7 +21,7 @@ warnings.filterwarnings("ignore")
 psychrolib.SetUnitSystem(psychrolib.SI)
 
 
-def save_var_latex(key, value):
+def save_var_latex(key, value, units=False):
     import csv
 
     dict_var = {}
@@ -34,14 +36,19 @@ def save_var_latex(key, value):
     except FileNotFoundError:
         pass
 
-    dict_var[key] = value
+    if units:
+        dict_var[key] = f"\\qty{{{value}}}{{{units}}}"
+    else:
+        dict_var[key] = f"\\num{{{value}}}"
 
     with open(file_path, "w") as f:
         for key in dict_var.keys():
             f.write(f"{key},{dict_var[key]}\n")
 
 
-def filter_data(df_):
+def filtering_processing(df_):
+
+    save_var_latex("all_entries_db", df.shape[0])
 
     pa_arr = []
     for i, row in df_.iterrows():
@@ -51,27 +58,63 @@ def filter_data(df_):
 
     # remove entries outside the Standards' applicability limits
     for key in applicability_limits.keys():
+        if "pmv" in key:
+            continue
         df_ = df_[
             (df_[key] >= applicability_limits[key][0])
             & (df_[key] <= applicability_limits[key][1])
         ]
 
-    return df_
+    two_nodes_results = two_nodes(
+        tdb=df_.ta,
+        tr=df_.tr,
+        v=df_.vel,
+        rh=df_.rh,
+        met=df_.met,
+        clo=df_.clo,
+    )
 
+    df_["pmv_gagge"] = two_nodes_results["pmv_gagge"]
+    df_["pmv_set"] = two_nodes_results["pmv_set"]
 
-def calculate_new_indices(df_):
+    df_["athb"] = athb(
+        tdb=df_.ta,
+        tr=df_.tr,
+        vr=df_.vel,
+        rh=df_.rh,
+        met=df_.met,
+        t_running_mean=df_.t_mot_isd,
[…]

### 2022-06-20 Manuscript/src/myglossary.tex
paper: working on first draft
paper: minor edits

@@
 
+\newglossaryentry{epmv}
+{
+name={ePMV},
+description={Adjusted Predicted Mean Votes with Expectancy Factor}
+}
+
+\newglossaryentry{apmv}
+{
+name={aPMV},
+description={Adaptive Predicted Mean Vote}
+}
+
+\newglossaryentry{pmvs}
+{
+name={PMV$_{SET}$},
+description={PMV calculated using the SET temperature inplace of $t_{db}$}
+}
+
+\newglossaryentry{pmvg}
+{
+name={PMV$_{Gagge}$},
+description={PMV calculated using the two-node heat balance model}
+}
+
 \newglossaryentry{db2}

### 2022-06-20 Manuscript/src/sections/introduction.tex
paper: minor edits

@@
 To overcome this problem, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \gls{pmv-ce}~\cite{ashrae552020}.
+The process used by the Standard is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
 For \ac{v} higher than \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
@@
 Consequently, for a given thermal environment, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+
+\begin{figure}[!htb]
+    \centering
+
+    \begin{tikzpicture}[node distance=1cm]
+        \node (start) [startstop] {\ac{pmv} calculation as per \gls{55}};
+        \node (dec1) [decision, below of=start, yshift=-.25cm] {\ac{met} $> \qty{1}{met}$ };
+        \node (pro1a) [process, right of=dec1, xshift=4cm] {\ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$};
+        \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\ac{met} $> \qty{1.2}{met}$ };
+        \node (pro2a) [process, right of=dec2, xshift=4cm] {\ac{clo} = \ac{clo}$(0.6 + 0.4/$\ac{met}$)$};
+        \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\ac{v} $> \qty{0.1}{\m\per\s}$ };
+        \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\ac{pmv}(\ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{met}, \ac{clo})};
+        \node (pro3) [process_large, below of=dec3, yshift=-.25cm, xshift=4cm] {Calculate \ac{ce}};
+        \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\ac{pmv}(\ac{tdb} - \ac{ce}, \ac{tr} - \ac{ce}, \ac{rh}, \ac{v} = \qty{0.1}{\m\per\s}, \ac{met}, \ac{clo})};
+
+        \draw [arrow] (start) -- (dec1);
+        \draw [arrow] (dec1) -- node[above,pos=0.3] {Yes} (pro1a);
+        \draw [arrow] (dec1) -- node[right,pos=0.3] {No} (dec2);
+        \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
+        \draw [arrow] (dec2) -- node[above,pos=0.3] {Yes} (pro2a);
+        \draw [arrow] (dec2) -- node[right,pos=0.3] {No} (dec3);
+        \draw [arrow] (pro2a) -- ($(pro2a.south)+(0,-0.5)$) -| (dec3);
+        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above,pos=0.3] {No} (end1);
+        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above,pos=0.3] {Yes} (pro3);
+        \draw [arrow] (pro3) -- (end2);
+    \end{tikzpicture}\label{fig:flowchart_pmv_ce}
+\end{figure}
+
 This is a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national and international building codes worldwide.
 To further complicate this issue, over the course of the last \num{50} years, several other formulations of the \ac{pmv} model have been proposed.
-Among the most notable formulations there are the:
-\begin{enumerate}
-    \item \gls{epmv}
-    \item \gls{apmv}
-    \item SET \ac{pmv}
-    \item Gagge \ac{pmv}
-    \item Marcel \ac{pmv}
-\end{enumerate}
+Among the most notable formulations there are the: \gls{pmvs}, \gls{pmvg}, \gls{epmv}, \gls{apmv}, and \ac{athb}.
+
+% todo pmv SET
+% todo pmv Gagge
 
@@
 In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
-To better account for occupants adaptation Yao, R. et al. (2009) developed the Adaptive Predicted Mean
- Vote Model~\gls{apmv}.
-They used a ``Black Box'' model to determine an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
+This arbitrary and non-scientific definition of the expectancy factor make the \gls{epmv} difficult to calculate and non-generalizable.
+Consequently, the \gls{epmv} has not gained any widespread adoption.
+
+Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
+They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
 The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
+This approach cannot be, therefore, generalizable and similarly to the \gls{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
 
-To our knowledge, no other study compared the accuracy all the aforementioned two aforementioned PMV models.
-One study compared the accuracy of the \gls{pmv-ce} and the \ac{pmv} model.
+% todo describe marcel's work
+
+To our knowledge, no other study compared the accuracy all the aforementioned aforementioned PMV models.
+\mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE ASHRAE Global Thermal Comfort Database II.
+While \textbf{COBEE paper} compared the accuracy of the \gls{pmv-ce} and the \ac{pmv} models.
 

### 2022-06-20 Manuscript/src/sections/methodology.tex
paper: working on first draft
paper: minor edits

@@
 
-    The \gls{db2} has over 107,000 independent recordings from more than 50 field investigations throughout the world and it is the largest of its kind.
+The \gls{db2} has \var{all_entries_db} independent recordings from various field investigations throughout the world and it is the largest of its kind.
 \Ac{tsv} were collected using a \gls{rhrn} survey and environmental parameters were measured and logged while participants completed their survey.
-    We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
+The Comfort DB contains data collected by researchers in cross-sectional field studies that have been published in peer reviewed journals.
+%We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
 %    All the results were calculated using the function pmv_ppd [6] included in pythermalcomfort v1.10.0 a Python package for thermal comfort calculations.
@@
 \subsection{Data Processing and Cleaning}\label{subsec:data-processing-and-cleaning}
-The Comfort DB contains data collected by researchers in cross-sectional field studies.
-However, not all researchers logged all the six parameters required to calculate the PMV.\@
-We, therefore, decided to remove all entries that did not have any of the following variables: tdb, relative humidity (RH), total clothing insulation (Icl), metabolic rate (M), and airspeed (V).
-These constraints filtered out approximately \qty{28}{\percent} of the data in the Comfort DB.\@
-We decided to keep only entries that did not have \ac{tr} data.
-However, since this is a required input, we used the following methodology to fill in missing data.
-If the \ac{to} was available \ac{tr} was calculated using the following equation: 2to – tdb.
-Alternatively, we assumed \ac{tr} equal to \ac{tdb}.
+Not all researchers logged all the six parameters required to calculate the PMV.\@
+We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, and \ac{met}.
+These constraints filtered out approximately \qty{28}{\percent} of the data in the Comfort DB.\@  % todo check the above value
+We decided to keep only entries that did not have \ac{tr} data but instead had the \ac{to} value.
+Since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: 2\ac{to} – \ac{tdb}.
+Alternatively, if not value of \ac{tr} was available we assumed \ac{tr} equal to \ac{tdb}.
 We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \qty{44}{\percent} drop in the total entries in the database and there is evidence that in many conditions is possible to assume assumed \ac{tr} equal to tdb [7].
-We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: https://github.com/FedericoTartarini/paper-cobee-2022.git so other users can test different assumptions.
+% todo check the above value and fix citation
 
 Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points which did not meet the inclusion criteria of both Standards.
+The rationale is that the model accuracy should only be tested in within its applicability limits.
 The limits we used to filter the data are as follows: \qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius}, \qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius}, \qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s}, \qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo}, 0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa, and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
 We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
-We used these inputs to calculate the \ac{pmv} values.
+We used these inputs to calculate the \ac{pmv} values.  % todo check with marcel if this is required in his model too
 
 The \ac{pmv} model is only applicable when its absolute value is lower than 2 \cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with |\ac{tsv}| or |\ac{pmv}| $\leq$ than 3.5.
+However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with |\ac{tsv}| or |\ac{pmv}| $\leq$ than \num{3.5}.
 We used Python to analyse and plot the results.
+
+We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \textbf{provide URL} so other users can test different assumptions.

### 2022-06-20 main.py

@@ -14,6 +14,13 @@ from sklearn.metrics import r2_score, mean_absolute_error
 import psychrolib
 import matplotlib.patches as patches
 from pythermalcomfort.models import two_nodes, athb
+from sklearn.linear_model import LogisticRegression
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.special import expit, logit
+from statsmodels.tools import add_constant
+import statsmodels.api as sm
+import pandas as pd
 
 warnings.filterwarnings("ignore")
 
@@ -46,9 +53,14 @@ def save_var_latex(key, value, units=False):
             f.write(f"{key},{dict_var[key]}\n")
 
 
-def filtering_processing(df_):
+def importing_filtering_processing(load_preprocessed=False):
 
-    save_var_latex("all_entries_db", df.shape[0])
+    df_ = pd.read_csv(r"./Data/db_measurements_v2.1.0.csv.gz", compression="gzip")
+
+    save_var_latex("all_entries_db", df_.shape[0])
+
+    if load_preprocessed:
+        return pd.read_pickle(r"./Data/db_analysis.pkl.gz", compression="gzip")
 
     pa_arr = []
     for i, row in df_.iterrows():
@@ -75,7 +87,13 @@ def filtering_processing(df_):
     )
 
     df_["pmv_gagge"] = two_nodes_results["pmv_gagge"]
+    df_["hb_gagge"] = two_nodes_results["pmv_gagge"] / (
+        0.303 * np.exp(-0.036 * df_["met"] * 58.2) + 0.028
+    )
     df_["pmv_set"] = two_nodes_results["pmv_set"]
+    df_["hb_set"] = two_nodes_results["pmv_set"] / (
+        0.303 * np.exp(-0.036 * df_["met"] * 58.2) + 0.028
+    )
 
     df_["athb"] = athb(
         tdb=df_.ta,
@@ -93,17 +111,22 @@ def filtering_processing(df_):
                 & (df_[key] <= applicability_limits[key][1])
             ]
 
-    df_["pmv_round"] = df_["pmv"].round()
-    df_["pmv_ce_round"] = df_["pmv_ce"].round()
+    # calculate rounded variables and differences
+    for model in models_to_test:
+        rounded_col = f"{model}_round"
+        diff_col = f"diff_ts_{model}"
+        df_[rounded_col] = df_[model].round()
+        df_[diff_col] = df_[["thermal_sensation", model]].diff(axis=1)[model]
[…]

### 2022-07-04 Manuscript/src/sections/introduction.tex
paper: revised intro, methodology

@@
 To overcome this problem, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \gls{pmv-ce}~\cite{ashrae552020}.
-The process used by the Standard is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
+The process used to calculate the \gls{pmv-ce} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
 For \ac{v} higher than \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
@@
 This is a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national and international building codes worldwide.
-To further complicate this issue, over the course of the last \num{50} years, several other formulations of the \ac{pmv} model have been proposed.
+To further complicate this issue, over the course of the last decades, several other formulations of the \ac{pmv} model have been proposed.
 Among the most notable formulations there are the: \gls{pmvs}, \gls{pmvg}, \gls{epmv}, \gls{apmv}, and \ac{athb}.
 
-% todo pmv SET
-% todo pmv Gagge
+% todo describe pmv SET
+% todo describe pmv Gagge
 
@@
 
-To our knowledge, no other study compared the accuracy all the aforementioned aforementioned PMV models.
-\mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE ASHRAE Global Thermal Comfort Database II.
-While \textbf{COBEE paper} compared the accuracy of the \gls{pmv-ce} and the \ac{pmv} models.
+To our knowledge, no other study compared the accuracy of all the aforementioned PMV models, aside from a conference paper we previously published that stands at the basis of this work.
+In particular, no previous study compared the accuracy of the \ac{pmv} and \ac{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
+Previous works were limited to the determination of the accuracy of the \ac{pmv} model.
+For example, some notable works include, the work from \mycite{Doherty1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
+They concluded that the \ac{pmv} model is accurate for simulations of resting subjects but its accuracy decreases as a function of metabolic rate.
+Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
+They used the ASHRAE Thermal Comfort database v1.0 and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
+Humphreys and Nicol also determined the validity of the \ac{pmv} for predicting comfort votes collected in field studies~\cite{Humphreys2002}.
+They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
+The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
+While, \mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
 
-This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} formulations developed over the course of the last years with self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.
+This study aims to specifically fill this knowledge gap by comparing the results of the major \ac{pmv} formulations developed over the course of the last decades with self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.
+Our results have significant and profound implications for both researchers and practitioners since it provides quantitative evidence on which formulation of the \ac{pmv} model is more accurate.

### 2022-07-04 Manuscript/src/sections/methodology.tex
paper: revised intro, methodology

@@
 
-The \gls{db2} has \var{all_entries_db} independent recordings from various field investigations throughout the world and it is the largest of its kind.
+The \gls{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world and it is the largest of its kind.
 \Ac{tsv} were collected using a \gls{rhrn} survey and environmental parameters were measured and logged while participants completed their survey.
 The Comfort DB contains data collected by researchers in cross-sectional field studies that have been published in peer reviewed journals.
-%We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
+% We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
 %    All the results were calculated using the function pmv_ppd [6] included in pythermalcomfort v1.10.0 a Python package for thermal comfort calculations.
 
-\subsection{Data Processing and Cleaning}\label{subsec:data-processing-and-cleaning}
+\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
 Not all researchers logged all the six parameters required to calculate the PMV.\@
-We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, and \ac{met}.
-These constraints filtered out approximately \qty{28}{\percent} of the data in the Comfort DB.\@  % todo check the above value
-We decided to keep only entries that did not have \ac{tr} data but instead had the \ac{to} value.
+We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
+These constraints filtered out approximately \var{entries_db_valid} of the data in the Comfort DB.\@
+We decided to keep entries that did not have \ac{tr}.
 Since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: 2\ac{to} – \ac{tdb}.
-Alternatively, if not value of \ac{tr} was available we assumed \ac{tr} equal to \ac{tdb}.
-We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \qty{44}{\percent} drop in the total entries in the database and there is evidence that in many conditions is possible to assume assumed \ac{tr} equal to tdb [7].
-% todo check the above value and fix citation
+Alternatively, if \ac{to} was not available we assumed \ac{tr} equal to \ac{tdb}.
+We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
+There is also evidence that in many conditions is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
 
@@
 The rationale is that the model accuracy should only be tested in within its applicability limits.
-The limits we used to filter the data are as follows: \qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius}, \qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius}, \qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s}, \qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo}, 0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa, and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
+The limits we used are as follows: \qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius}, \qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius}, \qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s}, \qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo}, 0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa, and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
 We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
@@
 
-The \ac{pmv} model is only applicable when its absolute value is lower than 2 \cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with |\ac{tsv}| or |\ac{pmv}| $\leq$ than \num{3.5}.
+Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
+However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with $|$\ac{tsv}$|$ or $|$\ac{pmv}$|$ $\leq$ than \num{3.5}.
+
+The values of \ac{pmv} and \gls{pmv-ce} are available inside the \gls{db2}.
+We calculated the values of \gls{pmvs}, \gls{pmvg}, and \ac{athb} using pythermalcomfort v2.5.1~\cite{Tartarini2020a}.
 We used Python to analyse and plot the results.
@@
 We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \textbf{provide URL} so other users can test different assumptions.
+
+\subsection{Model Validation}\label{subsec:model-validation}
+The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants sharing the same environment.
+While the \gls{db2} contains data from a sufficiently wide population no information is available on where exactly these data were collected.
+We only have access to building level data and no information about weather a group of occupants was sharing the same thermal environment.
+We, therefore, decided to first group the participants responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} model.
+This approach is useful to perform an overall assessment of the accuracy of the model, but it introduces an error due to the rounding of the \ac{pmv} models output and \ac{tsv} values.
+The latter were sometimes collected using a continuous scale and not a discrete one, in other words they could take any value in the range of \numrange{-3.0}{3.0}.
+To remove this rounding error, we then went on subtracted the \ac{tsv} value from the respective \ac{pmv} value.
+These deltas quantify the success of the model in predicting \ac{tsv}.
+However, on their own are a low precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
+These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{rh}).
+If the \ac{pmv} formulation is bias-free, the distribution of any batch derived from these differences would have a mean value that was little different from zero.
+The standard deviation would reflect the combined effect of the people's individual differences as well as any errors in the \ac{pmv} formulation~\cite{Humphreys2002}.
+According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned deltas are within \numrange{-0.25}{0.25}.
+% todo why not 0.5 since the delta between a PMV of 2.5 and a self-reported vote of either 2 or 3 (closest points to 2.5) is 0.5?
+
+\subsection{Sources of Error}\label{subsec:sources-of-error}
+One minor but not negligible source of error is the use of a discrete scale to assess thermal sensation.
+Most of the \ac{tsv} in the \gls{db2} were collected using a discrete scale while the \ac{pmv} output is continuous.
+For example, even if the model was 100~\% accurate and the estimated \ac{pmv} values was \num{2.5} the person could only report to be `warm (+2)' or `hot (+3)'.
+
+Measurement errors are also inevitably present in the dataset.
+These will add a random error to the model output.
+
+Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) which interacts with the surrounding environment.
+Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.
+Finally, it should be noted that the \ac{pmv} model has been developed considering based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermo regulatory system is always actively engaged to ensure a stable core temperature.
+However, as \mycite{Humphreys2002} note, these approximations and errors present in the \ac{pmv} model, if the model is accurate, should combine to produce a quasi-random error in the prediction.
+This would increase the standard deviation of the delta between the \ac{tsv} and \ac{pmv}, but if the model is sound, they should produce a negligible bias~\cite{Humphreys2002}.
+On the other hand, if the model is neither accurate nor representative of the real system a significant bias will arise.

### 2022-07-04 Manuscript/src/sections/results.tex
paper: revised intro, methodology

@@
 
-A total of \var{tot_surveys} entries met the inclusion criteria listed in the Methodology Section.
+\subsection{Comfort DB overview}\label{subsec:comfort-db-overview}
+
+A total of \var{entries_db_used} entries met the inclusion criteria listed in the Methodology Section.
 The distribution of the six input variables used to calculate the PMV indices is depicted in Figure~\ref{fig:dist_input_data}.
@@
 \end{figure}
+
+Figure~\ref{fig:dist_other_data} shows the distribution of the age, height, weight, and running mean outdoor temperature grouped by sex.
+Less than half (\num{23300}) of the total entries had information about the participant's sex.
+Out of those data were almost equally distributed among male \qty{52}{\percent} and females.
+Ages are not normally distributed, the same is true for the running mean outdoor temperature.
+About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old.
+While \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
+Mens were significantly heavier and taller than females.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/dist_other_data}
+    \caption{Distribution of age, height, weight, and running mean outdoor temperature.
+    The data are grouped by sex.}
+    \label{fig:dist_other_data}
+\end{figure}
+
+Figure~\ref{fig:bar_plot_tp_by_ts} shows on the left the percentage of thermal preference votes grouped by each \ac{tsv}.
+The left Figure also reports the number of points for each \ac{tsv}.
+Only \var{entries_with_tp} that had information about the thermal preference.
+Thus, on the right we show a bar chart depicting the distribution of all the \ac{tsv} votes, we have used in the analysis.
+The \gls{db2} in respect to the \ac{tsv} is unbalanced.
+Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
+While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
+In the \gls{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
+However, in thermal comfort research it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
+This finding challenges the above assumption.
+Moreover, these results provide additional evidence that the use of the thermal preference scale is better suited to determine how people perceive their thermal environment.
+The thermal sensation vote clearly allows to determine whether participants would take action or not to modify their current thermal environment.
+The \ac{tsv}, on the contrary, does not allow to determine which actions should be taken to increase thermal comfort conditions for occupants.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
+    \caption{The left Figure shows the percentage of thermal preference votes for each thermal sensation vote.
+    The numbers on the right side of each bar show the number of points for each \ac{tsv}.
+    The right figure shows the total number of data points grouped by \ac{tsv}.
+    Each bin in the Figure of the right has more data points than in the Figure on the left since information about the occupants thermal preference were not always available.}
+    \label{fig:bar_plot_tp_by_ts}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
+    \caption{}
+    \label{fig:bubble_models_vs_tsv}
+\end{figure}

### 2022-07-04 Manuscript/src/sections/titleAbstract.tex
paper: revised intro, methodology

@@
         The \ac{pmv} model was originally developed by Fanger to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
-        It has seen widespread adoption and it is currently used in the \gls{7730}.
-        However, to compensate for the model low prediction accuracy, several other formulations have been proposed, one of which is currently used in the \gls{55}.
-        This is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
-        We aim to solve this problem by comparing the accuracy of the main \ac{pmv} formulations against \var{tot_surveys} thermal sensation votes collected in buildings.
+        It has seen widespread adoption and in its original form it is currently used in the \gls{7730}.
+        However, the \ac{pmv} model has a low prediction accuracy and to compensate for that, several other formulations have been proposed, one of which is currently used in the \gls{55}.
+        Being the \gls{7730} and \gls{55} the most widely used Standards worldwide, this is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
+        We aim to solve this problem by comparing the accuracy of the main \ac{pmv} formulations against \var{entries_db_used} thermal sensation votes collected in buildings.
         % We found that for V>0.1m/s the \gls{pmv-ce} is significantly more accurate (48%) than the PMV (42%) in predicting thermal neutrality.

### 2022-07-04 main.py

@@ -17,6 +17,7 @@ from pythermalcomfort.models import two_nodes, athb
 from sklearn.linear_model import LogisticRegression
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
 from scipy.special import expit, logit
 from statsmodels.tools import add_constant
 import statsmodels.api as sm
@@ -26,6 +27,53 @@ warnings.filterwarnings("ignore")
 
 
 psychrolib.SetUnitSystem(psychrolib.SI)
+
+
+palette_tp = [
+    "#06A6EE",
+    "#31CAA8",
+    "#FF412C",
+]
+palette_tp = [
+    "#4F96FF",
+    "#60E693",
+    "#FF362B",
+]
+palette_tsv = [
+    "#2C45FE",
+    "#4F96FF",
+    "#6CDFFF",
+    "#60E693",
+    "#FFDF6B",
+    "#FFB36B",
+    "#FF362B",
+]
+
+# sns.palplot(palette_tsv)
+
+palette_primary = [
+    "#FFBA22",
+    "#FF2380",
+    "#6600AB",
+    "#FF7BB3",
+    "#003CAB",
+    "#83DFCB",
+    "#C20800",
+    "#C57FFF",
+    "#FF8D80",
+    "#6ACAF5",
+    "#006B46",
+    "#FF6756",
+    "#9F29FF",
+    "#8001D8",
+    "#FF4F99",
+    "#006BDB",
+    "#5AD5B9",
+    "#E8281E",
+    "#B254FF",
+    "#149678",
+    "#FFE146",
+]
 
 
 def save_var_latex(key, value, units=False):
@@ -57,7 +105,21 @@ def importing_filtering_processing(load_preprocessed=False):
 
     df_ = pd.read_csv(r"./Data/db_measurements_v2.1.0.csv.gz", compression="gzip")
 
-    save_var_latex("all_entries_db", df_.shape[0])
+    save_var_latex("entries_db_all", df_.shape[0])
+
+    # entries without ta, rh, v, clo, met
+    df_valid_input = df_.dropna(subset=["ta", "vel", "rh", "met", "clo"])
+    df_valid_input_no_tr = df_valid_input.dropna(subset=["tr"]).shape[0]
+    save_var_latex(
+        "entries_db_valid",
+        int(100 - df_valid_input.shape[0] / df_.shape[0] * 100),
+        "\\percent",
+    )
+    save_var_latex(
[…]

### 2022-07-06 main.py
feat: analysed data

@@ -22,6 +22,7 @@ from scipy.special import expit, logit
 from statsmodels.tools import add_constant
 import statsmodels.api as sm
 import pandas as pd
+from sklearn.metrics import f1_score
 
 warnings.filterwarnings("ignore")
 
@@ -365,48 +366,6 @@ def bar_chart(
     )
 
 
-def legend_pmv():
-    f, ax = plt.subplots()
-
-    colors = [
-        (33 / 255, 102 / 255, 172 / 255),
-        (103 / 255, 169 / 255, 207 / 255),
-        (209 / 255, 229 / 255, 240 / 255),
-        (153 / 255, 213 / 255, 148 / 255),
-        (253 / 255, 219 / 255, 199 / 255),
-        (239 / 255, 138 / 255, 98 / 255),
-        (178 / 255, 24 / 255, 43 / 255),
-    ]
-
-    ax.legend(
-        handles=[
-            patches.Patch(color=colors[ix], label=str(x))
-            for ix, x in enumerate(range(-3, 4))
-        ],
-        frameon=False,
-        mode="expand",
-        bbox_to_anchor=(0, 1.02, 1, 0.2),
-        loc="lower left",
-        ncol=7,
-    )
-    f.text(0.05, 0.95, "PMV = ", va="center", ha="left")
-    ax.grid(False)
-
-
-def distributions_pmv(v_lower=False):
-    # check difference in distribution for vel > 0.1
-    variables = ["pmv", "pmv_ce", "thermal_sensation"]
-    f, axs = plt.subplots(len(variables), 1, sharex=True)
-    if v_lower:
-        data = df[df["vel"] > 0.1]
-    else:
-        data = df.copy()
-    for ix, var in enumerate(variables):
-        sns.histplot(data=data, x=var, kde=True, stat="density", ax=axs[ix])
-        axs[ix].set(title=var)
-    plt.tight_layout()
-
-
 def scatter_plot_flip_x(data):
     f, axs = plt.subplots(1, 2, sharex=True, sharey=True, constrained_layout=True)
     sns.regplot(
@@ -692,7 +651,7 @@ def plot_bubble_models_vs_tsv():
     f, axs = plt.subplots(1, 5, sharex=True, sharey=True, constrained_layout=True)
     axs = axs.flatten()
 
-    for ix, model in enumerate(["pmv", "pmv_ce", "pmv_set", "pmv_gagge", "athb"]):
[…]

### 2023-04-18 Manuscript/src/sections/introduction.tex
feat: reviewed intro, specified that we will only analyse PMV and PMV_CE, cited Yao 2022

@@
 In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
+% todo also mention the other standards that use the PMV The PMV index has been widely used in national and international standards and guidance ASHRAE 55 [4], EN 15251 (new version EN 16798 [5]) [6], ISO 7730 [1], CIBSE Guide A [7] and GB/T 50785 [110] as  shown in  Table 5.
+This sentence is from Yao 2022.
 The \ac{pmv} model has been widely adopted worldwide both among researchers and practitioners.
 It is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
-However, it has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019}.
+A search within article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
+Out of those 1082 are scientific peer-review articles published in the research area of engineering, environmental science, social sciences, and energy.
+This highlights the extensive adaption and use of this model among the scientific community.
+
+To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model despite the fact that several research papers have highlighted that the \ac{pmv} has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019}.
+% todo cite more papers
 
 To overcome this problem, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \gls{pmv-ce}~\cite{ashrae552020}.
-The process used to calculate the \gls{pmv-ce} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
+The process used to calculate the \gls{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
 For \ac{v} higher than \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-The value of \ac{ce}, which is calculated using the \ac{set} equation, is subtracted from both the value of \ac{tdb} and \ac{tr}.
+The value of \ac{ce}, which is calculated using the \ac{set} equation, is then subtracted from both the value of \ac{tdb} and \ac{tr}.
 The resulting values become the new inputs into the \ac{pmv} model.
@@
 Consequently, for a given thermal environment, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
+This method, however, has its own limitation and throughout the years the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552020}.
 
@@
     \centering
+    \label{fig:flowchart_pmv_ce}
 
@@
         \draw [arrow] (pro3) -- (end2);
-    \end{tikzpicture}\label{fig:flowchart_pmv_ce}
+    \end{tikzpicture}
 \end{figure}
@@
 This is a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national and international building codes worldwide.
-To further complicate this issue, over the course of the last decades, several other formulations of the \ac{pmv} model have been proposed.
+
+\subsection{Other PMV formulations}\label{subsec:other-pmv-formulations}
+To further complicate this issue, several other formulations of the \ac{pmv} model have been proposed worldwide.
 Among the most notable formulations there are the: \gls{pmvs}, \gls{pmvg}, \gls{epmv}, \gls{apmv}, and \ac{athb}.
+\mycite{Yao2022} provides a comprehensive review of the difference \ac{pmv} formulations and the adaptive heat balance models based on \ac{pmv}.
+Hence, we invite our readers to read \mycite{Yao2022} paper to learn more about all the different variations that have been proposed.
 
-% todo describe pmv SET
-% todo describe pmv Gagge
+%% todo describe pmv SET
+%% todo describe pmv Gagge
+%
+%In 2002, Fanger, P. O. and Toftum, J. developed the \gls{epmv} an extension of the \ac{pmv}.
+%The \gls{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
+%The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
+%They, consequently, propose to reduce the metabolic rate used to calculate the \ac{pmv} by \qty{6.7}{\percent} for every scale unit of \ac{pmv} above neutral.
+%The resulting \ac{pmv} value is multiplied by the \ac{ef}.
+%The \ac{ef} for non-air-conditioned buildings can be determined by the length of warm weather throughout the year and whether such buildings may be compared to many others in the region that are air-conditioned.
+%If the temperature is hot all year or most of the year and there are no or few other air-conditioned buildings, \ac{ef} may be \num{0.5}, but if there are many other air-conditioned buildings, \ac{ef} may be \num{0.7}.
+%For non-air-conditioned buildings in locations where the temperature is warm solely during the summer and no or few buildings have air-conditioning \ac{ef} is \numrange{0.7}{0.8}, whereas it is \numrange{0.8}{.9} in areas where there are numerous air-conditioned buildings.
+%In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
+%This arbitrary and non-scientific definition of the expectancy factor make the \gls{epmv} difficult to calculate and non-generalizable.
+%Consequently, the \gls{epmv} has not gained any widespread adoption.
+%
+%Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
+%They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
+%The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
+%This approach cannot be, therefore, generalizable and similarly to the \gls{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
+%
+%% todo describe marcel's work
 
-In 2002, Fanger, P. O. and Toftum, J. developed the \gls{epmv} an extension of the \ac{pmv}.
-The \gls{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
-The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
-They, consequently, propose to reduce the metabolic rate used to calculate the \ac{pmv} by \qty{6.7}{\percent} for every scale unit of \ac{pmv} above neutral.
-The resulting \ac{pmv} value is multiplied by the \ac{ef}.
-The \ac{ef} for non-air-conditioned buildings can be determined by the length of warm weather throughout the year and whether such buildings may be compared to many others in the region that are air-conditioned.
-If the temperature is hot all year or most of the year and there are no or few other air-conditioned buildings, \ac{ef} may be \num{0.5}, but if there are many other air-conditioned buildings, \ac{ef} may be \num{0.7}.
-For non-air-conditioned buildings in locations where the temperature is warm solely during the summer and no or few buildings have air-conditioning \ac{ef} is \numrange{0.7}{0.8}, whereas it is \numrange{0.8}{.9} in areas where there are numerous air-conditioned buildings.
-In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
-This arbitrary and non-scientific definition of the expectancy factor make the \gls{epmv} difficult to calculate and non-generalizable.
-Consequently, the \gls{epmv} has not gained any widespread adoption.
+\subsection{Comparision of PMV formulations}\label{subsec:comparision-of-pmv-formulations}
+To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
 
-Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
-They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
-The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
-This approach cannot be, therefore, generalizable and similarly to the \gls{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
-
-% todo describe marcel's work
-
-To our knowledge, no other study compared the accuracy of all the aforementioned PMV models, aside from a conference paper we previously published that stands at the basis of this work.
-In particular, no previous study compared the accuracy of the \ac{pmv} and \ac{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
-Previous works were limited to the determination of the accuracy of the \ac{pmv} model.
-For example, some notable works include, the work from \mycite{Doherty1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
+Some notable works include, the work from \mycite{Doherty1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects but its accuracy decreases as a function of metabolic rate.
@@
 While, \mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
+Finally, while \mycite{Yao2022} compare the \ac{pmv} and \ac{pmv-ce} models, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
+In particular, they focus significantly how the models perform in different climates and when applied to people from different world regions, but for example they fail to report how the \ac{pmv} accuracy varies as a function of \ac{tsv} and how the bias of the model varies as a function of different environmental parameters.
 
-This study aims to specifically fill this knowledge gap by comparing the results of the major \ac{pmv} formulations developed over the course of the last decades with self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.
-Our results have significant and profound implications for both researchers and practitioners since it provides quantitative evidence on which formulation of the \ac{pmv} model is more accurate.
+\subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
+This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} and \ac{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
+We compared the results against the self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.
+Our results have significant and profound implications for both researchers, practitioners, and policymakers since it provides quantitative evidence on which formulation of the \ac{pmv} model is more accurate.
+This in turn will allow them to take an informed decision on which model it should be used in National and International Standards to predict thermal sensation.
+Policymakers would also benefit from our results since they can use them to review the current versions of the \gls{7730} and \gls{55} Standards.

### 2023-04-18 Manuscript/src/sections/titleAbstract.tex
feat: reviewed intro, specified that we will only analyse PMV and PMV_CE, cited Yao 2022

@@
     \begin{abstract}
-        The \ac{pmv} model was originally developed by Fanger to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
-        It has seen widespread adoption and in its original form it is currently used in the \gls{7730}.
-        However, the \ac{pmv} model has a low prediction accuracy and to compensate for that, several other formulations have been proposed, one of which is currently used in the \gls{55}.
+        The primary objectives of buildings is to shelter pople from the elements and provide a comfortable thermal environment.
+        The \ac{pmv} model was therefore developed by Fanger more than 50 years ago to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
+        This should enable to determine the optimal temperature range to be maintained indoors to satisfy the great majority of the occupants.
+        The \ac{pmv} has seen widespread adoption and is and has been used by practitioners to design and operate buildings and by scientists to predict thermal sensation.
+        It is still currently used in its original form in the \gls{7730} despite the fact that several research studies have highlighted that the \ac{pmv} model has a low prediction accuracy, especially when it tries to predict the thermal sensation of people who are not thermally neutral.
+        To compensate for that, several other formulations have been proposed, one of which one is currently used in the \gls{55}.
         Being the \gls{7730} and \gls{55} the most widely used Standards worldwide, this is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
-        We aim to solve this problem by comparing the accuracy of the main \ac{pmv} formulations against \var{entries_db_used} thermal sensation votes collected in buildings.
+        We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, currently being used a reference in the above-mentioned standard, against \var{entries_db_used} thermal sensation votes collected in buildings.
         % We found that for V>0.1m/s the \gls{pmv-ce} is significantly more accurate (48%) than the PMV (42%) in predicting thermal neutrality.
@@
         % Consequently, the PMV should only be used in conditions close to neutrality within the range |PMV|≤1.
-        % We open-sourced the source code.
+         We open-sourced the source code we used to calculate the \ac{pmv} and we used to analyse the data.
     \end{abstract}

### 2023-04-18 Manuscript/src/sections/introduction.tex
feat: reviewed methodology

@@
 Finally, while \mycite{Yao2022} compare the \ac{pmv} and \ac{pmv-ce} models, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
-In particular, they focus significantly how the models perform in different climates and when applied to people from different world regions, but for example they fail to report how the \ac{pmv} accuracy varies as a function of \ac{tsv} and how the bias of the model varies as a function of different environmental parameters.
+In particular, they focus significantly how the models perform in different climates and when applied to people from different world regions, but for example they fail to report how the \ac{pmv} accuracy varies as a function of \ac{tsv}.
+Moreover, their analysis only reported the mean bias of the different models.
+This as depicted by \mycite{Humphreys2000} does not provide sufficient insights and information in depicting which model is more accurate in predicting \ac{tsv} particularly because the ASHRAE Global Thermal Comfort Database II v2.1 is an unbalanced dataset.
 

### 2023-04-18 Manuscript/src/sections/methodology.tex
feat: reviewed methodology

@@
 
-The \gls{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world and it is the largest of its kind.
-\Ac{tsv} were collected using a \gls{rhrn} survey and environmental parameters were measured and logged while participants completed their survey.
+The \gls{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
+\Ac{tsv} were collected using a \gls{rhrn} survey and environmental parameters were measured and logged in the proximity of the participants when they completed the survey.
 The Comfort DB contains data collected by researchers in cross-sectional field studies that have been published in peer reviewed journals.
-% We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
-%    All the results were calculated using the function pmv_ppd [6] included in pythermalcomfort v1.10.0 a Python package for thermal comfort calculations.
+ We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
+Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.7.0 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
 
@@
 We decided to keep entries that did not have \ac{tr}.
-Since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: 2\ac{to} – \ac{tdb}.
+Since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: $2 t_{o} – t_{db}$.
 Alternatively, if \ac{to} was not available we assumed \ac{tr} equal to \ac{tdb}.
@@
 The rationale is that the model accuracy should only be tested in within its applicability limits.
-The limits we used are as follows: \qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius}, \qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius}, \qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s}, \qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo}, 0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa, and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
+The limits we used are as follows:
+\qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius},
+\qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius},
+\qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s},
+\qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo},
+0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa,
+and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
 We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
-We used these inputs to calculate the \ac{pmv} values.  % todo check with marcel if this is required in his model too
+We used these inputs to calculate the \ac{pmv} values.
 
@@
 
-The values of \ac{pmv} and \gls{pmv-ce} are available inside the \gls{db2}.
-We calculated the values of \gls{pmvs}, \gls{pmvg}, and \ac{athb} using pythermalcomfort v2.5.1~\cite{Tartarini2020a}.
 We used Python to analyse and plot the results.
@@
 We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \textbf{provide URL} so other users can test different assumptions.
+% todo add URL
 
@@
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned deltas are within \numrange{-0.25}{0.25}.
-% todo why not 0.5 since the delta between a PMV of 2.5 and a self-reported vote of either 2 or 3 (closest points to 2.5) is 0.5?
+% todo why not 0.5 since the delta between a PMV of 2.5 and a self-reported vote of either 2 or 3 (the closest points to 2.5) is 0.5?
 
-\subsection{Sources of Error}\label{subsec:sources-of-error}
-One minor but not negligible source of error is the use of a discrete scale to assess thermal sensation.
-Most of the \ac{tsv} in the \gls{db2} were collected using a discrete scale while the \ac{pmv} output is continuous.
-For example, even if the model was 100~\% accurate and the estimated \ac{pmv} values was \num{2.5} the person could only report to be `warm (+2)' or `hot (+3)'.
-
-Measurement errors are also inevitably present in the dataset.
-These will add a random error to the model output.
-
-Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) which interacts with the surrounding environment.
-Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.
-Finally, it should be noted that the \ac{pmv} model has been developed considering based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermo regulatory system is always actively engaged to ensure a stable core temperature.
-However, as \mycite{Humphreys2002} note, these approximations and errors present in the \ac{pmv} model, if the model is accurate, should combine to produce a quasi-random error in the prediction.
-This would increase the standard deviation of the delta between the \ac{tsv} and \ac{pmv}, but if the model is sound, they should produce a negligible bias~\cite{Humphreys2002}.
-On the other hand, if the model is neither accurate nor representative of the real system a significant bias will arise.
+\subsection{Performance Metrics}\label{subsec:performance-metrics}
+For classification problems that involve unbalanced dataset the simple accuracy is a misleading metric to use to quantify the performance of a predictive model~\cite{Chawla2005}.
+We, therefore, decided to use the F1 score when determining the performance of the \ac{pmv} in predicting the \ac{tsv} reported by a participant.
+The F1 score can be interpreted as a harmonic mean of the precision and recall, where an F1 score reaches its best value at 1 and worst score at 0.
+Precision (also called positive predictive value) is the fraction of relevant instances among the retrieved instances, while recall (also known as sensitivity) is the fraction of relevant instances that were retrieved.
+The F1-micro score calculate metrics globally by counting the total true positives, false negatives and false positives.
+The F1-macro calculate metrics for each label, and find their unweighted mean.
+The F1-weighted calculate metrics for each label, and find their average weighted by support (the number of true instances for each label).
+This alters ‘macro’ to account for label imbalance; it can result in an F-score that is not between precision and recall.
+% todo descibe also the other metrics we used like R2

### 2023-04-18 Manuscript/src/sections/titleAbstract.tex
feat: reviewed methodology

@@
 
-    \author[label1]{Federico Tartarini\corref{mycorrespondingauthor}}
+    \author[label1,label2]{Federico Tartarini\corref{mycorrespondingauthor}}
     \ead{federicotartarini@berkeley.edu}
-    \author[label2]{Stefano Schiavon}
+    \author[label3]{Stefano Schiavon}
 
     \address[label1]{Berkeley Education Alliance for Research in Singapore, Singapore}
-    \address[label2]{Center for the Built Environment, University of California, Berkeley, CA, USA}
+    \address[label2]{Heat and Health Research Incubator, Faculty of Health and Medicine, University of Sydney, Sydney, AU}
+    \address[label3]{Center for the Built Environment, University of California, Berkeley, CA, USA}
 

### 2023-04-18 Manuscript/src/sections/results.tex
feat: analysed data
feat: improved figures

--- 2022-06-09 Manuscript/src/sections/results.tex
@@
 
-A total of \var{tot_surveys} entries met the inclusion criteria listed in the Methodology Section.
+\subsection{Comfort DB overview}\label{subsec:comfort-db-overview}
+
+A total of \var{entries_db_used} entries met the inclusion criteria listed in the Methodology Section.
 The distribution of the six input variables used to calculate the PMV indices is depicted in Figure~\ref{fig:dist_input_data}.
 
-\begin{figure}
+\begin{figure}[htb!]
     \centering
@@
     \caption{Distribution of the input variables used to calculate the \ac{pmv} values.
+    The text in blue is showing the the 2.5th, 50th (median), and 97.5th percentiles.
     The data are shown using boxen-plots (letter-value plots).
@@
 \end{figure}
+
+Figure~\ref{fig:dist_input_data} also depicts the 2.5th, 50th (median), and 97.5th percentiles for all the input variables.
+We were not able to determine the accuracy of the \ac{pmv} formulations for inputs which were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of datapoints beyond those limits.
+In natural ventilated buildings \ac{rh} may exceed \var{rh_95_perc_max}, and we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} higher than \var{v_95_perc_max}.
+The latter could be an issue since the \gls{55} permits the use of \ac{v} up to \qty{0.8}{\m\per\sec} even if occupants do not have direct control over the local airspeed.
+
+Figure~\ref{fig:dist_other_data} shows the distribution of the age, height, weight, and running mean outdoor temperature grouped by sex.
+Less than half (\num{23300}) of the total entries had information about the participant's sex.
+Out of those data were almost equally distributed among male \qty{52}{\percent} and females.
+Ages are not normally distributed, the same is true for the running mean outdoor temperature.
+About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old, and only \qty{2.5}{\percent} were older than 60.
+This limits our analysis to healthy adults, and does not allow us to determine the accuracy of the \ac{pmv} model in predicting thermal sensation for children or older adults.
+Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
+Men were significantly heavier and taller than females.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/dist_other_data}
+    \caption{Distribution of age, height, weight, and running mean outdoor temperature.
+    The data are grouped by sex.
+    The text in blue is showing the the 2.5th, 50th (median), and 97.5th percentiles.}
+    \label{fig:dist_other_data}
+\end{figure}
+
+The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
+Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
+Thus, on the right we show a bar chart depicting the distribution of all the \ac{tsv} votes, we have used in the analysis.
+The \gls{db2} in respect to the \ac{tsv} is unbalanced.
+Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
+While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
+    \caption{The left Figure shows the percentage of \ac{tpv} for each thermal sensation vote.
+    The numbers on the right side of each bar show the number of points for each \ac{tsv}.
+    The right figure shows the total number of data points grouped by \ac{tsv}.
+    Each bin in the Figure of the right has more data points than in the Figure on the left since information about the occupants \ac{tpv} were not always available.}
+    \label{fig:bar_plot_tp_by_ts}
+\end{figure}
+
+\subsection{Comparison of PMV accuracy in predicting thermal sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
+The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, in this Section we are reporting the prediction accuracy of different \ac{pmv} formulations.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy}
+    \caption{}
+    \label{fig:bar_stacked_model_accuracy}
+\end{figure}
+
+Being \ac{tsv} the value reported by participants (i.e., ground truth) we grouped participants responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations in Figure~\ref{fig:bar_stacked_model_accuracy}.
+In other words, all \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
+The overall accuracy of the \gls{pmv-ce} and \ac{pmv} were \qty{34}{\percent} and \qty{33}{\percent}, respectively.
+Being the classification problem unbalanced the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
+A model that would have predicted always `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, which is the number of participants who voted \ac{tsv}=0.
+For this reason in Table~\ref{tab:f1} we report the F1-scores for all the models.
+The F1-macro score which is free from label imbalance depicts that the accuracy of both models across all metrics is marginally better to random guessing (i.e., \qty{14.3}{\percent}).
+
+\begin{table}[htb!]
+    \centering
+    \input{tables/f1.tex}\label{tab:f1}
+\end{table}
+
+Moreover, while the \gls{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment.
+Our results show that neither formulations can correctly classify participants who reported to be either `warm' or `hot' and the \gls{pmv-ce} is less accurate than the \ac{pmv} in predicting people who were `slightly warm'.
+Both models had an accuracy lower than \qty{1}{\percent} when predicting wither `warm', `hot', or `cold'.
+This value is significantly lower and worse than random guessing.
+
+Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variables.
+To compensate for this we plotted the \ac{pmv} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a lowess curve.
+The curve is calculated using the individual data and not the binned data.
+We binned the data only to aid the visualization of a large dataset.
+If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and a slope of 1.
+As depicted in Figure~\ref{fig:bar_stacked_model_accuracy}, while the regression line does not intercept is not zero, the intercepts for the \ac{pmv} is \qty{-0.23} and for the \gls{pmv-ce} is \qty{-0.24}.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
+    \caption{}
+    \label{fig:bubble_models_vs_tsv}
+\end{figure}
+
+\section{Model Bias}\label{sec:model-bias}
+\subsection{Model Overall Bias}\label{subsec:model-overall-bias}
+As previously mentioned in the Methodology section, the aim of the \ac{pmv} model is not to accurately predict each individual thermal response form participants.
+The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants sharing the same environment.
+Consequently, we calculated the overall bias of the model by subtracting the \ac{pmv} model prediction from the self-reported \ac{tsv}.
+Results are presented in Figure~\ref{fig:hist_discrepancies} and show that overall the \ac{pmv} model appear to be free from serius bias.
+
+% todo comparison models for v0.3 m/s
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
+    \caption{}
+    \label{fig:hist_discrepancies}
+\end{figure}
+
+\subsection{Model Bias as Function of Independent Variables, Thermal Sensation, Thermal Preference, and PMV}\label{subsec:model-bias-variable}
+To better indentify under which environmental conditions as well as how the model bias varied as a function of \ac{tsv}, \ac{tpv}, and \ac{pmv} in the Figures below we present the same results as shown in Figure~\ref{fig:hist_discrepancies} but this time we have binned the results into discrete intervals of the above-mentioned variables.
+In the respective figures we are depicting the bin interval and the number of points in each bin.
+We did not present the results if the number of points in the bin are lower than XXX.
+% todo see if I need to implement the above
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bias_pmv}
+    \caption{}
+    \label{fig:bias_pmv}
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bias_pmv_ce}
+    \caption{}
+    \label{fig:bias_pmv_ce}
+\end{figure}

### 2023-04-18 main.py

@@ -1,6 +1,6 @@
 import matplotlib as mpl
 
-mpl.use("Qt5Agg")  # or can use 'TkAgg', whatever you have/prefer
+# mpl.use("Qt5Agg")  # or can use 'TkAgg', whatever you have/prefer
 
 from matplotlib.colors import LinearSegmentedColormap
 import matplotlib.pyplot as plt
@@ -79,7 +79,7 @@ palette_primary = [
 ]
 
 
-def save_var_latex(key, value, units=False):
+def save_var_latex(key, value, units=False, round_var=False):
     import csv
 
     dict_var = {}
@@ -93,6 +93,9 @@ def save_var_latex(key, value, units=False):
                 dict_var[row[0]] = row[1]
     except FileNotFoundError:
         pass
+
+    if round_var:
+        value = round(value, round_var)
 
     if units:
         dict_var[key] = f"\\qty{{{value}}}{{{units}}}"
@@ -633,15 +636,41 @@ def plot_distribution_variable():
             xlabel=f"{var_names[var]} ({var_units[var]})",
             ylim=(applicability_limits[var][0], applicability_limits[var][1]),
         )
+        desc = df[var].describe(percentiles=[0.025, 0.25, 0.5, 0.75, 0.975])
         if var == "ta":
             axs[ix].set(
                 ylim=(applicability_limits["tr"][0], applicability_limits["tr"][1]),
             )
         if var == "clo":
             axs[ix].set(yticks=(np.arange(0, 1.8, 0.3)))
+
+        desc = desc.round(2)
+        if var in ["ta", "tr", "rh"]:
+            desc = desc.round(1)
+
+        axs[ix].text(0.5, desc["2.5%"], desc["2.5%"], size="small", c="b", va="center")
+        axs[ix].text(
+            0.5,
+            desc["97.5%"],
+            desc["97.5%"],
+            size="small",
+            c="b",
+            va="center",
+        )
+        axs[ix].text(0.5, desc["50%"], desc["50%"], size="small", c="b", va="center")
     sns.despine(bottom=True, left=True)
     plt.savefig("./Manuscript/src/figures/dist_input_data.png", dpi=300)
     plt.show()
+
+    desc = df[["ta", "tr", "rh", "vel", "clo", "met"]].describe(
+        percentiles=[0.025, 0.25, 0.5, 0.75, 0.975]
+    )
+
[…]

### 2023-07-11 Manuscript/src/myacronyms.tex
paper: minor edits
paper: working on first draft
feat: reviewed methodology
feat: almost completed first draft of the paper. Syncing with Overleaf

--- 2022-06-09 Manuscript/src/myacronyms.tex
@@
 
-    \acro{tdb}[$t_{db}$]{dry-bulb air temperature\acroextra{, \unit{\celsius}}}
+    \acro{tdb}[$t_{db}$]{dry-bulb air temperature\acroextra{, $^{\circ}$C}}
+    \acro{twb}[$t_{wb}$]{wet-bulb air temperature\acroextra{, $^{\circ}$C}}
+    \acro{ti}[$t_{i}$]{indoor air temperature\acroextra{, $^{\circ}$C}}
+    \acro{tout}[$t_{out}$]{outdoor air temperature\acroextra{, $^{\circ}$C}}
+    \acro{to}[$t_{o}$]{operative air temperature\acroextra{, $^{\circ}$C}}
+    \acro{tcl}[$t_{cl}$]{clothing temperature\acroextra{, $^{\circ}$C}}
+    \acro{tg}[$t_{g}$]{globe temperature\acroextra{, $^{\circ}$C}}
     \acro{rh}[$RH$]{relative humidity\acroextra{, \%}}
     \acro{v}[$V$]{average air speed\acroextra{, m/s}}
-    \acro{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, \unit{\celsius}}}
+    \acro{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, $^{\circ}$C}}
     \acro{clo}[$I_{cl}$]{total clothing insulation\acroextra{, clo}}
+    \acro{i-cl}[$i_{cl}$]{permeation efficiency of water vapor through the clothing layer}
     \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
 
+    \acro{athb}[ATHB]{Adaptive Thermal Heat Balance}
     \acro{pmv}[PMV]{Predicted Mean Vote}
-    \acro{pmv-ce}[PMV$_{CE}$]{PMV calculated in accordance with the ASHRAE 55:2020}
     \acro{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
-    \acro{set}[SET]{Standard Effective Temperature\acroextra{, \unit{\celsius}}}
-    \acro{ce}[CE]{Cooling Effect\acroextra{, \unit{\celsius}}}
+    \acro{set}[SET]{Standard Effective Temperature\acroextra{, $^{\circ}$C}}
+    \acro{ce}[CE]{Cooling Effect\acroextra{, $^{\circ}$C}}
+    \acro{tsv}[TSV]{Thermal Sensation Vote}
+    \acro{tpv}[TPV]{Thermal Preference Vote}
+    \acro{ef}[$e$]{expectancy factor}
+    \acro{af}[a]{adaptive coefficient}
+
+        \acro{i-cl}[$i_{cl}$]{permeation efficiency of water vapor through the clothing layer}
+    \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
+    \acro{work}[$W$]{rate of mechanical work accomplished\acroextra{, W/m\textsuperscript{2}}}
+    \acro{t-sk}[$t_{sk}$]{skin mean temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-sk-n}[$t_{sk,n}$]{neutral skin mean temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-cr}[$t_{cr}$]{core mean temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-re}[$t_{re}$]{rectal temperature\acroextra{, $^{\circ}$C}}
+    \acro{t-cr-n}[$t_{cr,n}$]{neutral core mean temperature\acroextra{, $^{\circ}$C}}
+    \acro{r-cl}[$R_{cl}$]{thermal resistance of clothing\acroextra{, m\textsuperscript{2}K/W}}
+    \acro{r-e-cl}[$R_{e,cl}$]{evaporative heat transfer resistance of clothing layer\acroextra{, m\textsuperscript{2}kPa/W}}
+    \acro{f-cl}[$f_{cl}$]{clothing area factor $A_{cl}/A_{body}$\acroextra{, m\textsuperscript{2}K/W}}
+    \acro{h}[$h$]{sum of convective $h_{c}$ and radiative $h_{r}$ heat transfer coefficients\acroextra{, W/(m\textsuperscript{2}K)}}
+    \acro{h-r}[$h_{r}$]{linear radiative heat transfer coefficient\acroextra{, W/(m\textsuperscript{2}K)}}
+    \acro{h-c}[$h_{c}$]{convective heat transfer coefficient\acroextra{, W/(m\textsuperscript{2}K)}}
+    \acro{h-e}[$h_{e}$]{evaporative heat transfer coefficient\acroextra{, W/(m\textsuperscript{2}kPa)}}
+    \acro{a}[$\alpha$]{fraction of the total body mass considered
+to be thermally in the skin compartment}
+        \acro{s-cr}[$S_{cr}$]{rate of heat storage in the core compartment\acroextra{, W/m\textsuperscript{2}}}
+    \acro{s-sk}[$S_{sk}$]{rate of heat storage in the skin compartment\acroextra{, W/m\textsuperscript{2}}}
+    \acro{s}[$S$]{rate of heat storage in the human body\acroextra{, W/m\textsuperscript{2}}}
+    \acro{e-res}[$E_{res}$]{rate of evaporative heat loss from respiration\acroextra{, W/m\textsuperscript{2}}}
+    \acro{e-dif}[$E_{dif}$]{rate of evaporative heat loss from moisture diffused through the skin\acroextra{, W/m\textsuperscript{2}}}
+    \acro{e-rsw}[$E_{rsw}$]{rate of evaporative heat loss from sweat evaporation\acroextra{, W/m\textsuperscript{2}}}
+    \acro{e-sk}[$E_{sk}$]{total rate of evaporative heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acro{e-max}[$E_{max}$]{maximum rate of evaporative heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acro{c-res}[$C_{res}$]{rate of convective heat loss from respiration\acroextra{, W/m\textsuperscript{2}}}
+    \acro{c-r}[$C + R$]{sensible heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acro{q-res}[$q_{res}$]{total rate of heat loss through respiration\acroextra{, W/m\textsuperscript{2}}}
+    \acro{q-sk}[$q_{sk}$]{total rate of heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acro{w}[$w$]{skin wettedness}
+    \acro{w-max}[$w_{max}$]{skin wettedness practical upper limit}
+    \acro{m-sweat}[$m_{rsw}$]{rate at which regulatory sweat is generated\acroextra{, mL/h\textsuperscript{2}}}
+    \acro{m-bl}[$m_{bl}$]{skin blood flow\acroextra{, L/(hm\textsuperscript{2})}}
+    \acro{c-sw}[$c_{sw}$]{driving coefficient for regulatory sweating\acroextra{, g/(hKm\textsuperscript{2})}}
 

### 2023-07-11 Manuscript/src/sections/conclusions.tex
feat: almost completed first draft of the paper. Syncing with Overleaf

%! Author = sbbfti
%! Date = 26/08/2021

\section{Conclusions}\label{sec:conclusions}
The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \gls{pmv-ce} model, as employed in the \gls{55} Standard.
Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \gls{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
These disparities shed light on the limitations and strengths of the \ac{pmv} model, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
The outcomes of this analysis have significant implications for researchers, practitioners, and policymakers involved in the field of thermal comfort assessment and standardization.

### 2023-07-11 Manuscript/src/sections/introduction.tex
feat: almost completed first draft of the paper. Syncing with Overleaf

@@
 Out of those 1082 are scientific peer-review articles published in the research area of engineering, environmental science, social sciences, and energy.
-This highlights the extensive adaption and use of this model among the scientific community.
+This highlights the extensive adoption and use of this model among the scientific community.
 
-To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model despite the fact that several research papers have highlighted that the \ac{pmv} has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019}.
-% todo cite more papers
+To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model despite the fact that several research papers have highlighted that the \ac{pmv} has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal,Humphreys2002,Doherty1988}.
 
@@
 The process used to calculate the \gls{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
-For \ac{v} higher than \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-The value of \ac{ce}, which is calculated using the \ac{set} equation, is then subtracted from both the value of \ac{tdb} and \ac{tr}.
+In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
+The value of \ac{ce} is calculated using the \ac{set} equation and it is then subtracted from both the measured \ac{tdb} and \ac{tr}.
 The resulting values become the new inputs into the \ac{pmv} model.
-The \ac{ce} accounts for convective heat losses from the person to the environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+Since the \ac{ce} accounts for the convective heat losses from the person to its environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
 While the other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
@@
 The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
-This method, however, has its own limitation and throughout the years the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552020}.
+This method, however, has its own limitation, as follows.
+Firstly, it should be noted that throughout the years the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552020}.
+Secondly, calculating the \gls{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
+Thirdly, we could not find any peer-reviewed scientific publication that described the benefits of using this model over the original \ac{pmv}.
 
@@
     \centering
-    \label{fig:flowchart_pmv_ce}
-
-    \begin{tikzpicture}[node distance=1cm]
+    \begin{tikzpicture}[node distance=1.2cm]
         \node (start) [startstop] {\ac{pmv} calculation as per \gls{55}};
         \node (dec1) [decision, below of=start, yshift=-.25cm] {\ac{met} $> \qty{1}{met}$ };
-        \node (pro1a) [process, right of=dec1, xshift=4cm] {\ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$};
+        \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$};
         \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\ac{met} $> \qty{1.2}{met}$ };
-        \node (pro2a) [process, right of=dec2, xshift=4cm] {\ac{clo} = \ac{clo}$(0.6 + 0.4/$\ac{met}$)$};
+        \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\ac{clo} = \ac{clo}$(0.6 + 0.4/$\ac{met}$)$};
         \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\ac{v} $> \qty{0.1}{\m\per\s}$ };
         \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\ac{pmv}(\ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{met}, \ac{clo})};
-        \node (pro3) [process_large, below of=dec3, yshift=-.25cm, xshift=4cm] {Calculate \ac{ce}};
+        \node (pro3) [process, below of=dec3, yshift=-.25cm, xshift=3cm] {Calculate \ac{ce}};
         \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\ac{pmv}(\ac{tdb} - \ac{ce}, \ac{tr} - \ac{ce}, \ac{rh}, \ac{v} = \qty{0.1}{\m\per\s}, \ac{met}, \ac{clo})};
@@
     \end{tikzpicture}
+    \caption{Flowchart depicting the steps needed to calculate the PMV model in accordance with the ASHRAE 55--2020 Standard.}
+    \label{fig:flowchart_pmv_ce}
 \end{figure}
 
-This is a source of confusion for many users worldwide since both the \gls{7730} and the \gls{55} are the widely adopted and incorporated in many national and international building codes worldwide.
+Choosing between the \ac{pmv} and \gls{pmv-ce} consequently is now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many national and international building codes worldwide.
 
@@
 To further complicate this issue, several other formulations of the \ac{pmv} model have been proposed worldwide.
-Among the most notable formulations there are the: \gls{pmvs}, \gls{pmvg}, \gls{epmv}, \gls{apmv}, and \ac{athb}.
-\mycite{Yao2022} provides a comprehensive review of the difference \ac{pmv} formulations and the adaptive heat balance models based on \ac{pmv}.
-Hence, we invite our readers to read \mycite{Yao2022} paper to learn more about all the different variations that have been proposed.
+Among the most notable formulations there are the: \gls{pmvs}~\cite{GaggeSET}, \gls{pmvg}~\cite{GaggeSET}, \gls{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
+\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describe some of the above-mentioned models.
+Hence, we invite our readers to read \mycite{Yao2022} paper to learn more about some of \ac{pmv} the different variations that have been proposed.
 
-%% todo describe pmv SET
-%% todo describe pmv Gagge
-%
 %In 2002, Fanger, P. O. and Toftum, J. developed the \gls{epmv} an extension of the \ac{pmv}.
@@
 %
-%% todo describe marcel's work
 
-\subsection{Comparision of PMV formulations}\label{subsec:comparision-of-pmv-formulations}
-To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
+\subsection{Comparison of ISO and ASHRAE PMV formulations}\label{subsec:comparision-of-pmv-formulations}
+To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \gls{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
+In particular, we could not find any study which answered the questions: ``which \ac{pmv} formulation should you use the \gls{7730} or \gls{55}?''.
 
@@
 While, \mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
-Finally, while \mycite{Yao2022} compare the \ac{pmv} and \ac{pmv-ce} models, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
+Finally, while \mycite{Yao2022} compare the \ac{pmv} and \gls{pmv-ce} models, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
 In particular, they focus significantly how the models perform in different climates and when applied to people from different world regions, but for example they fail to report how the \ac{pmv} accuracy varies as a function of \ac{tsv}.
@@
 \subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
-This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} and \ac{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
+This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} and \gls{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
 We compared the results against the self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.

### 2023-07-11 Manuscript/src/sections/methodology.tex
feat: almost completed first draft of the paper. Syncing with Overleaf

@@
  We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
-Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.7.0 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
+Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
 
@@
 
-We used Python to analyse and plot the results.
+We used Python to analyse and visualise the results.
 
@@
 The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants sharing the same environment.
-While the \gls{db2} contains data from a sufficiently wide population no information is available on where exactly these data were collected.
-We only have access to building level data and no information about weather a group of occupants was sharing the same thermal environment.
+While the \gls{db2} contains data from a sufficiently wide population no information is available on where and when exactly these data were collected.
+We only have access to building level data and no information about whether a group of occupants was sharing the same thermal environment.
 We, therefore, decided to first group the participants responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} model.
@@
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned deltas are within \numrange{-0.25}{0.25}.
+Binning data by the independent variables is a significant improvement over grouping the participants responses by their \ac{tsv}.
+This is because the \ac{pmv} model aim is, as previously mentioned, predicting the average thermal sensation of a large group of occupants sharing the same environment and not to correctly predict the average \ac{tsv} of people who reported the same \ac{tsv} vote.
 % todo why not 0.5 since the delta between a PMV of 2.5 and a self-reported vote of either 2 or 3 (the closest points to 2.5) is 0.5?
+
+Finally, we decided to compare the results of the two \ac{pmv} formulations with the results of the two-node model.
+The two-node model was developed by \mycite{GaggeSET} and it uses a more extensive set of equations to solve the heat balance equation depicted in Equation~\ref{eq:heat-balance}.
+We did this by passing as input to the \ac{pmv}, \gls{pmv-ce}, and two-node model a total of \qty{10000} randomly generated set of inputs. 
+Each set of input was generated using the Numpy random uniform function which picks a random value in within a predefined range. 
+We selected the range of each of the 6 input variables needed to calculate the \ac{pmv} to be equal to the 2.5th and 97.5th percentiles for all the input variables contained in the Comfort DB.
+The only exception was for \ac{v} which was randomly selected from this interval \qty{0.1}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s}, the rationale for this decision was that both the \ac{pmv} and \gls{pmv-ce} models use the same equations for \ac{v} $\leq$ \qty{0.1}{\m\per\s}.
+
+The two-node model, along with other physiological parameters and thermal comfort indices, it outputs the \gls{pmvg} and the amount of heat losses or gains form the body to the surrounding environment.
+Gagge's two-node model is also used by the \gls{pmv-ce} to calculate the \ac{ce}.
+\begin{equation}
+    M - W = C + R + E_{dif} + E_{rsw} + C_{res} + E_{res}\label{eq:heat-balance}
+\end{equation}
+Where \ac{met} minus the external work ($W$) must be dissipated either via sensible or latent heat losses.
+Sensible heat is transferred via conduction, convection and radiation (\ac{c-r} + \ac{c-res}).
+Latent heat loss occurs from the evaporation of sweat (\ac{e-rsw}), moisture diffused through the skin  (\ac{e-dif}), and respiration (\ac{e-res}).
+Since both thermal comfort standard deal with office spaces and moderate thermal environments we omitted the heat storage terms in Equation~\ref{eq:heat-balance}.
 

### 2023-07-11 Manuscript/src/sections/results.tex
feat: almost completed first draft of the paper. Syncing with Overleaf

@@
 %! Date = 10/06/2020
-
-\section{Results}\label{sec:results}
+\section{Results and Discussion}\label{sec:results}
 
 \subsection{Comfort DB overview}\label{subsec:comfort-db-overview}
-
-A total of \var{entries_db_used} entries met the inclusion criteria listed in the Methodology Section.
+While the \gls{db2} contains \var{entries_db_all}, only a total of \var{entries_db_used} entries met the inclusion criteria listed in the Methodology Section.
 The distribution of the six input variables used to calculate the PMV indices is depicted in Figure~\ref{fig:dist_input_data}.
 
-\begin{figure}[htb!]
+\begin{figure*}[htb!]
     \centering
@@
     \label{fig:dist_input_data}
-\end{figure}
+\end{figure*}
 
@@
 We were not able to determine the accuracy of the \ac{pmv} formulations for inputs which were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of datapoints beyond those limits.
-In natural ventilated buildings \ac{rh} may exceed \var{rh_95_perc_max}, and we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} higher than \var{v_95_perc_max}.
-The latter could be an issue since the \gls{55} permits the use of \ac{v} up to \qty{0.8}{\m\per\sec} even if occupants do not have direct control over the local airspeed.
+Moreover, the distribution of these data was not uniform across the Standard's applicability ranges.
+For example, only 2.5\% of the available dataset had a \ac{tdb} $\leq$ than \var{ta_95_perc_min}.
+This can be explained by the fact that only few buildings across the globe have operate at \ac{tdb} lower than this value.
+Which would be deemed to be unsatisfactory by a great number of occupants.
+Consequently, we were not able to validate the accuracy of both formulations beyond these limits, despite the fact that both thermal comfort Standards applicability limits extend to \qty{10}{\celsius}.
+This lack of data is particularly relevant for \ac{v}.
+The \ac{pmv} and \gls{pmv-ce} only differ when the value of \ac{v} exceeds \qty{0.1}{\m\per\sec}, however, due to the fact that most of the entries in the \gls{db2} were collected in buildings with low air movement we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} higher than \var{v_95_perc_max}.
 
@@
 Men were significantly heavier and taller than females.
-
-\begin{figure}[htb!]
+\begin{figure*}[htb!]
     \centering
@@
     \label{fig:dist_other_data}
-\end{figure}
-
+\end{figure*}
 The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
@@
 While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
-
-\begin{figure}[htb!]
+\begin{figure*}[htb!]
     \centering
@@
     \label{fig:bar_plot_tp_by_ts}
-\end{figure}
+\end{figure*}
+In thermal comfort research it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
+However, in the \gls{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
+This finding challenges the above assumption.
+This result suggests that the use of the thermal preference scale is better suited to determine how people perceive their thermal environment.
+The \ac{tpv} clearly allows to determine whether participants would take action or not to modify their current thermal environment.
+The \ac{tsv}, on the contrary, is ambiguous and does not allow to determine which actions should be taken to increase thermal comfort conditions for occupants.
 
@@
 
-\begin{figure}[htb!]
+\begin{figure*}[htb!]
     \centering
@@
     \label{fig:bar_stacked_model_accuracy}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \input{tables/f1.tex}\label{tab:f1}
+    \input{tables/f1.tex}
+    \caption{F1-score for the \ac{pmv} and \gls{pmv-ce} models.}
+    \label{tab:f1}
 \end{table}
@@
 
+The low prediction accuracy of both models formulations coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models applicability limits should be revisited and perhaps limited.
+For example, Standards applicability limits could be limited to \qty{18}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius} until more data are collected or the \ac{pmv} formulation is changed to better predict people who vote at the extremes of the thermal sensation scale.
+
 Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variables.
 To compensate for this we plotted the \ac{pmv} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a lowess curve.
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
+    \caption{}
+    \label{fig:bubble_models_vs_tsv}
+\end{figure*}
 The curve is calculated using the individual data and not the binned data.
@@
 
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
-    \caption{}
-    \label{fig:bubble_models_vs_tsv}
-\end{figure}
-
-\section{Model Bias}\label{sec:model-bias}
-\subsection{Model Overall Bias}\label{subsec:model-overall-bias}
+\subsection{Model Bias}\label{sec:model-bias}
+\subsubsection{Model Overall Bias}\label{subsec:model-overall-bias}
 As previously mentioned in the Methodology section, the aim of the \ac{pmv} model is not to accurately predict each individual thermal response form participants.
 The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants sharing the same environment.
-Consequently, we calculated the overall bias of the model by subtracting the \ac{pmv} model prediction from the self-reported \ac{tsv}.
-Results are presented in Figure~\ref{fig:hist_discrepancies} and show that overall the \ac{pmv} model appear to be free from serius bias.
+Consequently, while the above-mentioned analysis is informative in understanding the overall prediction error of the \ac{pmv} model it does compare individual \ac{tsv} to \ac{pmv}, hence it has some limitations, and it does not conclusively disprove the inefficacy of the \ac{pmv} model in predicting the average thermal sensation of a large group of occupants.
 
-% todo comparison models for v0.3 m/s
+We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} model prediction from the self-reported \ac{tsv}.
+Results are presented in Figure~\ref{fig:hist_discrepancies}.
+Overall both the \ac{pmv} and \gls{pmv-ce} models appear to be free from serius bias.
+However, the \gls{pmv-ce} had a higher bias than the \ac{pmv} when using the full dataset, the summary statistics are shown in Figure~\ref{fig:hist_discrepancies}.
+The median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0} when we only included those datapoints with \ac{v} higher than \qty{0.2}{\m\per\s}.
+While the opposite is true for the \gls{pmv-ce}, the median bias increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
 
-\begin{figure}[htb!]
+\begin{figure*}[htb!]
     \centering
@@
     \label{fig:hist_discrepancies}
-\end{figure}
+\end{figure*}
 
-\subsection{Model Bias as Function of Independent Variables, Thermal Sensation, Thermal Preference, and PMV}\label{subsec:model-bias-variable}
-To better indentify under which environmental conditions as well as how the model bias varied as a function of \ac{tsv}, \ac{tpv}, and \ac{pmv} in the Figures below we present the same results as shown in Figure~\ref{fig:hist_discrepancies} but this time we have binned the results into discrete intervals of the above-mentioned variables.
-In the respective figures we are depicting the bin interval and the number of points in each bin.
-We did not present the results if the number of points in the bin are lower than XXX.
-% todo see if I need to implement the above
+\subsubsection{Model Bias as Function of Independent Variables, Thermal Sensation, Thermal Preference, and PMV}\label{subsec:model-bias-variable}
+In the previous section we presented the overall bias of the two \ac{pmv} formulations.
+We then went on determining how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{met}, and \ac{clo}.
+Figure~\ref{fig:dist_input_data} shows the results.
+In each sub-figure we have binned the results into discrete intervals of equal width and the x-axis tick label is the center value of the bin interval.
+%and the width of the violin plot is linearly proportional to the number of points in that specific bin.
+In each sub-figure, we kept data only above and below the 2.5th and 97.5th percentiles.
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bias_models}
+    \caption{}
+    \label{fig:bias_models}
+\end{figure*}
+The results depict that across the four input variables the use of the \gls{pmv-ce} model did not improve the bias of the model.
+On the contrary, the bias of the \gls{pmv-ce} diverged more from 0 as the value of \ac{v} increased.
+This outcome is particularly negative since the \gls{pmv-ce} was specifically developed to be more accurate than the \ac{pmv} at high airspeed.
+These results are in agreement with those presented in the previous section.
 
-\begin{figure}[htb!]
+\subsection{Comparison between the \ac{pmv} formulations wih the two-node model}\label{subsec:comparison-between-the-ac{pmv}-formulations-wih-the-two-node-model}
+We then finally compared how the results of the \ac{pmv} and \gls{pmv-ce} compared with the results of the two-node model.
+Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the \ac{pmv} using a combination of 5000 randomly generated sets inputs and using these to calculate the \ac{pmv} and \gls{pmv-ce} (y-axis), and the respective \gls{pmvg} value (x-axis).
+The Figure also shows the locally weighted linear regression (LOWESS) curve.
+The results of the \ac{pmv} model had a higher agreement with the results of the two-node model, in particularly for \gls{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
+Depicting the failure of the \gls{pmv-ce} model in achieving a greater achievement with the two-node model, despite using it in the backend to calculate the cooling effect.
+\begin{figure*}[htb!]
     \centering
-    \includegraphics[width=\textwidth]{figures/bias_pmv}
+    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
     \caption{}
-    \label{fig:bias_pmv}
-\end{figure}
+    \label{fig:pmv_two_node_comparison}
+\end{figure*}
 
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/bias_pmv_ce}
-    \caption{}
-    \label{fig:bias_pmv_ce}
-\end{figure}
+\subsection{Sources of Error}\label{sec:sources-of-error}
+While the scope of this manuscript is solely to compare the two \ac{pmv} formulations and determine which of the two is more accurate and should be used in thermal comfort standards.
+In this section we are trying to identify some of the main reasons of why we observed in the predictive accuracy of the \ac{pmv} and \gls{pmv-ce} models when compared against the \ac{tsv} available in the \gls{db2}.
+
+\paragraph{Lack of Data}
+As previously mentioned several times the \ac{pmv} model aim to predict the average thermal sensation of a large group of occupants sharing the same environment.
+First, the term `large' is vague and it leads to confusion since it is not clear what is the minimum number of people required, is it 2, 5, 10, or more?
+Secondly, the data contained in the \gls{db2} do not contain information on where and when the thermal sensation votes where recorded and consequently we could not appropriately validate the accuracy of the model since in most of the analysis we ended up comparing single \ac{tsv} with the \ac{pmv} model output.
+The only exception to this is when we assessed the overall model bias.
+
+\paragraph{Heat Balance Equation and Calculation of the PMV value}
+ The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
+This in itself is a possible source of error since the model consider the human body to be a cilinder with constant width all uniformly covered with clothing were the heat is all generated in its core.
+This is great simplification since it ignores the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, it ignores vasoconstriction and vasodilation, and most importantly it fundamentally assumes that the human body is never in thermal neutrality.
+This is clearly not the case, since the human body in most conditions observed indoors in commercial buildings is capable of maintaining a constant core temperature.
+The \ac{pmv} model then maps the heat loss or gains, which are a proxi to the thermal stress that the human body has to undergone to maintain a constant internal temperature to a \ac{pmv} value which should represent a \ac{tsv} vote.
+Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature this equation was never updated.
+Moreover, the heat losses and gains are mapped equally to a thermal sensation vote, disregarding that the human body does not have the same ability to compensate for warm and cold conditions.
+It is, therefore, not surprising to observe that the \ac{pmv} model has an acceptable accuracy in determining when people are `neutral` which is equivalent to a \ac{pmv} of 0 and the body is dissipating all the internal metabolic heat production trough conduction, convection, or radiation.
+On the other hand the model is not capble of identifying people at either extremes of the \ac{tsv} scale.
+These people most likely were exposed to compensable conditions, which in other words means that their core temperature remain constant.
+However, their body either used vasocontriction to compensate for the heat losses, or a combination of vasodilation and sweating to compensate for heat gains.
+More research is need to determine whether the inacuracy of the model is related to either or a combination of the following factors: i) assuming that the heat lossess/gains estimated by the \ac{pmv} should represent the thermal stress the body has to udertake to compensate for unsatisfactory thermal comfort conditions is not correct;
+ii) the relationship between the heat lossess/gains and thermal stress is different from the one assumed by Fanger;
+iii) heat losses and gains correlation to thermal stress is not not the same and two equations should be used;
+iv) the mapping of the \ac{pmv} to the \ac{tsv} uses a different anchor from what participants use in reporting thermal sensation.
+For example, while the \ac{pmv} model may assume that a score of 3 signify the change between a compensable set of conditions to an uncompensable one, e.g., their skin wettedness reaches the max value ($w_{max}$).
+Participants in commercial buildings may, on the contrary, deem the environment to by `Hot' as soon as they start sweating, or in technical words their skin wettedness crosses a critical threshold which is only a fraction of $w_{max}$.
+It is likely that a combination of the above mentioned factors play a role, but only the latter has gained more momentum in the literature which led to development of several \ac{pmv} where the final value was multipled by a constant generally lower than one.
+While few of these indices tried to change the intercept.
+
+\paragraph{Measurement Errors}
+While all the data contained in the \gls{db2} have been published in peer-reviewed manuscript measurements errors are inevitably present in the dataset.
+This is due to the accuracy of the instrumentation used, the placement of the instrumentation, and the estimation of clothing, and activity using reference tables.
+In an ideal scenario, errors may be random and would cancel out and not affect the overall bias of the model but only the standard deviation of the error, however, this may not always be the case, for example if a researchers often underestimated the clothing insulation of participants or their activity levels.
+
+\paragraph{Thermal Sensation Scale}
+One minor but not negligible source of error is the use of a discrete scale to assess thermal sensation.
+Most of the \ac{tsv} in the \gls{db2} were collected using a discrete scale while the \ac{pmv} output is continuous.
+For example, even if the model was 100~\% accurate and the estimated \ac{pmv} values was \num{2.5} the person could only report to be `warm (+2)' or `hot (+3)'.
+% talk also about the research we published with Marcel and how participants perceive the thermal sensation scale
+
+\paragraph{Thermal Sensation Scale}
+Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) which interacts with the surrounding environment.
+Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.
+Finally, it should be noted that the \ac{pmv} model has been developed considering based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermo regulatory system is always actively engaged to ensure a stable core temperature.
+However, as \mycite{Humphreys2002} note, these approximations and errors present in the \ac{pmv} model, if the model is accurate, should combine to produce a quasi-random error in the prediction.
+This would increase the standard deviation of the delta between the \ac{tsv} and \ac{pmv}, but if the model is sound, they should produce a negligible bias~\cite{Humphreys2002}.
+On the other hand, if the model is neither accurate nor representative of the real system a significant bias will arise.

### 2023-07-11 Manuscript/src/sections/titleAbstract.tex
feat: almost completed first draft of the paper. Syncing with Overleaf

@@
-%! Author = sbbfti
-%! Date = 10/06/2020
-
 \begin{frontmatter}
 
-    \title{PMV formulations: which one is the most accurate?}
+    \title{PMV: which formulation should you use the ISO 7730:2005 or ASHRAE 55--2020?}
 
@@
     \begin{abstract}
-        The primary objectives of buildings is to shelter pople from the elements and provide a comfortable thermal environment.
+        The primary objectives of buildings is to shelter people from the elements and provide a comfortable thermal environment.
         The \ac{pmv} model was therefore developed by Fanger more than 50 years ago to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
-        This should enable to determine the optimal temperature range to be maintained indoors to satisfy the great majority of the occupants.
+        This should enable its users to determine the optimal temperature range to be maintained indoors to satisfy the great majority of the occupants.
+
         The \ac{pmv} has seen widespread adoption and is and has been used by practitioners to design and operate buildings and by scientists to predict thermal sensation.
@@
         To compensate for that, several other formulations have been proposed, one of which one is currently used in the \gls{55}.
-        Being the \gls{7730} and \gls{55} the most widely used Standards worldwide, this is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
+        Being the \gls{7730} and \gls{55} the most widely used Standards worldwide, this is a source of confusion for users who find it difficult to decide which model to use.
+
         We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, currently being used a reference in the above-mentioned standard, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        % We found that for V>0.1m/s the \gls{pmv-ce} is significantly more accurate (48%) than the PMV (42%) in predicting thermal neutrality.
-        % In 95% of the cases, the error was less than one point.
-        % The PMV and PMVCE accuracy were 35% and 36% for |PMV|≤2 and 43% and 44% for |PMV|≤1, respectively.
-        % Neither PMV model could accurately predict ‘cold’, ‘cool’, ‘warm’, or ‘hot’ thermal sensations.
-        % Consequently, the PMV should only be used in conditions close to neutrality within the range |PMV|≤1.
-         We open-sourced the source code we used to calculate the \ac{pmv} and we used to analyse the data.
+        We found that the \gls{pmv-ce} is not more accurate than the original \ac{pmv} in predicting thermal neutrality, despite its increased complexity and computational resources need to calculate the results.
+        Most importantly the bias of the \gls{pmv-ce} model was grater (a higher number is negative) than the \ac{pmv} when we only analysed the entries with \ac{v}$\geq$\qty{0.1}{\m\per\sec}.
+
+        The \ac{pmv} and \gls{pmv-ce} accuracy were XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 2$ and XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 1$, respectively.
+        Neither PMV model could accurately predict ‘cold’, ‘cool’, ‘warm’, or ‘hot’ thermal sensations.
+        Consequently, the PMV should only be used in conditions close to neutrality within the range $\mid$\ac{pmv}$\mid \leq 1$, and we suggest considering reducing the applicability limit of the Standards.
+
+         We open-sourced the source code we used to to analyse the data.
     \end{abstract}

### 2023-07-11 main.py

@@ -16,7 +16,6 @@ from pythermalcomfort.models import (
     athb,
     set_tmp,
     cooling_effect,
-    pmv,
 )
 from scipy import stats
 from scipy.stats import skewtest
@@ -24,6 +23,7 @@ from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score
 from sklearn.metrics import f1_score
 from sklearn.metrics import r2_score, mean_absolute_error
+from sklearn.linear_model import LinearRegression
 from itertools import product
 
 warnings.filterwarnings("ignore")
@@ -79,19 +79,7 @@ palette_primary = [
 ]
 
 
-@vectorize(
-    [
-        float64(
-            float64,
-            float64,
-            float64,
-            float64,
-            float64,
-            float64,
-            float64,
-        )
-    ],
-)
+@jit(nopython=True)
 def pmv_ppd_optimized(tdb, tr, vr, rh, met, clo, wme):
     pa = rh * 10 * math.exp(16.6536 - 4030.183 / (tdb + 235))
 
@@ -154,8 +142,9 @@ def pmv_ppd_optimized(tdb, tr, vr, rh, met, clo, wme):
 
     ts = 0.303 * math.exp(-0.036 * m) + 0.028
     _pmv = ts * (mw - hl1 - hl2 - hl3 - hl4 - hl5 - hl6)
+    heat_loss = mw - hl1 - hl2 - hl3 - hl4 - hl5 - hl6
 
-    return _pmv
+    return _pmv, heat_loss
 
 
 def pmv_ppd(tdb, tr, vr, rh, met, clo, wme=0, standard="ISO", **kwargs):
@@ -332,7 +321,9 @@ def pmv_ppd(tdb, tr, vr, rh, met, clo, wme=0, standard="ISO", **kwargs):
     tr = tr - ce
     vr = np.where(ce > 0, 0.1, vr)
 
-    pmv_array = pmv_ppd_optimized(tdb, tr, vr, rh, met, clo, wme)
+    (pmv_array, heat_losses) = np.vectorize(pmv_ppd_optimized, cache=True)(
+        tdb, tr, vr, rh, met, clo, wme
+    )
 
     ppd_array = 100.0 - 95.0 * np.exp(
         -0.03353 * np.power(pmv_array, 4.0) - 0.2179 * np.power(pmv_array, 2.0)
[…]

### 2023-08-29 Manuscript/src/myacronyms.tex
paper: reviewing Stefano's comments

@@
 
+    \acro{7730}[ISO~7730:2005]{ISO~7730:2005 is a thermal comfort standard developed by ISO}
+    \acro{55}[ASHRAE~55:2023]{ASHRAE~55:2023 is a thermal comfort standard developed by ANSI and ASHRAE}
+    \acro{pmv-ce}[PMV\textsubscript{CE}]{PMV calculated in accordance with the ASHRAE~55:2023}
+    \acro{epmv}[ePMV]{Adjusted Predicted Mean Votes with Expectancy Factor}
+    \acro{apmv}[aPMV]{Adaptive Predicted Mean Vote}
+    \acro{pmvs}[PMV$_{SET}$]{PMV calculated using the SET temperature in place of $t_{db}$}
+    \acro{pmvg}[PMV$_{Gagge}$]{PMV calculated using the two-node heat balance model}
+    \acro{db2}[DB$_{comfort}$]{ASHRAE Global Thermal Comfort Database II}
+    \acro{lowess}[LOWESS]{LOcally WEighted Scatterplot Smoothing}
+
 \end{acronym}

### 2023-08-29 Manuscript/src/sections/conclusions.tex
paper: reviewing Stefano's comments

@@
 \section{Conclusions}\label{sec:conclusions}
-The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \gls{pmv-ce} model, as employed in the \gls{55} Standard.
-Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \gls{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
-These disparities shed light on the limitations and strengths of the \ac{pmv} model, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
+The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} Standard.
+Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
+These disparities shed light on the limitations and strengths of the \ac{pmv} and \ac{pmv-ce} models, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
 The outcomes of this analysis have significant implications for researchers, practitioners, and policymakers involved in the field of thermal comfort assessment and standardization.

### 2023-08-29 Manuscript/src/sections/introduction.tex
paper: reviewing Stefano's comments

@@
 \section{Introduction}\label{sec:introduction}
-
 In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
@@
 This sentence is from Yao 2022.
-The \ac{pmv} model has been widely adopted worldwide both among researchers and practitioners.
-It is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
-A search within article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
-Out of those 1082 are scientific peer-review articles published in the research area of engineering, environmental science, social sciences, and energy.
-This highlights the extensive adoption and use of this model among the scientific community.
+The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
+It is the most extensively used thermal comfort index for estimating thermal sensation.
+A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
+Out of those 1082 are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
+This highlights this model's extensive adoption and use among the scientific community.
 
-To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model despite the fact that several research papers have highlighted that the \ac{pmv} has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal,Humphreys2002,Doherty1988}.
+To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988}.
 
-To overcome this problem, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \gls{pmv-ce}~\cite{ashrae552020}.
-The process used to calculate the \gls{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
+To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \ac{pmv-ce}~\cite{ashrae552023}.
+The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some choices were implemented. 
+A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}. 
+The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
 In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-The value of \ac{ce} is calculated using the \ac{set} equation and it is then subtracted from both the measured \ac{tdb} and \ac{tr}.
-The resulting values become the new inputs into the \ac{pmv} model.
+The value of \ac{ce} is calculated using the \ac{set} equation, and it is then subtracted from both the \ac{tdb} and \ac{tr}.
+The resulting values become the new inputs in the \ac{pmv} model.
 Since the \ac{ce} accounts for the convective heat losses from the person to its environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
-While the other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
-Consequently, for a given thermal environment, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
+Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
 The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
-This method, however, has its own limitation, as follows.
-Firstly, it should be noted that throughout the years the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552020}.
-Secondly, calculating the \gls{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
-Thirdly, we could not find any peer-reviewed scientific publication that described the benefits of using this model over the original \ac{pmv}.
+However, this method has limitations.
+First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
+Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
 
@@
     \end{tikzpicture}
-    \caption{Flowchart depicting the steps needed to calculate the PMV model in accordance with the ASHRAE 55--2020 Standard.}
+    \caption{Flowchart depicting the steps for the calculation of the PMV in accordance with the \acs{55} Standard.}
     \label{fig:flowchart_pmv_ce}
@@
 
-Choosing between the \ac{pmv} and \gls{pmv-ce} consequently is now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many national and international building codes worldwide.
+Choosing between the \ac{pmv} and \ac{pmv-ce} is consequently now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many building codes and standards worldwide.
 
-\subsection{Other PMV formulations}\label{subsec:other-pmv-formulations}
-To further complicate this issue, several other formulations of the \ac{pmv} model have been proposed worldwide.
-Among the most notable formulations there are the: \gls{pmvs}~\cite{GaggeSET}, \gls{pmvg}~\cite{GaggeSET}, \gls{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
-\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describe some of the above-mentioned models.
-Hence, we invite our readers to read \mycite{Yao2022} paper to learn more about some of \ac{pmv} the different variations that have been proposed.
+Several other formulations of the \ac{pmv} model have been proposed to complicate this issue further.
+Among the most notable formulations there are the: \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
+\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
 
-%In 2002, Fanger, P. O. and Toftum, J. developed the \gls{epmv} an extension of the \ac{pmv}.
-%The \gls{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
+%In 2002, Fanger, P. O. and Toftum, J. developed the \ac{epmv} an extension of the \ac{pmv}.
+%The \ac{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
 %The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
@@
 %In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
-%This arbitrary and non-scientific definition of the expectancy factor make the \gls{epmv} difficult to calculate and non-generalizable.
-%Consequently, the \gls{epmv} has not gained any widespread adoption.
+%This arbitrary and non-scientific definition of the expectancy factor make the \ac{epmv} difficult to calculate and non-generalizable.
+%Consequently, the \ac{epmv} has not gained any widespread adoption.
 %
@@
 %The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
-%This approach cannot be, therefore, generalizable and similarly to the \gls{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
+%This approach cannot be, therefore, generalizable and similarly to the \ac{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
 %
@@
 \subsection{Comparison of ISO and ASHRAE PMV formulations}\label{subsec:comparision-of-pmv-formulations}
-To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \gls{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
-In particular, we could not find any study which answered the questions: ``which \ac{pmv} formulation should you use the \gls{7730} or \gls{55}?''.
+To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models, which are used in the \gls{7730} and \gls{55} Standards.
 
-Some notable works include, the work from \mycite{Doherty1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
-They concluded that the \ac{pmv} model is accurate for simulations of resting subjects but its accuracy decreases as a function of metabolic rate.
+Some notable works that tried to determine the accuracy of the \ac{pmv} model include the manuscript from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
+They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
 Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
-They used the ASHRAE Thermal Comfort database v1.0 and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
-Humphreys and Nicol also determined the validity of the \ac{pmv} for predicting comfort votes collected in field studies~\cite{Humphreys2002}.
+They used the ASHRAE Global Thermal Comfort database I and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
+They also determined the validity of the \ac{pmv} for predicting comfort votes collected in field studies~\cite{Humphreys2002}.
 They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
 The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
-While, \mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
-Finally, while \mycite{Yao2022} compare the \ac{pmv} and \gls{pmv-ce} models, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
-In particular, they focus significantly how the models perform in different climates and when applied to people from different world regions, but for example they fail to report how the \ac{pmv} accuracy varies as a function of \ac{tsv}.
-Moreover, their analysis only reported the mean bias of the different models.
-This as depicted by \mycite{Humphreys2000} does not provide sufficient insights and information in depicting which model is more accurate in predicting \ac{tsv} particularly because the ASHRAE Global Thermal Comfort Database II v2.1 is an unbalanced dataset.
+\mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
+They found that the thermal sensation predicted by the PMV model, on average, is one full thermal sensation scale unit away from the subject’s responses, confirming the results of~\cite{Humphreys2002}.
+The accuracy of PMV was only \qty{34}{\percent}, the model has a slightly higher prediction accuracy for sensation close to neutrality, but the accuracy declined towards either end of the thermal sensation scale, and it overestimated both hot and cold sensations.
+They also found that the Predicted Percentage of Dissatisfied (PPD) failed to predict the percentage of unacceptable votes if the thermal sensation was predicted using the PMV model and suggested its removal from \gls{55} Standards.
+These results were confirmed based on the Chinese Thermal Comfort database~\cite{du_evaluation_2022}.
+\mycite{Yao2022} compared the \ac{pmv} and \ac{pmv-ce} models, however, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
+In particular, they focus significantly on how the models perform in different climates and when applied to people from different world regions, but for example, they do not report how the \ac{pmv} accuracy varies as a function of environmental and personal parameters, and \ac{tsv}.
+Their analysis only reports the mean bias of the different models.
+This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate in predicting \ac{tsv}.
+This is because the \ac{pmv} model has a higher prediction accuracy of sensation close to neutrality. 
+Being the ASHRAE Global Thermal Comfort Database II v2.1 an unbalanced dataset and most of the data points are neutral a more detailed analysis as the one we are presenting in this manuscript is needed to compare the accuracy of both formulations.
 
 \subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
-This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} and \gls{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
-We compared the results against the self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.
-Our results have significant and profound implications for both researchers, practitioners, and policymakers since it provides quantitative evidence on which formulation of the \ac{pmv} model is more accurate.
-This in turn will allow them to take an informed decision on which model it should be used in National and International Standards to predict thermal sensation.
-Policymakers would also benefit from our results since they can use them to review the current versions of the \gls{7730} and \gls{55} Standards.
+We aim to compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards.
+We used the self-reported \ac{tsv} recorded in the \ac{db2}.
+We aim to determine which \ac{pmv} model is most accurate, this will allow policymakers to make an informed decision on which model should be used in standards.

### 2023-08-29 Manuscript/src/sections/methodology.tex
paper: reviewing Stefano's comments

--- 2023-04-18 Manuscript/src/sections/methodology.tex
@@
-%! Author = sbbfti
-%! Date = 10/06/2020
-
 \section{Methodology}\label{sec:methodology}
-
-The \gls{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
-\Ac{tsv} were collected using a \gls{rhrn} survey and environmental parameters were measured and logged in the proximity of the participants when they completed the survey.
-The Comfort DB contains data collected by researchers in cross-sectional field studies that have been published in peer reviewed journals.
- We compared the results obtained by the \ac{pmv} and the \gls{pmv-ce} with the self-reported \ac{tsv}.
-Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.7.0 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
+The \ac{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
+\Ac{tsv} were collected using a right-here-right-now survey and environmental parameters were measured and logged in the proximity of participants when they completed the survey.
+The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
+ We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the self-reported \ac{tsv}.
+Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
 
 \subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
-Not all researchers logged all the six parameters required to calculate the PMV.\@
+Not all researchers logged all the six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
 We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
-These constraints filtered out approximately \var{entries_db_valid} of the data in the Comfort DB.\@
+These constraints filtered out approximately \var{entries_db_valid} votes.
 We decided to keep entries that did not have \ac{tr}.
-Since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: $2 t_{o} – t_{db}$.
-Alternatively, if \ac{to} was not available we assumed \ac{tr} equal to \ac{tdb}.
+However, since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: $2 t_{o} – t_{db}$.
+Alternatively, if the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
 We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
@@
 
-Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points which did not meet the inclusion criteria of both Standards.
-The rationale is that the model accuracy should only be tested in within its applicability limits.
+Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points that did not meet the inclusion criteria of both Standards.
+The rationale is that the models' accuracy should only be tested within their applicability ranges.
 The limits we used are as follows:
-\qty{10}{\celsius} $\leq$ \ac{tdb} $\leq$ \qty{30}{\celsius},
-\qty{10}{\celsius} $\leq$ \ac{tr} $\leq$ \qty{40}{\celsius},
-\qty{0}{\m\per\s} $\leq$ \ac{v} $\leq$ \qty{1}{\m\per\s},
-\qty{0}{clo} $\leq$ \ac{clo} $\leq$ \qty{1.5}{clo},
-0 Pa $\leq$ water vapour partial pressure $\leq$ 2000 Pa,
-and \qty{1}{met} $\leq$ \ac{met} $\leq$ \qty{4}{met}.
+\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
+\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
+\num{0}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s},
+\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
+\num{0}~$\leq$~water vapour partial pressure~$\leq$~\qty{2000}{\pascal},
+and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
 We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
-We used these inputs to calculate the \ac{pmv} values.
+We used these inputs to calculate the \ac{pmv} and \ac{pmv-ce} values.
 
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with $|$\ac{tsv}$|$ or $|$\ac{pmv}$|$ $\leq$ than \num{3.5}.
-
-We used Python to analyse and plot the results.
-
+However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with $|$\ac{tsv}$|$ or $|$\ac{pmv}$|$~$\leq$~\num{3}.
+We used Python to analyse and visualise the results.
 We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \textbf{provide URL} so other users can test different assumptions.  
@@
 \subsection{Model Validation}\label{subsec:model-validation}
-The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants sharing the same environment.
-While the \gls{db2} contains data from a sufficiently wide population no information is available on where exactly these data were collected.
-We only have access to building level data and no information about weather a group of occupants was sharing the same thermal environment.
-We, therefore, decided to first group the participants responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} model.
-This approach is useful to perform an overall assessment of the accuracy of the model, but it introduces an error due to the rounding of the \ac{pmv} models output and \ac{tsv} values.
-The latter were sometimes collected using a continuous scale and not a discrete one, in other words they could take any value in the range of \numrange{-3.0}{3.0}.
-To remove this rounding error, we then went on subtracted the \ac{tsv} value from the respective \ac{pmv} value.
-These deltas quantify the success of the model in predicting \ac{tsv}.
-However, on their own are a low precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
-These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{rh}).
-If the \ac{pmv} formulation is bias-free, the distribution of any batch derived from these differences would have a mean value that was little different from zero.
-The standard deviation would reflect the combined effect of the people's individual differences as well as any errors in the \ac{pmv} formulation~\cite{Humphreys2002}.
-According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned deltas are within \numrange{-0.25}{0.25}.
+The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants and it is applied from a singly occupied zone to an area with several hundred people.
+While the \ac{db2} contains data from a sufficiently wide population no information is available on where and when exactly these data were collected.
+We only have access to building-level data and no information about whether a group of occupants was sharing the same thermal environment.
+We grouped the participants' responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
+This approach is useful for performing an overall assessment of the accuracy of the models, but it introduces an error due to the rounding of both the \ac{pmv}, \ac{pmv-ce}, and \ac{tsv} values.
+The latter were sometimes collected using a continuous scale and not a discrete one, in other words, they could take any value in the range from -3.0 to 3.0.
+
+To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
+These differences, aka bias, quantify the success of the model in predicting \ac{tsv}.
+However, on their own are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
+These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{v}).
+If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
+The standard deviation would reflect the combined effect of the people's individual differences as well as any errors in formulation or in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
+According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between -0.25 and 0.25.
+Binning data by the independent variables is a significant improvement over grouping the participants' responses by their \ac{tsv}.
+This is because the \ac{pmv} model aim is, as previously mentioned, to predict the average thermal sensation of a large group of occupants sharing the same environment and not to correctly predict the average \ac{tsv} of people who reported the same \ac{tsv} vote, but where in different locations and moments of the year and day.
 % todo why not 0.5 since the delta between a PMV of 2.5 and a self-reported vote of either 2 or 3 (the closest points to 2.5) is 0.5?
 
+Finally, we decided to compare the results of \ac{pmv} and \ac{pmv-ce} formulations with the \ac{pmvg} calculated by the two-node model.
+The two-node model was developed by \mycite{GaggeSET} and it uses a more accurate set of equations to solve the heat balance equation.
+We compared the \ac{pmv} and \ac{pmv-ce} results to the \ac{pmvg} since Gagge's two-node model is used by the \ac{pmv-ce} to calculate the \ac{ce}.
+Hence, in principle, the results of the \ac{pmv-ce} model should closely match the \ac{pmvg} values.
+We did this by passing as input to the \ac{pmv}, \ac{pmv-ce}, and two-node models the same set of \num{10000} randomly generated set of inputs. 
+Each set of inputs was sampled assuming a random uniform distribution within a predefined range. 
+We selected the range of each of the 6 input variables needed to calculate the \ac{pmv} to be equal to the 2.5th and 97.5th percentiles for all the input variables contained in the \ac{db2}.
+The only exception was for the range of \ac{v} that was randomly selected from this interval \qty{0.1}{\m\per\s}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s}.
+The rationale for this decision was that both the \ac{pmv} and \ac{pmv-ce} models use the same equations for \ac{v}~$\leq$~\qty{0.1}{\m\per\s}.
+
 \subsection{Performance Metrics}\label{subsec:performance-metrics}
-For classification problems that involve unbalanced dataset the simple accuracy is a misleading metric to use to quantify the performance of a predictive model~\cite{Chawla2005}.
-We, therefore, decided to use the F1 score when determining the performance of the \ac{pmv} in predicting the \ac{tsv} reported by a participant.
-The F1 score can be interpreted as a harmonic mean of the precision and recall, where an F1 score reaches its best value at 1 and worst score at 0.
-Precision (also called positive predictive value) is the fraction of relevant instances among the retrieved instances, while recall (also known as sensitivity) is the fraction of relevant instances that were retrieved.
-The F1-micro score calculate metrics globally by counting the total true positives, false negatives and false positives.
-The F1-macro calculate metrics for each label, and find their unweighted mean.
-The F1-weighted calculate metrics for each label, and find their average weighted by support (the number of true instances for each label).
-This alters ‘macro’ to account for label imbalance; it can result in an F-score that is not between precision and recall.
-% todo descibe also the other metrics we used like R2
+Simple accuracy can be a misleading metric to use to quantify the performance of a classification predictive model that involves an unbalanced dataset~\cite{Chawla2005}.
+We then used the F1 score when determining the performance of the \ac{pmv} and \ac{pmv-ce} in predicting the \ac{tsv} reported by a participant.
+The F1 score is the harmonic mean of the precision and sensitivity, where an F1 score reaches its best value at 1 and the worst score at 0.
+Precision (also called positive predictive value) is the ratio of true positive overall positive results obtained from the test, while sensitivity (also known as recall or true positive rate) is the ratio of true positive over all the actual positive cases.
+The F1-micro score calculates metrics globally by counting the total true positives, false negatives and false positives.
+The F1-macro calculate metrics for each label and find their unweighted mean.
+The F1-weighted calculate metrics for each label and find their average weighted by support (the number of true instances for each label).
+% todo describe also the other metrics we used like R2

### 2023-08-29 Manuscript/src/sections/results.tex
paper: reviewing Stefano's comments

--- 2022-07-04 Manuscript/src/sections/results.tex
@@
 %! Date = 10/06/2020
+\section{Results and Discussion}\label{sec:results}
 
-\section{Results}\label{sec:results}
-
-\subsection{Comfort DB overview}\label{subsec:comfort-db-overview}
-
-A total of \var{entries_db_used} entries met the inclusion criteria listed in the Methodology Section.
-The distribution of the six input variables used to calculate the PMV indices is depicted in Figure~\ref{fig:dist_input_data}.
-
-\begin{figure}
+\subsection{\ac{db2} overview}\label{subsec:comfort-db-overview}
+While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
+The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
+The 2.5th, 50th (median), and 97.5th percentiles are shown in blue.
+% todo: remove the gridlines or marker them very short; remove outliers or make the line very thin; keep one decimal for clo values; remove decimal point for RH; annotation font should be the same as the label and same size, change only the color; You can use black or blue (HEX: 003262, RGB (0, 50, 98)) for the chart; one decimal for met not two; the chart is a bit dense of information
+\begin{figure*}[htb!]
     \centering
     \includegraphics[width=\textwidth]{figures/dist_input_data}
-    \caption{Distribution of the input variables used to calculate the \ac{pmv} values.
-    The data are shown using boxen-plots (letter-value plots).
-    They depict the median as the centerline and each successive level outward contains half of the remaining data.}
+    \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
+    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.
+    The data are shown using boxen plots (letter-value plots).
+    They depict the median as the center-line and each successive level outward contains half of the remaining data.}
     \label{fig:dist_input_data}
-\end{figure}
+\end{figure*}
+The input data distributions are far from uniformity (equal density) across the Standards' applicability ranges.
+We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
+For example, only 2.5\% of the available dataset had a \ac{tdb} $\leq$ than \var{ta_95_perc_min}, this may consequently be an issue since both thermal comfort Standards' applicability limits extend to \qty{10}{\celsius}.
+This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
+This is because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
+This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
+The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v} $\geq$ \qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
+Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} $\geq$ \var{v_95_perc_max}.
 
-Figure~\ref{fig:dist_other_data} shows the distribution of the age, height, weight, and running mean outdoor temperature grouped by sex.
-Less than half (\num{23300}) of the total entries had information about the participant's sex.
-Out of those data were almost equally distributed among male \qty{52}{\percent} and females.
+Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature grouped by sex.
+Less than half (\num{23300}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
+Men were significantly heavier and taller than females.
 Ages are not normally distributed, the same is true for the running mean outdoor temperature.
-About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old.
-While \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
-Mens were significantly heavier and taller than females.
-
-\begin{figure}
+About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old, and only \qty{2.5}{\percent} were older than 60.
+This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
+The \ac{db2} does not contain information about the health status of the participants.
+Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
+% todo: check that the colours are distinguishable for color blind and change them since I am using the same color palette to differentiate PMV and PMVCE; same comments as previous chart; use a boxen plot and have 4 for men and 4 for women.
+Tormt is missing the unit; age unot should be year.
+This figure can be merged with figure 2 because it is a visualization of the database.
+I would use blue (HEX: 003262, RGB (0, 50, 98)) and gold (HEX: FDB515, RGB (253, 181, 21)) as color because blue and green could be confounded by colorblind people.
+To be more consisten in the paper, I would use the letter value plots for this and the figure in section 3.3.2
+\begin{figure*}[htb!]
     \centering
@@
     \caption{Distribution of age, height, weight, and running mean outdoor temperature.
-    The data are grouped by sex.}
+    The data are grouped by sex.
+    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.}
     \label{fig:dist_other_data}
-\end{figure}
-
-Figure~\ref{fig:bar_plot_tp_by_ts} shows on the left the percentage of thermal preference votes grouped by each \ac{tsv}.
-The left Figure also reports the number of points for each \ac{tsv}.
-Only \var{entries_with_tp} that had information about the thermal preference.
-Thus, on the right we show a bar chart depicting the distribution of all the \ac{tsv} votes, we have used in the analysis.
-The \gls{db2} in respect to the \ac{tsv} is unbalanced.
+\end{figure*}
+The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
+Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
+Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
+The \ac{tsv} is unbalanced. 
 Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
 While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
-In the \gls{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
-However, in thermal comfort research it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
-This finding challenges the above assumption.
-Moreover, these results provide additional evidence that the use of the thermal preference scale is better suited to determine how people perceive their thermal environment.
-The thermal sensation vote clearly allows to determine whether participants would take action or not to modify their current thermal environment.
-The \ac{tsv}, on the contrary, does not allow to determine which actions should be taken to increase thermal comfort conditions for occupants.
-
-\begin{figure}
+% todo I really do not like this color, they are misleading.
+If someone want to be "warmer", then it should be shown as blue.
+I suggest to use the color pallet (figure 3) that Micheal used in the BCA paper.
+% todo TSV report words not numbers.
+% todo TSV chart the color here should not match the one on the thermal preference scale given that it is a different question.
+For example, a TSV=3 is red, like people that wants to be warmer, meaning that they are cold.
+Even with the inversion of the color palette, we cannot imply that TSV=3 mease wanted to be cooler.
+What about TSV=1, maybe people want no changer or cooler or warmer.
+Difficult to assume.
+Better to use two independet color schemes.
+\begin{figure*}[htb!]
     \centering
     \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
-    \caption{The left Figure shows the percentage of thermal preference votes for each thermal sensation vote.
+    \caption{The left Figure shows the percentage of \ac{tpv} for each thermal sensation vote.
     The numbers on the right side of each bar show the number of points for each \ac{tsv}.
     The right figure shows the total number of data points grouped by \ac{tsv}.
-    Each bin in the Figure of the right has more data points than in the Figure on the left since information about the occupants thermal preference were not always available.}
+    Each bin in the right figure has more data points than in the left figure since less data on \ac{tpv} were available.}
     \label{fig:bar_plot_tp_by_ts}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
+However, in the \ac{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
+This finding challenges the above assumption. 
+Assuming that `slightly warm' or `slightly cool' is comfortable leads to under-counting thermal discomfort.
+This result highlights one of the major limitations of the \ac{tsv} scale and suggests that the thermal preference scale is better suited to determine how people perceive their thermal environment.
+The \ac{tpv} more clearly allows an HVAC system to decide whether to change or not the thermal environment. 
+On the contrary, the \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
+
+\subsection{Comparison of PMV accuracy in predicting thermal sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
+The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are here comparing the prediction accuracy of the two formulations.
+% todo Use the words, not the numbers.
+Please use the color pallet from Toby's paper.
+It is hard to notice that these are two distint graph because the legend is moving accross.
+I suggest to move the legend above and have a thin grey line to separate the two group, another option would be to add a light grey background for the two PMV and show that they are distinct.
+% todo Move the percentage, on top and move it horizontally.
+% todo remove the white space between the boxes (horizontal lines)
+% todo TSV use words not numbers
+% todo Move this also above the figure, over each column. I would also add the term. Overall accuracy  PMV = 33%
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy}
+    \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants. 
+    The number of data points in each bin of the two charts is the same and this allows us to compare the accuracy of both models.
+    The overall prediction accuracy is shown above the chart while the model accuracy of predicting participants grouped by their \ac{tsv} is overlayed over each bar. 
+    A random model, randomly selecting a thermal sensation vote would have an overall accuracy of \qty{14}{\percent}}
+    \label{fig:bar_stacked_model_accuracy}
+\end{figure*}
+Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
+Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
+The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
+Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
+A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}=0.
+For this reason, in Table~\ref{tab:f1}, we report the F1-scores for all the models.
+The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}).
+% todo in the table I should also report the F1 scores if I filter out some of the TSV.
+The PMV model should only be used in the range -2 and 2 so at least we should report also the results in this range.
+\begin{table}[htb!]
+    \centering
+    \input{tables/f1.tex}
+    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models.}
+    \label{tab:f1}
+\end{table}
+
+The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used.
+However, our results show that the \ac{pmv-ce} is less accurate than the \ac{pmv} in predicting people who were `slightly warm'.
+Both models had an accuracy lower than \qty{1}{\percent} when predicting either `warm', `hot', or `cold'.
+This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
+The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
+Based on the results, we suggest that the Standards applicability limits could be reduced to the following range \qty{18}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{30}{\celsius} (instead of \qty{10}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius}),  until more data are collected.
+
+Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
+To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
+% todo I would not use grey for the data, it is too similar to the gridlines.
+A light color is fine, light brown (depending on the color pallet used in the rest of the paper) could work.
+The regression line could be done even thicker.
+Be more explicit in the graph of what is measured and what is calculated.
+The Y axis should be labeled as PMV
+
+\begin{figure*}[htb!]
     \centering
     \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
+    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
+    Raw data were then binned and a bubble chart is superimposed over the regression curve to aid the visualization of a large dataset.}
+    \label{fig:bubble_models_vs_tsv}
+\end{figure*}
+The curve, in the Figure, is calculated using the individual data and not the binned data.
+We binned the data to aid the visualization of a large dataset.
+If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and have a slope of 1.
+The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval. 
+The intercept is close to 0 and this is aligned with the results shown above that the model can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times.
+Both \ac{pmv} formulations always underpredicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv}; the slope of the curve was significantly lower than 1.
+We discuss in detail in sub-Section~\ref{subsec:sources-of-error} which factors may be the underlying cause of these errors.
+
+The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
+The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
+Consequently, while the above-mentioned analysis is needed to compare for the same sample size in each thermal sensation group the \ac{pmv} and \ac{pmv-ce} accuracies, it focuses primarily on comparing individual \ac{tsv} to \ac{pmv}.
+This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
+This was already done in \mycite{Cheung2019}, where they grouped people based on the same thermal environment (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified how many times the \ac{pmv} model correctly predicted \ac{tsv}.
+They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
+
+\subsection{Model Bias}\label{sec:model-bias}
+\subsubsection{Model Overall Bias}\label{subsec:model-overall-bias}
+We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and  \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
+Results are presented in Figure~\ref{fig:hist_discrepancies}.
+% todo You need to check of the four distibution are normally distributed, if so, report mean and SD, if not report median and IQR, no need to report all of them.
+I guess that Pswetest is an indication of this, so, for the entire dataset it seems that it is not sufficiently normally distributed, so report only median and IQR.
+The Pswetest This index is not reported, explained, and interpreted in the paper.
+I would like to have these figure a bit longer in the Y-axis.
+they are kind of squized.
+The word "delta" does not mean difference.
+So, I suggest to use "diffference" or just show the equation Bias=PMVi-TSVi.
+1) Can this chart be transformed in a letter value plot?
+The Caption is missing
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
     \caption{}
-    \label{fig:bubble_models_vs_tsv}
-\end{figure}
+    \label{fig:hist_discrepancies}
+\end{figure*} 
+Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
+Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
+The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s}; we also calculated the bias of the model, including data points with \ac{v} higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
+We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
+We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
+The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume (xxx).
+The threshold in the standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
+The results show that the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
+Surprisingly, we found that the median bias for the \ac{pmv-ce} increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
+These are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
+As far as we could find, there is no published paper proving a higher accuracy of the \ac{pmv-ce} compared to the \ac{pmv} model. 
+The \ac{pmv-ce} has lower accuracy while making the code more complex and computationally intensive to run.
+
+\subsubsection{Model Bias as Function of Four Input Variables}\label{subsec:model-bias-variable}
+In the previous section, we presented the overall bias of the two \ac{pmv} formulations.
+We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{met}, and \ac{clo}.
+Figure~\ref{fig:dist_input_data} shows the results.
+In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval. 
+The width of the violin plot is not proportional to the number of points in that specific bin.
+% todo This figure needs a lot of improvements.
+1) I would use the letter-value (or boxplot), they may be easier to interpret, provide less info and could be aligned with the first figure (in particular if  The actual density distribution it is not adding much value.
+2) The color for PMV and PMV_CE must be the same in the entire paper.
+3) The y-label should be moved to the x-label and the y-label should explicit.
+I would remove the decimal values and show only the five integers -2, -1, 0, 1 and 2.
+4) How is possible to have a velocity "centered" at 0?
+I would remove the red lines.
+If you want to underline the -0.5 and 0.5 range, then we should use the same strategy also for figure 7.
+The fact that you change graph time in the paper, it makes more difficult to interpret without adding a substantial amount of value.
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/bias_models}
+    \caption{}
+    \label{fig:bias_models}
+\end{figure*}
+The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all four input variables.
+On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 as the value of \ac{v} increased.
+These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.
+
+\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model}
+We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx.
+Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
+The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
+The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
+The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
+These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
+This is also an unexpected result and we will discuss why this may be in the following paragraph. 
+% todo When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
+1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.
+
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
+    \caption{}
+    \label{fig:pmv_two_node_comparison}
+\end{figure*}
+
+\subsection{Sources of Error}\label{subsec:sources-of-error}
+In this Section, we identify some of the main reasons that could explain the difference in the results from the two \ac{pmv} models when compared against the \ac{tsv} available in the \ac{db2}.
+Moreover, we tried to explain why we observed that the \ac{pmv-ce} model has a lower accuracy than the \ac{pmv} in predicting thermal sensation.
+The issues are not reported in order of importance and more research is needed to determine how to improve the overall accuracy of the \ac{pmv} model.
+
+\paragraph{Lack of Data}
+T.
+
+\paragraph{Heat Balance Equation and Calculation of the PMV value}
+The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
+This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
+This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
+However, the \ac{pmv} model erroneously assumes that the human body, aside from some edge cases, is either losing or gaining heat from its surrounding environment.
+This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
+Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
+
+The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
+Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
+In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.
+
+It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
+This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis.
+On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
+We showed that people tend to be closer to neutrality than what the models predicted (see figure XXX), this is an indication that they were exposed to compensable thermal conditions.
+Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.
+
+To continue using the \ac{pmv} model beyond an absolute value of 1, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
+i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
+ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress are constant and not xxx;
+iii) the regression coefficient between heat losses and heat gains and thermal stress are not the same for the heat losses and gains and two different equations should be used;
+iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
+For example, while the \ac{pmv} model may assume that a score of 3 (hot) signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches the max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
+Participants in commercial buildings may, on the contrary, deem the environment to be `Hot' as soon as they start sweating and their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.
+
+It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
+The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}
+
+\paragraph{Measurement Errors}
+All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts, however, measurement errors are inevitably present in the dataset.
+This is due to the fact that all measurement instrumentation is not always calibrated before use, the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the near proximity of the participant, and estimating clothing and metabolic rates from tables is a complex and prone to error task.
+In an ideal scenario, errors would be random and would cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
+However, this may not always be the case, for example, if a researcher often underestimated the clothing insulation of participants or their activity levels.
+
+\paragraph{Thermal Sensation Scale}
+One minor but not negligible source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
+For example, even if the model was 100~\% accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (+2)' or `hot (+3)'.
+Using a continuous, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.
+They also showed that the \ac{tsv}'s assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
+An alternative solution would be to transform the \ac{pmv} output to an ordinal and categorical seven-point scale.
+
+% \paragraph{Other neglected factors}
+% Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) that interacts with the surrounding environment.
+% Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.
+
+% todo write a paragraph trying to find the reason why the PMV CE model is less accurate.
+% todo write a paragraph about what we recommend for ASHRAE 55

### 2023-08-29 Manuscript/src/sections/titleAbstract.tex
paper: reviewing Stefano's comments

--- 2023-04-18 Manuscript/src/sections/titleAbstract.tex
@@
-%! Author = sbbfti
-%! Date = 10/06/2020
-
 \begin{frontmatter}
 
-    \title{PMV formulations: which one is the most accurate?}
+    \title{PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?}
 
@@
     \begin{abstract}
-        The primary objectives of buildings is to shelter pople from the elements and provide a comfortable thermal environment.
-        The \ac{pmv} model was therefore developed by Fanger more than 50 years ago to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
-        This should enable to determine the optimal temperature range to be maintained indoors to satisfy the great majority of the occupants.
-        The \ac{pmv} has seen widespread adoption and is and has been used by practitioners to design and operate buildings and by scientists to predict thermal sensation.
-        It is still currently used in its original form in the \gls{7730} despite the fact that several research studies have highlighted that the \ac{pmv} model has a low prediction accuracy, especially when it tries to predict the thermal sensation of people who are not thermally neutral.
-        To compensate for that, several other formulations have been proposed, one of which one is currently used in the \gls{55}.
-        Being the \gls{7730} and \gls{55} the most widely used Standards worldwide, this is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
-        We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, currently being used a reference in the above-mentioned standard, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        % We found that for V>0.1m/s the \gls{pmv-ce} is significantly more accurate (48%) than the PMV (42%) in predicting thermal neutrality.
-        % In 95% of the cases, the error was less than one point.
-        % The PMV and PMVCE accuracy were 35% and 36% for |PMV|≤2 and 43% and 44% for |PMV|≤1, respectively.
-        % Neither PMV model could accurately predict ‘cold’, ‘cool’, ‘warm’, or ‘hot’ thermal sensations.
-        % Consequently, the PMV should only be used in conditions close to neutrality within the range |PMV|≤1.
-         We open-sourced the source code we used to calculate the \ac{pmv} and we used to analyse the data.
+        The \ac{pmv} model was developed by Fanger more than 50 years ago to predict the mean thermal sensation of a group of people. 
+        It is based on a heat balance model of the human body.
+
+        \gls{55} states that the model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce satisfactory thermal environmental conditions for a majority of the occupants within the space''.
+
+        The \ac{pmv} has seen widespread adoption and has been used by professionals to design and operate buildings and by researchers to predict thermal sensation.
+        It is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
+        Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
+        \gls{7730} and \gls{55} are the most widely used Standards worldwide, and their differences are a source of confusion for users who find it difficult to decide which model to use.
+
+        We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
+        We found that the \ac{pmv-ce} has the same accuracy as the original \ac{pmv} in predicting thermal neutrality, despite its increased complexity and computational resources needed to calculate the results.
+        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}$\geq$\qty{0.1}{\m\per\sec}.
+
+        The \ac{pmv} and \ac{pmv-ce} accuracy were XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 2$ and XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 1$, respectively.
+        Neither PMV model could accurately predict `cold’, `cool’, `warm’, or `hot’ thermal sensations.
+        Consequently, the PMV should only be used in conditions close to neutrality within the range $\mid$\ac{pmv}$\mid \leq 1$, and we suggest reducing the applicability limit of the Standards to this range.
     \end{abstract}
@@
     \begin{keyword}
-        Personal thermal comfort model \sep Ecological momentary assessment \sep Skin temperature \sep Machine learning \sep Internet of Things (IoT)
+        Thermal comfort model \sep ISO 7730 \sep ASHRAE 55 \sep Thermal sensation \sep Accuracy
     \end{keyword}

### 2023-08-29 main.py
paper: fixed seconds unit

@@ -1602,7 +1602,7 @@ def plot_distribution_variable():
 
     # save_var_latex("rh_95_perc_min", desc["ta"]["2.5%"], "\\celsius", round_var=1)
     save_var_latex("rh_95_perc_max", desc["rh"]["97.5%"], "\\percent", round_var=2)
-    save_var_latex("v_95_perc_max", desc["vel"]["97.5%"], "\\m\\per\\sec", round_var=2)
+    save_var_latex("v_95_perc_max", desc["vel"]["97.5%"], "\\m\\per\\s", round_var=2)
 
     r2 = r2_score(df.ta, df.tr)
 
@@ -2764,7 +2764,7 @@ if __name__ == "__old_code__":
 
 if __name__ == "__testing_pmv_set_agreement__":
 
-    number_combinations = 5000
+    number_combinations = 10000
     # todo change the values below programmatically rather than a static input
     combinations = {
         "tdb": np.random.uniform(size=number_combinations, low=18.5, high=29),

### 2023-08-29 Manuscript/src/sections/results.tex
paper: updated Fig 2 and 3

@@
 The 2.5th, 50th (median), and 97.5th percentiles are shown in blue.
-% todo: remove the gridlines or marker them very short; remove outliers or make the line very thin; keep one decimal for clo values; remove decimal point for RH; annotation font should be the same as the label and same size, change only the color; You can use black or blue (HEX: 003262, RGB (0, 50, 98)) for the chart; one decimal for met not two; the chart is a bit dense of information
+% TODO:
+%  remove the gridlines or marker them very short
+%  the chart is a bit dense of information
 \begin{figure*}[htb!]
@@
 
-Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature grouped by sex.
+Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature.
 Less than half (\num{23300}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
-Men were significantly heavier and taller than females.
 Ages are not normally distributed, the same is true for the running mean outdoor temperature.
@@
 Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
-% todo: check that the colours are distinguishable for color blind and change them since I am using the same color palette to differentiate PMV and PMVCE; same comments as previous chart; use a boxen plot and have 4 for men and 4 for women.
-Tormt is missing the unit; age unot should be year.
-This figure can be merged with figure 2 because it is a visualization of the database.
-I would use blue (HEX: 003262, RGB (0, 50, 98)) and gold (HEX: FDB515, RGB (253, 181, 21)) as color because blue and green could be confounded by colorblind people.
-To be more consisten in the paper, I would use the letter value plots for this and the figure in section 3.3.2
+% todo:
+%  I merged men and women since not considering sex difference in the paper.
+%  This figure can be merged with figure 2 because it is a visualization of the database.
+%  To be more consisten in the paper, I would use the letter value plots for this and the figure in section 3.3.2
+% remove the gridlines or marker them very short;
 \begin{figure*}[htb!]

### 2023-08-29 main.py

@@ -53,6 +53,8 @@ palette_tsv = [
 ]
 
 # sns.palplot(palette_tsv)
+# sns.palplot(palette_tp)
+# sns.palplot(palette_primary)
 
 palette_primary = [
     "#FFBA22",
@@ -1564,7 +1566,9 @@ def plot_distribution_variable():
     f, axs = plt.subplots(1, 6, constrained_layout=True, figsize=(8, 3))
 
     for ix, var in enumerate(["ta", "tr", "rh", "vel", "clo", "met"]):
-        sns.boxenplot(y=var, data=df, ax=axs[ix], color="lightgray")
+        sns.boxenplot(
+            y=var, data=df, ax=axs[ix], color="#FDB515", showfliers=False, linewidth=0.5
+        )
         axs[ix].set(
             ylabel="",
             xlabel=f"{var_names[var]} ({var_units[var]})",
@@ -1579,19 +1583,22 @@ def plot_distribution_variable():
             axs[ix].set(yticks=(np.arange(0, 1.8, 0.3)))
 
         desc = desc.round(2)
-        if var in ["ta", "tr", "rh"]:
+        if var in ["ta", "tr", "rh", "met", "clo"]:
             desc = desc.round(1)
+        if var in ["rh"]:
+            desc = desc.astype(int)
 
-        axs[ix].text(0.5, desc["2.5%"], desc["2.5%"], size="small", c="b", va="center")
+        text_color = "#00B0Da"
+        axs[ix].text(0.5, desc["2.5%"], desc["2.5%"], c=text_color, va="center")
         axs[ix].text(
             0.5,
             desc["97.5%"],
             desc["97.5%"],
-            size="small",
-            c="b",
+            c=text_color,
             va="center",
         )
-        axs[ix].text(0.5, desc["50%"], desc["50%"], size="small", c="b", va="center")
+        axs[ix].text(0.5, desc["50%"], desc["50%"], c=text_color, va="center")
+
     sns.despine(bottom=True, left=True)
     plt.savefig("./Manuscript/src/figures/dist_input_data.png", dpi=300)
     plt.show()
@@ -1609,44 +1616,31 @@ def plot_distribution_variable():
     df["const"] = 1
     f, axs = plt.subplots(1, 4, figsize=(8, 3))
[…]

### 2023-08-30 main.py
paper: updated Fig 4 and 5

@@ -42,6 +42,11 @@ palette_tp = [
     "#60E693",
     "#FF362B",
 ]
+palette_tp = [
+    "#C40025",
+    "#008D3D",
+    "#0067B2",
+]
 palette_tsv = [
     "#2C45FE",
     "#4F96FF",
@@ -1728,7 +1733,9 @@ def plot_bar_tp_by_ts():
     gs = GridSpec(1, 3, figure=fig)
     ax1 = fig.add_subplot(gs[0, :-1])
     df_plot = df.groupby(x_var)[y_var].value_counts(normalize=True) * 100
-    df_plot.unstack(y_var).plot.barh(stacked=True, color=palette_tp, ax=ax1)
+    df_plot.unstack(y_var).plot.barh(
+        stacked=True, color=palette_tp, ax=ax1, linewidth=0
+    )
     ax1.set(xlabel="Percentage (%)", ylabel=var_names[x_var])
     ax1.legend(
         bbox_to_anchor=(0.5, 1.04),
@@ -1737,15 +1744,28 @@ def plot_bar_tp_by_ts():
         frameon=False,
         ncol=3,
     )
+    ax1.grid()
     ax1.grid(axis="y", ls="--")
     for ix, row in df_count.reset_index().iterrows():
         ax1.text(112, ix, int(row[y_var]), va="center", ha="right")
 
     ax2 = fig.add_subplot(gs[0, -1])
-    df.groupby(x_var)[x_var].count().plot.bar(color=palette_tsv, ax=ax2)
+    df.groupby(x_var)[x_var].count().plot.bar(color=palette_tsv, ax=ax2, linewidth=0)
     ax2.set(ylabel="", xlabel=var_names[x_var], title="Number of votes")
+    ax2.set_xticklabels(
+        [
+            "Cold",
+            "Cool",
+            "Sl. Cool",
+            "Neutral",
+            "Sl. Warm",
+            "Warm",
+            "Hot",
+        ],
+    )
     ax2.yaxis.tick_right()
-    ax2.grid(axis="x", ls="--")
+    ax2.grid()
+    ax2.grid(axis="y", ls="--")
 
     plt.savefig(f"./Manuscript/src/figures/bar_plot_tp_by_ts.png", dpi=300)
 
@@ -1760,9 +1780,7 @@ def plot_stacked_bar_predictions_ts(hb_models=False):
         fig_name = "bar_stacked_model_accuracy_hb"
 
     # Stacked boxplot
-    f, axs = plt.subplots(
[…]

### 2023-08-30 main.py
paper: updated more figures

@@ -1409,7 +1409,7 @@ def scatter_plot(data, ind="tsv", x_jitter=0):
             va="center",
         )
 
-        color = "#FDB515"
+        color = "#53626F"
         if model == "pmv":
             color = "#3B7EA1"
 
@@ -1459,7 +1459,7 @@ def plot_error_prediction(data):
         split=True,
         hue="model",
         inner="quartile",
-        palette=["#3B7EA1", "#FDB515"],
+        palette=["#3B7EA1", "#53626F"],
     )
     acceptable_error = 0.5
     axs.fill_between(
@@ -1496,7 +1496,7 @@ def plot_error_prediction(data):
     axs.legend(
         handles=[
             patches.Patch(color="#3B7EA1", label="PMV"),
-            patches.Patch(color="#FDB515", label="PMV$_{CE}$"),
+            patches.Patch(color="#53626F", label="PMV$_{CE}$"),
         ],
         frameon=False,
         loc=1,
@@ -1572,7 +1572,7 @@ def plot_distribution_variable():
 
     for ix, var in enumerate(["ta", "tr", "rh", "vel", "clo", "met"]):
         sns.boxenplot(
-            y=var, data=df, ax=axs[ix], color="#FDB515", showfliers=False, linewidth=0.5
+            y=var, data=df, ax=axs[ix], color="#53626F", showfliers=False, linewidth=0.5
         )
         axs[ix].set(
             ylabel="",
@@ -1622,7 +1622,7 @@ def plot_distribution_variable():
     f, axs = plt.subplots(1, 4, figsize=(8, 3))
     for ix, var in enumerate(["age", "ht", "wt", "t_mot_isd"]):
         sns.boxenplot(
-            y=var, data=df, ax=axs[ix], color="#FDB515", showfliers=False, linewidth=0.5
+            y=var, data=df, ax=axs[ix], color="#53626F", showfliers=False, linewidth=0.5
         )
         axs[ix].set(xlabel=var_names[var], xticks=[], ylabel="")
         if var == "age":
@@ -1652,6 +1652,7 @@ def plot_distribution_variable():
 
 
 def plot_bubble_models_vs_tsv():
+    plt.close("all")
     # Scatter thermal_sensation vs pmv prediction
     f, axs = plt.subplots(
         1, len(models_to_test), sharex=True, sharey=True, constrained_layout=True
@@ -1673,23 +1674,22 @@ def plot_bubble_models_vs_tsv():
[…]

### 2023-08-31 Manuscript/src/sections/results.tex
paper: updated Fig 4 and 5
paper: updated more figures
paper: updated description comparison TSV with TPV

@@
 While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
-% todo I really do not like this color, they are misleading.
-If someone want to be "warmer", then it should be shown as blue.
-I suggest to use the color pallet (figure 3) that Micheal used in the BCA paper.
-% todo TSV report words not numbers.
-% todo TSV chart the color here should not match the one on the thermal preference scale given that it is a different question.
+% todo:
+%  FT - I like my original colour more
+%  SS - TSV chart the color here should not match the one on the thermal preference scale given that it is a different question.
 For example, a TSV=3 is red, like people that wants to be warmer, meaning that they are cold.
@@
 Better to use two independet color schemes.
+%  FT - then also Michael color scheme is wrong since we cannot assume that someone that want to be warm must be cold, they could be already slightly warm but at the same time they may want to be warmed.
 \begin{figure*}[htb!]
@@
 In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
-However, in the \ac{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
+However, in the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
 This finding challenges the above assumption.
-Assuming that `slightly warm' or `slightly cool' is comfortable leads to under-counting thermal discomfort.
-This result highlights one of the major limitations of the \ac{tsv} scale and suggests that the thermal preference scale is better suited to determine how people perceive their thermal environment.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in this example.
+On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
+Of those who reported to be `cool' \qty{23}{\percent} of them wanted `no change' and \qty{5}{\percent} wanted to be `cooler'.
+Similarly, but in reverse, among those who reported to be `warm' \qty{13}{\percent} wanted `no change' and \qty{2}{\percent} wanted to be `warmer'.
+This result highlights one of the major limitations of the \ac{tsv} scale and of the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable.
+This suggests that the thermal preference scale is better suited to determine how people perceive their thermal environment.
 The \ac{tpv} more clearly allows an HVAC system to decide whether to change or not the thermal environment. 
@@
 The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are here comparing the prediction accuracy of the two formulations.
-% todo Use the words, not the numbers.
-Please use the color pallet from Toby's paper.
-It is hard to notice that these are two distint graph because the legend is moving accross.
-I suggest to move the legend above and have a thin grey line to separate the two group, another option would be to add a light grey background for the two PMV and show that they are distinct.
-% todo Move the percentage, on top and move it horizontally.
-% todo remove the white space between the boxes (horizontal lines)
-% todo TSV use words not numbers
-% todo Move this also above the figure, over each column. I would also add the term. Overall accuracy  PMV = 33%
+% todo:
+%  We can use Toby's or Michael but not both, even, though I like mine more
+%  It is hard to notice that these are two distint graph because the legend is moving accross.
+%  and have a thin grey line to separate the two group, another option would be to add a light grey background for the two PMV and show that they are distinct.
 \begin{figure*}[htb!]
@@
 The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}).
-% todo in the table I should also report the F1 scores if I filter out some of the TSV.
-The PMV model should only be used in the range -2 and 2 so at least we should report also the results in this range.
+% todo:
+%  in the table I should also report the F1 scores if I filter out some of the TSV.
+%  The PMV model should only be used in the range -2 and 2 so at least we should report also the results in this range.
 \begin{table}[htb!]
@@
 To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
-% todo I would not use grey for the data, it is too similar to the gridlines.
-A light color is fine, light brown (depending on the color pallet used in the rest of the paper) could work.
-The regression line could be done even thicker.
-Be more explicit in the graph of what is measured and what is calculated.
-The Y axis should be labeled as PMV
+% todo:
+%  Be more explicit in the graph of what is measured and what is calculated.
 
@@
 Results are presented in Figure~\ref{fig:hist_discrepancies}.
-% todo You need to check of the four distibution are normally distributed, if so, report mean and SD, if not report median and IQR, no need to report all of them.
-I guess that Pswetest is an indication of this, so, for the entire dataset it seems that it is not sufficiently normally distributed, so report only median and IQR.
-The Pswetest This index is not reported, explained, and interpreted in the paper.
-I would like to have these figure a bit longer in the Y-axis.
-they are kind of squized.
-The word "delta" does not mean difference.
-So, I suggest to use "diffference" or just show the equation Bias=PMVi-TSVi.
-1) Can this chart be transformed in a letter value plot?
-The Caption is missing
+% todo:
+%  You need to check of the four distibution are normally distributed, if so, report mean and SD, if not report median and IQR, no need to report all of them.
+%  I guess that Pswetest is an indication of this, so, for the entire dataset it seems that it is not sufficiently normally distributed, so report only median and IQR.
+%  The Pswetest This index is not reported, explained, and interpreted in the paper.
+%  Can this chart be transformed in a letter value plot?
+%  The Caption is missing
 \begin{figure*}[htb!]

### 2023-08-31 main.py

@@ -14,7 +14,6 @@ from matplotlib.gridspec import GridSpec
 from pythermalcomfort import p_sat_torr, check_standard_compliance_array
 from pythermalcomfort.models import (
     athb,
-    set_tmp,
     cooling_effect,
 )
 from scipy import stats
@@ -23,8 +22,6 @@ from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score
 from sklearn.metrics import f1_score
 from sklearn.metrics import r2_score, mean_absolute_error
-from sklearn.linear_model import LinearRegression
-from itertools import product
 
 warnings.filterwarnings("ignore")
 
@@ -1730,24 +1727,50 @@ def plot_bar_tp_by_ts():
         "\\percent",
     )
     fig = plt.figure(constrained_layout=True)
-    gs = GridSpec(1, 3, figure=fig)
+    gs = GridSpec(1, 3, figure=fig, wspace=0.15)
     ax1 = fig.add_subplot(gs[0, :-1])
     df_plot = df.groupby(x_var)[y_var].value_counts(normalize=True) * 100
     df_plot.unstack(y_var).plot.barh(
         stacked=True, color=palette_tp, ax=ax1, linewidth=0
     )
+    for ix, row in df_plot.unstack(y_var).round(0).reset_index(drop=True).iterrows():
+        x_shift = 0
+        print(ix)
+        for el in row:
+            print(el)
+            if el > 3:
+                ax1.text(
+                    x_shift + el / 2,
+                    ix,
+                    f"{int(el)}%",
+                    c="white",
+                    ha="center",
+                    va="center",
+                )
+            x_shift += el
     ax1.set(xlabel="Percentage (%)", ylabel=var_names[x_var])
     ax1.legend(
-        bbox_to_anchor=(0.5, 1.04),
+        bbox_to_anchor=(0.5, 0.975),
         loc="lower center",
         borderaxespad=0,
         frameon=False,
         ncol=3,
     )
     ax1.grid()
-    ax1.grid(axis="y", ls="--")
+    ax1.grid(axis="x", ls="--")
+    ax1.set_yticklabels(
+        [
+            "Cold",
+            "Cool",
+            "Sl. Cool",
+            "Neutral",
+            "Sl. Warm",
+            "Warm",
+            "Hot",
+        ],
+    )
     for ix, row in df_count.reset_index().iterrows():
[…]

### 2023-08-31 Manuscript/src/sections/results.tex
paper: updated per-mode units, added bar chart for v>=.2, updated table

@@
     \centering
-    \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy}
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0}
+        \caption{All data}
+     \end{subfigure}
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0.2}
+        \caption{Only data with \ac{v} $\geq$ \qty{0.2}{\m\per\s}}
+     \end{subfigure}
     \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants. 
@@
     The overall prediction accuracy is shown above the chart while the model accuracy of predicting participants grouped by their \ac{tsv} is overlayed over each bar. 
-    A random model, randomly selecting a thermal sensation vote would have an overall accuracy of \qty{14}{\percent}}
+    A random model, randomly selecting a thermal sensation vote would have an overall accuracy of \qty{14}{\percent}.}
     \label{fig:bar_stacked_model_accuracy}

### 2023-08-31 main.py

@@ -1793,10 +1793,10 @@ def plot_bar_tp_by_ts():
     plt.savefig(f"./Manuscript/src/figures/bar_plot_tp_by_ts.png", dpi=300)
 
 
-def plot_stacked_bar_predictions_ts(hb_models=False):
+def plot_stacked_bar_predictions_ts(hb_models=False, v_min=0):
     plt.close("all")
 
-    fig_name = "bar_stacked_model_accuracy"
+    fig_name = f"bar_stacked_model_accuracy_{v_min}"
     models = models_to_test
     if hb_models:
         models = [f"lr_hb_{x}" for x in models_to_test[:-1]]
@@ -1808,8 +1808,9 @@ def plot_stacked_bar_predictions_ts(hb_models=False):
 
     for ix, pmv in enumerate(models):
         var = f"{pmv}_round"
+        df_v = df[df["vel"] >= v_min]
         df_plot = (
-            df.groupby("thermal_sensation_round")[var]
+            df_v.groupby("thermal_sensation_round")[var]
             .value_counts(normalize=True)
             .unstack(var)
         )
@@ -1821,10 +1822,10 @@ def plot_stacked_bar_predictions_ts(hb_models=False):
                     df_plot[x] = np.nan
         df_plot = df_plot[df_plot.columns.sort_values()]
         df_plot.plot.bar(
-            stacked=True, color=palette_tsv, ax=axs[ix], rot=0, linewidth=0
+            stacked=True, color=palette_tsv, ax=axs[ix], rot=0, linewidth=0, width=0.85
         )
         accuracy = round(
-            accuracy_score(df[var].fillna(999), df["thermal_sensation_round"]) * 100
+            accuracy_score(df_v[var].fillna(999), df_v["thermal_sensation_round"]) * 100
         )
         axs[ix].set(xlabel="")
         axs[ix].set_title(f"Overall accuracy {var_names[pmv]} {accuracy}%")
@@ -1865,25 +1866,28 @@ def plot_stacked_bar_predictions_ts(hb_models=False):
         )
 
     plt.subplots_adjust(left=0.05, right=1, bottom=0.15, top=0.85)
-    f.supxlabel(var_names["thermal_sensation"])
     cax = plt.axes([0, 0.95, 1, 0.05])
     cax.axis("off")
 
-    cax.legend(
-        handles,
-        [
-            "Cold",
-            "Cool",
-            "Sl. Cool",
-            "Neutral",
-            "Sl. Warm",
-            "Warm",
-            "Hot",
-        ],
[…]

### 2023-09-01 Manuscript/src/sections/results.tex
paper: added reference and fixed sentence

@@
 We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
-The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume (xxx).
+The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
 The threshold in the standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
@@
 In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
-The width of the violin plot is not proportional to the number of points in that specific bin.
+All the violin plots have the same area to aid visualization, despite the fact that the number of points in each bin is not constant.
 % todo This figure needs a lot of improvements.

### 2023-09-01 Manuscript/src/sections/results.tex
Updates from Overleaf

--- 2023-08-31 Manuscript/src/sections/results.tex
@@
 
-The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used.
+The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
 However, our results show that the \ac{pmv-ce} is less accurate than the \ac{pmv} in predicting people who were `slightly warm'.

### 2023-09-01 Manuscript/src/sections/results.tex
paper: meeting with Stefano

@@
 % TODO:
-%  remove the gridlines or marker them very short
-%  the chart is a bit dense of information
+%  azzurrino or golden or marrone
+% remove numbers and stop at 97.5 with the box and 2.5
+% max ticks on the y line is 3 to 5, in some figure I have 10
 \begin{figure*}[htb!]
@@
 % todo:
-%  FT - I like my original colour more
-%  SS - TSV chart the color here should not match the one on the thermal preference scale given that it is a different question.
-For example, a TSV=3 is red, like people that wants to be warmer, meaning that they are cold.
-Even with the inversion of the color palette, we cannot imply that TSV=3 mease wanted to be cooler.
-What about TSV=1, maybe people want no changer or cooler or warmer.
-Difficult to assume.
-Better to use two independet color schemes.
-%  FT - then also Michael color scheme is wrong since we cannot assume that someone that want to be warm must be cold, they could be already slightly warm but at the same time they may want to be warmed.
+% barre piu spesse
+% y axis less tick
+% use K instead of 000
+% remove gridlines
+% use Toby's colour
+
 \begin{figure*}[htb!]
@@
 % todo:
-%  We can use Toby's or Michael but not both, even, though I like mine more
-%  It is hard to notice that these are two distint graph because the legend is moving accross.
+%  We can use Toby's 
+%  call charts a,b,c,d
 %  and have a thin grey line to separate the two group, another option would be to add a light grey background for the two PMV and show that they are distinct.
@@
 %  The PMV model should only be used in the range -2 and 2 so at least we should report also the results in this range.
+% consider limiting the table to PMV = 0 or PMV -1
 \begin{table}[htb!]
@@
 The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
-However, our results show that the \ac{pmv-ce} is less accurate than the \ac{pmv} in predicting people who were `slightly warm'.
 Both models had an accuracy lower than \qty{1}{\percent} when predicting either `warm', `hot', or `cold'.
@@
 The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
+% todo call it Standards implications, limit use PMV and of environmental variables
 Based on the results, we suggest that the Standards applicability limits could be reduced to the following range \qty{18}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{30}{\celsius} (instead of \qty{10}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius}),  until more data are collected.
@@
 \subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model}
+% todo consider moving to appendix and explain well why I have done this.
 We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx.

### 2023-09-01 Manuscript/src/sections/methodology.tex
paper: meeting with Stefano
paper: reviewed the paper and addressed most of Stefano's first round of comments

@@
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with $|$\ac{tsv}$|$ or $|$\ac{pmv}$|$~$\leq$~\num{3}.
+However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary we decided keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
 We used Python to analyse and visualise the results.
-We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \textbf{provide URL} so other users can test different assumptions.  
-% todo add URL
+We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \url{https://github.com/FedericoTartarini/paper-pmv-comparison} so other users can test different assumptions.
 
 \subsection{Model Validation}\label{subsec:model-validation}
-The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants and it is applied from a singly occupied zone to an area with several hundred people.
-While the \ac{db2} contains data from a sufficiently wide population no information is available on where and when exactly these data were collected.
-We only have access to building-level data and no information about whether a group of occupants was sharing the same thermal environment.
+The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, and it is applied from a singly occupied zone to an area with several hundred people.
 We grouped the participants' responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
@@
 The standard deviation would reflect the combined effect of the people's individual differences as well as any errors in formulation or in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
-According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between -0.25 and 0.25.
+According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
+However, we believe that this range is too narrow since \ac{tsv} is measured using an ordinal scale of integers while the output of the \ac{pmv} model is continuous.
+Consequently, we expanded the range to \num{-.5} and \num{0.5}.
+Thi is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
+We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally we would want to have the \ac{pmv} to have a higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
 Binning data by the independent variables is a significant improvement over grouping the participants' responses by their \ac{tsv}.
-This is because the \ac{pmv} model aim is, as previously mentioned, to predict the average thermal sensation of a large group of occupants sharing the same environment and not to correctly predict the average \ac{tsv} of people who reported the same \ac{tsv} vote, but where in different locations and moments of the year and day.
-% todo why not 0.5 since the delta between a PMV of 2.5 and a self-reported vote of either 2 or 3 (the closest points to 2.5) is 0.5?
+This is because the \ac{pmv} model aim is, as previously mentioned, to predict the average thermal sensation of a large group of occupants and not to correctly predict each individual \ac{tsv} of people who reported the same \ac{tsv} vote and where in different locations and moments of the year and day.
 

### 2023-09-01 main.py

@@ -45,13 +45,13 @@ palette_tp = [
     "#0067B2",
 ]
 palette_tsv = [
-    "#2C45FE",
-    "#4F96FF",
-    "#6CDFFF",
-    "#60E693",
-    "#FFDF6B",
-    "#FFB36B",
-    "#FF362B",
+    "#8DA2D8",
+    "#A9D4F5",
+    "#D1EAFA",
+    "#D2E8BF",
+    "#F0BECB",
+    "#DE7B6A",
+    "#A65558",
 ]
 
 # sns.palplot(palette_tsv)
@@ -1569,37 +1569,41 @@ def plot_distribution_variable():
 
     for ix, var in enumerate(["ta", "tr", "rh", "vel", "clo", "met"]):
         sns.boxenplot(
-            y=var, data=df, ax=axs[ix], color="#53626F", showfliers=False, linewidth=0.5
+            y=var,
+            data=df,
+            ax=axs[ix],
+            color=c_gold,
+            showfliers=False,
+            linewidth=0.5,
+            outlier_prop=0.05,
+            k_depth="proportion",
         )
         axs[ix].set(
             ylabel="",
             xlabel=f"{var_names[var]} ({var_units[var]})",
-            ylim=(applicability_limits[var][0], applicability_limits[var][1]),
         )
-        desc = df[var].describe(percentiles=[0.025, 0.25, 0.5, 0.75, 0.975])
-        if var == "ta":
+        if var == "ta" or var == "tr":
+            axs[ix].set(yticks=(np.linspace(15, 30, 4)), ylim=(15, 30))
+        if var == "rh":
             axs[ix].set(
-                ylim=(applicability_limits["tr"][0], applicability_limits["tr"][1]),
+                yticks=(np.linspace(20, 80, 4)),
+                ylim=(20, 80),
+            )
+        if var == "vel":
+            axs[ix].set(
+                yticks=(np.linspace(0, 0.6, 4)),
+                ylim=(0, 0.6),
             )
         if var == "clo":
-            axs[ix].set(yticks=(np.arange(0, 1.8, 0.3)))
-
-        desc = desc.round(2)
-        if var in ["ta", "tr", "rh", "met", "clo"]:
-            desc = desc.round(1)
-        if var in ["rh"]:
-            desc = desc.astype(int)
-
-        text_color = "#00B0Da"
-        axs[ix].text(0.5, desc["2.5%"], desc["2.5%"], c=text_color, va="center")
-        axs[ix].text(
-            0.5,
-            desc["97.5%"],
[…]

### 2023-09-01 Manuscript/src/sections/methodology.tex
Updates from Overleaf

--- 2023-08-29 Manuscript/src/sections/methodology.tex
@@
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with $|$\ac{tsv}$|$ or $|$\ac{pmv}$|$~$\leq$~\num{3}.
+However, since the thermal sensation was measured with at least a seven-point scale we decided also to show entries with $|$\ac{tsv}$|$ or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
 We used Python to analyse and visualise the results.
@@
 The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants and it is applied from a singly occupied zone to an area with several hundred people.
-While the \ac{db2} contains data from a sufficiently wide population no information is available on where and when exactly these data were collected.
-We only have access to building-level data and no information about whether a group of occupants was sharing the same thermal environment.
 We grouped the participants' responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
@@
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between -0.25 and 0.25.
+% todo we need to specify that we used the range 0.5 to cover the whole scale, this could potentially be an error up to 17% but is needed not to leave gaps, since thermal sensation is measured ordinally.
 Binning data by the independent variables is a significant improvement over grouping the participants' responses by their \ac{tsv}.

### 2023-09-01 Manuscript/src/sections/introduction.tex

@@
 In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
-% todo also mention the other standards that use the PMV The PMV index has been widely used in national and international standards and guidance ASHRAE 55 [4], EN 15251 (new version EN 16798 [5]) [6], ISO 7730 [1], CIBSE Guide A [7] and GB/T 50785 [110] as shown in  Table 5.
-This sentence is from Yao 2022.
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
 It is the most extensively used thermal comfort index for estimating thermal sensation.
+%ASHRAE's defines the \ac{pmv} as ``an index that predicts the mean value of the thermal sensation votes (self-reported perceptions) of a large group of persons on a sensation scale expressed from –3 to +3 corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{ashrae552023}.
 A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
@@
 The resulting values become the new inputs in the \ac{pmv} model.
-Since the \ac{ce} accounts for the convective heat losses from the person to its environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+Since the \ac{ce} accounts for the convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
 The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
 Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
-The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
+%The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
+
 However, this method has limitations.
@@
 Their analysis only reports the mean bias of the different models.
-This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate in predicting \ac{tsv}.
-This is because the \ac{pmv} model has a higher prediction accuracy of sensation close to neutrality. 
-Being the ASHRAE Global Thermal Comfort Database II v2.1 an unbalanced dataset and most of the data points are neutral a more detailed analysis as the one we are presenting in this manuscript is needed to compare the accuracy of both formulations.
+This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate \ac{tsv} since it does not explain how the model performs over a wide range of environmental and personal conditions.
+Their analysis also does not allow to determine the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv}.
+This is particularly for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral' when asked to cast their \ac{tsv}.
+Finally, the overall bias is reported for the whole dataset, despite the fact that the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
 

### 2023-09-01 Manuscript/src/sections/results.tex

@@
 While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
-The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
-The 2.5th, 50th (median), and 97.5th percentiles are shown in blue.
-% TODO:
-%  azzurrino or golden or marrone
-% remove numbers and stop at 97.5 with the box and 2.5
-% max ticks on the y line is 3 to 5, in some figure I have 10
+The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data} in which we excluded the data points below or above the 2.5th and 97.5th percentiles, respectively.
 \begin{figure*}[htb!]
@@
     \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
-    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.
+    We did not plot data points below or above the 2.5th and 97.5th percentiles, respectively.
     The data are shown using boxen plots (letter-value plots).
@@
 Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
-% todo:
-%  I merged men and women since not considering sex difference in the paper.
-%  This figure can be merged with figure 2 because it is a visualization of the database.
-%  To be more consisten in the paper, I would use the letter value plots for this and the figure in section 3.3.2
-% remove the gridlines or marker them very short;
 \begin{figure*}[htb!]
@@
 While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
-% todo:
-% barre piu spesse
-% y axis less tick
-% use K instead of 000
-% remove gridlines
-% use Toby's colour
-
 \begin{figure*}[htb!]
@@
 The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are here comparing the prediction accuracy of the two formulations.
+The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s};
+hence we are also presenting the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
+We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
+We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
+The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
+The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
 % todo:
-%  We can use Toby's 
 %  call charts a,b,c,d
-%  and have a thin grey line to separate the two group, another option would be to add a light grey background for the two PMV and show that they are distinct.
 \begin{figure*}[htb!]
@@
 Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
+% todo describe the results for V>0.2
 A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}=0.
@@
 The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}).
-% todo:
-%  in the table I should also report the F1 scores if I filter out some of the TSV.
-%  The PMV model should only be used in the range -2 and 2 so at least we should report also the results in this range.
-% consider limiting the table to PMV = 0 or PMV -1
 \begin{table}[htb!]
@@
 
+% todo describe the table results
 The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
@@
 % todo call it Standards implications, limit use PMV and of environmental variables
-Based on the results, we suggest that the Standards applicability limits could be reduced to the following range \qty{18}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{30}{\celsius} (instead of \qty{10}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius}),  until more data are collected.
-
+Based on our results we recommend restricting the use of the \ac{pmv} only to a values $\lvert \textrm{PMV}\lvert \leq 1.5$.
+In addition, we also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} Standards.
+Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
+Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
+In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
+Consequently, we suggest reducing both Standards applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
+\begin{table}[htb!]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        Variable & Proposed range \\
+        \midrule
+        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
+        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
+        \ac{rh} & \qtyrange{20}{80}{\percent} \\
+        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
+        \ac{met} & \qtyrange{1}{2}{met} \\
+        \ac{pmv} & \qtyrange{-1.5}{1.5}{} \\
+        % todo shall we also include limits on Age?
+        % todo shall we also include airspeed?
+        \bottomrule
+    \end{tabular}
+    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} Standards.}
+    \label{tab:ranges}
+\end{table}
 Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
 To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
-% todo:
-%  Be more explicit in the graph of what is measured and what is calculated.
-
 \begin{figure*}[htb!]
@@
     \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
-    Raw data were then binned and a bubble chart is superimposed over the regression curve to aid the visualization of a large dataset.}
+    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) is superimposed over the regression curve to aid the visualization of a large dataset.}
     \label{fig:bubble_models_vs_tsv}
@@
 This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
-This was already done in \mycite{Cheung2019}, where they grouped people based on the same thermal environment (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified how many times the \ac{pmv} model correctly predicted \ac{tsv}.
+This was already done in \mycite{Cheung2019}, where they grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified how many times the \ac{pmv} model correctly predicted \ac{tsv}.
 They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
@@
 % todo:
-%  You need to check of the four distibution are normally distributed, if so, report mean and SD, if not report median and IQR, no need to report all of them.
-%  I guess that Pswetest is an indication of this, so, for the entire dataset it seems that it is not sufficiently normally distributed, so report only median and IQR.
-%  The Pswetest This index is not reported, explained, and interpreted in the paper.
-%  Can this chart be transformed in a letter value plot?
+%  describe in the paper that we used test to determine that data are not normally distributed.
+%  We cannot change into a letter value plot since in that case it would be hard to highlight the number of responses that are within the range from -.5 to.5.
 %  The Caption is missing
@@
 Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
-The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s}; we also calculated the bias of the model, including data points with \ac{v} higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
-We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
-We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
-The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
-The threshold in the standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
 The results show that the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.

### 2023-09-05 Manuscript/src/sections/results.tex
paper: improved figure bias binned by var

@@
 We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{met}, and \ac{clo}.
+% todo add humidity ratio chart
+We did not include \ac{tr} in the analysis since \ac{tr} was not reported in several studies, hence, we estimated it as detailed in the Methodology.
+Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
 Figure~\ref{fig:dist_input_data} shows the results.
@@
 All the violin plots have the same area to aid visualization, despite the fact that the number of points in each bin is not constant.
-% todo This figure needs a lot of improvements.
-1) I would use the letter-value (or boxplot), they may be easier to interpret, provide less info and could be aligned with the first figure (in particular if  The actual density distribution it is not adding much value.
-2) The color for PMV and PMV_CE must be the same in the entire paper.
-3) The y-label should be moved to the x-label and the y-label should explicit.
-I would remove the decimal values and show only the five integers -2, -1, 0, 1 and 2.
-4) How is possible to have a velocity "centered" at 0?
-I would remove the red lines.
-If you want to underline the -0.5 and 0.5 range, then we should use the same strategy also for figure 7.
-The fact that you change graph time in the paper, it makes more difficult to interpret without adding a substantial amount of value.
+% todo:
+%  I need to have the red lines for the same inverse reason why I cannot use the boxenplot in the previous chart
+
 \begin{figure*}[htb!]
@@
     \includegraphics[width=\textwidth]{figures/bias_models}
-    \caption{}
+    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.}
     \label{fig:bias_models}

### 2023-09-05 main.py

@@ -2237,7 +2237,7 @@ def plot_bias_distribution_by_variable_binned():
         # "top",
         # "t_mot_isd",
         "vel",
-        # "rh",  # todo report pa rather than RH
+        "rh",
         "clo",
         "met",
         # "thermal_sensation",
@@ -2246,11 +2246,12 @@ def plot_bias_distribution_by_variable_binned():
     ]
 
     bins = {
-        "ta": np.arange(18.5, 30.5, 1),
-        "tr": np.arange(18.5, 31.5, 1),
-        "top": np.arange(17.5, 30.5, 1),
-        "vel": np.arange(-0.05, 0.50, 0.1),
-        "rh": np.arange(17.5, 75, 5),
+        "ta": np.arange(18, 32, 2),
+        "tr": np.arange(18, 31.5, 2),
+        "top": np.arange(17.5, 30.5, 2),
+        "vel": np.arange(-0.00001, 0.7, 0.2),
+        "rh": np.arange(10, 100, 20),
+        "hr": np.arange(0, 20, 2),
         "clo": np.arange(0.2, 1.5, 0.2),
         "met": np.arange(0.9, 2, 0.2),
         "thermal_sensation": np.arange(-3.5, 4.5, 1),
@@ -2264,6 +2265,15 @@ def plot_bias_distribution_by_variable_binned():
     # todo only look at high velocities
 
     df_analysis = df.copy()
+
+    # # Code to calculate the Humidity Ratio
+    # hr_arr = []
+    # for i, row in df_analysis.iterrows():
+    #     hr_arr.append(
+    #         psychrolib.GetHumRatioFromRelHum(row["ta"], row["rh"] / 100, 101325) * 1000
+    #     )
+    # df_analysis["hr"] = hr_arr
+
     # for i, var in enumerate(variables):
     #     if var != "thermal_preference":
     #         df_analysis = df_analysis[
@@ -2275,7 +2285,7 @@ def plot_bias_distribution_by_variable_binned():
 
     # for ix, model in enumerate(models_to_test):
     #     color = palette_primary[ix]
-    f, axs = plt.subplots(2, 2, constrained_layout=True, figsize=(7, 5))
+    f, axs = plt.subplots(5, 1, figsize=(7, 9))
     axs = axs.flatten()
     for i, var in enumerate(variables):
         # plot bias distribution
@@ -2324,47 +2334,49 @@ def plot_bias_distribution_by_variable_binned():
[…]

### 2023-09-05 Manuscript/src/sections/results.tex
accepted some of the comments from Stefano

@@
 Moreover, we tried to explain why we observed that the \ac{pmv-ce} model has a lower accuracy than the \ac{pmv} in predicting thermal sensation.
-The issues are not reported in order of importance and more research is needed to determine how to improve the overall accuracy of the \ac{pmv} model.
-
-\paragraph{Lack of Data}
-T.
+The issues are reported in order of importance of what we believe may affect the accuracy of the model.
+More research is needed to determine how to improve the overall accuracy of the \ac{pmv} model.
 

### 2023-09-05 Manuscript/src/sections/results.tex
paper: converted comments into todos

@@
 
-\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model}
+\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model} % todo Stefano Schiavon: call it Gagge PMV model in line with the text below
 % todo consider moving to appendix and explain well why I have done this.
-We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx.
+Stefano Schiavon: The logic and need of this paragraph needs to be better explained.
+Why did you add?
+Why can we assume that the PMV-SET is better?
+Who developed the PMV-SET?
+We should discuss if we need this paragraph and if it add something that is important and not reported in other sections.
+In other to do this, we need to explain why we think the PMV-SET is important even if it is not present in standards.
+
+We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx. % todo provide information
+% todo: Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
 Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
@@
 The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
+% todo Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
 The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
+% todo Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from?
+How can a PMV model be described by an air speed?
 These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
@@
 This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
+% todo Stefano Schiavon: Please add references suportting this.
 Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
@@
 Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
+% todo Stefano Schiavon: This is vague, Which papers or data is available to update that equation/mapping?
 In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.
@@
 On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
-We showed that people tend to be closer to neutrality than what the models predicted (see figure XXX), this is an indication that they were exposed to compensable thermal conditions.
+% todo we need to check this since we may only recommend the use of the PMV model for thermal neutrality
+We showed that people tend to be closer to neutrality than what the models predicted (see Figure XXX), this is an indication that they were exposed to compensable thermal conditions.
+% todo add figure number
 Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.
@@
 i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
-ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress are constant and not xxx;
-iii) the regression coefficient between heat losses and heat gains and thermal stress are not the same for the heat losses and gains and two different equations should be used;
+% todo Stefano Schiavon: The difference between these two points ii) and iii) are not clear, can you add more explanation?
+ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress;
+iii) the regression coefficient between heat losses and heat gains and thermal stress is not linear and two different equations should be used;
 iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
@@
 It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
+% to do check that the fourth below still refers to the anchors issue above
 The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}

### 2023-09-05 Manuscript/src/sections/results.tex
paper: merged overleaf changes with local

@@
 This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
-However, the \ac{pmv} model erroneously assumes that the human body, aside from some edge cases, is either losing or gaining heat from its surrounding environment.
+Moreover, the \ac{pmv} model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
 This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
@@
 Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
+% todo talk also about how sweat is calculated
 
 The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
-Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
+Fanger created this equation based on a limited dataset collected in a laboratory, and despite the growing body of literature, this equation was never updated or revisited~\cite{Fanger1970}.
 % todo Stefano Schiavon: This is vague, Which papers or data is available to update that equation/mapping?
-In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.
+In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions, respectively~\cite{romanovsky_thermoregulation_2018}.
 
 It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
-This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis.
+This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis~\cite{ romanovsky_thermoregulation_2018}.
 On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).

### 2023-09-11 Manuscript/src/sections/appendix.tex
fix: some pmv values were missing since were outside the applicability limits of the standard

--- 2023-09-05 Manuscript/src/sections/results.tex
@@
-%! Author = sbbfti
-%! Date = 10/06/2020
-\section{Results and Discussion}\label{sec:results}
+\appendix
 
-\subsection{\ac{db2} overview}\label{subsec:comfort-db-overview}
-While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
-The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data} in which we excluded the data points below or above the 2.5th and 97.5th percentiles, respectively.
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/dist_input_data}
-    \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
-    We did not plot data points below or above the 2.5th and 97.5th percentiles, respectively.
-    The data are shown using boxen plots (letter-value plots).
-    They depict the median as the center-line and each successive level outward contains half of the remaining data.}
-    \label{fig:dist_input_data}
-\end{figure*}
-The input data distributions are far from uniformity (equal density) across the Standards' applicability ranges.
-We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
-For example, only 2.5\% of the available dataset had a \ac{tdb} $\leq$ than \var{ta_95_perc_min}, this may consequently be an issue since both thermal comfort Standards' applicability limits extend to \qty{10}{\celsius}.
-This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
-This is because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
-This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
-The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v} $\geq$ \qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
-Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} $\geq$ \var{v_95_perc_max}.
-
-Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature.
-Less than half (\num{23300}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
-Ages are not normally distributed, the same is true for the running mean outdoor temperature.
-About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old, and only \qty{2.5}{\percent} were older than 60.
-This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
-The \ac{db2} does not contain information about the health status of the participants.
-Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/dist_other_data}
-    \caption{Distribution of age, height, weight, and running mean outdoor temperature.
-    The data are grouped by sex.
-    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.}
-    \label{fig:dist_other_data}
-\end{figure*}
-The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
-Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
-Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
-The \ac{tsv} is unbalanced. 
-Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
-While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
-    \caption{The left Figure shows the percentage of \ac{tpv} for each thermal sensation vote.
-    The numbers on the right side of each bar show the number of points for each \ac{tsv}.
-    The right figure shows the total number of data points grouped by \ac{tsv}.
-    Each bin in the right figure has more data points than in the left figure since less data on \ac{tpv} were available.}
-    \label{fig:bar_plot_tp_by_ts}
-\end{figure*}
-
-In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
-However, in the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
-This finding challenges the above assumption.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in this example.
-On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
-Of those who reported to be `cool' \qty{23}{\percent} of them wanted `no change' and \qty{5}{\percent} wanted to be `cooler'.
-Similarly, but in reverse, among those who reported to be `warm' \qty{13}{\percent} wanted `no change' and \qty{2}{\percent} wanted to be `warmer'.
-This result highlights one of the major limitations of the \ac{tsv} scale and of the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable.
-This suggests that the thermal preference scale is better suited to determine how people perceive their thermal environment.
-The \ac{tpv} more clearly allows an HVAC system to decide whether to change or not the thermal environment. 
-On the contrary, the \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
-
-\subsection{Comparison of PMV accuracy in predicting thermal sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
-The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are here comparing the prediction accuracy of the two formulations.
-The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s};
-hence we are also presenting the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
-We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
-We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
-The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
-The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
-% todo:
-%  call charts a,b,c,d
-\begin{figure*}[htb!]
-    \centering
-    \begin{subfigure}[b]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0}
-        \caption{All data}
-     \end{subfigure}
-    \begin{subfigure}[b]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0.2}
-        \caption{Only data with \ac{v} $\geq$ \qty{0.2}{\m\per\s}}
-     \end{subfigure}
-    \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants. 
-    The number of data points in each bin of the two charts is the same and this allows us to compare the accuracy of both models.
-    The overall prediction accuracy is shown above the chart while the model accuracy of predicting participants grouped by their \ac{tsv} is overlayed over each bar. 
-    A random model, randomly selecting a thermal sensation vote would have an overall accuracy of \qty{14}{\percent}.}
-    \label{fig:bar_stacked_model_accuracy}
-\end{figure*}
-Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
-Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
-The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
-Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
-% todo describe the results for V>0.2
-A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}=0.
-For this reason, in Table~\ref{tab:f1}, we report the F1-scores for all the models.
-The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}).
-\begin{table}[htb!]
-    \centering
-    \input{tables/f1.tex}
-    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models.}
-    \label{tab:f1}
-\end{table}
-
-% todo describe the table results
-The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
-Both models had an accuracy lower than \qty{1}{\percent} when predicting either `warm', `hot', or `cold'.
-This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
-The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
-% todo call it Standards implications, limit use PMV and of environmental variables
-Based on our results we recommend restricting the use of the \ac{pmv} only to a values $\lvert \textrm{PMV}\lvert \leq 1.5$.
-In addition, we also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} Standards.
-Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
-Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
-In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
-Consequently, we suggest reducing both Standards applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
-\begin{table}[htb!]
-    \centering
-    \begin{tabular}{cc}
-        \toprule
-        Variable & Proposed range \\
-        \midrule
-        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
-        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
-        \ac{rh} & \qtyrange{20}{80}{\percent} \\
-        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
-        \ac{met} & \qtyrange{1}{2}{met} \\
-        \ac{pmv} & \qtyrange{-1.5}{1.5}{} \\
-        % todo shall we also include limits on Age?
-        % todo shall we also include airspeed?
-        \bottomrule
-    \end{tabular}
-    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} Standards.}
-    \label{tab:ranges}
-\end{table}
-Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
-To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
-    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
-    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) is superimposed over the regression curve to aid the visualization of a large dataset.}
-    \label{fig:bubble_models_vs_tsv}
-\end{figure*}
-The curve, in the Figure, is calculated using the individual data and not the binned data.
-We binned the data to aid the visualization of a large dataset.
-If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and have a slope of 1.
-The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval. 
-The intercept is close to 0 and this is aligned with the results shown above that the model can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times.
-Both \ac{pmv} formulations always underpredicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv}; the slope of the curve was significantly lower than 1.
-We discuss in detail in sub-Section~\ref{subsec:sources-of-error} which factors may be the underlying cause of these errors.
-
-The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
-The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
-Consequently, while the above-mentioned analysis is needed to compare for the same sample size in each thermal sensation group the \ac{pmv} and \ac{pmv-ce} accuracies, it focuses primarily on comparing individual \ac{tsv} to \ac{pmv}.
-This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
-This was already done in \mycite{Cheung2019}, where they grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified how many times the \ac{pmv} model correctly predicted \ac{tsv}.
-They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
-
-\subsection{Model Bias}\label{sec:model-bias}
-\subsubsection{Model Overall Bias}\label{subsec:model-overall-bias}
-We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and  \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
-Results are presented in Figure~\ref{fig:hist_discrepancies}.
-% todo:
-%  describe in the paper that we used test to determine that data are not normally distributed.
-%  We cannot change into a letter value plot since in that case it would be hard to highlight the number of responses that are within the range from -.5 to.5.
-%  The Caption is missing
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
-    \caption{}
-    \label{fig:hist_discrepancies}
-\end{figure*} 
-Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
-Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
-The results show that the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
-Surprisingly, we found that the median bias for the \ac{pmv-ce} increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
-These are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
-As far as we could find, there is no published paper proving a higher accuracy of the \ac{pmv-ce} compared to the \ac{pmv} model. 
-The \ac{pmv-ce} has lower accuracy while making the code more complex and computationally intensive to run.
-
-\subsubsection{Model Bias as Function of Four Input Variables}\label{subsec:model-bias-variable}
-In the previous section, we presented the overall bias of the two \ac{pmv} formulations.
-We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{met}, and \ac{clo}.
-% todo add humidity ratio chart
-We did not include \ac{tr} in the analysis since \ac{tr} was not reported in several studies, hence, we estimated it as detailed in the Methodology.
-Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
-Figure~\ref{fig:dist_input_data} shows the results.
-In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
-All the violin plots have the same area to aid visualization, despite the fact that the number of points in each bin is not constant.
-% todo:
-%  I need to have the red lines for the same inverse reason why I cannot use the boxenplot in the previous chart
-
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/bias_models}
-    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.}
-    \label{fig:bias_models}
-\end{figure*}
-The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all four input variables.
-On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 as the value of \ac{v} increased.
-These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.
-
-\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model}
-% todo consider moving to appendix and explain well why I have done this.
-We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx.
+\section{Comparison between the PMV formulations with the \ac{pmvg}}\label{sec:comparison-between-the-pmv-formulations-with-the-two-node-model}
+We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the \ac{pmvg}.
+% SS provide more information why this was done
+% Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
 Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
@@
 The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
+% Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
 The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
+% Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from? How can a PMV model be described by an air speed?
 These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
 This is also an unexpected result and we will discuss why this may be in the following paragraph.
-% todo When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
+% SS When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
 1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.
@@
 \end{figure*}
-
-\subsection{Sources of Error}\label{subsec:sources-of-error}
-In this Section, we identify some of the main reasons that could explain the difference in the results from the two \ac{pmv} models when compared against the \ac{tsv} available in the \ac{db2}.
-Moreover, we tried to explain why we observed that the \ac{pmv-ce} model has a lower accuracy than the \ac{pmv} in predicting thermal sensation.
-The issues are reported in order of importance of what we believe may affect the accuracy of the model.
-More research is needed to determine how to improve the overall accuracy of the \ac{pmv} model.
-
-\paragraph{Heat Balance Equation and Calculation of the PMV value}
-The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
-This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
-This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
-However, the \ac{pmv} model erroneously assumes that the human body, aside from some edge cases, is either losing or gaining heat from its surrounding environment.
-This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
-Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
-
-The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
-Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
-In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.
-
-It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
-This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis.
-On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
-We showed that people tend to be closer to neutrality than what the models predicted (see figure XXX), this is an indication that they were exposed to compensable thermal conditions.
-Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.
-
-To continue using the \ac{pmv} model beyond an absolute value of 1, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
-i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
-ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress are constant and not xxx;
-iii) the regression coefficient between heat losses and heat gains and thermal stress are not the same for the heat losses and gains and two different equations should be used;
-iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
-For example, while the \ac{pmv} model may assume that a score of 3 (hot) signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches the max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
-Participants in commercial buildings may, on the contrary, deem the environment to be `Hot' as soon as they start sweating and their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.
-
-It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
-The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}
-
-\paragraph{Measurement Errors}
-All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts, however, measurement errors are inevitably present in the dataset.
-This is due to the fact that all measurement instrumentation is not always calibrated before use, the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the near proximity of the participant, and estimating clothing and metabolic rates from tables is a complex and prone to error task.
-In an ideal scenario, errors would be random and would cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
-However, this may not always be the case, for example, if a researcher often underestimated the clothing insulation of participants or their activity levels.
-
-\paragraph{Thermal Sensation Scale}
-One minor but not negligible source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
-For example, even if the model was 100~\% accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (+2)' or `hot (+3)'.
-Using a continuous, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.
-They also showed that the \ac{tsv}'s assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
-An alternative solution would be to transform the \ac{pmv} output to an ordinal and categorical seven-point scale.
-
-% \paragraph{Other neglected factors}
-% Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) that interacts with the surrounding environment.
-% Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.
-
-% todo write a paragraph trying to find the reason why the PMV CE model is less accurate.
-% todo write a paragraph about what we recommend for ASHRAE 55

### 2023-09-11 Manuscript/src/sections/conclusions.tex
paper: converted comments into todos
fix: some pmv values were missing since were outside the applicability limits of the standard

@@
 The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} Standard.
+% todo sentence below is too vague
 Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
+% todo Stefano Schiavon: Also this is too vague.
+Remember that the large majority of the readers, will just look at title, abstract and conclusions.
+The conclusions must be informative and actionable to be useful.
+These are too general.
 These disparities shed light on the limitations and strengths of the \ac{pmv} and \ac{pmv-ce} models, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
+% todo Stefano Schiavon: Almost every article could end up with such sentence. Instead of us saying that this analysis "have significant implications", let list them and let the reader decide that.
 The outcomes of this analysis have significant implications for researchers, practitioners, and policymakers involved in the field of thermal comfort assessment and standardization.
+% todo add also as a result the fact that thermal sensation -1 to 1 does not mean comfortable
+% todo report also this result: Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
+The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
+
+\subsection{Implications for the ASHRAE 55 and ISO 7730 Standards}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
+The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
+Both models had an accuracy lower than \qty{1}{\percent} when predicting either `warm', `hot', or `cold'.
+This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
+The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
+Based on our results we recommend restricting the use of the \ac{pmv} only to a values $\lvert \textrm{PMV}\lvert \leq 1.5$.
+In addition, we also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} Standards.
+Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
+Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
+In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
+Consequently, we suggest reducing both Standards applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
+\begin{table}[htb!]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        Variable & Proposed range \\
+        \midrule
+        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
+        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
+        \ac{rh} & \qtyrange{20}{80}{\percent} \\
+        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
+        \ac{met} & \qtyrange{1}{2}{met} \\
+        \ac{pmv} & \qtyrange{-1.5}{1.5}{} \\
+        % todo shall we also include limits on Age?
+        % todo shall we also include airspeed?
+        \bottomrule
+    \end{tabular}
+    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} Standards.}
+    \label{tab:ranges}
+\end{table}
+
+% todo add more implications for Standards

### 2023-09-11 Manuscript/src/sections/introduction.tex
fix: some pmv values were missing since were outside the applicability limits of the standard

@@
 \section{Introduction}\label{sec:introduction}
-In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
+\mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
+The \ac{pmv} is as an index that predicts the mean value of the thermal sensation votes (self-reported perceptions) of a large group of persons on a sensation scale expressed from –3 to +3 corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
 It is the most extensively used thermal comfort index for estimating thermal sensation.
-%ASHRAE's defines the \ac{pmv} as ``an index that predicts the mean value of the thermal sensation votes (self-reported perceptions) of a large group of persons on a sensation scale expressed from –3 to +3 corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{ashrae552023}.
-A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
-Out of those 1082 are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
+A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{1600} documents.
+Out of those \num{1082} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
 This highlights this model's extensive adoption and use among the scientific community.
@@
 
-To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \ac{pmv-ce}~\cite{ashrae552023}.
-The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some choices were implemented. 
+To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
+The best source to see how \ac{pmv-ce} works is the Standard itself, but the standard does not explain why some choices were implemented.
 A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
@@
 Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
+Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
+However, while the \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
+We could not find a rationale for this choice in the available literature.
 
@@
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
+% todo use this finding to support limit to PMV range
 Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
@@
 They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
+% todo use this finding to support limit to PMV range
 The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
@@
 \mycite{Yao2022} compared the \ac{pmv} and \ac{pmv-ce} models, however, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
-In particular, they focus significantly on how the models perform in different climates and when applied to people from different world regions, but for example, they do not report how the \ac{pmv} accuracy varies as a function of environmental and personal parameters, and \ac{tsv}.
-Their analysis only reports the mean bias of the different models.
-This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate \ac{tsv} since it does not explain how the model performs over a wide range of environmental and personal conditions.
-Their analysis also does not allow to determine the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv}.
-This is particularly for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral' when asked to cast their \ac{tsv}.
-Finally, the overall bias is reported for the whole dataset, despite the fact that the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
+They focus significantly on how the models perform in different climates and when applied to people from different world regions and their analysis only reports the mean bias of the different models.
+This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate since it does not explain how the model performs over a wide range of environmental, personal conditions, and \ac{tsv}.
+Reporting the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv} is particularly for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral'.
+Finally, they only reported the overall bias for the whole dataset, despite the fact that the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
 
@@
 We used the self-reported \ac{tsv} recorded in the \ac{db2}.
-We aim to determine which \ac{pmv} model is most accurate, this will allow policymakers to make an informed decision on which model should be used in standards.
+We aim to determine which \ac{pmv} model is most accurate and which are its applicability limits.
+This will allow policymakers to make an informed decision on which model should be used in standards.

### 2023-09-11 Manuscript/src/sections/methodology.tex
fix: some pmv values were missing since were outside the applicability limits of the standard

--- 2023-09-01 Manuscript/src/sections/methodology.tex
@@
 We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
-These constraints filtered out approximately \var{entries_db_valid} votes.
+This filtered out approximately \var{entries_db_valid} votes.
 We decided to keep entries that did not have \ac{tr}.
 However, since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: $2 t_{o} – t_{db}$.
-Alternatively, if the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
+If the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
 We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
@@
 \subsection{Model Validation}\label{subsec:model-validation}
-The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, and it is applied from a singly occupied zone to an area with several hundred people.
-We grouped the participants' responses by their \ac{tsv} and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
+The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, but it is commonly used both in singly occupied zones and in areas with several hundred people.
+We first grouped participants' responses by their \ac{tsv}, and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
 This approach is useful for performing an overall assessment of the accuracy of the models, but it introduces an error due to the rounding of both the \ac{pmv}, \ac{pmv-ce}, and \ac{tsv} values.
@@
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
-However, we believe that this range is too narrow since \ac{tsv} is measured using an ordinal scale of integers while the output of the \ac{pmv} model is continuous.
+However, we believe that this range is too narrow since \ac{tsv} is measured using an ordinal scale of integers.
 Consequently, we expanded the range to \num{-.5} and \num{0.5}.
@@
 
-Finally, we decided to compare the results of \ac{pmv} and \ac{pmv-ce} formulations with the \ac{pmvg} calculated by the two-node model.
-The two-node model was developed by \mycite{GaggeSET} and it uses a more accurate set of equations to solve the heat balance equation.
-We compared the \ac{pmv} and \ac{pmv-ce} results to the \ac{pmvg} since Gagge's two-node model is used by the \ac{pmv-ce} to calculate the \ac{ce}.
-Hence, in principle, the results of the \ac{pmv-ce} model should closely match the \ac{pmvg} values.
-We did this by passing as input to the \ac{pmv}, \ac{pmv-ce}, and two-node models the same set of \num{10000} randomly generated set of inputs. 
-Each set of inputs was sampled assuming a random uniform distribution within a predefined range. 
-We selected the range of each of the 6 input variables needed to calculate the \ac{pmv} to be equal to the 2.5th and 97.5th percentiles for all the input variables contained in the \ac{db2}.
-The only exception was for the range of \ac{v} that was randomly selected from this interval \qty{0.1}{\m\per\s}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s}.
-The rationale for this decision was that both the \ac{pmv} and \ac{pmv-ce} models use the same equations for \ac{v}~$\leq$~\qty{0.1}{\m\per\s}.
+%In \ref{sec:comparison-between-the-pmv-formulations-with-the-two-node-model} we also compare the results of \ac{pmv} and \ac{pmv-ce} formulations with the \ac{pmvg} calculated by the two-node model.
+%The two-node model was developed by \mycite{GaggeSET} and it uses a more accurate set of equations to solve the heat balance equation.
+%We compared the \ac{pmv} and \ac{pmv-ce} results to the \ac{pmvg} since Gagge's two-node model is used by the \ac{pmv-ce} to calculate the \ac{ce}.
+%Hence, in principle, the results of the \ac{pmv-ce} model should closely match the \ac{pmvg} values.
+%We did this by passing as input to the \ac{pmv}, \ac{pmv-ce}, and two-node models the same set of \num{10000} randomly generated set of inputs.
+%Each set of inputs was sampled assuming a random uniform distribution within a predefined range.
+%We selected the range of each of the 6 input variables needed to calculate the \ac{pmv} to be equal to the 2.5th and 97.5th percentiles for all the input variables contained in the \ac{db2}.
+%The only exception was for the range of \ac{v} that was randomly selected from this interval \qty{0.1}{\m\per\s}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s}.
+%The rationale for this decision was that both the \ac{pmv} and \ac{pmv-ce} models use the same equations for \ac{v}~$\leq$~\qty{0.1}{\m\per\s}.
+%
+%Interestingly, \mycite{GaggeSET} also describe how to calculate the \ac{pmvs} which uses the \ac{set} temperature to calculate the \ac{pmv} value in place of \ac{tdb} and \ac{tr}.
+%This would solve some of the uncertainties that the \ac{pmv-ce} model introduces as detailed above.
+%However, it is not possible for us to comment on why the \ac{pmv-ce} model was included in the \gls{55} instead of the \ac{pmvs} since no peer-reviewed paper or official document describe the rationale for this decision.
 
@@
 Simple accuracy can be a misleading metric to use to quantify the performance of a classification predictive model that involves an unbalanced dataset~\cite{Chawla2005}.
-We then used the F1 score when determining the performance of the \ac{pmv} and \ac{pmv-ce} in predicting the \ac{tsv} reported by a participant.
+We then also report the F1 score when determining the performance of the \ac{pmv} and \ac{pmv-ce} in predicting the \ac{tsv}.
 The F1 score is the harmonic mean of the precision and sensitivity, where an F1 score reaches its best value at 1 and the worst score at 0.
 Precision (also called positive predictive value) is the ratio of true positive overall positive results obtained from the test, while sensitivity (also known as recall or true positive rate) is the ratio of true positive over all the actual positive cases.
-The F1-micro score calculates metrics globally by counting the total true positives, false negatives and false positives.
+We report three F1-scores in the paper.
+The F1-micro score calculates metrics globally by counting the total true positives, false negatives, and false positives.
 The F1-macro calculate metrics for each label and find their unweighted mean.
 The F1-weighted calculate metrics for each label and find their average weighted by support (the number of true instances for each label).
-% todo describe also the other metrics we used like R2

### 2023-09-11 Manuscript/src/sections/results.tex
fix: some pmv values were missing since were outside the applicability limits of the standard

@@
-%! Author = sbbfti
-%! Date = 10/06/2020
 \section{Results and Discussion}\label{sec:results}
@@
 While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
-The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data} in which we excluded the data points below or above the 2.5th and 97.5th percentiles, respectively.
+The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
 \begin{figure*}[htb!]
@@
     \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
-    We did not plot data points below or above the 2.5th and 97.5th percentiles, respectively.
     The data are shown using boxen plots (letter-value plots).
@@
 We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
-For example, only 2.5\% of the available dataset had a \ac{tdb} $\leq$ than \var{ta_95_perc_min}, this may consequently be an issue since both thermal comfort Standards' applicability limits extend to \qty{10}{\celsius}.
+For example, only \qty{2.5}{\percent} of the available dataset had a \ac{tdb}~$\leq$ than \var{ta_95_perc_min}, this is a problem since both thermal comfort Standards' applicability limits extend to \qty{10}{\celsius}.
 This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
-This is because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
+Mainly because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
 This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
-The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v} $\geq$ \qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
-Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} $\geq$ \var{v_95_perc_max}.
+The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
+Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
+% todo v_95_perc_max these variables should be calculated dropping nan
 
@@
 \end{figure*}
+
 The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
@@
 Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
-The \ac{tsv} is unbalanced. 
+The \ac{tsv} dataset is unbalanced.
 Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
@@
 This finding challenges the above assumption.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in this example.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in our dataset.
 On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
@@
 \subsection{Comparison of PMV accuracy in predicting thermal sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
-The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are here comparing the prediction accuracy of the two formulations.
+The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are comparing the predicted value from the two models with measured \ac{tsv}.
 The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s};
-hence we are also presenting the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
+hence we also present the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
 We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
-We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
+We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was change to \qty{0.1}{\m\per\s} only in the \gls{55}.
 The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
 The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
-% todo:
-%  call charts a,b,c,d
+% todo call charts a,b,c,d
 \begin{figure*}[htb!]
@@
         \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0.2}
-        \caption{Only data with \ac{v} $\geq$ \qty{0.2}{\m\per\s}}
+        \caption{Only data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}}
      \end{subfigure}
@@
 Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
-% todo describe the results for V>0.2
-A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}=0.
-For this reason, in Table~\ref{tab:f1}, we report the F1-scores for all the models.
-The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}).
+A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
+For both models the simple accuracy decreases when we only include entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+We want to remind the reader that the number of datapoints in each bin is different.
+
+In Table~\ref{tab:f1}, we report the F1-scores for both models, using the whole dataset, only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all the dataset.
+For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-scores of the \ac{pmv-ce} are at best \num{0.01} point higher than \ac{pmv} which does not justify its superiority.
+On the contrary the F1-macro score of the \ac{pmv} is \num{.23}, \qty{44}{\percent} higher than the \ac{pmv-ce} which is \num{.16} for subset of data with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+These results imply that the \ac{pmv} is more accurate than the \ac{pmv-ce} within the range in which both model perform better, i.e., closer to thermo neutrality.
 \begin{table}[htb!]
@@
     \input{tables/f1.tex}
-    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models.}
+    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models for three datasets.}
     \label{tab:f1}
@@
 
-% todo describe the table results
-The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
-Both models had an accuracy lower than \qty{1}{\percent} when predicting either `warm', `hot', or `cold'.
-This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
-The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
-% todo call it Standards implications, limit use PMV and of environmental variables
-Based on our results we recommend restricting the use of the \ac{pmv} only to a values $\lvert \textrm{PMV}\lvert \leq 1.5$.
-In addition, we also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} Standards.
-Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
-Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
-In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
-Consequently, we suggest reducing both Standards applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
-\begin{table}[htb!]
-    \centering
-    \begin{tabular}{cc}
-        \toprule
-        Variable & Proposed range \\
-        \midrule
-        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
-        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
-        \ac{rh} & \qtyrange{20}{80}{\percent} \\
-        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
-        \ac{met} & \qtyrange{1}{2}{met} \\
-        \ac{pmv} & \qtyrange{-1.5}{1.5}{} \\
-        % todo shall we also include limits on Age?
-        % todo shall we also include airspeed?
-        \bottomrule
-    \end{tabular}
-    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} Standards.}
-    \label{tab:ranges}
-\end{table}
 Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
@@
 The curve, in the Figure, is calculated using the individual data and not the binned data.
-We binned the data to aid the visualization of a large dataset.
+We only binned the data to aid the visualization of a large dataset.
 If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and have a slope of 1.
 The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval. 
-The intercept is close to 0 and this is aligned with the results shown above that the model can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times.
-Both \ac{pmv} formulations always underpredicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv}; the slope of the curve was significantly lower than 1.
+This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times, respectively.
+Both \ac{pmv} formulations always under-predicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv}; the slope of the curve was significantly lower than 1.
 We discuss in detail in sub-Section~\ref{subsec:sources-of-error} which factors may be the underlying cause of these errors.
@@
 
-\subsection{Model Bias}\label{sec:model-bias}
-\subsubsection{Model Overall Bias}\label{subsec:model-overall-bias}
+\subsection{Model Bias}\label{subsec:model-bias}
+\subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
 We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and  \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
 Results are presented in Figure~\ref{fig:hist_discrepancies}.
-% todo:
-%  describe in the paper that we used test to determine that data are not normally distributed.
 %  We cannot change into a letter value plot since in that case it would be hard to highlight the number of responses that are within the range from -.5 to.5.
-%  The Caption is missing
 \begin{figure*}[htb!]
@@
 Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
-The results show that the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
+When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
 Surprisingly, we found that the median bias for the \ac{pmv-ce} increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
 These are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
-As far as we could find, there is no published paper proving a higher accuracy of the \ac{pmv-ce} compared to the \ac{pmv} model. 
-The \ac{pmv-ce} has lower accuracy while making the code more complex and computationally intensive to run.
+The \ac{pmv-ce} has a higher bias while making the code more complex and computationally intensive to run.
 
-\subsubsection{Model Bias as Function of Four Input Variables}\label{subsec:model-bias-variable}
+\subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
 In the previous section, we presented the overall bias of the two \ac{pmv} formulations.
-We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{met}, and \ac{clo}.
-% todo add humidity ratio chart
-We did not include \ac{tr} in the analysis since \ac{tr} was not reported in several studies, hence, we estimated it as detailed in the Methodology.
+We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo}.
+We did not include \ac{tr} in the analysis since \ac{tr} was not measured in several studies, hence, we estimated it as detailed in the Methodology.
 Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
@@
 All the violin plots have the same area to aid visualization, despite the fact that the number of points in each bin is not constant.
-% todo:
 %  I need to have the red lines for the same inverse reason why I cannot use the boxenplot in the previous chart
-
+% todo in the plot do not report the mean value but the range instead since for example V max is 0.46m/s
+% todo report p-values
 \begin{figure*}[htb!]
@@
 These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.
-
-\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model} % todo Stefano Schiavon: call it Gagge PMV model in line with the text below
-% todo consider moving to appendix and explain well why I have done this.
-Stefano Schiavon: The logic and need of this paragraph needs to be better explained.
-Why did you add?
-Why can we assume that the PMV-SET is better?
-Who developed the PMV-SET?
-We should discuss if we need this paragraph and if it add something that is important and not reported in other sections.
-In other to do this, we need to explain why we think the PMV-SET is important even if it is not present in standards.
-
-We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx. % todo provide information
-% todo: Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
-Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
-The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
-The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
-% todo Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
-The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
-% todo Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from?
-How can a PMV model be described by an air speed?
-These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
-This is also an unexpected result and we will discuss why this may be in the following paragraph. 
-% todo When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
-1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.
-
-\begin{figure*}[htb!]
-    \centering
-    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
-    \caption{}
-    \label{fig:pmv_two_node_comparison}
-\end{figure*}
+Both model had a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{20}{\celsius}; \ac{clo}~$\leq$~\qty{.6}{clo} and $\geq$~\qty{1.2}{clo}.
+Surprisingly again the bias of the \ac{pmv-ce} model was significantly worse than the one of the \ac{pmv} for \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+% todo report the p-value
 
@@
 
-The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
+The \ac{pmv} model uses the overall heat losses or gains to calculate a \ac{pmv} value which should represent a \ac{tsv} vote.
+Heat losses or gains are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature.
 Fanger created this equation based on a limited dataset collected in a laboratory, and despite the growing body of literature, this equation was never updated or revisited~\cite{Fanger1970}.
@@
 This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis~\cite{ romanovsky_thermoregulation_2018}.
-On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
+On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral, and slightly warm).
 % todo we need to check this since we may only recommend the use of the PMV model for thermal neutrality
@@
 For example, even if the model was 100~\% accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (+2)' or `hot (+3)'.
-Using a continuous, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.
+Using a continuous scale, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.
 They also showed that the \ac{tsv}'s assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
@@
 % todo write a paragraph trying to find the reason why the PMV CE model is less accurate.
-% todo write a paragraph about what we recommend for ASHRAE 55

### 2023-09-11 Manuscript/src/sections/titleAbstract.tex
paper: reviewed the paper and addressed most of Stefano's first round of comments
fix: some pmv values were missing since were outside the applicability limits of the standard

@@
 
-    \title{PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?}
+    \title{PMV: Comparing the prediction accuracy of ISO 7730:2005 and ASHRAE 55:2023}
+    % alternatives PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?
+    % alternatives ISO 7730:2005 vs ASHRAE 55:2023: A comparison on the accuracy of the two implementations of the PMV model
 
@@
 
-        The \ac{pmv} has seen widespread adoption and has been used by professionals to design and operate buildings and by researchers to predict thermal sensation.
+        Being relatively simple to calculate and being included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
         It is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
         Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
-        \gls{7730} and \gls{55} are the most widely used Standards worldwide, and their differences are a source of confusion for users who find it difficult to decide which model to use.
+        \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their differences are a source of confusion for users who find it difficult to decide which model to use.
 
@@
         We found that the \ac{pmv-ce} has the same accuracy as the original \ac{pmv} in predicting thermal neutrality, despite its increased complexity and computational resources needed to calculate the results.
-        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}$\geq$\qty{0.1}{\m\per\sec}.
-
+        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}$\geq$\qty{0.2}{\m\per\sec}.
+        % todo add values below and complete the Abstract
         The \ac{pmv} and \ac{pmv-ce} accuracy were XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 2$ and XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 1$, respectively.

### 2023-09-11 main.py

@@ -1095,6 +1095,20 @@ def importing_filtering_processing(load_preprocessed=False):
     df_["pmv_gagge"] = two_nodes_results["pmv_gagge"]
     df_["pmv_set"] = two_nodes_results["pmv_set"]
 
+    results_pmv_ppd = pmv_ppd(
+        tdb=df_.ta,
+        tr=df_.tr,
+        vr=df_.vel_r,
+        rh=df_.rh,
+        met=df_.met,
+        clo=df_.clo,
+        standard="ISO",
+        limit_inputs=False,
+    )
+
+    df_["pmv"] = results_pmv_ppd["pmv"]
+    df_["ppd"] = results_pmv_ppd["ppd"]
+
     # estimate thermal sensation using toby's model
     df_["pmv_toby"] = list(
         pd.cut(
@@ -1575,7 +1589,6 @@ def plot_distribution_variable():
             color=c_gold,
             showfliers=False,
             linewidth=0.5,
-            outlier_prop=0.05,
             k_depth="proportion",
         )
         axs[ix].set(
@@ -1583,26 +1596,26 @@ def plot_distribution_variable():
             xlabel=f"{var_names[var]} ({var_units[var]})",
         )
         if var == "ta" or var == "tr":
-            axs[ix].set(yticks=(np.linspace(15, 30, 4)), ylim=(15, 30))
+            axs[ix].set(yticks=(np.linspace(10, 34, 5)), ylim=(10, 34))
         if var == "rh":
             axs[ix].set(
-                yticks=(np.linspace(20, 80, 4)),
-                ylim=(20, 80),
+                yticks=(np.linspace(10, 90, 5)),
+                ylim=(10, 90),
             )
         if var == "vel":
             axs[ix].set(
-                yticks=(np.linspace(0, 0.6, 4)),
-                ylim=(0, 0.6),
+                yticks=(np.linspace(0, 1, 5)),
+                ylim=(0, 1),
             )
         if var == "clo":
             axs[ix].set(
-                yticks=(np.linspace(0.3, 1.5, 4)),
-                ylim=(0.3, 1.5),
+                yticks=(np.linspace(0.1, 1.6, 4)),
+                ylim=(0.1, 1.6),
             )
         if var == "met":
             axs[ix].set(
-                yticks=(np.linspace(0.8, 2, 5)),
-                ylim=(0.8, 2),
+                yticks=(np.linspace(0.8, 2.8, 5)),
[…]

### 2023-09-12 Manuscript/src/sections/conclusions.tex
feat: finalised second draft to send to Stefano

--- 2023-07-11 Manuscript/src/sections/conclusions.tex
@@
 \section{Conclusions}\label{sec:conclusions}
-The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \gls{pmv-ce} model, as employed in the \gls{55} Standard.
-Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \gls{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
-These disparities shed light on the limitations and strengths of the \ac{pmv} model, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
-The outcomes of this analysis have significant implications for researchers, practitioners, and policymakers involved in the field of thermal comfort assessment and standardization.
+The present study compares the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} Standard.
+The \ac{pmv-ce} was specifically developed with the aim of more accurately estimating latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
+We determined the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models by comparing their results to the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
+
+The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
+Both models had an accuracy lower than \qty{2}{\percent} when predicting either `warm', `hot', or `cold'.
+This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
+When only comparing the results of the two models for entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} -- range in which the two models differ -- the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
+For example, the median bias of the \ac{pmv} model is \var{bias_median_pmv_0.2} while the median bias for the \ac{pmv-ce} is \var{bias_median_pmv_ce_0.2}.
+
+Our results also show that the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.
+In the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
+This finding challenges the above assumption.
+
+\subsection{Implications for the ASHRAE 55 and ISO 7730 Standards}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
+Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion wordwide since both models are widely adopted and incorporated in many building codes and Standards worldwide.
+The \gls{55} Standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that clearly explain the benefit of using it.
+The \ac{pmv-ce} is more complex, computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
+Hence we do not recommend its use.
+
+The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
+Based on our results we recommend restricting the use of the \ac{pmv} only to a values $\lvert \textrm{PMV}\lvert \leq 0.5$.
+Or in other words, we only recommend its use to determine if an environment may be deemed comfortable by a large group of occupants.
+The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classy if people are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
+Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023} provide corroborating evidence that support our results.
+
+We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} Standards.
+Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
+Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
+In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
+Consequently, we suggest reducing both Standards applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
+\begin{table}[htb!]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        Variable & Proposed range \\
+        \midrule
+        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
+        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
+        \ac{rh} & \qtyrange{20}{80}{\percent} \\
+        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
+        \ac{met} & \qtyrange{1}{2}{met} \\
+        % todo shall we also include limits on Age?
+        % todo shall we also include airspeed?
+        \bottomrule
+    \end{tabular}
+    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} Standards.}
+    \label{tab:ranges}
+\end{table}
+\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs that are consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
+The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
+
+The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in our dataset, and it is not a valid assumption.
+We recommend the use of the thermal preference scale to determine how people perceive their thermal environment, this scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.

### 2023-09-12 Manuscript/src/sections/introduction.tex
feat: finalised second draft to send to Stefano

@@
 
-To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988}.
+To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.
 
 To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
-The best source to see how \ac{pmv-ce} works is the Standard itself, but the standard does not explain why some choices were implemented.
+The best source to see how \ac{pmv-ce} works is the Standard itself, but the Standard does not explain why some choices were implemented.
 A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
@@
     \end{tikzpicture}
-    \caption{Flowchart depicting the steps for the calculation of the PMV in accordance with the \acs{55} Standard.}
+    \caption{Flowchart depicting the steps for the calculation of the PMV in accordance with the \gls{55} Standard.}
     \label{fig:flowchart_pmv_ce}
@@
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
-% todo use this finding to support limit to PMV range
 Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
@@
 They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
-% todo use this finding to support limit to PMV range
 The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
@@
 Finally, they only reported the overall bias for the whole dataset, despite the fact that the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
+\mycite{tartarini_prediction_2023} also published a conference paper on this topic.
+They concluded that the \ac{pmv} and \ac{pmv-ce} models accuracy is \qty{35}{\percent} and \qty{36}{\percent} for $\mid$\ac{pmv}$\mid \leq 2$, \qty{43}{\percent} and \qty{44}{\percent} for $\mid$\ac{pmv}$\mid \leq 1$ in predicting the self-reported thermal sensation of people, respectively.
+However, their analysis is limited to providing the precision of the model alone, without reporting the sensitivity (also commonly referred as recall) and they do not provide the bias of the two models.
 

### 2023-09-12 Manuscript/src/sections/results.tex
feat: finalised second draft to send to Stefano

@@
 Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
-% todo v_95_perc_max these variables should be calculated dropping nan
 
@@
 A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
-For both models the simple accuracy decreases when we only include entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+For both models the simple accuracy decreases by one point when we only include entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
 We want to remind the reader that the number of datapoints in each bin is different.
@@
 The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all the dataset.
-For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-scores of the \ac{pmv-ce} are at best \num{0.01} point higher than \ac{pmv} which does not justify its superiority.
-On the contrary the F1-macro score of the \ac{pmv} is \num{.23}, \qty{44}{\percent} higher than the \ac{pmv-ce} which is \num{.16} for subset of data with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
-These results imply that the \ac{pmv} is more accurate than the \ac{pmv-ce} within the range in which both model perform better, i.e., closer to thermo neutrality.
+For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models improved and the former outperformed the latter.
+This is an unexpected result since the \ac{pmv-ce}, to justify its use, should significantly outperform the \ac{pmv} model when analyzing this subset of data.
+Finally, the F1-micro and weighted scores of both models increased by about 0.1 when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+However, the macro score, within this range, remained practically unchanged.
 \begin{table}[htb!]
@@
 If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and have a slope of 1.
-The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval. 
+The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.25} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
 This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times, respectively.
-Both \ac{pmv} formulations always under-predicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv}; the slope of the curve was significantly lower than 1.
+Both \ac{pmv} formulations always under-predicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
+the slope of the curve was significantly lower than 1.
 We discuss in detail in sub-Section~\ref{subsec:sources-of-error} which factors may be the underlying cause of these errors.
@@
 This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
-This was already done in \mycite{Cheung2019}, where they grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified how many times the \ac{pmv} model correctly predicted \ac{tsv}.
+\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
 They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
@@
 Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
+We report the median and inter-quantile ranges since data are not normally distributed.
 When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
 Surprisingly, we found that the median bias for the \ac{pmv-ce} increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
-These are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
+These, together with those presented in Table~\ref{tab:f1}, are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
 The \ac{pmv-ce} has a higher bias while making the code more complex and computationally intensive to run.
@@
 %  I need to have the red lines for the same inverse reason why I cannot use the boxenplot in the previous chart
-% todo in the plot do not report the mean value but the range instead since for example V max is 0.46m/s
-% todo report p-values
+% todo Shall I report the p-values?
 \begin{figure*}[htb!]
@@
     \includegraphics[width=\textwidth]{figures/bias_models}
-    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.}
+    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models plotted as a function of the input variables.
+    The x ticks labels show the bin range.
+    The $]$ means that the right value is included in the range.
+    }
     \label{fig:bias_models}
 \end{figure*}
-The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all four input variables.
-On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 as the value of \ac{v} increased.
+The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all input variables.
+On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 and was significantly worse (\num{-.37} versus \num{-0.07}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.45}{\m\per\s}.
 These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.
-Both model had a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{20}{\celsius}; \ac{clo}~$\leq$~\qty{.6}{clo} and $\geq$~\qty{1.2}{clo}.
-Surprisingly again the bias of the \ac{pmv-ce} model was significantly worse than the one of the \ac{pmv} for \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-% todo report the p-value
+Both model have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
 
@@
 Moreover, the \ac{pmv} model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
-This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
-% todo Stefano Schiavon: Please add references suportting this.
+This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature~\cite{romanovsky_thermoregulation_2018}.
 Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
-% todo talk also about how sweat is calculated
+The \ac{pmv} model also assumes that the human body is constantly loosing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
+This amount of heat loss is included in the equation even if the cylinder has a negative heat gain, or in other words the person is estimated to be feeling cold.
 
@@
 Heat losses or gains are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature.
-Fanger created this equation based on a limited dataset collected in a laboratory, and despite the growing body of literature, this equation was never updated or revisited~\cite{Fanger1970}.
-% todo Stefano Schiavon: This is vague, Which papers or data is available to update that equation/mapping?
+Fanger created this equation based on a limited dataset collected in a laboratory and this equation was never updated or revisited~\cite{Fanger1970}.
 In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions, respectively~\cite{romanovsky_thermoregulation_2018}.
 
-It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
+It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral'.
 This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis~\cite{ romanovsky_thermoregulation_2018}.
-On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral, and slightly warm).
-% todo we need to check this since we may only recommend the use of the PMV model for thermal neutrality
-We showed that people tend to be closer to neutrality than what the models predicted (see Figure XXX), this is an indication that they were exposed to compensable thermal conditions.
-% todo add figure number
-Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.
+On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the thermal neutrality zone.
+The models' inability to account for these regulatory mechanisms is a fundamental limitation.
 
-To continue using the \ac{pmv} model beyond an absolute value of 1, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
-i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
-% todo Stefano Schiavon: The difference between these two points ii) and iii) are not clear, can you add more explanation?
-ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress;
-iii) the regression coefficient between heat losses and heat gains and thermal stress is not linear and two different equations should be used;
+To continue using the \ac{pmv} model beyond an absolute value of \num{0.5}, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
+i) verify the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
+ii) if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
+iii) currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different set of  equations should be used;
 iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
-For example, while the \ac{pmv} model may assume that a score of 3 (hot) signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches the max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
-Participants in commercial buildings may, on the contrary, deem the environment to be `Hot' as soon as they start sweating and their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.
+For example, while the \ac{pmv} model may assume that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
+Participants in commercial buildings may, on the contrary, deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.
 
@@
 % to do check that the fourth below still refers to the anchors issue above
-The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}
+The latter hypothesis -- iv) anchoring -- has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
+Supporting evidence to this issue is provided my \mycite{schweiker2020evaluating} who showed that the distances between the anchors are not constant and a lay person may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
 
@@
 % Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.
-
-% todo write a paragraph trying to find the reason why the PMV CE model is less accurate.

### 2023-09-12 Manuscript/src/sections/titleAbstract.tex
feat: finalised second draft to send to Stefano

@@
     \begin{abstract}
-        The \ac{pmv} model was developed by Fanger more than 50 years ago to predict the mean thermal sensation of a group of people. 
-        It is based on a heat balance model of the human body.
+%        The \ac{pmv} model was developed by Fanger more than 50 years ago to predict the mean thermal sensation of a group of people.
+%        It is based on a heat balance model of the human body.
 
-        \gls{55} states that the model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce satisfactory thermal environmental conditions for a majority of the occupants within the space''.
+%        Being relatively simple to calculate and being included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
+        The \ac{pmv} model, developed by Fanger, is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
+        Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
+        \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion for users who find it difficult to decide which model to use.
 
-        Being relatively simple to calculate and being included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
-        It is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
-        Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
-        \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their differences are a source of confusion for users who find it difficult to decide which model to use.
+        We compared the accuracy of the two main \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
+        We found that the use of the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, despite its increased complexity and computational resources needed to calculate the results.
+        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\sec}.
 
-        We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that the \ac{pmv-ce} has the same accuracy as the original \ac{pmv} in predicting thermal neutrality, despite its increased complexity and computational resources needed to calculate the results.
-        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}$\geq$\qty{0.2}{\m\per\sec}.
-        % todo add values below and complete the Abstract
-        The \ac{pmv} and \ac{pmv-ce} accuracy were XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 2$ and XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 1$, respectively.
-        Neither PMV model could accurately predict `cold’, `cool’, `warm’, or `hot’ thermal sensations.
-        Consequently, the PMV should only be used in conditions close to neutrality within the range $\mid$\ac{pmv}$\mid \leq 1$, and we suggest reducing the applicability limit of the Standards to this range.
+        We, therefore, recommend using the original \ac{pmv} instead of the \ac{pmv-ce}.
+        In addition, we suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
+        This recommendation already aligns with the \gls{55} Standard's definition which states that the \ac{pmv} model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce \textit{satisfactory thermal environmental conditions} for a majority of the occupants within the space''.
+        The \ac{pmv} model cannot determine the degree by which people are dissatisfied.
+        A value higher or lower than \num{.5} or \num{-.5}, respectively only indicates that people are on the `warm' or `cold' side.
+        Finally, our results show that the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.
+        We, hence, recommend in thermal comfort research the use of the thermal preference scale to determine how people perceive their thermal environment.
     \end{abstract}
@@
     \begin{keyword}
-        Thermal comfort model \sep ISO 7730 \sep ASHRAE 55 \sep Thermal sensation \sep Accuracy
+        Thermal comfort model \sep ISO 7730 \sep ASHRAE 55 \sep Thermal sensation \sep Thermal preference
     \end{keyword}

### 2023-09-12 main.py

@@ -1101,7 +1101,7 @@ def importing_filtering_processing(load_preprocessed=False):
         vr=df_.vel_r,
         rh=df_.rh,
         met=df_.met,
-        clo=df_.clo,
+        clo=df_.clo_d,
         standard="ISO",
         limit_inputs=False,
     )
@@ -2258,14 +2258,18 @@ def plot_bias_distribution_by_variable_binned():
     ]
 
     bins = {
-        "ta": np.arange(18, 32, 2),
-        "tr": np.arange(18, 31.5, 2),
-        "top": np.arange(17.5, 30.5, 2),
-        "vel": np.arange(-0.00001, 0.7, 0.2),
-        "rh": np.arange(10, 100, 20),
+        # vel 0.0058333333333333 0.45
+        # rh 21.27 71.5
+        # clo 0.3000000000000001 1.3099999999999998
+        # met 1.0 1.877133105802048
+        "ta": [18.5, 21.0, 24.0, 27.0, 29.0],
+        # "tr": np.arange(18, 31.5, 2),
+        # "top": np.arange(17.5, 30.5, 2),
+        "vel": [0.0, 0.15, 0.3, 0.46],
+        "rh": [21, 40, 60, 72],
         "hr": np.arange(0, 20, 2),
-        "clo": np.arange(0.2, 1.5, 0.2),
-        "met": np.arange(0.9, 2, 0.2),
+        "clo": [0.3, 0.5, 0.7, 0.9, 1.1, 1.31],
+        "met": [1.0, 1.2, 1.4, 1.6, 1.9],
         "thermal_sensation": np.arange(-3.5, 4.5, 1),
         "thermal_preference": np.arange(-3.5, 4.5, 1),
         "pmv": np.arange(-3.5, 4.5, 1),
@@ -2275,6 +2279,7 @@ def plot_bias_distribution_by_variable_binned():
     # plt.close("all")
 
     df_analysis = df.copy()
+    df_analysis.loc[df_analysis.vel == 0, "vel"] = 0.0000001
 
     f, axs = plt.subplots(5, 1, figsize=(7, 9))
     axs = axs.flatten()
@@ -2315,6 +2320,8 @@ def plot_bias_distribution_by_variable_binned():
             )
 
             df_plot[var] = pd.cut(df_plot[var], bins=bins[var])
+
+        print(df_plot.groupby([var, "model"])["diff_ts"].median())
 
         # try:
[…]

### 2024-06-21 Manuscript/src/sections/results.tex
feat: overleaf version

@@
 This finding challenges the above assumption.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in our dataset.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset.
 On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
-Of those who reported to be `cool' \qty{23}{\percent} of them wanted `no change' and \qty{5}{\percent} wanted to be `cooler'.
-Similarly, but in reverse, among those who reported to be `warm' \qty{13}{\percent} wanted `no change' and \qty{2}{\percent} wanted to be `warmer'.
+Of those who reported being `cool' \qty{23}{\percent} of them wanted `no change' and \qty{5}{\percent} wanted to be `cooler'.
+Similarly, but in reverse, among those who reported being `warm' \qty{13}{\percent} wanted `no change' and \qty{2}{\percent} wanted to be `warmer'.
 This result highlights one of the major limitations of the \ac{tsv} scale and of the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable.
@@
 We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
-We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was change to \qty{0.1}{\m\per\s} only in the \gls{55}.
+We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
 The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
@@
 A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
-For both models the simple accuracy decreases by one point when we only include entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-We want to remind the reader that the number of datapoints in each bin is different.
+For both models, the simple accuracy decreases by one point when we only include entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+We want to remind the reader that the number of data points in each bin is different.
 
 In Table~\ref{tab:f1}, we report the F1-scores for both models, using the whole dataset, only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
-The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all the dataset.
+The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
 For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models improved and the former outperformed the latter.
@@
 Finally, the F1-micro and weighted scores of both models increased by about 0.1 when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+% todo compare how the models compare for PMV < |0.5| not |1.5|
 However, the macro score, within this range, remained practically unchanged.
@@
     \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
-    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) is superimposed over the regression curve to aid the visualization of a large dataset.}
+    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.}
     \label{fig:bubble_models_vs_tsv}
@@
 Results are presented in Figure~\ref{fig:hist_discrepancies}.
-%  We cannot change into a letter value plot since in that case it would be hard to highlight the number of responses that are within the range from -.5 to.5.
+%  We cannot change into a letter value plot since in that case, it would be hard to highlight the number of responses that are within the range from -.5 to 5.
 \begin{figure*}[htb!]
@@
 In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
-All the violin plots have the same area to aid visualization, despite the fact that the number of points in each bin is not constant.
+All the violin plots have the same area to aid visualization, even though the number of points in each bin is not constant.
 %  I need to have the red lines for the same inverse reason why I cannot use the boxenplot in the previous chart
@@
 Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
-The \ac{pmv} model also assumes that the human body is constantly loosing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
-This amount of heat loss is included in the equation even if the cylinder has a negative heat gain, or in other words the person is estimated to be feeling cold.
+The \ac{pmv} model also assumes that the human body is constantly losing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
+This amount of heat loss is included in the equation even if the cylinder has a negative heat gain, or in other words, the person is estimated to be feeling cold.
 
 The \ac{pmv} model uses the overall heat losses or gains to calculate a \ac{pmv} value which should represent a \ac{tsv} vote.
-Heat losses or gains are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature.
+Heat losses or gains are only a proxy for the thermal stress that the human body has to undergo to maintain a constant internal temperature.
 Fanger created this equation based on a limited dataset collected in a laboratory and this equation was never updated or revisited~\cite{Fanger1970}.
@@
 ii) if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
-iii) currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different set of  equations should be used;
+iii) currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different sets of  equations should be used;
 iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
@@
 
-It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
-% to do check that the fourth below still refers to the anchors issue above
+A combination of all the above-mentioned factors likely plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
 The latter hypothesis -- iv) anchoring -- has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
-Supporting evidence to this issue is provided my \mycite{schweiker2020evaluating} who showed that the distances between the anchors are not constant and a lay person may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
+Supporting evidence to this issue is provided by \mycite{schweiker2020evaluating} who showed that the distances between the anchors are not constant and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
 
@@
 All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts, however, measurement errors are inevitably present in the dataset.
-This is due to the fact that all measurement instrumentation is not always calibrated before use, the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the near proximity of the participant, and estimating clothing and metabolic rates from tables is a complex and prone to error task.
+This is because all measurement instrumentation is not always calibrated before use, the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the proximity of the participant, and estimating clothing and metabolic rates from tables is a complex and prone to error task.
 In an ideal scenario, errors would be random and would cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.

### 2024-06-21 main.py
feat: new figures

@@ -1,7 +1,7 @@
 import math
 import warnings
+from itertools import product
 
-from numba import jit, vectorize, float64
 import matplotlib as mpl
 import matplotlib.patches as patches
 import matplotlib.pyplot as plt
@@ -11,17 +11,19 @@ import psychrolib
 import seaborn as sns
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.gridspec import GridSpec
+from numba import jit
 from pythermalcomfort import p_sat_torr, check_standard_compliance_array
 from pythermalcomfort.models import (
     athb,
     cooling_effect,
 )
 from scipy import stats
-from scipy.stats import skewtest
 from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score
 from sklearn.metrics import f1_score
 from sklearn.metrics import r2_score, mean_absolute_error
+
+# mpl.use("TkAgg")
 
 warnings.filterwarnings("ignore")
 
@@ -2540,6 +2542,224 @@ def pmv_jiayu_fed(tdb, tr, vr, rh, met, clo, wme=0):
     clf.predict([[0, 1, 0.6]])
 
 
+def compare_pmv_disc():
+    def generate_t_rh_combinations(t_range, rh_range):
+        all_combinations = list(product(t_range, rh_range))
+        return pd.DataFrame(all_combinations, columns=["t", "rh"])
+
+    def heatmap(
+        data, row_labels, col_labels, ax=None, cbar_kw=None, cbarlabel="", **kwargs
+    ):
+        """
+        Create a heatmap from a numpy array and two lists of labels.
+
+        Parameters
+        ----------
+        data
+            A 2D numpy array of shape (M, N).
+        row_labels
+            A list or array of length M with the labels for the rows.
+        col_labels
+            A list or array of length N with the labels for the columns.
+        ax
+            A `matplotlib.axes.Axes` instance to which the heatmap is plotted.  If
+            not provided, use current axes or create a new one.  Optional.
+        cbar_kw
+            A dictionary with arguments to `matplotlib.Figure.colorbar`.  Optional.
+        cbarlabel
+            The label for the colorbar.  Optional.
+        **kwargs
+            All other arguments are forwarded to `imshow`.
+        """
+
+        if ax is None:
+            ax = plt.gca()
+
+        if cbar_kw is None:
[…]

### 2024-06-21 test_f1_score.py

@@ -204,3 +204,63 @@ preds = log_reg_2.predict(np.arange(0, 50, 0.5).reshape(-1, 1))
 plt.figure()
 plt.plot(np.arange(0, 50, 0.5), preds)
 plt.show()
+
+
+from pythermalcomfort.models import pmv_ppd
+from pythermalcomfort.utilities import v_relative, clo_dynamic
+
+tdb = 25
+tr = 25
+rh = 50
+v = 0.1
+met = 1.4
+clo = 0.5
+# calculate relative air speed
+v_r = v_relative(v=v, met=met)
+# calculate dynamic clothing
+clo_d = clo_dynamic(clo=clo, met=met, standard="ashrae")
+results = pmv_ppd(tdb=tdb, tr=tr, vr=v_r, rh=rh, met=met, clo=clo_d, standard="ashrae")
+print(results)
+print(results["pmv"])
+
+tdb = 19.6
+tr = 19.6
+rh = 86
+v = 0.1
+met = 1.1
+clo = 1
+# calculate relative air speed
+v_r = v_relative(v=v, met=met)
+# calculate dynamic clothing
+clo_d = clo_dynamic(clo=clo, met=met, standard="ashrae")
+results = pmv_ppd(tdb=tdb, tr=tr, vr=v_r, rh=rh, met=met, clo=clo_d, standard="ashrae")
+print(results)
+print(results["pmv"])
+
+tdb = 19.6
+tr = 19.6
+rh = 86
+v = 0.2
+met = 1.1
+clo = 1
+# calculate relative air speed
+v_r = v_relative(v=v, met=met)
+# calculate dynamic clothing
+clo_d = clo_dynamic(clo=clo, met=met, standard="ashrae")
+results = pmv_ppd(tdb=tdb, tr=tr, vr=v_r, rh=rh, met=met, clo=clo_d, standard="ashrae")
+print(results)
+print(results["pmv"])
+
+tdb = 29.4
+tr = 29.4
+rh = 61
+v = 0.1
+met = 1.9
+clo = 0.5
+# calculate relative air speed
+v_r = v_relative(v=v, met=met)
+# calculate dynamic clothing
+clo_d = clo_dynamic(clo=clo, met=met, standard="ashrae")
+results = pmv_ppd(tdb=tdb, tr=tr, vr=v_r, rh=rh, met=met, clo=clo_d, standard="ashrae")
+print(results)
+print(results["pmv"])

### 2024-06-21 Manuscript/src/sections/introduction.tex

@@
 \mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
-The \ac{pmv} is as an index that predicts the mean value of the thermal sensation votes (self-reported perceptions) of a large group of persons on a sensation scale expressed from –3 to +3 corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
+The \ac{pmv} is an index that predicts the mean value of the thermal sensation votes (self-reported perceptions) of a large group of persons on a sensation scale expressed from –3 to +3 corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
@@
 
-To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.
+To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.
 
@@
     \end{tikzpicture}
-    \caption{Flowchart depicting the steps for the calculation of the PMV in accordance with the \gls{55} Standard.}
+    \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{55} Standard.}
     \label{fig:flowchart_pmv_ce}
@@
 %In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
-%This arbitrary and non-scientific definition of the expectancy factor make the \ac{epmv} difficult to calculate and non-generalizable.
+%This arbitrary and non-scientific definition of the expectancy factor makes the \ac{epmv} difficult to calculate and non-generalizable.
 %Consequently, the \ac{epmv} has not gained any widespread adoption.
@@
 %Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
-%They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
+%They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn used to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
 %The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
@@
 This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate since it does not explain how the model performs over a wide range of environmental, personal conditions, and \ac{tsv}.
-Reporting the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv} is particularly for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral'.
-Finally, they only reported the overall bias for the whole dataset, despite the fact that the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
+Reporting the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv} is particularly important for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral'.
+Finally, they only reported the overall bias for the whole dataset, even though the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
 \mycite{tartarini_prediction_2023} also published a conference paper on this topic.
 They concluded that the \ac{pmv} and \ac{pmv-ce} models accuracy is \qty{35}{\percent} and \qty{36}{\percent} for $\mid$\ac{pmv}$\mid \leq 2$, \qty{43}{\percent} and \qty{44}{\percent} for $\mid$\ac{pmv}$\mid \leq 1$ in predicting the self-reported thermal sensation of people, respectively.
-However, their analysis is limited to providing the precision of the model alone, without reporting the sensitivity (also commonly referred as recall) and they do not provide the bias of the two models.
+However, their analysis is limited to providing the precision of the model alone, without reporting the sensitivity (also commonly referred to as recall) and they do not provide the bias of the two models.
 

### 2024-06-21 Manuscript/src/sections/methodology.tex

@@
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary we decided keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
+However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
 We used Python to analyse and visualise the results.
@@
 Consequently, we expanded the range to \num{-.5} and \num{0.5}.
-Thi is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
-We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally we would want to have the \ac{pmv} to have a higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
+This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
+We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
 Binning data by the independent variables is a significant improvement over grouping the participants' responses by their \ac{tsv}.
@@
 %This would solve some of the uncertainties that the \ac{pmv-ce} model introduces as detailed above.
-%However, it is not possible for us to comment on why the \ac{pmv-ce} model was included in the \gls{55} instead of the \ac{pmvs} since no peer-reviewed paper or official document describe the rationale for this decision.
+%However, it is not possible for us to comment on why the \ac{pmv-ce} model was included in the \gls{55} instead of the \ac{pmvs} since no peer-reviewed paper or official document describes the rationale for this decision.
 
@@
 Precision (also called positive predictive value) is the ratio of true positive overall positive results obtained from the test, while sensitivity (also known as recall or true positive rate) is the ratio of true positive over all the actual positive cases.
-We report three F1-scores in the paper.
+We report three F1 scores in the paper.
 The F1-micro score calculates metrics globally by counting the total true positives, false negatives, and false positives.
@@
 The F1-weighted calculate metrics for each label and find their average weighted by support (the number of true instances for each label).
+
+% Ed's comments below
+% The primary sticking point is with the input data.  The following issue seems fundamental to me, something you would have to address in a published paper about this.
+
+% Air speed measurements in the ASHRAE Field Dataset rarely conform to ASHRAE or ISO standards.
+Only the studies done with instrumented carts with anemometer arrays are likely to have measured at three heights as required.
+We know that very few of the studies have done this.
+
+% If you only measure at one height instead of three, you are very likely to record higher values than the averaged velocities that are required for the standards’ comfort models.
+This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level.
+The measurement will therefore most often be an overestimate for inputting into the model, and the associated sensation prediction will be cooler than it should be.
+
+% So if you have a systematic bias toward higher airspeeds in the statistical data you are using, a comfort model that inherently predicts stronger rates of air speed cooling (CE) will overpredict more than a model that inherently predicts weaker rates (PMV).
+The weaker one will appear to be predicting correctly, or in any event something closer to correct.
+The stronger-predicting model could appear ‘biased’ when compared to the weaker-predicting model.
+
+% This is one of the points of Pearl’s causal inference book—in assigning bias, you have to find the source of the bias, which in this case may be in the dataset, and not in either of the models being evaluated.
+
+% The DB2 dataset is very useful for temperature, but it is really not suitable for detailed evaluation of air speed.
+Air speed is far more variable and harder to measure accurately than temperature.
+I suppose someone could wade through all the studies and observations in the database and observe the types of conditions under which the elevated air speeds were recorded.
+Were they consistent over time and geographic location?
+Did they average over sufficient time as specified in standards?
+Any sense of the accuracy of the anemometers used to measure them?
+What kind of buildings were they in—or any indications about the thermal sensitivity of the occupancies being surveyed?
+You can see that all these questions suggest possible sources of bias, all which would be need to identified and assessed, to judge whether information from such surveys is suitable for evaluating a systematic measurement system like a comfort model.
+It seems a really big task with uncertain benefits.
+
+% Models vs laboratory study data:
+
+% In our 2015 Corrective Power paper, Hui assembled the results from many laboratory studies of cooling devices, including fans.
+In this paper you can see the studies of Tanabe where his subjects were exposed to a uniform plug flow, where the air movement measurement represents the whole body as required by the standard.
+You can compare his results to the CE model and find that the CE model slightly underpredicts the subjective cooling votes he measured.
+
+% If you then look at the cooling found by Yongchao Zhai where the air movement was directed only at the upper body and measured there, the CE model slightly overpredicts his measured results.
+This is of course appropriate.
+It is seen in other recent studies of local fan cooling.
+
+% Bigger view of this:
+
+% I think the comparison of models must be done using data from controlled experiments in the laboratory.
+A lot of this was well-examined 50 years ago.
+It hasn’t been followed up as intensely since then because it wasn’t considered that important by the profession.
+What’s different now is that 1) people are recognizing climate warming, and 2) it’s a lot easier to do comfort and physiological modeling and learn from it, using the Web.
+That’s why there is now an opportunity that people can design better buildings using more sophisticated comfort and climate information.

### 2024-06-21 Manuscript/src/sections/titleAbstract.tex
feat: overleaf version
fix: minor typos

@@
     % alternatives PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?
-    % alternatives ISO 7730:2005 vs ASHRAE 55:2023: A comparison on the accuracy of the two implementations of the PMV model
+    % alternatives ISO 7730:2005 vs ASHRAE 55:2023: A comparison of the accuracy of the two implementations of the PMV model
 
@@
 
-%        Being relatively simple to calculate and being included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
+%        Being relatively simple to calculate and included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
         The \ac{pmv} model, developed by Fanger, is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
         Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
-        \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion for users who find it difficult to decide which model to use.
+        The \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion for many users who find it difficult to decide which model to use.
 
@@
 
-        We, therefore, recommend using the original \ac{pmv} instead of the \ac{pmv-ce}.
-        In addition, we suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
+        We recommend using the original \ac{pmv} instead of the \ac{pmv-ce} and we suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
         This recommendation already aligns with the \gls{55} Standard's definition which states that the \ac{pmv} model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce \textit{satisfactory thermal environmental conditions} for a majority of the occupants within the space''.
-        The \ac{pmv} model cannot determine the degree by which people are dissatisfied.
+        The \ac{pmv} model cannot determine the degree to which people are dissatisfied.
         A value higher or lower than \num{.5} or \num{-.5}, respectively only indicates that people are on the `warm' or `cold' side.

### 2024-08-15 Manuscript/src/myglossary.tex
feat: reviewed methodology
feat: code to analyse data from studies with three air speed measurements

--- 2022-06-09 Manuscript/src/myglossary.tex
@@
 \usepackage[nonumberlist,nogroupskip]{glossaries}
+% \usepackage{glossaries}
 
@@
 {
-name={ASHRAE~55--2020},
-description={ASHRAE~55--2020 is a thermal comfort standard developed by ANSI and ASHRAE}
+name={ASHRAE~55:2023},
+description={ASHRAE~55:2023 is a thermal comfort standard developed by ANSI and ASHRAE}
 }
 
-\newglossaryentry{rhrn}
-{
-name={right-here-right-now},
-description={Right here right now thermal comfort survey}
-}
-
-\newglossaryentry{pmv-ce}
-{
-name={PMV\textsubscript{CE}},
-description={PMV calculated in accordance with the ASHRAE~55--2020}
-}
-
-\newglossaryentry{db2}
-{
-name={Comfort DB},
-description={ASHRAE Global Thermal Comfort Database II}
-}
-
-\makenoidxglossaries
+% \makenoidxglossaries

### 2024-08-15 Manuscript/src/sections/conclusions.tex
feat: code to analyse data from studies with three air speed measurements

@@
-%! Author = sbbfti
-%! Date = 26/08/2021
-
 \section{Conclusions}\label{sec:conclusions}
@@
 \subsection{Implications for the ASHRAE 55 and ISO 7730 Standards}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
-Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion wordwide since both models are widely adopted and incorporated in many building codes and Standards worldwide.
-The \gls{55} Standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that clearly explain the benefit of using it.
-The \ac{pmv-ce} is more complex, computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
+Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion worldwide since both models are widely adopted and incorporated in many building codes and Standards worldwide.
+The \gls{55} Standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that clearly explains the benefit of using it.
+The \ac{pmv-ce} is more complex, and computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
 Hence we do not recommend its use.
@@
 The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
-Based on our results we recommend restricting the use of the \ac{pmv} only to a values $\lvert \textrm{PMV}\lvert \leq 0.5$.
-Or in other words, we only recommend its use to determine if an environment may be deemed comfortable by a large group of occupants.
+Based on our results we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
+In other words, we only recommend its use to determine if an environment may be deemed comfortable by a large group of occupants.
 The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classy if people are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
@@
 Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
-Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
+Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
 In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
-Consequently, we suggest reducing both Standards applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
+Consequently, we suggest reducing both Standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
 \begin{table}[htb!]
@@
 The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort in our dataset, and it is not a valid assumption.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset, and it is not a valid assumption.
 We recommend the use of the thermal preference scale to determine how people perceive their thermal environment, this scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.

### 2024-08-15 Manuscript/src/sections/introduction.tex
fix: minor typos
feat: code to analyse data from studies with three air speed measurements

@@
 \mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
-The \ac{pmv} is an index that predicts the mean value of the thermal sensation votes (self-reported perceptions) of a large group of persons on a sensation scale expressed from –3 to +3 corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
+The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
@@
 This highlights this model's extensive adoption and use among the scientific community.
-
 To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.
@@
 In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-The value of \ac{ce} is calculated using the \ac{set} equation, and it is then subtracted from both the \ac{tdb} and \ac{tr}.
+The value of \ac{ce} is calculated using the \ac{set} equation.
+The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
 The resulting values become the new inputs in the \ac{pmv} model.
-Since the \ac{ce} accounts for the convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+Since the \ac{ce} accounts for  convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
 The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
@@
 Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-However, while the \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
+However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
 We could not find a rationale for this choice in the available literature.
@@
 To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models, which are used in the \gls{7730} and \gls{55} Standards.
-
 Some notable works that tried to determine the accuracy of the \ac{pmv} model include the manuscript from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
@@
 They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
-The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
+The \ac{pmv} was free from bias only when it was used to predict thermal neutrality~\cite{Humphreys2002}.
 \mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
-They found that the thermal sensation predicted by the PMV model, on average, is one full thermal sensation scale unit away from the subject’s responses, confirming the results of~\cite{Humphreys2002}.
-The accuracy of PMV was only \qty{34}{\percent}, the model has a slightly higher prediction accuracy for sensation close to neutrality, but the accuracy declined towards either end of the thermal sensation scale, and it overestimated both hot and cold sensations.
+They found that the thermal sensation predicted by the PMV model, on average, is one full thermal sensation scale unit away from the subject’s responses, confirming the results of~\mycite{Humphreys2002}.
+\mycite{Cheung2019} also concluded that the accuracy of \ac{pmv} was only \qty{34}{\percent}, the model has a slightly higher prediction accuracy for sensation close to neutrality, but the accuracy declined towards either end of the thermal sensation scale, and it overestimated both hot and cold sensations.
 They also found that the Predicted Percentage of Dissatisfied (PPD) failed to predict the percentage of unacceptable votes if the thermal sensation was predicted using the PMV model and suggested its removal from \gls{55} Standards.
@@
 Reporting the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv} is particularly important for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral'.
-Finally, they only reported the overall bias for the whole dataset, even though the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
-\mycite{tartarini_prediction_2023} also published a conference paper on this topic.
-They concluded that the \ac{pmv} and \ac{pmv-ce} models accuracy is \qty{35}{\percent} and \qty{36}{\percent} for $\mid$\ac{pmv}$\mid \leq 2$, \qty{43}{\percent} and \qty{44}{\percent} for $\mid$\ac{pmv}$\mid \leq 1$ in predicting the self-reported thermal sensation of people, respectively.
+Finally, \mycite{Yao2022} only reported the overall bias for the whole dataset, even though the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
+\mycite{tartarini_prediction_2023} concluded that the \ac{pmv} and \ac{pmv-ce} models accuracy is \qty{35}{\percent} and \qty{36}{\percent} for $\mid$\ac{pmv}$\mid \leq 2$, \qty{43}{\percent} and \qty{44}{\percent} for $\mid$\ac{pmv}$\mid \leq 1$ in predicting the self-reported thermal sensation of people, respectively.
 However, their analysis is limited to providing the precision of the model alone, without reporting the sensitivity (also commonly referred to as recall) and they do not provide the bias of the two models.
@@
 \subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
-We aim to compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards.
-We used the self-reported \ac{tsv} recorded in the \ac{db2}.
+In this paper, we aim to compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards, respectively.
+We used the self-reported \ac{tsv} recorded in the \acf{db2}.
 We aim to determine which \ac{pmv} model is most accurate and which are its applicability limits.
-This will allow policymakers to make an informed decision on which model should be used in standards.
+This will allow \ac{pmv} users - engineers, architects, and educators - to make an informed decision on which model they should use and will help policymakers to select the most appropriate model to be used in Standards.

### 2024-08-15 Manuscript/src/sections/methodology.tex
feat: code to analyse data from studies with three air speed measurements

@@
 \subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
-Not all researchers logged all the six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
+Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
 We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
-This filtered out approximately \var{entries_db_valid} votes.
+This filtered out approximately \var{entries_db_valid} of the data.
 We decided to keep entries that did not have \ac{tr}.
@@
 However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
+
+In addition, it should also be noted that while all the data included in the \ac{db2} is published in peer-reviewed papers, not all entries were collected using the same methodology.
+For example, while both thermal comfort Standards recommend measuring \ac{v} at three different heights, this was not always done.
+This could introduce a bias in the data.
+Since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
+\begin{enumerate}
+    \item all the data points,
+    \item all the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s},
+    \item only the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
+    These results are reported in the supplementary material.
+\end{enumerate}
+We believe that this approach will help us understand the impact of the data quality on the results of the models.
+In particular, by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the Standards.
+
 We used Python to analyse and visualise the results.
@@
 To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
-These differences, aka bias, quantify the success of the model in predicting \ac{tsv}.
+These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
 However, on their own are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.

### 2024-08-15 Manuscript/src/sections/results.tex
feat: code to analyse data from studies with three air speed measurements

@@
 The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
-Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
+Consequently, we also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
 
@@
 The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
+In the supplementary material, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
+% todo add results to supplementary material and reference to section in the supplementary material
+
+Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
 % todo call charts a,b,c,d
@@
 \end{figure*}
-Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
 Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
@@
 % todo compare how the models compare for PMV < |0.5| not |1.5|
+% todo check the results
 However, the macro score, within this range, remained practically unchanged.
@@
     \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
-    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.}
+    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
+    The black dashed line represents the ideal line where the slope is 1 and the intercept is 0.}
     \label{fig:bubble_models_vs_tsv}
@@
 We only binned the data to aid the visualization of a large dataset.
-If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and have a slope of 1.
+If the model is to accurately predict the thermal sensation of people, the regression line (black continuous line) should pass through the origin of the cartesian plane and have a slope of 1.
+This ideal line is represented by the black dashed line.
 The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.25} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
@@
     \label{fig:hist_discrepancies}
+%    todo add caption
 \end{figure*} 
@@
 However, this may not always be the case, for example, if a researcher often underestimated the clothing insulation of participants or their activity levels.
+% todo add an additional sentence also about the airspeed error measurement
 

### 2024-08-15 Manuscript/src/sections/titleAbstract.tex
feat: code to analyse data from studies with three air speed measurements

@@
         Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
-        The \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion for many users who find it difficult to decide which model to use.
+        The \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion for many users who often struggle to determine which model is most appropriate for their needs.
 
@@
 
-        We recommend using the original \ac{pmv} instead of the \ac{pmv-ce} and we suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
+        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce} and we suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
         This recommendation already aligns with the \gls{55} Standard's definition which states that the \ac{pmv} model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce \textit{satisfactory thermal environmental conditions} for a majority of the occupants within the space''.
         The \ac{pmv} model cannot determine the degree to which people are dissatisfied.
-        A value higher or lower than \num{.5} or \num{-.5}, respectively only indicates that people are on the `warm' or `cold' side.
+        A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
         Finally, our results show that the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.

### 2024-08-15 main.py

@@ -12,7 +12,8 @@ import seaborn as sns
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.gridspec import GridSpec
 from numba import jit
-from pythermalcomfort import p_sat_torr, check_standard_compliance_array
+from pythermalcomfort.psychrometrics import p_sat_torr
+from pythermalcomfort.utilities import check_standard_compliance_array
 from pythermalcomfort.models import (
     athb,
     cooling_effect,
@@ -1673,7 +1674,104 @@ def plot_distribution_variable():
     plt.show()
 
 
+def analyse_studies_high_v():
+    # [['contributor', 'publication', 'region', 'country', 'city']]
+    df.dropna(subset='vel_l').groupby(['contributor', 'publication']).sum().sort_values(by='publication', ascending=False)
+    df_publications = df[df['vel']>0.2].dropna(subset='vel_l').groupby(['contributor', 'publication'])['vel'].count().sort_values(ascending=False)
+    print(df_publications.to_markdown())
+
+    df_perc = (
+        (
+            df.groupby(["country", "contributor"])["record_id"]
+            .count()
+            .sort_values(ascending=False)
+            / df.shape[0]
+            * 100
+        )
+        .to_frame()
+        .rename(columns={"record_id": "percentage"})
+        .round()
+    )
+    df_count = (
+        (
+            df.groupby(["country", "contributor"])["record_id"]
+            .count()
+            .sort_values(ascending=False)
+        )
+        .to_frame()
+        .rename(columns={"record_id": "count"})
+        .round(1)
+    )
+    df_sources = pd.concat([df_count, df_perc], axis=1).head(10)
+    df_sources.loc["Total"] = df_sources.sum()
+    print(df_sources.reset_index().to_markdown())
+
+    top_contributors = (
+        df.groupby(["country", "contributor"])["record_id"]
+        .count()
+        .sort_values(ascending=False)
+        .head(10)
+        .reset_index()["contributor"]
+        .to_list()
+    )
+
+    df_top = df[df["contributor"].isin(top_contributors)]
+
+    sns.boxenplot(df_top, y="contributor", x="vel")
+    plt.tight_layout()
+    plt.show()
+
+    # only those with v > 0.2
[…]

### 2024-08-28 Manuscript/src/sections/appendix.tex
feat: wrote code to analyse data from studies with three air speed measurements

@@
 
-\section{Comparison between the PMV formulations with the \ac{pmvg}}\label{sec:comparison-between-the-pmv-formulations-with-the-two-node-model}
-We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the \ac{pmvg}.
-% SS provide more information why this was done
-% Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
-Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
-The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
-The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
-% Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
-The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
-% Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from? How can a PMV model be described by an air speed?
-These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
-This is also an unexpected result and we will discuss why this may be in the following paragraph.
-% SS When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
-1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.
+\section{Analysis of the dataset with air speed measurements at three heights}\label{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}
 
+In this section, we analyzed the dataset with air speed measurements at three heights.
+Only eight papers in the \ac{db2} had air speed measurements at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+A summary of the papers is provided in Table~\ref{tab:three_heights_papers}.
+The table also shows the number of measurements for each paper and how much (percentage) each study contributed to the dataset.
+\begin{table}[htb!]
+    \centering
+    \input{tables/publications_with_three_heights.tex}
+    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
+    \label{tab:three_heights_papers}
+\end{table}
+The paper by Zhang et al.\ and Cena et al.\ contributed the most to the dataset with \qty{58}{\percent} and \qty{23}{\percent}, respectively.
+These \num{1184} where used to compare the \ac{pmv} and \ac{pmv-ce} models using the same methodology as in Section~\ref{sec:results}.
+The air speed distribution at the three heights for each paper is shown in Figure~\ref{fig:boxenplot_wind_speed_three_heights}.
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/boxenplot_wind_speed_three_heights}
+    \caption{Wind speeds distribution at three heights.}
+    \label{fig:boxenplot_wind_speed_three_heights}
+\end{figure}
+The distribution of air speed at the three heights varies between the papers, hence, it is difficult to make generalizations.
+However, it should be noted that in most of the studies the air speed measured at the chest height was a good representation of the air speed at the other two heights.
+This assumption is not valid for Brenton et al.\ (1994) and Donnini et al.\ (1996) it should be noted that these two studies had a only \num{11} and \num{7} measurements, respectively.
+In this section we are not showing the F1-score for the \ac{pmv} and \ac{pmv-ce} models for the different subsets of data since we have already presented the results in Table~\ref{tab:f1}.
+We would like to remind the reader that the F1-score for \ac{pmv-ce} model was lower than the \ac{pmv} model for this subset of data, i.e, \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and \ac{v} measured at three heights.
+As previously discussed this is an unexpected result since the \ac{pmv} should be significantly more accurate than the \ac{pmv-ce} model for this subset of data.
+In this section we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since we have a very limited number of measurements for this subset of data moreover this analysis has some limitations as previously discussed.
+
+We, however, plotted the \ac{pmv} and \ac{pmv-ce} values against the \ac{tsv} values and we plotted the \ac{lowess} curves to visualize the prevailing data trend.
+The results are shown in Figure~\ref{fig:bubble_models_vs_tsv_three_heights}.
 \begin{figure*}[htb!]
     \centering
-    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
-    \caption{}
-    \label{fig:pmv_two_node_comparison}
+    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv_three_heights}
+    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data for the dataset with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and measured at three heights.
+    Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
+    The brown dashed line represents the identity line where the slope is 1 and the intercept is 0.
+    Comparison between the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights.}
+    \label{fig:bubble_models_vs_tsv_three_heights}
 \end{figure*}
+The results show that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
+The intercept of the \ac{pmv} model is \num{-.2} while the intercept of the \ac{pmv-ce} model is \num{-.39}.
+Moreover, the \ac{pmv-ce} \ac{lowess} curve is only marginally higher than \num{0} despite partcipants reporting a \ac{tsv} of \num{3}.
+
+Despite the fact that these results already showing that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for this subset of data, we also calculated the overall bias of the \ac{pmv} and \ac{pmv-ce} models.
+The results are shown in Figure~\ref{fig:hist_discrepancies_three_heights}.
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/hist_discrepancies_three_heights}
+    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights.}
+    \label{fig:hist_discrepancies_three_heights}
+\end{figure}
+The \ac{pmv} model has a bias of \num{-.1} while the \ac{pmv-ce} model has a bias of \num{-.28}.
+Moreover, the distribution of the discrepancies for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
+These results allows us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
+Consequently, we recommend using the \ac{pmv} model instead of the \ac{pmv-ce}.
+
+%\section{Comparison between the PMV formulations with the \ac{pmvg}}\label{sec:comparison-between-the-pmv-formulations-with-the-two-node-model}
+%We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the \ac{pmvg}.
+%% SS provide more information why this was done
+%% Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
+%Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
+%The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
+%The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
+%% Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
+%The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
+%% Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from? How can a PMV model be described by an air speed?
+%These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
+%This is also an unexpected result and we will discuss why this may be in the following paragraph.
+%% SS When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
+1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.
+%
+%\begin{figure*}[htb!]
+%    \centering
+%    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
+%    \caption{}
+%    \label{fig:pmv_two_node_comparison}
+%\end{figure*}
+%
+%% todo include the folling
+%%I think it is reasonable to have an appendix where we do the analysis for the dataset with only three heights.
+%%Given that we have only five authors, we can provide a bit more info about those, maybe in a table format, (paper DOI, publication year, measuring heights, type of anemometer, accuracy, compliance with ISO 7726 or ASHRAE 55, type of building, how long they measure, etc) and have a graph of air speed letter-plot as you did in the email to Ed.
+For example, for the Yufeng's paper he specified that he followed ISO 7726 (so very good), he used an omni-directional anemometer with a range of 0-5 m/s and accuracy of +-0.02 m/s between 0 and 0.99 m/s and +-0.1 m/s between 1 and 5 m/s).
+The measurements were taken at 0.1, 0.6 and 1.1 m.

### 2024-08-28 Manuscript/src/sections/introduction.tex
feat: wrote code to analyse data from studies with three air speed measurements

@@
 Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+To quantify by how much the two models differ, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models and we plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
 %The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
+    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models.
+    The chart shows how the comfort regions change as a function of \ac{v}.
+    We assumed $M$~=~1.2 met, $i_{cl}$~=~0.5 clo.
+    \label{fig:comfort_regios_pmv_pmvce}}
+\end{figure}
 

### 2024-08-28 Manuscript/src/sections/results.tex
feat: wrote code to analyse data from studies with three air speed measurements

@@
 This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
-The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
+The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
 Consequently, we also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
@@
 Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature.
-Less than half (\num{23300}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
+Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
 Ages are not normally distributed, the same is true for the running mean outdoor temperature.
@@
 
-In Table~\ref{tab:f1}, we report the F1-scores for both models, using the whole dataset, only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
-The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
-For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models improved and the former outperformed the latter.
-This is an unexpected result since the \ac{pmv-ce}, to justify its use, should significantly outperform the \ac{pmv} model when analyzing this subset of data.
-Finally, the F1-micro and weighted scores of both models increased by about 0.1 when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
-% todo compare how the models compare for PMV < |0.5| not |1.5|
-% todo check the results
-However, the macro score, within this range, remained practically unchanged.
+In Table~\ref{tab:f1}, we report the F1-scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$, and iv) only keeping datapoints that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
 \begin{table}[htb!]
@@
     \input{tables/f1.tex}
-    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models for three datasets.}
+    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
     \label{tab:f1}
 \end{table}
+The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
+For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models only slightly improved.
+We then analyzed the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{v} measured at three heights.
+The F1-macro score of the \ac{pmv-ce} increased, however, it was lower than the F1-macro score of the \ac{pmv} model.
+This is an unexpected result since the \ac{pmv-ce}, to justify its use, should significantly outperform the \ac{pmv} model when analyzing this subset of data.
+All the F1 scores significantly increased when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
 
@@
     Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
-    The black dashed line represents the ideal line where the slope is 1 and the intercept is 0.}
+    The brown dashed line represents the identity line where the slope is 1 and the intercept is 0.}
     \label{fig:bubble_models_vs_tsv}
@@
 We only binned the data to aid the visualization of a large dataset.
-If the model is to accurately predict the thermal sensation of people, the regression line (black continuous line) should pass through the origin of the cartesian plane and have a slope of 1.
-This ideal line is represented by the black dashed line.
-The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.25} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
-This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times, respectively.
+If the model is to accurately predict the thermal sensation of people, the regression line (black continuous line) should pass through the origin of the cartesian plane and have a slope of 1, also called identity line which is represented by the brown dashed line.
+The intercepts for the \ac{pmv} is \qty{-0.22} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
+This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{54}{\percent} and \qty{58}{\percent} of the times, respectively.
 Both \ac{pmv} formulations always under-predicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
@@
 \subsection{Model Bias}\label{subsec:model-bias}
+
 \subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
 We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
-Results are presented in Figure~\ref{fig:hist_discrepancies}.
+Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
+Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
+The results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+We report the median and inter-quantile ranges since data are not normally distributed.
 %  We cannot change into a letter value plot since in that case, it would be hard to highlight the number of responses that are within the range from -.5 to 5.
@@
     \includegraphics[width=\textwidth]{figures/hist_discrepancies}
-    \caption{}
+    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+    The x-axis shows the discrepancy between the predicted and self-reported thermal sensation vote.
+    We have binned the data into intervals of equal width.
+    We colored the bars in the range from \num{-.5} to \num{.5} to highlight the number of responses that are within this range.
+    The chart also shows the median and inter-quantile ranges of the data.
+    }
     \label{fig:hist_discrepancies}
-%    todo add caption
 \end{figure*}
-Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
-Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
-We report the median and inter-quantile ranges since data are not normally distributed.
 When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
@@
 \subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
-In the previous section, we presented the overall bias of the two \ac{pmv} formulations.
-We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo}.
-We did not include \ac{tr} in the analysis since \ac{tr} was not measured in several studies, hence, we estimated it as detailed in the Methodology.
-Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
-Figure~\ref{fig:dist_input_data} shows the results.
-In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
-All the violin plots have the same area to aid visualization, even though the number of points in each bin is not constant.
-%  I need to have the red lines for the same inverse reason why I cannot use the boxenplot in the previous chart
-% todo Shall I report the p-values?
+After presenting the overall bias of the two \ac{pmv} formulations, we analyzed the bias as a function of the five input variables \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo} and we are presenting the results in Figure~\ref{fig:bias_models}.
 \begin{figure*}[htb!]
@@
     The $]$ means that the right value is included in the range.
+    The red dashed lines show the acceptable range of bias.
+    Please note that each bin has a different number of data points.
     }
@@
 \end{figure*}
+We did not include \ac{tr} in the analysis since \ac{tr} was not measured in several studies, hence, we estimated it as detailed in the Methodology.
+Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
+In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
+All the boxen-plots have the same area to aid visualization, even though the number of points in each bin is not constant.
 The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all input variables.
@@
 The \ac{pmv} model uses the overall heat losses or gains to calculate a \ac{pmv} value which should represent a \ac{tsv} vote.
-Heat losses or gains are only a proxy for the thermal stress that the human body has to undergo to maintain a constant internal temperature.
+Heat losses or gains are only a proxy for the thermoregulation effort that the human body has to undergo to maintain a constant internal temperature.
 Fanger created this equation based on a limited dataset collected in a laboratory and this equation was never updated or revisited~\cite{Fanger1970}.
@@
 To continue using the \ac{pmv} model beyond an absolute value of \num{0.5}, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced:
-i) verify the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
-ii) if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
-iii) currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different sets of  equations should be used;
-iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
+\begin{enumerate}
+    \item verify the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
+    \item if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
+    \item currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different sets of equations may be used;
+    \item the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
     For example, while the \ac{pmv} model may assume that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
     Participants in commercial buildings may, on the contrary, deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.
+\end{enumerate}
 
 A combination of all the above-mentioned factors likely plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
-The latter hypothesis -- iv) anchoring -- has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
+The latter hypothesis -- 4) anchoring -- has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
 Supporting evidence to this issue is provided by \mycite{schweiker2020evaluating} who showed that the distances between the anchors are not constant and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
@@
 However, this may not always be the case, for example, if a researcher often underestimated the clothing insulation of participants or their activity levels.
-% todo add an additional sentence also about the airspeed error measurement
+Moreover, most of the data in the \ac{db2} also only contains entries where the \ac{v} was measured at a single height.
+This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body and this is not always the case.
+Measuring the air speed at a single height may lead to record higher values than the averaged velocities that are required for the standards’ comfort models.
+This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently the associated sensation prediction will be cooler than it should be.
+To account for this error, we only included the studies where the \ac{v} was measured at three different heights in the supplementary material.
+%todo add reference to the supplementary material
+However, these results were not significantly different from those presented in this paper and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.
+It should also be noted that in buildings air speed is rarely measured, and it is never measured at three different heights.
 
@@
 Using a continuous scale, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.
-They also showed that the \ac{tsv}'s assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
+They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
 An alternative solution would be to transform the \ac{pmv} output to an ordinal and categorical seven-point scale.
-
-% \paragraph{Other neglected factors}
-% Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) that interacts with the surrounding environment.
-% Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.

### 2024-08-28 main.py

@@ -23,8 +23,10 @@ from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score
 from sklearn.metrics import f1_score
 from sklearn.metrics import r2_score, mean_absolute_error
+from scipy import optimize
+import scipy
 
-# mpl.use("TkAgg")
+mpl.use("TkAgg")
 
 warnings.filterwarnings("ignore")
 
@@ -61,6 +63,8 @@ palette_tsv = [
 # sns.palplot(palette_tp)
 # sns.palplot(palette_primary)
 
+percentiles_to_show = [0.025, 0.25, 0.5, 0.75, 0.975]
+
 palette_primary = [
     "#FFBA22",
     "#FF2380",
@@ -84,6 +88,93 @@ palette_primary = [
     "#149678",
     "#FFE146",
 ]
+
+sns.set_context("paper")
+mpl.rcParams["figure.figsize"] = [8.0, 3.5]
+sns.set_style(
+    "whitegrid",
+    {
+        "grid.color": ".85",
+        "grid.linewidth": "1",
+        "grid.linestyle": "--",
+        "axes.spines.left": False,
+        "axes.spines.bottom": False,
+        "axes.spines.right": False,
+        "axes.spines.top": False,
+    },
+)
+
+c_gold = "#FDB515"
+c_brown = "#6C3302"
+c_blue = "#3B7EA1"
+c_green = "#53626F"
+
+plt.rc("axes.spines", top=False, right=False, left=False)
+plt.rcParams["font.family"] = "sans-serif"
+plt.rcParams["font.size"] = 8.8
+
+applicability_limits = {
+    "ta": [10, 30],
+    "tr": [10, 40],
+    "vel": [0, 1],
+    "clo": [0, 1.5],
+    "met": [1, 4],
+    "thermal_sensation": [-3.5, 3.5],
+    "pmv": [-3.49999, 3.5],
+    "pmv_ce": [-3.49999, 3.5],
+    "pmv_set": [-3.49999, 3.5],
+    "pmv_gagge": [-3.49999, 3.5],
+    "pmv_toby": [-3.49999, 3.5],
+    "rh": [0, 100],
+    "pa": [0, 2700],
+}
+
+var_names = {
+    "ta": r"$t_{db}$",
+    "tr": r"$\overline{t_{r}}$",
+    "top": r"$t_{o}$",
+    "vel": r"$V$",
+    "rh": r"RH",
[…]

### 2024-08-29 Manuscript/src/sections/appendix.tex
feat: updated the results using v_r instead of v

@@
 In this section, we analyzed the dataset with air speed measurements at three heights.
-Only eight papers in the \ac{db2} had air speed measurements at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+Only a limited subset of papers in the \ac{db2} had air speed measurements at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
 A summary of the papers is provided in Table~\ref{tab:three_heights_papers}.
@@
 \end{table}
-The paper by Zhang et al.\ and Cena et al.\ contributed the most to the dataset with \qty{58}{\percent} and \qty{23}{\percent}, respectively.
-These \num{1184} where used to compare the \ac{pmv} and \ac{pmv-ce} models using the same methodology as in Section~\ref{sec:results}.
+The paper by Cena et al.\ (1999), de Dear et al. (1994), and Zhang et al.\ (2013 and 2010) contributed the most to the dataset.
+These \num{3063} entries where used to compare the \ac{pmv} and \ac{pmv-ce} models using the same methodology as in Section~\ref{sec:results}.
 The air speed distribution at the three heights for each paper is shown in Figure~\ref{fig:boxenplot_wind_speed_three_heights}.
@@
 However, it should be noted that in most of the studies the air speed measured at the chest height was a good representation of the air speed at the other two heights.
-This assumption is not valid for Brenton et al.\ (1994) and Donnini et al.\ (1996) it should be noted that these two studies had a only \num{11} and \num{7} measurements, respectively.
+This assumption is not valid for Brenton et al.\ (1994) however this study contributed only \qty{1}{\percent} of the dataset.
+The adjustment of \ac{v} based on the \ac{met} significantly increased the base value of the air speed measured by all the studies.
+With the value of \ac{v} being \var{delta_v_mean} higher than the average of the other wind speeds combined.
+The discussion on weather this increase is justified is beyond the scope of this paper since both Standards require the user to adjust the air speed based on the \ac{met} value.
+
 In this section we are not showing the F1-score for the \ac{pmv} and \ac{pmv-ce} models for the different subsets of data since we have already presented the results in Table~\ref{tab:f1}.
-We would like to remind the reader that the F1-score for \ac{pmv-ce} model was lower than the \ac{pmv} model for this subset of data, i.e, \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and \ac{v} measured at three heights.
-As previously discussed this is an unexpected result since the \ac{pmv} should be significantly more accurate than the \ac{pmv-ce} model for this subset of data.
-In this section we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since we have a very limited number of measurements for this subset of data moreover this analysis has some limitations as previously discussed.
+We would like to remind the reader that the F1-macro score for \ac{pmv-ce}~=~\num{.16} was lower than the \ac{pmv}~=~\num{.17} for this subset of data, i.e, \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and \ac{v} measured at three heights.
+As previously discussed this is an unexpected result since the \ac{pmv-ce} should be significantly more accurate than the \ac{pmv} model for this subset of data to justify its use.
+In this section we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since we have a very limited number of measurements for this subset of data moreover this analysis has several limitations as previously discussed.
 
@@
 The results show that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
-The intercept of the \ac{pmv} model is \num{-.2} while the intercept of the \ac{pmv-ce} model is \num{-.39}.
-Moreover, the \ac{pmv-ce} \ac{lowess} curve is only marginally higher than \num{0} despite partcipants reporting a \ac{tsv} of \num{3}.
+The intercept of the \ac{pmv} model is \num{-.31} while the intercept of the \ac{pmv-ce} model is \num{-.41}.
+Moreover, the \ac{pmv-ce} \ac{lowess} curve is only marginally higher than \num{0} despite participants reporting a \ac{tsv} of \num{3}.
 
-Despite the fact that these results already showing that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for this subset of data, we also calculated the overall bias of the \ac{pmv} and \ac{pmv-ce} models.
+Finally, we are showing calculated the overall bias of the \ac{pmv} and \ac{pmv-ce} models.
 The results are shown in Figure~\ref{fig:hist_discrepancies_three_heights}.
@@
 \end{figure}
-The \ac{pmv} model has a bias of \num{-.1} while the \ac{pmv-ce} model has a bias of \num{-.28}.
-Moreover, the distribution of the discrepancies for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
-These results allows us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
+The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.38}.
+The inter-quartile range for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
+These results allow us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
 Consequently, we recommend using the \ac{pmv} model instead of the \ac{pmv-ce}.
@@
 %\end{figure*}
-%
-%% todo include the folling
-%%I think it is reasonable to have an appendix where we do the analysis for the dataset with only three heights.
-%%Given that we have only five authors, we can provide a bit more info about those, maybe in a table format, (paper DOI, publication year, measuring heights, type of anemometer, accuracy, compliance with ISO 7726 or ASHRAE 55, type of building, how long they measure, etc) and have a graph of air speed letter-plot as you did in the email to Ed.
-For example, for the Yufeng's paper he specified that he followed ISO 7726 (so very good), he used an omni-directional anemometer with a range of 0-5 m/s and accuracy of +-0.02 m/s between 0 and 0.99 m/s and +-0.1 m/s between 1 and 5 m/s).
-The measurements were taken at 0.1, 0.6 and 1.1 m.

### 2024-08-29 Manuscript/src/sections/conclusions.tex
feat: updated the results using v_r instead of v

@@
 The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
-Both models had an accuracy lower than \qty{2}{\percent} when predicting either `warm', `hot', or `cold'.
-This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
+Both models had an accuracy lower than \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
+This value is significantly lower and worse than random guessing, this means that the models are actively misleading the users.
 When only comparing the results of the two models for entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} -- range in which the two models differ -- the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
@@
         \ac{tr} & \qtyrange{17}{30}{\celsius} \\
-        \ac{rh} & \qtyrange{20}{80}{\percent} \\
+        \ac{rh} & \qtyrange{10}{80}{\percent} \\
         \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
         \ac{met} & \qtyrange{1}{2}{met} \\
-        % todo shall we also include limits on Age?
-        % todo shall we also include airspeed?
         \bottomrule

### 2024-08-29 Manuscript/src/sections/introduction.tex
feat: updated the results using v_r instead of v

@@
 It is the most extensively used thermal comfort index for estimating thermal sensation.
-A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{1600} documents.
-Out of those \num{1082} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
+A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{2050} documents.
+Out of those \num{1200} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
 This highlights this model's extensive adoption and use among the scientific community.
@@
 The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
-In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-The value of \ac{ce} is calculated using the \ac{set} equation.
-The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
-The resulting values become the new inputs in the \ac{pmv} model.
-Since the \ac{ce} accounts for  convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
-The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
-Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
-To quantify by how much the two models differ, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models and we plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
-%The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
 \begin{figure}[!htb]
     \centering
-    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
-    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models.
-    The chart shows how the comfort regions change as a function of \ac{v}.
-    We assumed $M$~=~1.2 met, $i_{cl}$~=~0.5 clo.
-    \label{fig:comfort_regios_pmv_pmvce}}
-\end{figure}
-
-However, this method has limitations.
-First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
-Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
-Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
-We could not find a rationale for this choice in the available literature.
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tikzpicture}[node distance=1.2cm]
-        \node (start) [startstop] {\ac{pmv} calculation as per \gls{55}};
-        \node (dec1) [decision, below of=start, yshift=-.25cm] {\ac{met} $> \qty{1}{met}$ };
-        \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$};
-        \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\ac{met} $> \qty{1.2}{met}$ };
-        \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\ac{clo} = \ac{clo}$(0.6 + 0.4/$\ac{met}$)$};
-        \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\ac{v} $> \qty{0.1}{\m\per\s}$ };
-        \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\ac{pmv}(\ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{met}, \ac{clo})};
-        \node (pro3) [process, below of=dec3, yshift=-.25cm, xshift=3cm] {Calculate \ac{ce}};
-        \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\ac{pmv}(\ac{tdb} - \ac{ce}, \ac{tr} - \ac{ce}, \ac{rh}, \ac{v} = \qty{0.1}{\m\per\s}, \ac{met}, \ac{clo})};
+    \begin{tikzpicture}[node distance=1cm and 3cm, transform shape]
+        \node (start) [startstop] {\acs{pmv} calculation as per \gls{55}};
+        \node (dec1) [decision, below of=start, yshift=-.25cm] {\acs{met} $> \qty{1}{met}$ };
+        \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\acs{v} = \acs{v}$+ 0.3 ($\acs{met}$-1)$};
+        \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\acs{met} $> \qty{1.2}{met}$ };
+        \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\acs{clo} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$};
+        \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\acs{v} $> \qty{0.1}{\m\per\s}$ };
+        \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\acs{pmv}(\acs{tdb}, \acs{tr}, \acs{rh}, \acs{v}, \acs{met}, \acs{clo})};
+        \node (pro3) [process, below of=dec3, yshift=-.25cm, xshift=3cm] {Calculate \acs{ce}};
+        \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\acs{pmv}(\acs{tdb} - \acs{ce}, \acs{tr} - \acs{ce}, \acs{rh}, \acs{v} = \qty{0.1}{\m\per\s}, \acs{met}, \acs{clo})};
 
@@
 \end{figure}
+In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
+The reader should note that \ac{v} is not the measured value.
+The value of \ac{v} measured by the sensors, should be adjusted based on the \ac{met} of the occupant if \ac{met}~$>$~\qty{1}{met} using this equation \ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$.
+The value of \ac{ce} is calculated using the \ac{set} equation.
+The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
+The resulting values become the new inputs in the \ac{pmv} model.
+Since the \ac{ce} accounts for  convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+The other three input parameters (\ac{clo}, \ac{rh}, \ac{met}) remain unchanged.
+Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
+To show to the reader by how much the output of the two models differ, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clo}~=~\qty{0.5}{clo}.
+We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
+    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models.
+    The chart shows how the comfort regions change as a function of \ac{v}.
+    We assumed $M$~=~1.2 met, $i_{cl}$~=~0.5 clo.
+    \label{fig:comfort_regios_pmv_pmvce}}
+\end{figure}
+The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} the comfort regions estimated using the \ac{pmv-ce} is wider than the one calculated using the \ac{pmv} and shifted towards the warmer temperatures.
 
-Choosing between the \ac{pmv} and \ac{pmv-ce} is consequently now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many building codes and standards worldwide.
+The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}.
+However, this method has its own limitations.
+First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
+Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
+Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
+However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
+We could not find a rationale for this choice in the available literature.
 
@@
 \mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
+Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models since they are the most widely used thermal comfort models worldwide and are incorporated in the \gls{7730} and \gls{55} Standards.
+
+Choosing between the \ac{pmv} and \ac{pmv-ce} is consequently now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many building codes and standards worldwide.
 
@@
 \subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
-In this paper, we aim to compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards, respectively.
+In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards, respectively.
+The paper objective is to quantitatively answer the following questions:
+which model should you use?
+Which model is more accurate?
 We used the self-reported \ac{tsv} recorded in the \acf{db2}.

### 2024-08-29 Manuscript/src/sections/methodology.tex
feat: wrote code to analyse data from studies with three air speed measurements
feat: updated the results using v_r instead of v

@@
  We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the self-reported \ac{tsv}.
-Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
 
@@
 This filtered out approximately \var{entries_db_valid} of the data.
-We decided to keep entries that did not have \ac{tr}.
-However, since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: $2 t_{o} – t_{db}$.
+We decided to keep entries that did not have \ac{tr} and we planned to calculate it using the \ac{tdb} and \ac{to} values using the following equation: $2 t_{o} – t_{db}$.
 If the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
-We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
-There is also evidence that in many conditions is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
+However, all the entries that had \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv} also had \ac{tr}.
+Consequently, we did not need to make any assumptions about the \ac{tr} values.
+In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the \ac{v} measured by the researchers.
+This is because both \ac{pmv} formulations use the \ac{v} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
+%We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
+%There is also evidence that in many conditions is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
 
@@
 and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
-We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
-We used these inputs to calculate the \ac{pmv} and \ac{pmv-ce} values.
+The adjusted total clothing insulation and relative airspeed, as required by both Standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.
 
@@
 For example, while both thermal comfort Standards recommend measuring \ac{v} at three different heights, this was not always done.
-This could introduce a bias in the data.
-Since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
+This could introduce a bias in the measurements and consequently in the results of the models.
+Moreover, since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
 \begin{enumerate}
@@
     \item only the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
-    These results are reported in the supplementary material.
+    These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
 \end{enumerate}
-We believe that this approach will help us understand the impact of the data quality on the results of the models.
+We believe that comparing the results from these three datasets will help us understand the impact of the data quality on the results of the models.
 In particular, by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the Standards.
@@
 We used Python to analyse and visualise the results.
+The \ac{pmv} results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
 We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \url{https://github.com/FedericoTartarini/paper-pmv-comparison} so other users can test different assumptions.

### 2024-08-29 Manuscript/src/sections/results.tex
feat: updated the results using v_r instead of v

@@
 This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
-The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
-Consequently, we also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
+The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}.
+The median value in the \ac{db2} was \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
+We also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
 
 Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature.
-Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
-Ages are not normally distributed, the same is true for the running mean outdoor temperature.
-About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old, and only \qty{2.5}{\percent} were older than 60.
-This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
-The \ac{db2} does not contain information about the health status of the participants.
-Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
 \begin{figure*}[htb!]
@@
 \end{figure*}
+Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
+Ages are not normally distributed, the same is true for the running mean outdoor temperature.
+About half of the participants (\var{percentage_age_20_35}) were aged between \num{20} and \num{35} years old, and only \var{percentage_age_over_60} were older than 60.
+This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
+The \ac{db2} does not contain information about the health status of the participants.
+Approximately \var{running_mean_10_25} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
 
 The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
-Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
-Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
-The \ac{tsv} dataset is unbalanced.
-Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
-While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
 \begin{figure*}[htb!]
@@
 \end{figure*}
+Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
+Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
+The \ac{tsv} dataset is unbalanced.
+Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
+While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
 
@@
 This finding challenges the above assumption.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under counting thermal discomfort in our dataset.
 On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
@@
 The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are comparing the predicted value from the two models with measured \ac{tsv}.
-The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s};
-hence we also present the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
+However, since the two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s};
+we also present the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s}.
+In this subset of data the differences in the outputs of the two models are more pronounced.
 We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
@@
 The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
-In the supplementary material, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
-% todo add results to supplementary material and reference to section in the supplementary material
+In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
+We do not present these results here since they are similar to the results presented in this section and lead to the same conclusions.
 
 Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
-% todo call charts a,b,c,d
 \begin{figure*}[htb!]
@@
 Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
-The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
+The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively when using the whole dataset.
+The overall accuracy of both \ac{pmv} models increased to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+Interestingly, the \ac{pmv-ce} model outperformed the \ac{pmv} model for \ac{tsv}~=~0 and -1.
+However, the \ac{pmv} model outperformed the \ac{pmv-ce} model when considering \ac{tsv}~=~1.
+We want to remind the reader that the number of data points in each \ac{tsv} bin is different.
+
 Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
 A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
-For both models, the simple accuracy decreases by one point when we only include entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-We want to remind the reader that the number of data points in each bin is different.
 
-In Table~\ref{tab:f1}, we report the F1-scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$, and iv) only keeping datapoints that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+In Table~\ref{tab:f1}, we report the F1-scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping datapoints that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
 \begin{table}[htb!]
@@
 We then analyzed the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{v} measured at three heights.
-The F1-macro score of the \ac{pmv-ce} increased, however, it was lower than the F1-macro score of the \ac{pmv} model.
-This is an unexpected result since the \ac{pmv-ce}, to justify its use, should significantly outperform the \ac{pmv} model when analyzing this subset of data.
+Both F1-macro scores decreased and the \ac{pmv-ce} one \num{0.16} is lower than the value for the \ac{pmv}~=~\num{0.17}.
+This is an unexpected result since the \ac{pmv-ce}, to justify its use, should significantly outperform the \ac{pmv} model when analyzing data that had \ac{v} measured at three heights and arguably a lower bias in the \ac{v} value.
 All the F1 scores significantly increased when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+This highlights that the \ac{pmv} models are more accurate in predicting the thermal sensation of people when the thermal sensation is close to neutral.
 
@@
 
+\subsection{Model Bias}\label{subsec:model-bias}
 The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
@@
 They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
-
-\subsection{Model Bias}\label{subsec:model-bias}
 
@@
 When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
-Surprisingly, we found that the median bias for the \ac{pmv-ce} increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
-These, together with those presented in Table~\ref{tab:f1}, are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
+The median bias for the \ac{pmv-ce} also decreased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it was still higher than the one for the \ac{pmv}.
+These results, together with those presented in Table~\ref{tab:f1}, disprove the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v} than the \ac{pmv}.
 The \ac{pmv-ce} has a higher bias while making the code more complex and computationally intensive to run.
@@
 \end{figure*}
-We did not include \ac{tr} in the analysis since \ac{tr} was not measured in several studies, hence, we estimated it as detailed in the Methodology.
-Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
+%We did not include \ac{tr} in the analysis since \ac{tr} was not measured in several studies, hence, we estimated it as detailed in the Methodology.
+%Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
 In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
@@
 The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all input variables.
-On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 and was significantly worse (\num{-.37} versus \num{-0.07}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.45}{\m\per\s}.
+On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 and was significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
 These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.
@@
 This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently the associated sensation prediction will be cooler than it should be.
-To account for this error, we only included the studies where the \ac{v} was measured at three different heights in the supplementary material.
-%todo add reference to the supplementary material
+To account for this error, we only included the studies where the \ac{v} was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
 However, these results were not significantly different from those presented in this paper and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.
@@
 One minor but not negligible source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
-For example, even if the model was 100~\% accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (+2)' or `hot (+3)'.
+For example, even if the model was \qty{100}{\percent} accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (\num{2})' or `hot (\num{3})'.
 Using a continuous scale, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.

### 2024-08-29 Manuscript/src/sections/titleAbstract.tex
feat: updated the results using v_r instead of v

@@
 
-    \title{PMV: Comparing the prediction accuracy of ISO 7730:2005 and ASHRAE 55:2023}
+    \title{Which PMV model should you use? Comparison of the ISO 7730:2005 and ASHRAE 55:2023 PMV models}
     % alternatives PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?
@@
     \author[label1,label2]{Federico Tartarini\corref{mycorrespondingauthor}}
-    \ead{federicotartarini@berkeley.edu}
+    \ead{federico.tartarini@sydney.edu.au}
     \author[label3]{Stefano Schiavon}
 
-    \address[label1]{Berkeley Education Alliance for Research in Singapore, Singapore}
-    \address[label2]{Heat and Health Research Incubator, Faculty of Health and Medicine, University of Sydney, Sydney, AU}
+    \address[label1]{University of Sydney, Sydney, AU}
+    \address[label2]{Berkeley Education Alliance for Research in Singapore, Singapore}
     \address[label3]{Center for the Built Environment, University of California, Berkeley, CA, USA}
@@
 %        Being relatively simple to calculate and included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
-        The \ac{pmv} model, developed by Fanger, is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
-        Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
-        The \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion for many users who often struggle to determine which model is most appropriate for their needs.
+        The \ac{pmv} model, developed by Fanger, is still currently used in its original form in the \gls{7730}.
+        It has a low prediction accuracy, especially outside thermal neutrality.
+        Several other formulations have been proposed to compensate for its limitations, one is the \ac{pmv-ce}.
+        The \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion.
+%        for many users who often struggle to determine which model is most appropriate for their needs.
+        Which model is more accurate?
+        Which model should you use?
 
-        We compared the accuracy of the two main \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that the use of the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, despite its increased complexity and computational resources needed to calculate the results.
-        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\sec}.
-
-        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce} and we suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
-        This recommendation already aligns with the \gls{55} Standard's definition which states that the \ac{pmv} model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce \textit{satisfactory thermal environmental conditions} for a majority of the occupants within the space''.
-        The \ac{pmv} model cannot determine the degree to which people are dissatisfied.
+        We compared the accuracy of the two \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
+        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce}, since the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, despite its increased complexity.
+%        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\sec}.
+        We suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} model cannot determine the degree to which people are dissatisfied.
+%        This recommendation already aligns with the \gls{55} Standard's definition which states that the \ac{pmv} model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce \textit{satisfactory thermal environmental conditions} for a majority of the occupants within the space''.
         A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
-        Finally, our results show that the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.
-        We, hence, recommend in thermal comfort research the use of the thermal preference scale to determine how people perceive their thermal environment.
+        Finally, the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.
+        We, hence, recommend in thermal comfort research the use of the thermal preference scale instead of the thermal sensation scale.
     \end{abstract}

### 2024-08-29 main.py

@@ -26,6 +26,8 @@ from sklearn.metrics import r2_score, mean_absolute_error
 from scipy import optimize
 import scipy
 
+from matplotlib.lines import Line2D
+
 mpl.use("TkAgg")
 
 warnings.filterwarnings("ignore")
@@ -116,7 +118,7 @@ plt.rcParams["font.size"] = 8.8
 applicability_limits = {
     "ta": [10, 30],
     "tr": [10, 40],
-    "vel": [0, 1],
+    "vel_r": [0, 1],
     "clo": [0, 1.5],
     "met": [1, 4],
     "thermal_sensation": [-3.5, 3.5],
@@ -133,7 +135,7 @@ var_names = {
     "ta": r"$t_{db}$",
     "tr": r"$\overline{t_{r}}$",
     "top": r"$t_{o}$",
-    "vel": r"$V$",
+    "vel_r": r"$V$",
     "rh": r"RH",
     "clo": r"$I_{cl}$",
     "met": r"$M$",
@@ -165,7 +167,7 @@ var_names = {
 var_units = {
     "ta": r"$^{\circ}$C",
     "tr": r"$^{\circ}$C",
-    "vel": r"m/s",
+    "vel_r": r"m/s",
     "rh": r"%",
     "clo": r"clo",
     "met": r"met",
@@ -1154,7 +1156,7 @@ def importing_filtering_processing(load_preprocessed=False):
     save_var_latex("entries_db_all", df_.shape[0])
 
     # entries without ta, rh, v, clo, met
-    df_valid_input = df_.dropna(subset=["ta", "vel", "rh", "met", "clo"])
+    df_valid_input = df_.dropna(subset=["ta", "vel_r", "rh", "met", "clo"])
     df_valid_input_no_tr = df_valid_input.dropna(subset=["tr"]).shape[0]
     save_var_latex(
         "entries_db_valid",
@@ -1212,6 +1214,20 @@ def importing_filtering_processing(load_preprocessed=False):
 
     df_["pmv"] = results_pmv_ppd["pmv"]
     df_["ppd"] = results_pmv_ppd["ppd"]
+
+    results_pmv_ppd = pmv_ppd(
+        tdb=df_.ta,
+        tr=df_.tr,
+        vr=df_.vel_r,
+        rh=df_.rh,
+        met=df_.met,
+        clo=df_.clo_d,
+        standard="ashrae",
+        limit_inputs=False,
+    )
+
[…]

### 2024-09-10 Manuscript/src/sections/appendix.tex
paper: Stefano's comments

@@
     \input{tables/publications_with_three_heights.tex}
-    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
+    \caption{The number of entries used in this analysis grouped by the author/publication.}
     \label{tab:three_heights_papers}
 \end{table}
-The paper by Cena et al.\ (1999), de Dear et al. (1994), and Zhang et al.\ (2013 and 2010) contributed the most to the dataset.
-These \num{3063} entries where used to compare the \ac{pmv} and \ac{pmv-ce} models using the same methodology as in Section~\ref{sec:results}.
+The paper by Cena et al.\ (1999), de Dear et al. (1994), and Zhang et al.\ (2013 and 2010) contributed the most to the dataset (\qty{79}{\percent}).
+These \num{3063} entries were used to compare the \ac{pmv} and \ac{pmv-ce} models using the same methodology as in Section~\ref{sec:results}.
 The air speed distribution at the three heights for each paper is shown in Figure~\ref{fig:boxenplot_wind_speed_three_heights}.
+% todo improvements:
+% - it is Bauman, F. not Fred Baumann. The year is missing. 
+% -The year is missing for the paper of Benton and Brager. 
+% - I would remove the outlier from the graph, the circles. 
+% - I would make the figure a bit bigger vertically, it is too compact. 
+% -capitalize the Head, Chest, Ankle and Relative
+% -order the paper based on their contribution to the dataset
+% - call it air speed, not wind speed. 
 \begin{figure}[htb!]
@@
     \includegraphics[width=\textwidth]{figures/boxenplot_wind_speed_three_heights}
-    \caption{Wind speeds distribution at three heights.}
+    \caption{Air speeds distribution at three heights.}
     \label{fig:boxenplot_wind_speed_three_heights}
@@
 The distribution of air speed at the three heights varies between the papers, hence, it is difficult to make generalizations.
-However, it should be noted that in most of the studies the air speed measured at the chest height was a good representation of the air speed at the other two heights.
-This assumption is not valid for Brenton et al.\ (1994) however this study contributed only \qty{1}{\percent} of the dataset.
+% todo What do you mean by this? Is the average value close to the chest value? If so, it would be nice to quantify, at least for the overall average of them. 
+However, it should be noted that in most of the studies, the air speed measured at the chest height was a good representation of the air speed at the other two heights.
+This assumption is not valid for Benton et al.\ (1994); however, this study contributed only \qty{1}{\percent} of the dataset.
 The adjustment of \ac{v} based on the \ac{met} significantly increased the base value of the air speed measured by all the studies.
-With the value of \ac{v} being \var{delta_v_mean} higher than the average of the other wind speeds combined.
-The discussion on weather this increase is justified is beyond the scope of this paper since both Standards require the user to adjust the air speed based on the \ac{met} value.
+With the value of \ac{v} being \var{delta_v_mean} higher than the average of the other air speeds combined.
+The discussion on whether this increase is justified is beyond the scope of this paper, since both standards require the user to adjust the air speed based on the \ac{met} value.
 
-In this section we are not showing the F1-score for the \ac{pmv} and \ac{pmv-ce} models for the different subsets of data since we have already presented the results in Table~\ref{tab:f1}.
-We would like to remind the reader that the F1-macro score for \ac{pmv-ce}~=~\num{.16} was lower than the \ac{pmv}~=~\num{.17} for this subset of data, i.e, \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and \ac{v} measured at three heights.
-As previously discussed this is an unexpected result since the \ac{pmv-ce} should be significantly more accurate than the \ac{pmv} model for this subset of data to justify its use.
-In this section we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since we have a very limited number of measurements for this subset of data moreover this analysis has several limitations as previously discussed.
+In this section, we are not showing the F1-score for the \ac{pmv} and \ac{pmv-ce} models for the different subsets of data, since we have already presented the results in Table~\ref{tab:f1}.
+The F1-macro score for \ac{pmv-ce}~=~\num{.16} was lower than the \ac{pmv}~=~\num{.17} for this subset of data, i.e, \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and \ac{v} measured at three heights.
+As previously discussed, this is an unexpected result, since the \ac{pmv-ce} should be more accurate than the \ac{pmv} model for this subset of data to justify its use.
+In this Appendix, we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since we have too little data for this analysis.
 
@@
     \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv_three_heights}
-    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data for the dataset with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and measured at three heights.
+    \caption{Comparison between the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights. 
+    The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data for the dataset with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and measured at three heights.
     Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
-    The brown dashed line represents the identity line where the slope is 1 and the intercept is 0.
-    Comparison between the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights.}
+    The brown dashed line represents the identity line, where the slope is 1 and the intercept is 0.}
     \label{fig:bubble_models_vs_tsv_three_heights}
@@
 
-Finally, we are showing calculated the overall bias of the \ac{pmv} and \ac{pmv-ce} models.
-The results are shown in Figure~\ref{fig:hist_discrepancies_three_heights}.
+The overall bias of the \ac{pmv} and \ac{pmv-ce} models is shown in Figure~\ref{fig:hist_discrepancies_three_heights}.
 \begin{figure}[htb!]
@@
 The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.38}.
-The inter-quartile range for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
+The interquartile range for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
 These results allow us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
-Consequently, we recommend using the \ac{pmv} model instead of the \ac{pmv-ce}.
-
-%\section{Comparison between the PMV formulations with the \ac{pmvg}}\label{sec:comparison-between-the-pmv-formulations-with-the-two-node-model}
-%We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the \ac{pmvg}.
-%% SS provide more information why this was done
-%% Stefano Schiavon: Please provide more details of how these inputs were generated, within which ranges, etc.
-%Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
-%The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
-%The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
-%% Stefano Schiavon: Beside the visualization, can you discuss here also the quantitative results to support the conclusion?
-%The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
-%% Stefano Schiavon: This is unclear and cannot be deduced from the figure, from where does it come from? How can a PMV model be described by an air speed?
-%These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
-%This is also an unexpected result and we will discuss why this may be in the following paragraph.
-%% SS When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed.
-1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.
-%
-%\begin{figure*}[htb!]
-%    \centering
-%    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
-%    \caption{}
-%    \label{fig:pmv_two_node_comparison}
-%\end{figure*}

### 2024-09-10 Manuscript/src/sections/conclusions.tex
paper: Stefano's comments

@@
 \section{Conclusions}\label{sec:conclusions}
-The present study compares the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} Standard.
-The \ac{pmv-ce} was specifically developed with the aim of more accurately estimating latent and sensible heat losses from the skin to the environment when air movement is used~\cite{arens_moving_2009}.
-We determined the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models by comparing their results to the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
+The present study compares the accuracy of the \ac{pmv} model, as used in the \gls{7730} standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} standard.
+The \ac{pmv-ce} was intentionally developed to more accurately estimate the effect of elevated air movement~\cite{arens_moving_2009}.
+We determined the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models by comparing their results to the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1.
 
+% todo  could you add the values of F1-macro here as well?
 The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
 Both models had an accuracy lower than \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
-This value is significantly lower and worse than random guessing, this means that the models are actively misleading the users.
-When only comparing the results of the two models for entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} -- range in which the two models differ -- the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
+This value is lower than random guessing (\qty{14}{\percent}), this means that the models are misleading the users.
+When only comparing the results of the two models for entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
 For example, the median bias of the \ac{pmv} model is \var{bias_median_pmv_0.2} while the median bias for the \ac{pmv-ce} is \var{bias_median_pmv_ce_0.2}.
 
-Our results also show that the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.
+Our results also show that the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable is incorrect and leads to under-counting thermal discomfort.
 In the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
-This finding challenges the above assumption.
 
-\subsection{Implications for the ASHRAE 55 and ISO 7730 Standards}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
-Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion worldwide since both models are widely adopted and incorporated in many building codes and Standards worldwide.
-The \gls{55} Standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that clearly explains the benefit of using it.
+\subsection{Implications for the ASHRAE 55 and ISO 7730 standards}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
+As mentioned previously, researchers, educators and practitioners worldwide may wonder which model they should use.
+The \gls{55} standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that proves it has superior prediction performances.
 The \ac{pmv-ce} is more complex, and computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
-Hence we do not recommend its use.
+Hence, we do not recommend its use.
 
-The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
-Based on our results we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
-In other words, we only recommend its use to determine if an environment may be deemed comfortable by a large group of occupants.
-The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classy if people are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
-Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023} provide corroborating evidence that support our results.
+The very low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits may need to be reconsidered.
+Based on our results, we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
+In other words, we only recommend it for determining whether an environment may be deemed comfortable by a large group of occupants.
+The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classify if individuals are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
+Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988} provide corroborating evidence that supports our results.
 
-We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} Standards.
+We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} standards.
 Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
-Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} Standards when these environmental variables are rarely observed in buildings across the world.
+Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
 In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
-Consequently, we suggest reducing both Standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
+Consequently, we suggest reducing both standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
 \begin{table}[htb!]
@@
     \end{tabular}
-    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} Standards.}
+    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} standards.}
     \label{tab:ranges}
 \end{table}
-\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs that are consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
+\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
 The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
@@
 The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset, and it is not a valid assumption.
-We recommend the use of the thermal preference scale to determine how people perceive their thermal environment, this scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
+We recommend the use of the thermal preference scale to determine how people perceive their thermal environment. 
+This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.

### 2024-09-10 Manuscript/src/sections/introduction.tex
paper: Stefano's comments

@@
 \section{Introduction}\label{sec:introduction}
-\mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
+\mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} standard~\cite{iso7730} in its original form.
 The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
@@
 
-To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
-The best source to see how \ac{pmv-ce} works is the Standard itself, but the Standard does not explain why some choices were implemented.
+To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
+The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some choices were implemented.
 A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
@@
         \draw [arrow] (start) -- (dec1);
-        \draw [arrow] (dec1) -- node[above,pos=0.3] {Yes} (pro1a);
-        \draw [arrow] (dec1) -- node[right,pos=0.3] {No} (dec2);
+        \draw [arrow] (dec1) -- node[above, pos=0.3] {Yes} (pro1a);
+        \draw [arrow] (dec1) -- node[right, pos=0.3] {No} (dec2);
         \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
-        \draw [arrow] (dec2) -- node[above,pos=0.3] {Yes} (pro2a);
-        \draw [arrow] (dec2) -- node[right,pos=0.3] {No} (dec3);
+        \draw [arrow] (dec2) -- node[above, pos=0.3] {Yes} (pro2a);
+        \draw [arrow] (dec2) -- node[right, pos=0.3] {No} (dec3);
         \draw [arrow] (pro2a) -- ($(pro2a.south)+(0,-0.5)$) -| (dec3);
-        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above,pos=0.3] {No} (end1);
-        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above,pos=0.3] {Yes} (pro3);
+        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above, pos=0.3] {No} (end1);
+        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above, pos=0.3] {Yes} (pro3);
         \draw [arrow] (pro3) -- (end2);
     \end{tikzpicture}
-    \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{55} Standard.}
+    \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{55} standard.}
     \label{fig:flowchart_pmv_ce}
@@
 In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-The reader should note that \ac{v} is not the measured value.
-The value of \ac{v} measured by the sensors, should be adjusted based on the \ac{met} of the occupant if \ac{met}~$>$~\qty{1}{met} using this equation \ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$.
+\ac{v} is not the measured value by air speed sensors, according to the standard, the measured value should be adjusted based on the \ac{met} of the occupant if \ac{met}~$>$~\qty{1}{met} using this equation \ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$.
 The value of \ac{ce} is calculated using the \ac{set} equation.
@@
 The resulting values become the new inputs in the \ac{pmv} model.
-Since the \ac{ce} accounts for  convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+Since the \ac{ce} accounts for convective and evaporative heat losses from the person to its environment, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
 The other three input parameters (\ac{clo}, \ac{rh}, \ac{met}) remain unchanged.
 Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
-To show to the reader by how much the output of the two models differ, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clo}~=~\qty{0.5}{clo}.
+To illustrate the extent of the differences between the outputs of the two models, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clo}~=~\qty{0.5}{clo}.
 We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
@@
     The chart shows how the comfort regions change as a function of \ac{v}.
-    We assumed $M$~=~1.2 met, $i_{cl}$~=~0.5 clo.
+    We assumed \ac{met}~=~\qty{1.2}{met}, \ac{clo}~=~\qty{0.5}{clo}.
     \label{fig:comfort_regios_pmv_pmvce}}
 \end{figure}
-The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} the comfort regions estimated using the \ac{pmv-ce} is wider than the one calculated using the \ac{pmv} and shifted towards the warmer temperatures.
+The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} for \ac{rh}~=~\qty{50}{\percent} and shifted towards the warmer temperatures.
+For \ac{v}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.
 
 The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}.
-However, this method has its own limitations.
+However, this method has some limitations.
 First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
+Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set}.
 Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
 Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
+However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective and latent heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
 We could not find a rationale for this choice in the available literature.
 
-Several other formulations of the \ac{pmv} model have been proposed to complicate this issue further.
+Several other formulations of the \ac{pmv} model have been proposed.
 Among the most notable formulations there are the: \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
 \mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
-Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models since they are the most widely used thermal comfort models worldwide and are incorporated in the \gls{7730} and \gls{55} Standards.
+Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models since they are incorporated in the most widely referenced thermal comfort standards worldwide (\gls{7730} and \gls{55}).
 
-Choosing between the \ac{pmv} and \ac{pmv-ce} is consequently now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many building codes and standards worldwide.
-
-%In 2002, Fanger, P. O. and Toftum, J. developed the \ac{epmv} an extension of the \ac{pmv}.
-%The \ac{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
-%The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
-%They, consequently, propose to reduce the metabolic rate used to calculate the \ac{pmv} by \qty{6.7}{\percent} for every scale unit of \ac{pmv} above neutral.
-%The resulting \ac{pmv} value is multiplied by the \ac{ef}.
-%The \ac{ef} for non-air-conditioned buildings can be determined by the length of warm weather throughout the year and whether such buildings may be compared to many others in the region that are air-conditioned.
-%If the temperature is hot all year or most of the year and there are no or few other air-conditioned buildings, \ac{ef} may be \num{0.5}, but if there are many other air-conditioned buildings, \ac{ef} may be \num{0.7}.
-%For non-air-conditioned buildings in locations where the temperature is warm solely during the summer and no or few buildings have air-conditioning \ac{ef} is \numrange{0.7}{0.8}, whereas it is \numrange{0.8}{.9} in areas where there are numerous air-conditioned buildings.
-%In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
-%This arbitrary and non-scientific definition of the expectancy factor makes the \ac{epmv} difficult to calculate and non-generalizable.
-%Consequently, the \ac{epmv} has not gained any widespread adoption.
-%
-%Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
-%They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn used to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
-%The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
-%This approach cannot be, therefore, generalizable and similarly to the \ac{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
-%
+Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
+For example, the WELL certification allows both compliance with \gls{7730} and \gls{55}.
 
 \subsection{Comparison of ISO and ASHRAE PMV formulations}\label{subsec:comparision-of-pmv-formulations}
-To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models, which are used in the \gls{7730} and \gls{55} Standards.
-Some notable works that tried to determine the accuracy of the \ac{pmv} model include the manuscript from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
+To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models, which are used in the \gls{7730} and \gls{55} standards.
+Some notable works that tried to determine the accuracy of the \ac{pmv} model include the work from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
@@
 \mycite{Cheung2019} also concluded that the accuracy of \ac{pmv} was only \qty{34}{\percent}, the model has a slightly higher prediction accuracy for sensation close to neutrality, but the accuracy declined towards either end of the thermal sensation scale, and it overestimated both hot and cold sensations.
-They also found that the Predicted Percentage of Dissatisfied (PPD) failed to predict the percentage of unacceptable votes if the thermal sensation was predicted using the PMV model and suggested its removal from \gls{55} Standards.
+They also found that the Predicted Percentage of Dissatisfied (PPD) failed to predict the percentage of unacceptable votes if the thermal sensation was predicted using the PMV model and suggested its removal from the thermal comfort standard.
 These results were confirmed based on the Chinese Thermal Comfort database~\cite{du_evaluation_2022}.
 \mycite{Yao2022} compared the \ac{pmv} and \ac{pmv-ce} models, however, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
-They focus significantly on how the models perform in different climates and when applied to people from different world regions and their analysis only reports the mean bias of the different models.
+They focus significantly on how the models perform in different climates and when applied to people from different world regions, and their analysis only reports the mean bias of the different models.
 This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate since it does not explain how the model performs over a wide range of environmental, personal conditions, and \ac{tsv}.
@@
 Finally, \mycite{Yao2022} only reported the overall bias for the whole dataset, even though the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
-\mycite{tartarini_prediction_2023} concluded that the \ac{pmv} and \ac{pmv-ce} models accuracy is \qty{35}{\percent} and \qty{36}{\percent} for $\mid$\ac{pmv}$\mid \leq 2$, \qty{43}{\percent} and \qty{44}{\percent} for $\mid$\ac{pmv}$\mid \leq 1$ in predicting the self-reported thermal sensation of people, respectively.
-However, their analysis is limited to providing the precision of the model alone, without reporting the sensitivity (also commonly referred to as recall) and they do not provide the bias of the two models.
 
-\subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
-In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards, respectively.
-The paper objective is to quantitatively answer the following questions:
-which model should you use?
-Which model is more accurate?
-We used the self-reported \ac{tsv} recorded in the \acf{db2}.
-We aim to determine which \ac{pmv} model is most accurate and which are its applicability limits.
-This will allow \ac{pmv} users - engineers, architects, and educators - to make an informed decision on which model they should use and will help policymakers to select the most appropriate model to be used in Standards.
+\subsection{Objectives}\label{subsec:aim-and-objectives}
+In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} standards, respectively.
+We used the \ac{tsv} recorded in the \acf{db2}.
+We aim to determine which \ac{pmv} model is most accurate and the models' applicability limits.

### 2024-09-10 Manuscript/src/sections/methodology.tex
paper: Stefano's comments

@@
 The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
- We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the self-reported \ac{tsv}.
+ We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the \ac{tsv}.
 
@@
 Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
-We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
+We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
 This filtered out approximately \var{entries_db_valid} of the data.
-We decided to keep entries that did not have \ac{tr} and we planned to calculate it using the \ac{tdb} and \ac{to} values using the following equation: $2 t_{o} – t_{db}$.
-If the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
-However, all the entries that had \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv} also had \ac{tr}.
-Consequently, we did not need to make any assumptions about the \ac{tr} values.
 In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the \ac{v} measured by the researchers.
 This is because both \ac{pmv} formulations use the \ac{v} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
-%We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
-%There is also evidence that in many conditions is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
 
-Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points that did not meet the inclusion criteria of both Standards.
+Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points that did not meet the inclusion criteria of both standards.
 The rationale is that the models' accuracy should only be tested within their applicability ranges.
@@
 \num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
-\num{0}~$\leq$~water vapour partial pressure~$\leq$~\qty{2000}{\pascal},
+\num{0}~$\leq$~water vapor partial pressure~$\leq$~\qty{2000}{\pascal},
 and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
-The adjusted total clothing insulation and relative airspeed, as required by both Standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.
+The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.
 
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
+However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
 
-In addition, it should also be noted that while all the data included in the \ac{db2} is published in peer-reviewed papers, not all entries were collected using the same methodology.
-For example, while both thermal comfort Standards recommend measuring \ac{v} at three different heights, this was not always done.
-This could introduce a bias in the measurements and consequently in the results of the models.
+In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same methodology.
+For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
+This could introduce a bias in the measurements and, consequently, in the results of the models.
 Moreover, since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
@@
 \end{enumerate}
-We believe that comparing the results from these three datasets will help us understand the impact of the data quality on the results of the models.
-In particular, by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the Standards.
+Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models, because by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the standards.
 
-We used Python to analyse and visualise the results.
+We used Python to analyze and visualize the results.
 The \ac{pmv} results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
-We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \url{https://github.com/FedericoTartarini/paper-pmv-comparison} so other users can test different assumptions.
+We are committed to reproducible research, hence we have shared the source code and the dataset we used publicly at this URL: \url{https://github.com/FedericoTartarini/paper-pmv-comparison} so other users can test different assumptions.
 
@@
 These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
-However, on their own are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
+However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
 These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{v}).
 If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
-The standard deviation would reflect the combined effect of the people's individual differences as well as any errors in formulation or in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
+The standard deviation would reflect the combined effect of the people's individual differences, as well as any errors in formulation or in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
@@
 We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
-Binning data by the independent variables is a significant improvement over grouping the participants' responses by their \ac{tsv}.
-This is because the \ac{pmv} model aim is, as previously mentioned, to predict the average thermal sensation of a large group of occupants and not to correctly predict each individual \ac{tsv} of people who reported the same \ac{tsv} vote and where in different locations and moments of the year and day.
-
-%In \ref{sec:comparison-between-the-pmv-formulations-with-the-two-node-model} we also compare the results of \ac{pmv} and \ac{pmv-ce} formulations with the \ac{pmvg} calculated by the two-node model.
-%The two-node model was developed by \mycite{GaggeSET} and it uses a more accurate set of equations to solve the heat balance equation.
-%We compared the \ac{pmv} and \ac{pmv-ce} results to the \ac{pmvg} since Gagge's two-node model is used by the \ac{pmv-ce} to calculate the \ac{ce}.
-%Hence, in principle, the results of the \ac{pmv-ce} model should closely match the \ac{pmvg} values.
-%We did this by passing as input to the \ac{pmv}, \ac{pmv-ce}, and two-node models the same set of \num{10000} randomly generated set of inputs.
-%Each set of inputs was sampled assuming a random uniform distribution within a predefined range.
-%We selected the range of each of the 6 input variables needed to calculate the \ac{pmv} to be equal to the 2.5th and 97.5th percentiles for all the input variables contained in the \ac{db2}.
-%The only exception was for the range of \ac{v} that was randomly selected from this interval \qty{0.1}{\m\per\s}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s}.
-%The rationale for this decision was that both the \ac{pmv} and \ac{pmv-ce} models use the same equations for \ac{v}~$\leq$~\qty{0.1}{\m\per\s}.
-%
-%Interestingly, \mycite{GaggeSET} also describe how to calculate the \ac{pmvs} which uses the \ac{set} temperature to calculate the \ac{pmv} value in place of \ac{tdb} and \ac{tr}.
-%This would solve some of the uncertainties that the \ac{pmv-ce} model introduces as detailed above.
-%However, it is not possible for us to comment on why the \ac{pmv-ce} model was included in the \gls{55} instead of the \ac{pmvs} since no peer-reviewed paper or official document describes the rationale for this decision.
+Binning data by the independent variables is a better representation over grouping the participants' responses by their \ac{tsv} when the aim is to predict the average thermal sensation of a large group of occupants and not to correctly predict each individual \ac{tsv} of people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.
 
@@
 The F1 score is the harmonic mean of the precision and sensitivity, where an F1 score reaches its best value at 1 and the worst score at 0.
-Precision (also called positive predictive value) is the ratio of true positive overall positive results obtained from the test, while sensitivity (also known as recall or true positive rate) is the ratio of true positive over all the actual positive cases.
+Precision (also called positive predictive value) is the ratio of true positives over all the positive results obtained from the test. 
+Sensitivity (also known as recall or true positive rate) is the ratio of true positive results over all the actual positive cases.
 We report three F1 scores in the paper.
 The F1-micro score calculates metrics globally by counting the total true positives, false negatives, and false positives.
-The F1-macro calculate metrics for each label and find their unweighted mean.
+The F1-macro calculates metrics for each label and finds their unweighted mean.
 The F1-weighted calculate metrics for each label and find their average weighted by support (the number of true instances for each label).
-
-% Ed's comments below
-% The primary sticking point is with the input data.  The following issue seems fundamental to me, something you would have to address in a published paper about this.
-
-% Air speed measurements in the ASHRAE Field Dataset rarely conform to ASHRAE or ISO standards.
-Only the studies done with instrumented carts with anemometer arrays are likely to have measured at three heights as required.
-We know that very few of the studies have done this.
-
-% If you only measure at one height instead of three, you are very likely to record higher values than the averaged velocities that are required for the standards’ comfort models.
-This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level.
-The measurement will therefore most often be an overestimate for inputting into the model, and the associated sensation prediction will be cooler than it should be.
-
-% So if you have a systematic bias toward higher airspeeds in the statistical data you are using, a comfort model that inherently predicts stronger rates of air speed cooling (CE) will overpredict more than a model that inherently predicts weaker rates (PMV).
-The weaker one will appear to be predicting correctly, or in any event something closer to correct.
-The stronger-predicting model could appear ‘biased’ when compared to the weaker-predicting model.
-
-% This is one of the points of Pearl’s causal inference book—in assigning bias, you have to find the source of the bias, which in this case may be in the dataset, and not in either of the models being evaluated.
-
-% The DB2 dataset is very useful for temperature, but it is really not suitable for detailed evaluation of air speed.
-Air speed is far more variable and harder to measure accurately than temperature.
-I suppose someone could wade through all the studies and observations in the database and observe the types of conditions under which the elevated air speeds were recorded.
-Were they consistent over time and geographic location?
-Did they average over sufficient time as specified in standards?
-Any sense of the accuracy of the anemometers used to measure them?
-What kind of buildings were they in—or any indications about the thermal sensitivity of the occupancies being surveyed?
-You can see that all these questions suggest possible sources of bias, all which would be need to identified and assessed, to judge whether information from such surveys is suitable for evaluating a systematic measurement system like a comfort model.
-It seems a really big task with uncertain benefits.
-
-% Models vs laboratory study data:
-
-% In our 2015 Corrective Power paper, Hui assembled the results from many laboratory studies of cooling devices, including fans.
-In this paper you can see the studies of Tanabe where his subjects were exposed to a uniform plug flow, where the air movement measurement represents the whole body as required by the standard.
-You can compare his results to the CE model and find that the CE model slightly underpredicts the subjective cooling votes he measured.
-
-% If you then look at the cooling found by Yongchao Zhai where the air movement was directed only at the upper body and measured there, the CE model slightly overpredicts his measured results.
-This is of course appropriate.
-It is seen in other recent studies of local fan cooling.
-
-% Bigger view of this:
-
-% I think the comparison of models must be done using data from controlled experiments in the laboratory.
-A lot of this was well-examined 50 years ago.
-It hasn’t been followed up as intensely since then because it wasn’t considered that important by the profession.
-What’s different now is that 1) people are recognizing climate warming, and 2) it’s a lot easier to do comfort and physiological modeling and learn from it, using the Web.
-That’s why there is now an opportunity that people can design better buildings using more sophisticated comfort and climate information.
+For the interpretation of our results, F1-macro is arguably the most important because it is not affected by the unbalanced data distribution (most of the data are for thermal neutrality) and gives equal importance to predicting each of the thermal sensation votes.
+F1-micro is the least important because it is heavenly skewed by the unbalance of the classes.

### 2024-09-10 Manuscript/src/sections/results.tex
paper: Stefano's comments

@@
 \section{Results and Discussion}\label{sec:results}
-
-\subsection{\ac{db2} overview}\label{subsec:comfort-db-overview}
 While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
@@
 \end{figure*}
-The input data distributions are far from uniformity (equal density) across the Standards' applicability ranges.
+The input data distributions are far from uniformity (equal density) across the standards' applicability ranges.
 We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
-For example, only \qty{2.5}{\percent} of the available dataset had a \ac{tdb}~$\leq$ than \var{ta_95_perc_min}, this is a problem since both thermal comfort Standards' applicability limits extend to \qty{10}{\celsius}.
+For example, only \qty{2.5}{\percent} of the available dataset had a \ac{tdb}~$\leq$ than \var{ta_95_perc_min}, this is a problem since both thermal comfort standards' applicability limits extend to \qty{10}{\celsius}.
 This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
@@
 \end{figure*}
-Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
+Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex, and are almost equally distributed among males (\qty{52}{\percent}) and females.
 Ages are not normally distributed, the same is true for the running mean outdoor temperature.
@@
 The \ac{db2} does not contain information about the health status of the participants.
-Approximately \var{running_mean_10_25} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
+Approximately \var{running_mean_10_25} of the running mean outdoor temperature values are between \qtyrange{10}{25}{\celsius}.
 
@@
     \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
-    \caption{The left Figure shows the percentage of \ac{tpv} for each thermal sensation vote.
+    \caption{The graph on the left shows the percentage of \ac{tpv} for each thermal sensation vote.
     The numbers on the right side of each bar show the number of points for each \ac{tsv}.
-    The right figure shows the total number of data points grouped by \ac{tsv}.
-    Each bin in the right figure has more data points than in the left figure since less data on \ac{tpv} were available.}
+    The graph on the right shows the total number of data points grouped by \ac{tsv}.
+    Each bin in the right figure has more data points than in the left figure, since less data on \ac{tpv} were available.}
     \label{fig:bar_plot_tp_by_ts}
@@
 
-In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
+In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable~\cite{Fanger1970, schweiker2020evaluating}.
 However, in the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
 This finding challenges the above assumption.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under counting thermal discomfort in our dataset.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort.
 On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
@@
 This result highlights one of the major limitations of the \ac{tsv} scale and of the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable.
-This suggests that the thermal preference scale is better suited to determine how people perceive their thermal environment.
+This shows that the thermal preference scale is better suited to determine how people perceive their thermal environment.
 The \ac{tpv} more clearly allows an HVAC system to decide whether to change or not the thermal environment.
@@
 we also present the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s}.
-In this subset of data the differences in the outputs of the two models are more pronounced.
+In this subset of data, the differences in the outputs of the two models are more pronounced.
 We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
@@
 In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
-We do not present these results here since they are similar to the results presented in this section and lead to the same conclusions.
+We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.
 
 Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
-\begin{figure*}[htb!]
+\begin{figure}[htb!]
     \centering
@@
         \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0}
-        \caption{All data}
     \end{subfigure}
+\par\bigskip % force a bit of vertical whitespace
     \begin{subfigure}[b]{\textwidth}
@@
         \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy_0.2}
-        \caption{Only data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}}
     \end{subfigure}
     \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants.
-    The number of data points in each bin of the two charts is the same and this allows us to compare the accuracy of both models.
-    The overall prediction accuracy is shown above the chart while the model accuracy of predicting participants grouped by their \ac{tsv} is overlayed over each bar.
-    A random model, randomly selecting a thermal sensation vote would have an overall accuracy of \qty{14}{\percent}.}
+    The number of data points in each bin of the two charts is the same, and this allows us to compare the accuracy of both models.
+    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.
+    A random model, randomly selecting a thermal sensation vote, would have an overall accuracy of \qty{14}{\percent}.}
     \label{fig:bar_stacked_model_accuracy}
-\end{figure*}
+\end{figure}
 Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
-The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively when using the whole dataset.
+The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
 The overall accuracy of both \ac{pmv} models increased to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
@@
 However, the \ac{pmv} model outperformed the \ac{pmv-ce} model when considering \ac{tsv}~=~1.
-We want to remind the reader that the number of data points in each \ac{tsv} bin is different.
+The number of data points in each \ac{tsv} bin is different.
 
-Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
+Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy, since a default strategy of guessing the majority class would lead to a high overall model accuracy.
 A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
 
-In Table~\ref{tab:f1}, we report the F1-scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping datapoints that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+In Table~\ref{tab:f1}, we report the F1-scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
 \begin{table}[htb!]
@@
 We then analyzed the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{v} measured at three heights.
-Both F1-macro scores decreased and the \ac{pmv-ce} one \num{0.16} is lower than the value for the \ac{pmv}~=~\num{0.17}.
-This is an unexpected result since the \ac{pmv-ce}, to justify its use, should significantly outperform the \ac{pmv} model when analyzing data that had \ac{v} measured at three heights and arguably a lower bias in the \ac{v} value.
+Both F1-macro scores decreased and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
+This is an unexpected result since the \ac{pmv-ce} should outperform the \ac{pmv} model when \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and even more when \ac{v} were measured at three heights. 
+The justification of the introduction of the \ac{pmv-ce} was that it should better take into account the effect of elevated air speed.
 All the F1 scores significantly increased when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
@@
 We only binned the data to aid the visualization of a large dataset.
-If the model is to accurately predict the thermal sensation of people, the regression line (black continuous line) should pass through the origin of the cartesian plane and have a slope of 1, also called identity line which is represented by the brown dashed line.
+For the model to accurately predict people’s thermal sensation, the regression line (black continuous line) should pass through the origin of the Cartesian plane and have a slope of 1. 
+This ideal prediction is referred to as the identity line, shown as the brown dashed line in the figure.
 The intercepts for the \ac{pmv} is \qty{-0.22} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
@@
 The median bias for the \ac{pmv-ce} also decreased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it was still higher than the one for the \ac{pmv}.
-These results, together with those presented in Table~\ref{tab:f1}, disprove the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v} than the \ac{pmv}.
-The \ac{pmv-ce} has a higher bias while making the code more complex and computationally intensive to run.
 
@@
     The red dashed lines show the acceptable range of bias.
-    Please note that each bin has a different number of data points.
+    Each bin has a different number of data points.
     }
@@
 \end{figure*}
-%We did not include \ac{tr} in the analysis since \ac{tr} was not measured in several studies, hence, we estimated it as detailed in the Methodology.
-%Furthermore, in many buildings is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.
 In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
@@
 On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 and was significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
-These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.
+
+These results, together with those presented in Table~\ref{tab:f1} and in the previous section, prove that the \ac{pmv-ce} model has lower accuracy and higher bias than the \ac{pmv} at elevated air speed while making the model more complex and computationally intensive to run.
+
 Both model have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
@@
 \subsection{Sources of Error}\label{subsec:sources-of-error}
-In this Section, we identify some of the main reasons that could explain the difference in the results from the two \ac{pmv} models when compared against the \ac{tsv} available in the \ac{db2}.
+In this section, we try to identify some of the main reasons that could explain the difference in the results from the two \ac{pmv} models when compared against the \ac{tsv} available in the \ac{db2}.
 Moreover, we tried to explain why we observed that the \ac{pmv-ce} model has a lower accuracy than the \ac{pmv} in predicting thermal sensation.
@@
 \paragraph{Heat Balance Equation and Calculation of the PMV value}
-The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
+The \ac{pmv} model uses simplified heat balance equations to estimate the heat losses and gains from the human body to its surrounding environment.
 This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
-This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
-Moreover, the \ac{pmv} model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
-This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature~\cite{romanovsky_thermoregulation_2018}.
-Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
+This is a significant simplification, as it overlooks several important factors, such as the uneven coverage of clothing across different parts of the body, the non-uniformity of metabolic heat generation, and the variations of mass-to-surface ratios.
+It also ignores vasoconstriction and vasodilation.
+Additionally, the \ac{pmv} model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
+In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature~\cite{romanovsky_thermoregulation_2018}.
+Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer; however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
 The \ac{pmv} model also assumes that the human body is constantly losing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
@@
 Heat losses or gains are only a proxy for the thermoregulation effort that the human body has to undergo to maintain a constant internal temperature.
-Fanger created this equation based on a limited dataset collected in a laboratory and this equation was never updated or revisited~\cite{Fanger1970}.
-In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions, respectively~\cite{romanovsky_thermoregulation_2018}.
+Fanger created this equation based on a limited dataset collected in a laboratory, and this equation was never updated or revisited~\cite{Fanger1970}.
+In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions~\cite{romanovsky_thermoregulation_2018}.
 
@@
 On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the thermal neutrality zone.
-The models' inability to account for these regulatory mechanisms is a fundamental limitation.
+The model's inability to account for these regulatory mechanisms is a fundamental limitation.
 
@@
     \item the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
-    For example, while the \ac{pmv} model may assume that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
-    Participants in commercial buildings may, on the contrary, deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.
+    For example, while the \ac{pmv} model may assume that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge~\cite{GaggeSET}.
+    On the contrary, participants indoors may deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold, which is only a fraction of $w_{max}$.
 \end{enumerate}
@@
 A combination of all the above-mentioned factors likely plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
-The latter hypothesis -- 4) anchoring -- has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
-Supporting evidence to this issue is provided by \mycite{schweiker2020evaluating} who showed that the distances between the anchors are not constant and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
+The latter hypothesis -- 4) anchoring -- has gained significant momentum in the literature, which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
+Supporting evidence was provided by \mycite{schweiker2020evaluating}, who showed that the distances between the anchors are not constant and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
 
 \paragraph{Measurement Errors}
-All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts, however, measurement errors are inevitably present in the dataset.
-This is because all measurement instrumentation is not always calibrated before use, the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the proximity of the participant, and estimating clothing and metabolic rates from tables is a complex and prone to error task.
-In an ideal scenario, errors would be random and would cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
-However, this may not always be the case, for example, if a researcher often underestimated the clothing insulation of participants or their activity levels.
+All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts; however, measurement errors are inevitably present in the dataset.
+Instrumentation is not always calibrated and the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the participant's proximity, and estimating clothing and metabolic rates from tables is a complex and prone-to-error task.
+In an ideal scenario, errors would be randomly distributed and cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
+However, this may not always be the case, for example, if a researcher often underestimates participants' clothing insulation or activity levels.
 Moreover, most of the data in the \ac{db2} also only contains entries where the \ac{v} was measured at a single height.
-This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body and this is not always the case.
+This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body, and this is not always the case.
 Measuring the air speed at a single height may lead to record higher values than the averaged velocities that are required for the standards’ comfort models.
@@
 To account for this error, we only included the studies where the \ac{v} was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
-However, these results were not significantly different from those presented in this paper and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.
-It should also be noted that in buildings air speed is rarely measured, and it is never measured at three different heights.
+However, these results were not significantly different from those presented in this paper, and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.
 
 \paragraph{Thermal Sensation Scale}
-One minor but not negligible source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
+One minor, but not negligible, source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
 For example, even if the model was \qty{100}{\percent} accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (\num{2})' or `hot (\num{3})'.
@@
 They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
-An alternative solution would be to transform the \ac{pmv} output to an ordinal and categorical seven-point scale.
+An alternative solution would be to transform the \ac{pmv} output to an ordinal seven-point scale.

### 2024-09-10 Manuscript/src/sections/titleAbstract.tex
paper: Stefano's comments

@@
 
-    \title{Which PMV model should you use? Comparison of the ISO 7730:2005 and ASHRAE 55:2023 PMV models}
-    % alternatives PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?
-    % alternatives ISO 7730:2005 vs ASHRAE 55:2023: A comparison of the accuracy of the two implementations of the PMV model
+    \title{Comparative analysis of PMV Models accuracy implemented in the ISO 7730:2005 and ASHRAE 55:2023}
 
-    \author[label1,label2]{Federico Tartarini\corref{mycorrespondingauthor}}
+    \author[label1]{Federico Tartarini\corref{mycorrespondingauthor}}
     \ead{federico.tartarini@sydney.edu.au}
@@
 
-    \address[label1]{University of Sydney, Sydney, AU}
-    \address[label2]{Berkeley Education Alliance for Research in Singapore, Singapore}
+    \address[label1]{The University of Sydney, Sydney, AU}
     \address[label3]{Center for the Built Environment, University of California, Berkeley, CA, USA}
@@
     \begin{abstract}
-%        The \ac{pmv} model was developed by Fanger more than 50 years ago to predict the mean thermal sensation of a group of people.
-%        It is based on a heat balance model of the human body.
-
-%        Being relatively simple to calculate and included in both the \gls{7730} and \gls{55} Standards, the \ac{pmv} has seen widespread adoption and has been used by professionals worldwide to design and operate buildings and by researchers to predict thermal sensation.
         The \ac{pmv} model, developed by Fanger, is still currently used in its original form in the \gls{7730}.
         It has a low prediction accuracy, especially outside thermal neutrality.
-        Several other formulations have been proposed to compensate for its limitations, one is the \ac{pmv-ce}.
-        The \gls{7730} and \gls{55} are the most widely used thermal comfort Standards worldwide, and their different \ac{pmv} formulations are a source of confusion.
-%        for many users who often struggle to determine which model is most appropriate for their needs.
+        Several other formulations have been proposed to compensate for its limitations. 
+        One is the \ac{pmv-ce}.
+        The \gls{7730} and \gls{55} are the most widely referenced thermal comfort standards worldwide, and their different \ac{pmv} formulations are a source of confusion.
         Which model is more accurate?
-        Which model should you use?
 
         We compared the accuracy of the two \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce}, since the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, despite its increased complexity.
-%        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\sec}.
+        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce}, since the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, even in the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\sec}.
         We suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} model cannot determine the degree to which people are dissatisfied.
-%        This recommendation already aligns with the \gls{55} Standard's definition which states that the \ac{pmv} model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce \textit{satisfactory thermal environmental conditions} for a majority of the occupants within the space''.
         A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
-        Finally, the classic assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable leads to under-counting thermal discomfort in our dataset.
-        We, hence, recommend in thermal comfort research the use of the thermal preference scale instead of the thermal sensation scale.
+        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally comfortable leads to under-counting thermal discomfort, therefore, we recommend using the thermal preference scale instead of the thermal sensation scale.
     \end{abstract}

### 2024-09-23 Manuscript/src/sections/titleAbstract.tex
feat(abstract): reviewed text

@@
         Several other formulations have been proposed to compensate for its limitations. 
-        One is the \ac{pmv-ce}.
+        One is the \ac{pmv-ce} standard.
         The \gls{7730} and \gls{55} are the most widely referenced thermal comfort standards worldwide, and their different \ac{pmv} formulations are a source of confusion.
@@
         We compared the accuracy of the two \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce}, since the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, even in the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\sec}.
+        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce}, since the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, even in the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
         We suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} model cannot determine the degree to which people are dissatisfied.
         A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
-        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally comfortable leads to under-counting thermal discomfort, therefore, we recommend using the thermal preference scale instead of the thermal sensation scale.
+        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort, therefore, we recommend using the thermal preference scale instead of the thermal sensation scale.
     \end{abstract}
@@
     \begin{keyword}
-        Thermal comfort model \sep ISO 7730 \sep ASHRAE 55 \sep Thermal sensation \sep Thermal preference
+        Thermal comfort model \sep Predicted Mean Vote \sep Thermal comfort \sep Thermal sensation \sep Thermal preference
     \end{keyword}

### 2024-09-23 test_f1_score.py
feat(introduction): reviewed text

@@ -56,7 +56,7 @@ ax[2].set(ylabel="w", xlim=(20, 40), xlabel="tdb")
 ax[1].set(ylabel="pmv", xlim=(20, 40), xlabel="set")
 ax[3].set(ylabel="pmv", xlim=(20, 40), xlabel="tdb")
 plt.show()
-plt.savefig("./Manuscript/src/figures/PMV and w as a function of SET.png", dpi=300)
+plt.savefig("./Manuscript/src/figures/PMV and w as a function of SET.pdf")
 
 # SMOTE example https://towardsdatascience.com/smote-fdce2f605729
 # F1 score example https://towardsdatascience.com/the-f1-score-bec2bbc38aa6

### 2024-09-23 Manuscript/src/sections/results.tex
feat(results): reviewed text

@@
 \section{Results and Discussion}\label{sec:results}
-While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
+While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries meet the inclusion criteria listed in the Methodology section.
 The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
@@
 The input data distributions are far from uniformity (equal density) across the standards' applicability ranges.
-We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
+We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that are lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
 For example, only \qty{2.5}{\percent} of the available dataset had a \ac{tdb}~$\leq$ than \var{ta_95_perc_min}, this is a problem since both thermal comfort standards' applicability limits extend to \qty{10}{\celsius}.
@@
 The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}.
-The median value in the \ac{db2} was \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
+The median value in the \ac{db2} is \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
 We also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
@@
     \includegraphics[width=\textwidth]{figures/dist_other_data}
-    \caption{Distribution of age, height, weight, and running mean outdoor temperature.
-    The data are grouped by sex.
-    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.}
+    \caption{Distribution of age, height, weight, and running mean outdoor temperature.}
     \label{fig:dist_other_data}
@@
 \end{figure*}
-Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
+Only \var{entries_with_tp} entries have information about both \ac{tsv} and \ac{tpv}.
 Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
@@
 In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable~\cite{Fanger1970, schweiker2020evaluating}.
-However, in the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
+However, \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
 This finding challenges the above assumption.
@@
 On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
-Of those who reported being `cool' \qty{23}{\percent} of them wanted `no change' and \qty{5}{\percent} wanted to be `cooler'.
+Of those who reported being `cool', \qty{23}{\percent} of them wanted `no change' and \qty{5}{\percent} wanted to be `cooler'.
 Similarly, but in reverse, among those who reported being `warm' \qty{13}{\percent} wanted `no change' and \qty{2}{\percent} wanted to be `warmer'.
-This result highlights one of the major limitations of the \ac{tsv} scale and of the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable.
-This shows that the thermal preference scale is better suited to determine how people perceive their thermal environment.
-The \ac{tpv} more clearly allows an HVAC system to decide whether to change or not the thermal environment.
+This result highlights one of the major limitations of the \ac{tsv} scale.
+The thermal preference scale is better suited to determine how people perceive their thermal environment and more clearly provides a signal to decide whether to change or not the thermal environment.
 On the contrary, the \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
 
-\subsection{Comparison of PMV accuracy in predicting thermal sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
-The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are comparing the predicted value from the two models with measured \ac{tsv}.
-However, since the two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s};
-we also present the results for a subset of the data where \ac{v} is higher than \qty{0.2}{\m\per\s}.
+\subsection{Comparison of PMV Accuracy in Predicting Thermal Sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
+The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we compared the predicted value from the two models with measured \ac{tsv}.
+However, since the two models only differ when \ac{v}~$>$~\qty{0.1}{\m\per\s};
+we also present the results for a subset of the data where \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
 In this subset of data, the differences in the outputs of the two models are more pronounced.
 We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
-We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
-The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
-The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
-In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
-We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.
 
@@
     \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants.
-    The number of data points in each bin of the two charts is the same, and this allows us to compare the accuracy of both models.
-    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.
-    A random model, randomly selecting a thermal sensation vote, would have an overall accuracy of \qty{14}{\percent}.}
+    The number of data points in each bin of the two charts at the top (a and b) and at the bottom (c and d) is the same, and this allows us to compare the accuracy of both models.
+    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.}
     \label{fig:bar_stacked_model_accuracy}
 \end{figure}
-Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
-The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
-The overall accuracy of both \ac{pmv} models increased to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-Interestingly, the \ac{pmv-ce} model outperformed the \ac{pmv} model for \ac{tsv}~=~0 and -1.
-However, the \ac{pmv} model outperformed the \ac{pmv-ce} model when considering \ac{tsv}~=~1.
+Both \ac{pmv} formulations underestimate the conditions at which participants are thermally dissatisfied with their environment.
+The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
+The overall accuracy of both \ac{pmv} models increass to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.
+However, the \ac{pmv} model outperforms the \ac{pmv-ce} model when considering \ac{tsv}~=~1.
 The number of data points in each \ac{tsv} bin is different.
+\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
+They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
 
 Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy, since a default strategy of guessing the majority class would lead to a high overall model accuracy.
-A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
+A model that always predicts `neutral' would achieve an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}~=~0.
 
-In Table~\ref{tab:f1}, we report the F1-scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+In Table~\ref{tab:f1}, we report the F1 scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and iv) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
 \begin{table}[htb!]
     \centering
-    \input{tables/f1.tex}
-    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
+    \input{tables/f1}
+    \caption{F1 scores for the \ac{pmv} and \ac{pmv-ce} models for different subsets of data.}
     \label{tab:f1}
@@
 We then analyzed the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{v} measured at three heights.
-Both F1-macro scores decreased and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
-This is an unexpected result since the \ac{pmv-ce} should outperform the \ac{pmv} model when \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and even more when \ac{v} were measured at three heights. 
-The justification of the introduction of the \ac{pmv-ce} was that it should better take into account the effect of elevated air speed.
-All the F1 scores significantly increased when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
-This highlights that the \ac{pmv} models are more accurate in predicting the thermal sensation of people when the thermal sensation is close to neutral.
+Both F1-macro scores decrease and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
+Our results show that the assumption that the \ac{pmv-ce} better takes into account the effect of elevated air speed is false.
+All the F1 scores significantly increase when analyzing the dataset with $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+This highlights that the \ac{pmv} models are only accurate in predicting the thermal sensation of people when the thermal sensation is close to neutral.
 
-Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
+\subsection{Model Bias}\label{subsec:model-bias}
+The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
+The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
+The results presented so far are primarily comparing individual \ac{tsv} to \ac{pmv}.
+This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
+Moreover, grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
 To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
@@
 This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{54}{\percent} and \qty{58}{\percent} of the times, respectively.
-Both \ac{pmv} formulations always under-predicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
+Both \ac{pmv} formulations always under-predict the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
 the slope of the curve was significantly lower than 1.
-We discuss in detail in sub-Section~\ref{subsec:sources-of-error} which factors may be the underlying cause of these errors.
-
-\subsection{Model Bias}\label{subsec:model-bias}
-The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
-The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
-Consequently, while the above-mentioned analysis is needed to compare for the same sample size in each thermal sensation group the \ac{pmv} and \ac{pmv-ce} accuracies, it focuses primarily on comparing individual \ac{tsv} to \ac{pmv}.
-This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
-\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
-They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
 
@@
 Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
-The results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+The results highlight that the \ac{pmv-ce} has a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
 We report the median and inter-quantile ranges since data are not normally distributed.
@@
     \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-    The x-axis shows the discrepancy between the predicted and self-reported thermal sensation vote.
+    The x-axis shows the discrepancy (bias) between the predicted and self-reported thermal sensation vote.
     We have binned the data into intervals of equal width.
@@
 \end{figure*}
-When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
-The median bias for the \ac{pmv-ce} also decreased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it was still higher than the one for the \ac{pmv}.
+When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
+The median bias for the \ac{pmv-ce} also decreases to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it is higher than the one for the \ac{pmv}.
 
 \subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
-After presenting the overall bias of the two \ac{pmv} formulations, we analyzed the bias as a function of the five input variables \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo} and we are presenting the results in Figure~\ref{fig:bias_models}.
+Here we present the model bias as a function of the five input variables \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo}.
+Results are shown in Figure~\ref{fig:bias_models}.
 \begin{figure*}[htb!]
@@
     The x ticks labels show the bin range.
-    The $]$ means that the right value is included in the range.
+    The square bracket($]$) means that the right value is included in the range.
     The red dashed lines show the acceptable range of bias.
@@
 \end{figure*}
-In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval.
+In each sub-figure, we have binned the results into discrete intervals.
+The x-axis tick label shows the bin range.
 All the boxen-plots have the same area to aid visualization, even though the number of points in each bin is not constant.
-The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all input variables.
-On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 and was significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
+The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
+On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
+Both model have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
 
-These results, together with those presented in Table~\ref{tab:f1} and in the previous section, prove that the \ac{pmv-ce} model has lower accuracy and higher bias than the \ac{pmv} at elevated air speed while making the model more complex and computationally intensive to run.
-
-Both model have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
+These findings, along with those shown in Table~\ref{tab:f1} and earlier figures, demonstrate that the \ac{pmv-ce} model has lower accuracy and higher bias compared to the \ac{pmv} at elevated air speeds, even though the \ac{pmv-ce} model is more complex and requires more computational resources.
 
@@
 The issues are reported in order of importance of what we believe may affect the accuracy of the model.
-More research is needed to determine how to improve the overall accuracy of the \ac{pmv} model.
 
-\paragraph{Heat Balance Equation and Calculation of the PMV value}
+\paragraph{Heat Balance Equation and Calculation of the PMV Value}
 The \ac{pmv} model uses simplified heat balance equations to estimate the heat losses and gains from the human body to its surrounding environment.
@@
 In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature~\cite{romanovsky_thermoregulation_2018}.
-Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer; however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.
+Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer; however, this never precisely occurs.
+The human thermoregulatory system, in healthy humans, is constantly actively engaged to ensure a stable core temperature.
 The \ac{pmv} model also assumes that the human body is constantly losing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
@@
 
-It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral'.
+It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy only in determining when people are `neutral'.
 This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis~\cite{ romanovsky_thermoregulation_2018}.
@@
     \item the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
-    For example, while the \ac{pmv} model may assume that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge~\cite{GaggeSET}.
+    The \ac{pmv} model assumes that a score of 3 (`hot') signifies the change between a compensable set of conditions to an un-compensable one, e.g., skin wettedness reaches its max value ($w_{max}$), as suggested by Gagge~\cite{GaggeSET}.
     On the contrary, participants indoors may deem the environment to be `hot' as soon as their \ac{w} crosses a critical threshold, which is only a fraction of $w_{max}$.
@@
 However, this may not always be the case, for example, if a researcher often underestimates participants' clothing insulation or activity levels.
-Moreover, most of the data in the \ac{db2} also only contains entries where the \ac{v} was measured at a single height.
+Moreover, most of the data in the \ac{db2} also only contains entries where the air speed was measured at a single height.
 This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body, and this is not always the case.
@@
 This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently the associated sensation prediction will be cooler than it should be.
-To account for this error, we only included the studies where the \ac{v} was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
+To account for this error, we only included the studies where the air speed was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
 However, these results were not significantly different from those presented in this paper, and the \ac{pmv-ce} model still had a higher bias than the \ac{pmv} model.
@@
 One minor, but not negligible, source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
-For example, even if the model was \qty{100}{\percent} accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (\num{2})' or `hot (\num{3})'.
-Using a continuous scale, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}.
+For example, even if the model is \qty{100}{\percent} accurate and the estimated \ac{pmv} value is \num{2.5} the person can only report being `warm (\num{2})' or `hot (\num{3})'.
+Using a continuous scale, may in principle be a more robust approach since the distances between the anchors are not constant~\cite{schweiker2019scales, schweiker2020evaluating}.
 They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.

### 2024-09-23 Manuscript/src/sections/conclusions.tex
feat(conclusions): reviewed text

@@
 \section{Conclusions}\label{sec:conclusions}
-The present study compares the accuracy of the \ac{pmv} model, as used in the \gls{7730} standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} standard.
+The present study compares the accuracy of the \ac{pmv} model, as used in the \gls{7730} standard, and its modified version, the \ac{pmv-ce} model, included in the \gls{55} standard.
 The \ac{pmv-ce} was intentionally developed to more accurately estimate the effect of elevated air movement~\cite{arens_moving_2009}.
@@
 
-% todo  could you add the values of F1-macro here as well?
-The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
-Both models had an accuracy lower than \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
-This value is lower than random guessing (\qty{14}{\percent}), this means that the models are misleading the users.
+The overall accuracy and F1-macro score of the \ac{pmv} is \qty{33}{\percent} and \num{.16}, respectively.
+While the overall accuracy and F1-macro score of the \ac{pmv-ce} is \qty{34}{\percent} and \num{.16}, respectively.
+Both models have an accuracy lower or equal than \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
+This value is lower than random guessing (\qty{14}{\percent}).
+This means that the models are misleading the users.
 When only comparing the results of the two models for entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
@@
 
-\subsection{Implications for the ASHRAE 55 and ISO 7730 standards}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
-As mentioned previously, researchers, educators and practitioners worldwide may wonder which model they should use.
+\subsection{Implications for the \gls{55} and \gls{7730} Standards and \ac{pmv} Users}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
+Researchers, educators and practitioners worldwide may wonder which model they should use.
 The \gls{55} standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that proves it has superior prediction performances.
@@
 
-The very low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits may need to be reconsidered.
+The very low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits should be reconsidered.
 Based on our results, we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
-In other words, we only recommend it for determining whether an environment may be deemed comfortable by a large group of occupants.
+In other words, we only recommend using it for determining whether an environment may be deemed thermally neutral by a large group of occupants.
 The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classify if individuals are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
@@
 Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
-In addition, the lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
+The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
 Consequently, we suggest reducing both standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
@@
         \toprule
-        Variable & Proposed range \\
+        Variable & New Proposed range \\
         \midrule
@@
         \ac{met} & \qtyrange{1}{2}{met} \\
+        \ac{pmv} & \numrange{-.5}{.5}{} \\
         \bottomrule
@@
 The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
+We then recommend tha the \gls{55} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
 
 The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
+Assuming that people who are `slightly warm' or `slightly cool' are thermally neutral leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
 We recommend the use of the thermal preference scale to determine how people perceive their thermal environment. 

### 2024-09-26 Manuscript/src/sections/appendix.tex
feat(code): figures PDF, changed fig size, updated chart three heights

@@
 
-\section{Analysis of the dataset with air speed measurements at three heights}\label{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}
+\section{Analysis of the Dataset with Air Speed Measurements at Three Heights}\label{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}
 
-In this section, we analyzed the dataset with air speed measurements at three heights.
-Only a limited subset of papers in the \ac{db2} had air speed measurements at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-A summary of the papers is provided in Table~\ref{tab:three_heights_papers}.
+In this section, we show the results of the analysis of the dataset with air speed measurements at three heights.
+Only \num{3063} entries in the \ac{db2} had air speed measurements at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+A summary of the papers that collected these data is provided in Table~\ref{tab:three_heights_papers}.
 The table also shows the number of measurements for each paper and how much (percentage) each study contributed to the dataset.
@@
     \input{tables/publications_with_three_heights.tex}
-    \caption{The number of entries used in this analysis grouped by the author/publication.}
+    \caption{The number of entries used in this analysis grouped by publication.
+    We did not include papers with less than \num{5} entries.}
     \label{tab:three_heights_papers}
 \end{table}
-The paper by Cena et al.\ (1999), de Dear et al. (1994), and Zhang et al.\ (2013 and 2010) contributed the most to the dataset (\qty{79}{\percent}).
-These \num{3063} entries were used to compare the \ac{pmv} and \ac{pmv-ce} models using the same methodology as in Section~\ref{sec:results}.
+The paper by Cena et al.\ (1999), de Dear et al.\ (1994), and Zhang et al.\ (2013 and 2010) contributed the most to the dataset (\qty{79}{\percent}).
 The air speed distribution at the three heights for each paper is shown in Figure~\ref{fig:boxenplot_wind_speed_three_heights}.
-% todo improvements:
-% - it is Bauman, F. not Fred Baumann. The year is missing. 
-% -The year is missing for the paper of Benton and Brager. 
-% - I would remove the outlier from the graph, the circles. 
-% - I would make the figure a bit bigger vertically, it is too compact. 
-% -capitalize the Head, Chest, Ankle and Relative
-% -order the paper based on their contribution to the dataset
-% - call it air speed, not wind speed. 
 \begin{figure}[htb!]
@@
     \includegraphics[width=\textwidth]{figures/boxenplot_wind_speed_three_heights}
-    \caption{Air speeds distribution at three heights.}
+    \caption{Measured air speeds distribution at three heights and relative air speed adjusted based on the \ac{met} value.
+    We did not include papers with less than \num{5} entries.
+    We are also not showing the outliers.}
     \label{fig:boxenplot_wind_speed_three_heights}
 \end{figure}
-The distribution of air speed at the three heights varies between the papers, hence, it is difficult to make generalizations.
-% todo What do you mean by this? Is the average value close to the chest value? If so, it would be nice to quantify, at least for the overall average of them. 
-However, it should be noted that in most of the studies, the air speed measured at the chest height was a good representation of the air speed at the other two heights.
-This assumption is not valid for Benton et al.\ (1994); however, this study contributed only \qty{1}{\percent} of the dataset.
 The adjustment of \ac{v} based on the \ac{met} significantly increased the base value of the air speed measured by all the studies.
@@
 
-In this section, we are not showing the F1-score for the \ac{pmv} and \ac{pmv-ce} models for the different subsets of data, since we have already presented the results in Table~\ref{tab:f1}.
-The F1-macro score for \ac{pmv-ce}~=~\num{.16} was lower than the \ac{pmv}~=~\num{.17} for this subset of data, i.e, \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and \ac{v} measured at three heights.
-As previously discussed, this is an unexpected result, since the \ac{pmv-ce} should be more accurate than the \ac{pmv} model for this subset of data to justify its use.
-In this Appendix, we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since we have too little data for this analysis.
+The F1-macro score for \ac{pmv-ce}~=~\num{.16} is lower than for the \ac{pmv}~=~\num{.17} as previously shown in Table~\ref{tab:f1}.
+The \ac{pmv-ce} should be more accurate than the \ac{pmv} model for this subset of data to justify its use.
+In this Appendix, we are not showing the accuracy of the \ac{pmv} and \ac{pmv-ce} model as a function of the \ac{tsv} since this method has limitations.
 
@@
 \end{figure*}
-The results show that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
+The \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
 The intercept of the \ac{pmv} model is \num{-.31} while the intercept of the \ac{pmv-ce} model is \num{-.41}.

### 2024-09-26 Manuscript/src/sections/introduction.tex
feat(code): figures PDF, changed fig size, updated chart three heights

@@
 To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
-The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some choices were implemented.
+The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented.
 A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
@@
 In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-\ac{v} is not the measured value by air speed sensors, according to the standard, the measured value should be adjusted based on the \ac{met} of the occupant if \ac{met}~$>$~\qty{1}{met} using this equation \ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$.
+\ac{v} is not the measured value by air speed sensors.
+According to the standard, the measured value should be adjusted using this equation \ac{v} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
 The value of \ac{ce} is calculated using the \ac{set} equation.
@@
     \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
-    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models.
-    The chart shows how the comfort regions change as a function of \ac{v}.
-    We assumed \ac{met}~=~\qty{1.2}{met}, \ac{clo}~=~\qty{0.5}{clo}.
+    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{v}.
+    We set \ac{met}~=~\qty{1.2}{met}, \ac{clo}~=~\qty{0.5}{clo}.
     \label{fig:comfort_regios_pmv_pmvce}}
 \end{figure}
-The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} for \ac{rh}~=~\qty{50}{\percent} and shifted towards the warmer temperatures.
+The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
 For \ac{v}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.
 
+\subsection{\ac{pmv-ce} Limitations}\label{subsec:pmv-ce-limitations}
 The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}.
 However, this method has some limitations.
-First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set}.
-Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
+Firstly, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+Secondly, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set}.
+Thirdly, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
 Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective and latent heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
+However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
 We could not find a rationale for this choice in the available literature.
 
+\subsection{Other \ac{pmv} Formulations}\label{subsec:other-pmv-formulations}
 Several other formulations of the \ac{pmv} model have been proposed.
-Among the most notable formulations there are the: \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
+Among the most notable formulations there are the \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
 \mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
@@
 
-Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
-For example, the WELL certification allows both compliance with \gls{7730} and \gls{55}.
-
-\subsection{Comparison of ISO and ASHRAE PMV formulations}\label{subsec:comparision-of-pmv-formulations}
-To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models, which are used in the \gls{7730} and \gls{55} standards.
+\subsection{Comparison of \ac{pmv} and \ac{pmv-ce}}\label{subsec:comparision-of-pmv-formulations}
+To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models.
 Some notable works that tried to determine the accuracy of the \ac{pmv} model include the work from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
@@
 Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
-They used the ASHRAE Global Thermal Comfort database I and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
+They used the ASHRAE Global Thermal Comfort database~I and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
 They also determined the validity of the \ac{pmv} for predicting comfort votes collected in field studies~\cite{Humphreys2002}.
@@
 They also found that the Predicted Percentage of Dissatisfied (PPD) failed to predict the percentage of unacceptable votes if the thermal sensation was predicted using the PMV model and suggested its removal from the thermal comfort standard.
-These results were confirmed based on the Chinese Thermal Comfort database~\cite{du_evaluation_2022}.
+These results were confirmed by the analysis of the Chinese Thermal Comfort database~\cite{du_evaluation_2022}.
 \mycite{Yao2022} compared the \ac{pmv} and \ac{pmv-ce} models, however, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
-They focus significantly on how the models perform in different climates and when applied to people from different world regions, and their analysis only reports the mean bias of the different models.
+They focused significantly on how the models perform in different climates and when applied to people from different world regions, and their analysis only reports the mean bias of the different models.
 This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate since it does not explain how the model performs over a wide range of environmental, personal conditions, and \ac{tsv}.
@@
 \subsection{Objectives}\label{subsec:aim-and-objectives}
+Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
+For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards.
 In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} standards, respectively.

### 2024-09-26 Manuscript/src/sections/methodology.tex
feat(code): figures PDF, changed fig size, updated chart three heights

@@
 The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
- We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the \ac{tsv}.
 
@@
 This filtered out approximately \var{entries_db_valid} of the data.
-In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the \ac{v} measured by the researchers.
+In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
 This is because both \ac{pmv} formulations use the \ac{v} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
 
-Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points that did not meet the inclusion criteria of both standards.
+Both the \gls{55} and \gls{7730} specify a set of applicability limits.
+We hence filtered out those data points that did not meet the inclusion criteria of both standards.
 The rationale is that the models' accuracy should only be tested within their applicability ranges.
@@
 Moreover, since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
-\begin{enumerate}
-    \item all the data points,
-    \item all the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s},
-    \item only the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
+\begin{enumerate}[ {}1{)} ]
+    \item all the data points with \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv};
+    \item the data points in 1) with \ac{v}~$\geq$~\qty{0.2}{\m\per\s};
+    \item the data points in 1) with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
     These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
 \end{enumerate}
-Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models, because by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the standards.
+Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models.
+By only including those studies that measured air speed at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards.
+We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
+The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
+The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
+In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
+We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.
 
@@
 If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
-The standard deviation would reflect the combined effect of the people's individual differences, as well as any errors in formulation or in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
+The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
-However, we believe that this range is too narrow since \ac{tsv} is measured using an ordinal scale of integers.
+However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers.
 Consequently, we expanded the range to \num{-.5} and \num{0.5}.
@@
 We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
-Binning data by the independent variables is a better representation over grouping the participants' responses by their \ac{tsv} when the aim is to predict the average thermal sensation of a large group of occupants and not to correctly predict each individual \ac{tsv} of people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.
+When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
+Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.
 
 \subsection{Performance Metrics}\label{subsec:performance-metrics}
-Simple accuracy can be a misleading metric to use to quantify the performance of a classification predictive model that involves an unbalanced dataset~\cite{Chawla2005}.
+Simple accuracy is a misleading metric to use to quantify the performance of a classification predictive model that involves an unbalanced dataset~\cite{Chawla2005}.
 We then also report the F1 score when determining the performance of the \ac{pmv} and \ac{pmv-ce} in predicting the \ac{tsv}.

### 2024-09-26 main.py

@@ -8,25 +8,23 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import psychrolib
+import scipy
 import seaborn as sns
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.gridspec import GridSpec
 from numba import jit
-from pythermalcomfort.psychrometrics import p_sat_torr
-from pythermalcomfort.utilities import check_standard_compliance_array
 from pythermalcomfort.models import (
     athb,
     cooling_effect,
 )
+from pythermalcomfort.psychrometrics import p_sat_torr
+from pythermalcomfort.utilities import check_standard_compliance_array
+from scipy import optimize
 from scipy import stats
 from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score
 from sklearn.metrics import f1_score
 from sklearn.metrics import r2_score, mean_absolute_error
-from scipy import optimize
-import scipy
-
-from matplotlib.lines import Line2D
 
 mpl.use("TkAgg")
 
@@ -92,7 +90,7 @@ palette_primary = [
 ]
 
 sns.set_context("paper")
-mpl.rcParams["figure.figsize"] = [8.0, 3.5]
+mpl.rcParams["figure.figsize"] = [7.0, 3.5]
 sns.set_style(
     "whitegrid",
     {
@@ -1467,9 +1465,7 @@ def bar_chart(
     if figletter:
         plt.gcf().text(0.05, 0.95, f"{figletter})", weight="bold")
 
-    plt.savefig(
-        f"./Manuscript/src/figures/bar_plot_{ind}_Vmin_{data.vel_r.min()}.png", dpi=300
-    )
+    plt.savefig(f"./Manuscript/src/figures/bar_plot_{ind}_Vmin_{data.vel_r.min()}.pdf")
 
 
 def scatter_plot_flip_x(data):
@@ -1564,7 +1560,7 @@ def scatter_plot(data, ind="tsv", x_jitter=0):
         axs[ix].grid(axis="x")
 
     plt.tight_layout()
-    plt.savefig("./Manuscript/src/figures/scatter_tsv_pmv.png", dpi=300)
+    plt.savefig("./Manuscript/src/figures/scatter_tsv_pmv.pdf")
 
 
 def plot_error_prediction(data):
@@ -1693,13 +1689,12 @@ def plot_error_prediction(data):
             )
 
     plt.savefig(
[…]

### 2024-10-04 Manuscript/src/sections/conclusions.tex
paper: Stefano's review after meeting with Ed

@@
 While the overall accuracy and F1-macro score of the \ac{pmv-ce} is \qty{34}{\percent} and \num{.16}, respectively.
-Both models have an accuracy lower or equal than \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
+Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm,' `hot,' or `cold.'
 This value is lower than random guessing (\qty{14}{\percent}).
@@
 The \gls{55} standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that proves it has superior prediction performances.
-The \ac{pmv-ce} is more complex, and computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
-Hence, we do not recommend its use.
+The \ac{pmv-ce} is more complex and computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
+Hence, we suggest researchers should calculate the \ac{pmv} values according to the \gls{7730} standards.
 
@@
 The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
-We then recommend tha the \gls{55} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
+We then recommend that the \gls{55} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
 
@@
 This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.
+
+% todo Federico could you add an acknowledgement section? Paul Raftery

### 2024-10-04 Manuscript/src/sections/results.tex
paper: Stefano's review after meeting with Ed

@@
     \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants. 
+    In a) and b) for the entire dataset, while in c and d for the dataset when \ac{v}~$\geq$~\qty{0.2}{\m\per\s}. 
+    In a) and c) are shown the results for the \ac{pmv} model, and in b) and d) for the \ac{pmv-ce} model. 
     The number of data points in each bin of the two charts at the top (a and b) and at the bottom (c and d) is the same, and this allows us to compare the accuracy of both models.
-    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.}
+    The overall prediction accuracy is shown above the chart, while the model accuracy of predicting participants grouped by their \ac{tsv} is shown over each bar.
+    A random model would have an accuracy of \qty{14}{\percent}.}
     \label{fig:bar_stacked_model_accuracy}
@@
 On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
-Both model have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
+Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
 

### 2024-10-04 Manuscript/src/sections/titleAbstract.tex
paper: Stefano's review after meeting with Ed

@@
         We compared the accuracy of the two \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        Based on our results, we recommend using the original \ac{pmv} instead of the \ac{pmv-ce}, since the \ac{pmv-ce} provides no significant improvement over the original \ac{pmv} in predicting thermal sensation, even in the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-        We suggest reducing the applicability limit of the \ac{pmv} within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} model cannot determine the degree to which people are dissatisfied.
+        We found that both models have low and similar prediction accuracy (\qty{34}{\percent}) and the \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+        We suggest reducing the applicability limit of the models within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} models cannot determine the degree to which people are dissatisfied.
+        Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm', `hot', or `cold'. 
+        This value is lower than random guessing (\qty{14}{\percent}), meaning they are misleading the users.
         A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
         Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort, therefore, we recommend using the thermal preference scale instead of the thermal sensation scale.
+        In the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
     \end{abstract}

### 2024-10-04 Manuscript/src/sections/methodology.tex

@@
 \num{0}~$\leq$~water vapor partial pressure~$\leq$~\qty{2000}{\pascal},
-and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
+and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met
+% todo add the number below, Federico, please specify how many datapoint were removed from the used dataset due to this criteria.
+Besically, we want to say how many points are out because they are not in the applicability range of the standard.
+This criteria removed XXX datapoints. 
 The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.
@@
 For the interpretation of our results, F1-macro is arguably the most important because it is not affected by the unbalanced data distribution (most of the data are for thermal neutrality) and gives equal importance to predicting each of the thermal sensation votes.
-F1-micro is the least important because it is heavenly skewed by the unbalance of the classes.
+F1-micro is the least important because it is heavily skewed by the unbalance of the classes.

### 2024-10-04 main.py
feat(code): updated the code analysis three speeds

@@ -1813,9 +1813,13 @@ def plot_distribution_variable():
     )
 
 
-def analyse_studies_with_three_speed_measurements(data_: pd.DataFrame = None):
+def analyse_studies_with_three_speed_measurements(
+    data_: pd.DataFrame = None,
+    air_speed_limit: float = 0.2,
+    temperature_limit: float = 24.0,
+):
     # [['contributor', 'publication', 'region', 'country', 'city']]
-    df_three_heights = data_[data_["vel_r"] >= 0.2].dropna(subset="vel_l")
+    df_three_heights = data_[data_["vel_r"] >= air_speed_limit].dropna(subset="vel_l")
     df_three_heights.replace(
         {
             "https://doi.org/10.1016/j.buildenv.2010.05.024": "Zhang, Y. et al. (2010).",
@@ -1875,7 +1879,7 @@ def analyse_studies_with_three_speed_measurements(data_: pd.DataFrame = None):
         index=False,
     )
 
-    authors_to_show = df_publications['publication'].tolist()
+    authors_to_show = df_publications["publication"].tolist()
 
     df_delta_v = df_three_heights["vel_r"] - df_three_heights[
         ["vel_h", "vel_m", "vel_l"]
@@ -1918,14 +1922,25 @@ def analyse_studies_with_three_speed_measurements(data_: pd.DataFrame = None):
     # )
     # plt.show()
 
-    df_three_heights[df_three_heights["publication"].isin(
-        [
-            'Zhang, Y., Chen, H., Meng, Q. (2013)', 'Zhang, Y. et al. (2010)', 'Tartarini, F., Cooper, P., Fleming, R. (2018)'
-        ]
-    )].to_csv("data_three_heights.csv")
+    df_three_heights[
+        # df_three_heights["publication"].isin(
+        #     [
+        #         "Zhang, Y., Chen, H., Meng, Q. (2013)",
+        #         "Zhang, Y. et al. (2010)",
+        #         "Tartarini, F., Cooper, P., Fleming, R. (2018)",
+        #     ]
+        # )
+        # &
+        # (
+                df_three_heights["ta"] > temperature_limit
+        # )
+    ].to_csv("data_three_heights.csv")
 
     plt.close("all")
-    df_plot = df_three_heights.loc[df_three_heights['publication'].isin(authors_to_show), ["vel_h", "vel_m", "vel_l", "vel_r", "publication"]]
[…]

### 2024-10-04 Manuscript/src/cover-letter.tex
feat(paper): created cover letter

\documentclass[11pt,a4paper,roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
% \usepackage[spanish,es-lcroman]{babel}

% moderncv themes
\moderncvstyle{classic}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{black}                              % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
% \renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using

\renewcommand\labelitemi{•}

% adjust the page margins
\usepackage[scale=0.8]{geometry}
\usepackage{enumitem}
\usepackage[range-phrase=--, range-units=single, per-mode=symbol, range-phrase = ~--~]{siunitx}
\usepackage{mathtools}
\setlist[enumerate]{align=left}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

% todo change text below
% personal data
\name{}{Dr. Federico Tartarini}
% \title{Postdoctoral scholar}                               % optional, remove / comment the line if not wanted
\address{Senior Lecturer, The University of Sydney}{Wilkinson Building, 148 A36, Camperdown}{Sydney, NSW, Australia, 2050} % optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
% \phone[mobile]{000-000-000-000}                   % optional, remove / comment the line if not wanted
% \phone[fixed]{+2~(345)~678~901}                    % optional, remove / comment the line if not wanted
% \phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
% todo change text below
\email{federico.tartarini@sydney.edu.au}                               % optional, remove / comment the line if not wanted
%\homepage{www.johndoe.com}                         % optional, remove / comment the line if not wanted
% \extrainfo{Berkeley Education Alliance for Research in Singapore}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}
%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
% todo change text below
\recipient{Cover letter to the Editor}{of the journal of Building and Environment}
\date{\today}
% todo change text below
\opening{Dear Editors,}

\makelettertitle

% todo change text below
We submit an original research article titled ``Comparative analysis of PMV Models accuracy implemented in the ISO 7730:2005 and ASHRAE 55:2023'' for consideration for publication in Building and Environment.

The study compares the accuracy of the PMV model, as used in the ISO 7730:2005 standard, and its modified version, the PMV$_{CE}$ model, included in the ASHRAE 55:2023 standard.
The PMV$_{CE}$ was intentionally developed to more accurately estimate the effect of elevated air movement.
The two model formulations only differ when the air speed is greater than \qty{0.1}{\m\per\s}.
We determined the predictive accuracy of the PMV and PMV$_{CE}$ models by comparing their results to 49,282 thermal sensation votes available in the ASHRAE Global Thermal Comfort Database II v2.1.

We found that both models have low and similar prediction accuracy ($ \approx$\qty{34}{\percent}) and the PMV$_{CE}$ has higher bias than PMV in predicting thermal sensation in the subset of data with air speed greater than \qty{0.2}{\m\per\s}.
Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm,' `hot,' or `cold.'
This value is lower than random guessing (\qty{14}{\percent}), meaning they are misleading the users.

Based on our results, we recommend restricting the use of the PMV only to values $\lvert \textrm{PMV} \lvert \leq 0.5$ since the PMV models cannot determine the degree to which people are dissatisfied.
Moreover, we recommend using the PMV model over the PMV$_{CE}$ model, as the latter is more complex and computationally intensive to run, while not being more accurate than the original PMV model.
Moreover, we challenge the assumption that individuals who feel `slightly warm' or `slightly cool' are thermally neutral.

We believe our findings will significantly contribute to the ongoing discourse on thermal comfort modeling and have practical applications in building design, engineering, and architecture.
Particularly because the PMV model is used by both the ISO 7730:2005 and ASHRAE 55:2023 standards.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.

All authors have approved the manuscript and agree with its submission to Building and Environment.

We have no conflicts of interest to disclose.

Thank you for considering our manuscript for publication.
Please do not hesitate to contact me if I can be of any further assistance.
Please address all correspondence to Dr. Federico Tartarini at \href{mailto:federico.tartarini@sydney.edu.au}{federico.tartarini@sydney.edu.au}.
We look forward to your positive response.

\bigskip

% todo change text below
Dr. Federico Tartarini, PhD\\
Senior Lecturer\\
The University of Sydney\\

\end{document}

### 2024-10-04 Manuscript/src/sections/introduction.tex
feat(paper): added another limitation PMV-ce

@@
 Secondly, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set}.
-Thirdly, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
+Thirdly, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}
 Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
 However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
+Assuming \ac{rh} to remain constant is incorrect since the \ac{ce} is subtracted from both the \ac{tdb} and \ac{tr} to calculate the \ac{pmv-ce}.
+Consequently, the \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} and \ac{tr} values.
 We could not find a rationale for this choice in the available literature.

### 2024-10-04 Manuscript/src/sections/methodology.tex
feat(paper): added another limitation PMV-ce

--- 2024-09-26 Manuscript/src/sections/methodology.tex
@@
 and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
+These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} datapoints.
+Consequently, the final dataset used for the analysis contains \var{entries_db_used} entries.
 The adjusted total clothing insulation and relative airspeed, as required by both standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.

### 2024-10-04 Manuscript/src/sections/titleAbstract.tex
feat(abstract): highlights and graphical abstract

@@
 
+    \begin{graphicalabstract}
+        \includegraphics[width=\linewidth]{figures/graphical_abstract}
+    \end{graphicalabstract}
+
+    % Research highlights
+    \begin{highlights}
+        \item We compared the accuracy the \ac{pmv} model in the \gls{7730} and \gls{55}.
+        \item Both models have low and similar prediction accuracy (\qty{34}{\percent}).
+        \item The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation.
+        \item We recommend limiting the applicability range of the models to $\mid$\ac{pmv}$\mid \leq 0.5$.
+        \item Individuals who feel slightly warm/cool are not thermally neutral.
+    \end{highlights}
+
     \begin{keyword}

### 2024-10-04 main.py
feat(main): can change the hist figsize

@@ -1292,6 +1292,7 @@ def importing_filtering_processing(load_preprocessed=False):
         ].diff(axis=1)[f"lr_hb_{model}"]
 
     save_var_latex("entries_db_used", df_.shape[0])
+    save_var_latex("entries_db_filtered_by_limit_inputs", df_valid_input.shape[0] - df_.shape[0])
 
     df_.to_pickle(r"./Data/db_analysis.pkl.gz", compression="gzip")
 
@@ -2546,6 +2547,7 @@ def plot_bias_distribution_whole_db(
     data_: pd.DataFrame = None,
     fig_name: str = "hist_discrepancies",
     fig_letters: list[str] = ["a)", "b)", "c)", "d)"],
+    fig_size: tuple[int, int] = (7, 4.5),
 ):
 
     models = models_to_test
@@ -2560,7 +2562,7 @@ def plot_bias_distribution_whole_db(
         sharex=True,
         sharey="row",
         constrained_layout=True,
-        figsize=(7, 4.5),
+        figsize=fig_size,
     )
 
     for row, v in enumerate([0, 0.2]):
@@ -3509,6 +3511,7 @@ if __name__ == "__plot__":
 
     # plot bias distribution
     plot_bias_distribution_whole_db(data_=df.copy())
+    plot_bias_distribution_whole_db(data_=df.copy(), fig_size=(5, 5), fig_name="bias_whole_db_graphical_abstract")
     # plot_bias_distribution_whole_db(hb_models=True)
 
     # # plot bias by building

### 2024-10-04 main.py
feat(main): removed unused code from import data

@@ -122,9 +122,6 @@ applicability_limits = {
     "thermal_sensation": [-3.5, 3.5],
     "pmv": [-3.49999, 3.5],
     "pmv_ce": [-3.49999, 3.5],
-    "pmv_set": [-3.49999, 3.5],
-    "pmv_gagge": [-3.49999, 3.5],
-    "pmv_toby": [-3.49999, 3.5],
     "rh": [0, 100],
     "pa": [0, 2700],
 }
@@ -174,7 +171,7 @@ var_units = {
 models_to_test = [
     "pmv",
     "pmv_ce",
-]  # ["pmv", "pmv_ce", "pmv_set", "pmv_gagge", "athb", "pmv_toby"]
+]
 
 
 @jit(nopython=True)
@@ -1168,9 +1165,7 @@ def importing_filtering_processing(load_preprocessed=False):
     )
 
     if load_preprocessed:
-        data_ = pd.read_pickle(r"./Data/db_analysis.pkl.gz", compression="gzip")
-        df_meta = pd.read_csv("./Data/db_metadata.csv")
-        return pd.merge(data_, df_meta, on="building_id", how="left")
+        return pd.read_pickle(r"./Data/db_analysis.pkl.gz", compression="gzip")
 
     pa_arr = []
     for i, row in df_.iterrows():
@@ -1187,18 +1182,6 @@ def importing_filtering_processing(load_preprocessed=False):
             & (df_[key] <= applicability_limits[key][1])
         ]
 
-    two_nodes_results = two_nodes(
-        tdb=df_.ta,
-        tr=df_.tr,
-        v=df_.vel,
-        rh=df_.rh,
-        met=df_.met,
-        clo=df_.clo,
-    )
-
-    df_["pmv_gagge"] = two_nodes_results["pmv_gagge"]
-    df_["pmv_set"] = two_nodes_results["pmv_set"]
-
     results_pmv_ppd = pmv_ppd(
         tdb=df_.ta,
         tr=df_.tr,
@@ -1206,7 +1189,7 @@ def importing_filtering_processing(load_preprocessed=False):
         rh=df_.rh,
         met=df_.met,
         clo=df_.clo_d,
-        standard="ISO",
+        standard="iso",
         limit_inputs=False,
     )
 
@@ -1227,24 +1210,6 @@ def importing_filtering_processing(load_preprocessed=False):
     df_["pmv_ce"] = results_pmv_ppd["pmv"]
[…]

### 2024-10-04 main.py
feat(data): updated input data

@@ -3399,8 +3399,6 @@ if __name__ == "__plot__":
     # filter data outside standard applicability limits
     df = importing_filtering_processing(load_preprocessed=True)
 
-    analyse_studies_with_three_speed_measurements(data_=df.copy())
-
     compare_pmv_pmv_ce_comfort_region()
 
     # Figure 1 and 2
@@ -3470,6 +3468,8 @@ if __name__ == "__plot__":
 
     # plot bias by each variable
     plot_bias_distribution_by_variable_binned()
+
+    analyse_studies_with_three_speed_measurements(data_=df.copy())
 
     compare_pmv_disc()
 

### 2024-10-04 main.py
feat(figures): created figure thermal acceptability and thermal comfort

@@ -14,7 +14,6 @@ from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.gridspec import GridSpec
 from numba import jit
 from pythermalcomfort.models import (
-    athb,
     cooling_effect,
 )
 from pythermalcomfort.psychrometrics import p_sat_torr
@@ -1230,7 +1229,9 @@ def importing_filtering_processing(load_preprocessed=False):
     )
 
     save_var_latex("entries_db_used", df_.shape[0])
-    save_var_latex("entries_db_filtered_by_limit_inputs", df_valid_input.shape[0] - df_.shape[0])
+    save_var_latex(
+        "entries_db_filtered_by_limit_inputs", df_valid_input.shape[0] - df_.shape[0]
+    )
 
     df_meta = pd.read_csv("./Data/db_metadata.csv")
     df_ = pd.merge(df_, df_meta, on="building_id", how="left")
@@ -1874,7 +1875,8 @@ def analyse_studies_with_three_speed_measurements(
         # )
         # &
         # (
-                df_three_heights["ta"] > temperature_limit
+        df_three_heights["ta"]
+        > temperature_limit
         # )
     ].to_csv("data_three_heights.csv")
 
@@ -2209,6 +2211,71 @@ def plot_bar_tp_by_ts(data_: pd.DataFrame() = None) -> None:
     ax2.yaxis.tick_right()
     ax2.grid(None)
     plt.savefig(f"./Manuscript/src/figures/bar_plot_tp_by_ts.pdf")
+    plt.show()
+
+
+def plot_bar(
+    data_: pd.DataFrame() = None,
+    x_var: str = "thermal_sensation_round",
+    y_var: str = "thermal_comfort",
+) -> None:
+    # list of colors from matplotlib colormap
+    n = data_[y_var].nunique()
+    binary_cmap = plt.get_cmap("cividis")
+    binary_colors = [binary_cmap(i / (n - 1)) for i in range(n)]
+
+    df_count = data_.groupby(x_var)[[y_var]].count()
+
+    fig, ax1 = plt.subplots(constrained_layout=True)
+    df_plot = data_.groupby(x_var)[y_var].value_counts(normalize=True) * 100
+    df_plot.unstack(y_var).plot.barh(
+        stacked=True, ax=ax1, linewidth=0, width=0.85, color=binary_colors
+    )
+    for ix, row in df_plot.unstack(y_var).round(0).reset_index(drop=True).iterrows():
[…]

### 2024-10-04 main.py
feat(main): re-arranged code

@@ -2255,7 +2255,14 @@ def plot_bar(
         ncol=6,
     )
     if y_var == "thermal_comfort":
-        legend_texts = ['1 - Very uncomfortable', '2', '3', '4', '5', '6 - Very Comfortable']
+        legend_texts = [
+            "1 - Very uncomfortable",
+            "2",
+            "3",
+            "4",
+            "5",
+            "6 - Very Comfortable",
+        ]
         for text, new_label in zip(lg.get_texts(), legend_texts):
             text.set_text(new_label)
     ax1.grid(None)
@@ -2272,7 +2279,6 @@ def plot_bar(
     )
     for ix, row in df_count.reset_index().iterrows():
         ax1.text(102, ix, int(row[y_var]), va="center", ha="left")
-
 
     plt.savefig(f"./Manuscript/src/figures/bar_plot_{y_var}_by_{x_var}.pdf")
 
@@ -3474,8 +3480,14 @@ if __name__ == "__plot__":
     plot_bar_tp_by_ts(data_=df.copy())
 
     plot_bar(data_=df.copy(), x_var="thermal_sensation_round", y_var="thermal_comfort")
-    plot_bar(data_=df.copy(), x_var="thermal_sensation_round", y_var="thermal_acceptability")
-    plot_bar(data_=df.copy(), x_var="thermal_sensation_round", y_var="air_movement_preference")
+    plot_bar(
+        data_=df.copy(), x_var="thermal_sensation_round", y_var="thermal_acceptability"
+    )
+    plot_bar(
+        data_=df.copy(),
+        x_var="thermal_sensation_round",
+        y_var="air_movement_preference",
+    )
 
     # plot model accuracy using bar chart
     plot_stacked_bar_predictions_ts(data_=df.copy(), v_min=0, fig_letters=["a)", "b)"])
@@ -3490,14 +3502,6 @@ if __name__ == "__plot__":
         fig_letters=["a)", "b)"],
         fig_letters_height=0.95,
     )
-    # # only for entries with velocity >= 0.2 m/s and three heights
-    # plot_stacked_bar_predictions_ts(
-    #     data_=pd.read_csv("data_three_heights.csv"),
-    #     v_min=0.2,
-    #     fig_name="bar_stacked_model_accuracy_0.2_three_heights_filtered",
-    #     fig_letters=["a)", "b)"],
-    #     fig_letters_height=0.95,
-    # )
 
     # print Markdown table of f1-scores
[…]

### 2024-10-04 Manuscript/src/sections/introduction.tex
feat(introduction): added limitations and inaccuracies of the pmv-ce

--- 2024-09-26 Manuscript/src/sections/introduction.tex
@@
 \subsection{\ac{pmv-ce} Limitations}\label{subsec:pmv-ce-limitations}
-The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}.
-However, this method has some limitations.
-Firstly, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-Secondly, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set}.
-Thirdly, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
-Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
+The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}~\cite{huang_applicability_2014}.
+\mycite{huang_applicability_2014} showed the \ac{set} model is more accurate than the \ac{pmv} model in predicting thermal sensation votes when subjects are exposed to air movement that are not uniformly distributed across the body.
+However, they never compared the accuracy of the \ac{pmv-ce} with the \ac{pmv} model.
+Using the \ac{set} model to change the inputs in the \ac{pmv} model has several limitations and inaccuracies listed below.
+However, we could not find a rationale for the assumptions used to calculate the \ac{pmv-ce} in the available literature.
+\begin{enumerate}
+    \item Despite the fact that the \ac{ce} is used to partially estimate the heat losses from the skin to the environment, the \ac{pmv-ce} model is still using the underlying equations of the \ac{pmv} model.
+    Hence, if \mycite{huang_applicability_2014} claim that the \ac{pmv} model is inaccurate when the skin has begun to sweat, then the \ac{pmv-ce} model would be equally inaccurate to the \ac{pmv} when the following inputs are passed to the model: \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.15}{\m\per\s}, \ac{met}~=~\qty{1.4}{met}, \ac{clo}~=~\qty{0.5}{clo}.
+    This hypothetical person would be sweating and the \ac{pmv-ce} would only account for the small increase in cooling effect due to the air movement of \qty{0.15}{\m\per\s}, however, it would still not properly account for the evaporative heat loss caused by the hot and humid environment.
+    \item It should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+    \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
+    \item We could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv} model.
+    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
     However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
-We could not find a rationale for this choice in the available literature.
+    \item The \ac{pmv-ce} assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is incorrect since the value of \ac{rh} is a function of the \ac{tdb}.
+    The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.
+\end{enumerate}
 

### 2024-10-29 Manuscript/src/cover-letter.tex
paper: added limitations pmv ce

@@
 
-The study compares the accuracy of the PMV model, as used in the ISO 7730:2005 standard, and its modified version, the PMV$_{CE}$ model, included in the ASHRAE 55:2023 standard.
-The PMV$_{CE}$ was intentionally developed to more accurately estimate the effect of elevated air movement.
+The study compares the accuracy of the PMV models, as used in the ISO 7730:2005 standard, and its modified version, the PMV$_{CE}$ model, included in the ASHRAE 55:2023 standard.
+The PMV$_{CE}$ was intentionally developed to estimate the effect of elevated air movement more accurately.
 The two model formulations only differ when the air speed is greater than \qty{0.1}{\m\per\s}.
@@
 Moreover, we recommend using the PMV model over the PMV$_{CE}$ model, as the latter is more complex and computationally intensive to run, while not being more accurate than the original PMV model.
-Moreover, we challenge the assumption that individuals who feel `slightly warm' or `slightly cool' are thermally neutral.
+Moreover, we challenge the assumption that individuals who feel `slightly warm' or `slightly cool' should be considered thermally comfortable.
 
@@
 All authors have approved the manuscript and agree with its submission to Building and Environment.
+
+We suggest the following reviewers:
+\begin{itemize}
+    \item Prof. Runming Yao, r.yao@reading.ac.uk
+    \item Prof. Jørn Toftum, jt@byg.dtu.dk
+    \item Prof. Shin-ici Tanabe, tanabe@tanabe.arch.waseda.ac.jp
+    \item Assoc. Prof. Peixian Li, lipx@tongji.edu.cn
+    \item Prof. Marcel Schweiker, mschweiker@ukaachen.de
+\end{itemize}
 

### 2024-10-29 Manuscript/src/sections/introduction.tex
paper: added limitations pmv ce

--- 2024-09-26 Manuscript/src/sections/introduction.tex
@@
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
-It is the most extensively used thermal comfort index for estimating thermal sensation.
+It is the most extensively used thermal comfort index.
 A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{2050} documents.
@@
 This highlights this model's extensive adoption and use among the scientific community.
-To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.
+To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, Humphreys2002, doherty_evaluation_1988}.
 
-To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
+\subsection{\ac{pmv-ce} model}\label{subsec:pmv-ce-limitations}
+The \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
 The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented.
-A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
+A partial justification of the model is mainly described in \mycite{arens_moving_2009} and secondarily in \mycite{yang_cooling_2015}.
 The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
@@
 
-\subsection{\ac{pmv-ce} Limitations}\label{subsec:pmv-ce-limitations}
-The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}.
-However, this method has some limitations.
-Firstly, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-Secondly, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set}.
-Thirdly, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
-Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
+The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate heat losses from the skin to the environment due to sweat~\cite{huang_applicability_2014}.
+However, the \ac{pmv} does not, on purpose, estimate the heat losses trough sweating.
+By not estimating the heat losses through sweating, the model calculates the hypothetical heat gain that a cylinder would have based on the inputs, and it converts these gains value into a \ac{pmv} value.
+% \mycite{huang_applicability_2014}, using 19 averaged data points obtained from a set of experiments involving 30 subjects, showed the \ac{set} model could be used to estimate the thermal sensation with two different ways of estimating the airspeed.
+% They developed a model to relate the SET and the TSV. 
+% They also plotted the \ac{pmv} versus the TSV and provided a visual analysis of its performance.
+They concluded that the SET is better than the \ac{pmv} model in predicting thermal sensation votes when subjects are exposed to air movements, but they did not report the performance of the \ac{pmv} model.
+% Moreover, they did not compare the accuracy of the \ac{pmv-ce} with the \ac{pmv} model.
+
+Using the \ac{set} model to change the inputs in the \ac{pmv} model has several concerns, we listed them below.
+\begin{enumerate}
+    \item Although the \ac{ce} is used to partially account for the latent heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model is still using the underlying equations of the \ac{pmv} model.
+    Hence, for example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.15}{\m\per\s}, it has a \ac{met} of \qty{1}{met} and a \ac{clo} of \qty{0.5}{clo}.
+    According to the Gagge's two-node model, which is the model used to calculate \ac{set}, this hypothetical person would be sweating, and they would have \qty{30}{\percent} of their skin covered with sweat.
+    Based on the results from \mycite{huang_applicability_2014}, the \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating.
+    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this scenario, is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement of \qty{0.15}{\m\per\s} which is equal to \qty{.17}{\celsius}.
+    However, even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating since the new inputs to the \ac{pmv-ce} model are \ac{tdb}~=~\ac{tr}~=~\qty{28.8}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.1}{\m\per\s}, \ac{met}~=~\qty{1}{met}, and \ac{clo}~=~\qty{0.5}{clo}.
+    \item The previous example, shows the second limitation of the \ac{pmv-ce}. 
+    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is incorrect, since the value of \ac{rh} is a function of the \ac{tdb}.
+    The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.
+    \item It should be noted that over the years, the critical value of \ac{v} has been changed from \qty{0.15}{\m\per\s} in \mycite{ashrae552023} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+    \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
+    \item We could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the ASHRAE standard over the original \ac{pmv} model.
+    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
     However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
-We could not find a rationale for this choice in the available literature.
+\end{enumerate}
 

### 2024-10-29 Manuscript/src/sections/results.tex
feat(code): updated the code analysis three speeds
paper: added limitations pmv ce

@@
 On the contrary, the \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
+As illustrated in Figure 5c, over \qty{80}{\percent} of occupants who feel `slightly warm' or `slightly cool' consider the space thermally acceptable. 
+However, thermal comfort is defined as ``the state of mind that expresses satisfaction with the thermal environment,'' rather than simply finding the space acceptable. 
+This distinction highlights that while many may tolerate slight deviations from their preferred temperature, true thermal comfort involves a higher level of satisfaction with the surrounding thermal conditions.
 
@@
 The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
-The overall accuracy of both \ac{pmv} models increass to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+The overall accuracy of both \ac{pmv} models increase to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
 The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.

### 2024-10-29 Manuscript/src/sections/titleAbstract.tex
paper: added limitations pmv ce

--- 2024-10-04 Manuscript/src/sections/titleAbstract.tex
@@
 
-    \address[label1]{The University of Sydney, Sydney, AU}
-    \address[label3]{Center for the Built Environment, University of California, Berkeley, CA, USA}
+    \address[label1]{School of Architecture, Design, and Planning, The University of Sydney, Sydney, AU}
+    \address[label3]{Center for the Built Environment, Department of Architecture, Department of Civil and Environmental Engineering, University of California, Berkeley, CA, USA}
 
@@
 
+    \begin{graphicalabstract}
+        \includegraphics[width=\linewidth]{figures/graphical_abstract}
+    \end{graphicalabstract}
+
+    % Research highlights
+    \begin{highlights}
+        \item We compared the accuracy of the \ac{pmv} models in \gls{7730} and \gls{55}.
+        \item Both models have low and similar prediction accuracy ($\approx$\qty{34}{\percent}).
+        \item The \ac{pmv-ce} has a higher bias than \ac{pmv} in predicting thermal sensation.
+        \item We recommend limiting the applicability range of the models to $\mid$\ac{pmv}$\mid \leq 0.5$.
+        \item It is misleading to classify slightly warm/cool individuals as thermally comfortable.
+    \end{highlights}
+
     \begin{keyword}

### 2024-10-30 Manuscript/src/sections/titleAbstract.tex
feat(abstract): minor change

@@
         A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
-        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort, therefore, we recommend using the thermal preference scale instead of the thermal sensation scale.
+        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort.
         In the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
+        Therefore, we recommend using the three point thermal preference scale instead of the thermal sensation scale.
     \end{abstract}

### 2024-10-30 Manuscript/src/sections/appendix.tex
feat(code): analysis three height and t threshold

@@
 \end{figure}
-The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.38}.
+The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.39}.
 The interquartile range for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
 These results allow us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
+
+\subsection{Dataset with \ac{v} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}}\label{subsec:dataset-with-v-measured-at-three-heights-and-tdb-geq-24-celsius}
+Here we are showing the results for the subset of data with \ac{v} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}.
+We are also reporting these results since the \ac{pmv-ce} aims to more accurately predict the thermal sensation, compared to the \ac{pmv}, when air movement is used to cool the occupants.
+The results are shown in Figure~\ref{fig:hist_discrepancies_three_heights_limit_t}.
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/hist_discrepancies_three_heights_limit_t}
+    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}.}
+    \label{fig:hist_discrepancies_three_heights_limit_t}
+\end{figure*}
+Even for this subset of data, the \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
+The \ac{pmv} model has a bias of \num{-.08} while the \ac{pmv-ce} model has a bias of \num{-.31}.
+This result is consistent with the previous analysis and confirms that the \ac{pmv} model is more accurate than the \ac{pmv-ce}.

### 2024-10-30 main.py

@@ -13,9 +13,7 @@ import seaborn as sns
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.gridspec import GridSpec
 from numba import jit
-from pythermalcomfort.models import (
-    cooling_effect,
-)
+from pythermalcomfort.models import cooling_effect, two_nodes
 from pythermalcomfort.psychrometrics import p_sat_torr
 from pythermalcomfort.utilities import check_standard_compliance_array
 from scipy import optimize
@@ -32,6 +30,11 @@ warnings.filterwarnings("ignore")
 
 psychrolib.SetUnitSystem(psychrolib.SI)
 
+
+results = two_nodes(29, 29, 0.15, 80, 1, 0.5, 0)
+results["w"]
+results = cooling_effect(29, 29, 0.15, 80, 1, 0.5, 0)
+print(results)
 
 palette_tp = [
     "#06A6EE",
@@ -1759,8 +1762,7 @@ def plot_distribution_variable():
 def analyse_studies_with_three_speed_measurements(
     data_: pd.DataFrame = None,
     air_speed_limit: float = 0.2,
-    temperature_limit: float = 24.0,
-):
+) -> pd.DataFrame:
     # [['contributor', 'publication', 'region', 'country', 'city']]
     df_three_heights = data_[data_["vel_r"] >= air_speed_limit].dropna(subset="vel_l")
     df_three_heights.replace(
@@ -1865,21 +1867,6 @@ def analyse_studies_with_three_speed_measurements(
     # )
     # plt.show()
 
-    df_three_heights[
-        # df_three_heights["publication"].isin(
-        #     [
-        #         "Zhang, Y., Chen, H., Meng, Q. (2013)",
-        #         "Zhang, Y. et al. (2010)",
-        #         "Tartarini, F., Cooper, P., Fleming, R. (2018)",
-        #     ]
-        # )
-        # &
-        # (
-        df_three_heights["ta"]
-        > temperature_limit
-        # )
-    ].to_csv("data_three_heights.csv")
-
     plt.close("all")
     df_plot = df_three_heights.loc[
         df_three_heights["publication"].isin(authors_to_show),
@@ -1938,6 +1925,8 @@ def analyse_studies_with_three_speed_measurements(
     # ax.grid(False)
     plt.savefig("./Manuscript/src/figures/boxenplot_wind_speed_three_heights.pdf")
     plt.show()
+
[…]

### 2024-10-30 Manuscript/src/sections/introduction.tex
feat(introduction): wrote limitations pmv ce

@@
 \mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} standard~\cite{iso7730} in its original form.
-The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
+The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories `cold,' `cool,' `slightly cool,' `neutral,' `slightly warm,' `warm,' and `hot.'~\cite{iso7730, ashrae552023}.
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
 It is the most extensively used thermal comfort index.
-A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{2050} documents.
+A search within the article title, abstract, and keyword on Scopus with the search term `predicted mean vote' returns approximately \num{2050} documents.
 Out of those \num{1200} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
@@
 
-The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate heat losses from the skin to the environment due to sweat~\cite{huang_applicability_2014}.
-However, the \ac{pmv} does not, on purpose, estimate the heat losses trough sweating.
-By not estimating the heat losses through sweating, the model calculates the hypothetical heat gain that a cylinder would have based on the inputs, and it converts these gains value into a \ac{pmv} value.
+The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does accurately estimate the convective and evaporative heat losses from the skin to the environment~\cite{huang_applicability_2014}.
+It is true that the \ac{pmv} does not accurately estimate the evaporative heat losses trough sweating.
+The \ac{pmv} model assumes that sweating is minimal and constant, for a specific \ac{met}, hence if sweating would be required to maintain the body in thermal equilibrium the \ac{pmv} model return a positive \ac{pmv} value.
+The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
+However, despite this limitation of the \ac{pmv} model, we have identified several limitations of the \ac{pmv-ce} model, which we list below.
 % \mycite{huang_applicability_2014}, using 19 averaged data points obtained from a set of experiments involving 30 subjects, showed the \ac{set} model could be used to estimate the thermal sensation with two different ways of estimating the airspeed.
@@
 
-Using the \ac{set} model to change the inputs in the \ac{pmv} model has several concerns, we listed them below.
 \begin{enumerate}
-    \item Although the \ac{ce} is used to partially account for the latent heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model is still using the underlying equations of the \ac{pmv} model.
-    Hence, for example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.15}{\m\per\s}, it has a \ac{met} of \qty{1}{met} and a \ac{clo} of \qty{0.5}{clo}.
-    According to the Gagge's two-node model, which is the model used to calculate \ac{set}, this hypothetical person would be sweating, and they would have \qty{30}{\percent} of their skin covered with sweat.
+    \item We could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the ASHRAE standard over the original \ac{pmv} model.
+    \item Although the \ac{ce} is used to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model.
+    For example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.15}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clo} of \qty{0.5}{clo}.
+    According to the Gagge's two-node model, which is used to calculate the \ac{set}, this hypothetical person would have \qty{30}{\percent} of their skin covered with sweat.
     Based on the results from \mycite{huang_applicability_2014}, the \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating.
-    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this scenario, is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement of \qty{0.15}{\m\per\s} which is equal to \qty{.17}{\celsius}.
-    However, even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating since the new inputs to the \ac{pmv-ce} model are \ac{tdb}~=~\ac{tr}~=~\qty{28.8}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.1}{\m\per\s}, \ac{met}~=~\qty{1}{met}, and \ac{clo}~=~\qty{0.5}{clo}.
-    \item The previous example, shows the second limitation of the \ac{pmv-ce}. 
-    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is incorrect, since the value of \ac{rh} is a function of the \ac{tdb}.
+    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this scenario, is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement (delta between \qty{0.15}{\m\per\s} and \qty{0.1}{\m\per\s}) which is equal to \qty{.17}{\celsius}.
+    Even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating.
+    Therefore, in this scenario, the \ac{pmv-ce} model also cannot accurately predict the \ac{tsv}.
+    \item The previous example, highlights another limitation of the \ac{pmv-ce}.
+    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is incorrect, since the value of \ac{rh} is dependent on \ac{tdb}.
     The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.
-    \item It should be noted that over the years, the critical value of \ac{v} has been changed from \qty{0.15}{\m\per\s} in \mycite{ashrae552023} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-    \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
-    \item We could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the ASHRAE standard over the original \ac{pmv} model.
+    \item It should be noted that over the years, the critical value of \ac{v} has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
     \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
     However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
+    \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
 \end{enumerate}

### 2024-10-30 Manuscript/src/sections/results.tex
feat(results): added figure thermal acceptability

@@
 
-The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
+The percentage of \ac{tpv} grouped by each \ac{tsv} is shown in Figure~\ref{fig:bar_plot_tp_by_ts} a).
 \begin{figure*}[htb!]
     \centering
+    \begin{subfigure}[b]{\textwidth}
+        \centering
         \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
-    \caption{The graph on the left shows the percentage of \ac{tpv} for each thermal sensation vote.
+        \caption{Figure a) shows the percentage of \ac{tpv} for each thermal sensation vote.
     The numbers on the right side of each bar show the number of points for each \ac{tsv}.
-    The graph on the right shows the total number of data points grouped by \ac{tsv}.
+        Figure b) shows the total number of data points grouped by \ac{tsv}.
     Each bin in the right figure has more data points than in the left figure, since less data on \ac{tpv} were available.}
     \label{fig:bar_plot_tp_by_ts}
+    \end{subfigure}
+    \par\bigskip % force a bit of vertical whitespace
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/bar_plot_thermal_acceptability_by_thermal_sensation_round}
+    \caption{
+        Figure c) shows the percentage of participants who consider the space thermally acceptable for each thermal sensation vote.}
+        \label{fig:thermal_acceptability_by_ts}
+    \end{subfigure}
+    \caption{The distribution of \ac{tpv} and thermal acceptability grouped by each \ac{tsv}.}
 \end{figure*}
@@
 The \ac{tsv} dataset is unbalanced.
-Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
-While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
-
+Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `neutral'.
+While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `hot' or `cold'.
 In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable~\cite{Fanger1970, schweiker2020evaluating}.
+Fanger, in his seminal book on thermal comfort wrote that ``The dissatisfied are defined here as those who vote \num{-2} (cool) or \num{-3} (cold), \num{2} (warm) or \num{3} (hot).
+One could perhaps object that those voting \num{-1} or \num{1} were not included also, but as evidenced by Gagge et al., real discomfort is first expressed by those voting higher than \num{2} or lower than \num{-2}.''~\cite{Fanger1970}.
 However, \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
@@
 On the contrary, the \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
-As illustrated in Figure 5c, over \qty{80}{\percent} of occupants who feel `slightly warm' or `slightly cool' consider the space thermally acceptable. 
+As illustrated in Figure~\ref{fig:thermal_acceptability_by_ts}, over \qty{80}{\percent} of occupants who feel `slightly warm' or `slightly cool' consider the space thermally acceptable.
 However, thermal comfort is defined as ``the state of mind that expresses satisfaction with the thermal environment,'' rather than simply finding the space acceptable. 
 This distinction highlights that while many may tolerate slight deviations from their preferred temperature, true thermal comfort involves a higher level of satisfaction with the surrounding thermal conditions.
+We did not analyze the thermal comfort data since in the \ac{db2} it is unclear which scale was used to collect the data.
 

### 2024-12-05 Manuscript/src/sections/appendix.tex
stefano: final review before submission

--- 2024-09-26 Manuscript/src/sections/appendix.tex
@@
 \end{figure}
-The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.38}.
+The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.39}.
 The interquartile range for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
 These results allow us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
+
+\subsection{Dataset with \ac{v} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}}\label{subsec:dataset-with-v-measured-at-three-heights-and-tdb-geq-24-celsius}
+We are showing here the results for the subset of data with \ac{v} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}.
+We are reporting these results since the \ac{pmv-ce} aims to more accurately predict the thermal sensation, compared to the \ac{pmv}, when air movement is used to cool the occupants.
+The results are shown in Figure~\ref{fig:hist_discrepancies_three_heights_limit_t}.
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/hist_discrepancies_three_heights_limit_t}
+    \caption{Bias of the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}.}
+    \label{fig:hist_discrepancies_three_heights_limit_t}
+\end{figure*}
+
+The \ac{pmv} model has a bias of \num{-.08} while the \ac{pmv-ce} model has a bias of \num{-.31}.
+Even for this subset of data, the \ac{pmv} model is more accurate than the \ac{pmv-ce} model; this result is consistent with the previous analysis and confirms that the \ac{pmv} model is more accurate than the \ac{pmv-ce}.

### 2024-12-05 Manuscript/src/sections/introduction.tex
stefano: final review before submission

@@
 
-The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does accurately estimate the convective and evaporative heat losses from the skin to the environment~\cite{huang_applicability_2014}.
-It is true that the \ac{pmv} does not accurately estimate the evaporative heat losses trough sweating.
-The \ac{pmv} model assumes that sweating is minimal and constant, for a specific \ac{met}, hence if sweating would be required to maintain the body in thermal equilibrium the \ac{pmv} model return a positive \ac{pmv} value.
+The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment~\cite{huang_applicability_2014}.
+The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}. 
+Hence, if sweating is required to maintain the body in thermal equilibrium, the \ac{pmv} model assumes that the body would gain heat.
 The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
@@
     \item The previous example, highlights another limitation of the \ac{pmv-ce}.
-    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is incorrect, since the value of \ac{rh} is dependent on \ac{tdb}.
+    The assumption of keeping \ac{rh} constant after adjusting \ac{tdb} and \ac{tr} is thermodynamically incorrect since the value of \ac{rh} is dependent on \ac{tdb}.
     The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.

### 2024-12-05 Manuscript/src/sections/titleAbstract.tex
stefano: final review before submission

@@
         We compared the accuracy of the two \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that both models have low and similar prediction accuracy (\qty{34}{\percent}) and the \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+        We found that both models have low and similar prediction accuracy (\qty{34}{\percent}). 
+        The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{v}~$\geq$~\qty{0.2}{\m\per\s}.
         We suggest reducing the applicability limit of the models within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} models cannot determine the degree to which people are dissatisfied.

### 2024-12-05 Manuscript/src/sections/titleAbstract.tex
submission: first submission energy and buildings

@@
     \begin{abstract}
-        The \ac{pmv} model, developed by Fanger, is still currently used in its original form in the \gls{7730}.
+        The \ac{pmv} model, developed by Fanger, is included in the ISO 7730:2005.
         It has a low prediction accuracy, especially outside thermal neutrality.
@@
         The \gls{7730} and \gls{55} are the most widely referenced thermal comfort standards worldwide, and their different \ac{pmv} formulations are a source of confusion.
-        Which model is more accurate?
 
-        We compared the accuracy of the two \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that both models have low and similar prediction accuracy (\qty{34}{\percent}). 
+        We compared the accuracy of the two \ac{pmv} models, against \var{entries_db_used} thermal sensation votes collected in buildings.
+        We found that both models have low and similar prediction accuracy ($\approx$\qty{34}{\percent}).
         The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{v}~$\geq$~\qty{0.2}{\m\per\s}.
-        We suggest reducing the applicability limit of the models within the range $\mid$\ac{pmv}$\mid \leq 0.5$ since the \ac{pmv} models cannot determine the degree to which people are dissatisfied.
         Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
-        This value is lower than random guessing (\qty{14}{\percent}), meaning they are misleading the users.
+        This value is lower than random guessing (\qty{14}{\percent}), meaning they are misleading users.
         A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} only indicates that people's thermal sensation is on the `warm' or `cold' side, respectively.
+        Hence, we suggest reducing the applicability limit of both models within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
         Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort.
-        In the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
-        Therefore, we recommend using the three point thermal preference scale instead of the thermal sensation scale.
+        In the \acl{db2}, \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
     \end{abstract}

### 2024-12-05 main.py

@@ -3666,361 +3666,3 @@ if __name__ == "__plot__":
         #
         # # Display plot in 'SciView'.
         # p.show()
-
-
-if __name__ == "__old_code__":
-    # accuracies calculation
-    for limit in [3, 2, 1]:
-        data = df[df["thermal_sensation_round"].abs() <= limit]
-        data_iso = data[df["pmv_round"].abs() <= limit]
-        data_ash = data[df["pmv_ce_round"].abs() <= limit]
-
-        acc_iso = (
-            data_iso[
-                data_iso["thermal_sensation_round"] == data_iso["pmv_round"]
-            ].shape[0]
-            / data_iso.shape[0]
-        )
-        save_var_latex(f"Overall PMV ISO accuracy - limit {limit}", int(acc_iso * 100))
-        acc_ash = (
-            data_ash[
-                data_ash["thermal_sensation_round"] == data_ash["pmv_ce_round"]
-            ].shape[0]
-            / data_ash.shape[0]
-        )
-        save_var_latex(
-            f"Overall PMV ASHRAE accuracy - limit {limit}", int(acc_ash * 100)
-        )
-
-    # logistic regression models
-    plt.figure()
-    sns.boxenplot(df.pmv_gagge_hb)
-    print(df.groupby("thermal_preference")["ta"].count())
-
-    # gagge heat balance vs thermal sensation
-    clf = LogisticRegression(random_state=0).fit(
-        df["pmv_gagge_hb"].values.reshape(-1, 1), df["thermal_sensation_round"]
-    )
-    set_range = np.arange(-60, 60, 0.5)
-    prob = clf.predict_proba(set_range.reshape(-1, 1))
-    df_prob = pd.DataFrame(prob, columns=sorted(df["thermal_sensation_round"].unique()))
-    plt.figure()
-    for col in df_prob.columns:
-        plt.plot(set_range, df_prob[col], label=col)
-    plt.tight_layout()
-    plt.legend()
-
-    # gagge heat balance vs thermal preference
-    df_dropna = df[["pmv_set_hb", "thermal_preference"]].dropna().sample(frac=1)
-    df_log = pd.DataFrame()
-    for preference in df_dropna.thermal_preference.unique():
-        _df = df_dropna.query("thermal_preference == @preference").head(10000)
[…]

### 2025-01-10 Manuscript/src/response-letter.tex
feat: started adding reviewers comments to response letter.

\documentclass[a4paper, 10pt]{letter}

% package to highlight text
\usepackage{color,soul}

\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
}

\usepackage{graphicx}

% Name of sender
\name{Federico Tartarini}

% Signature of the sender
\signature{Federico Tartarini}

\newcommand{\response}[1]{\textcolor{blue}{\textbf{Author response:} #1}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO}: #1}}

%--------------------------------------------------------------------------%

\begin{document}

% Name and address of the receiver
    \begin{letter}
    {
        Response to reviewers' comments.
    }

% Opening statement
        \opening{Dear Athanasios Tzempelikos and Reviewers,}

% Letter body

        We would like to thank you for your time and your valuable feedback.
        Please find below our answers to all your comments.

        \textbf{Reviewer 1}
        Dear Authors,
        This manuscript is well-written and showcases your strong capabilities in addressing thermal comfort issues. 
        I have just a few comments to share with you Wishing you a Merry Christmas and a Happy New Year.

        \begin{enumerate}

            \item P4 L17. Icl is the basic clothing insulation. Total clothing insulation accounts for the air insulation and the clothing area factor (see ISO 9920).

            \response{
            Here our response
            }

            \item P4 L34-38. "The rationale for the development of the PMVCE model is that the original PMV
formulation does not accurately estimate convective and evaporative heat losses from
the skin to the environment [10]." If this is the rationale, why CE is subtracted from both
the dry-bulb air temperature (tdb) and mean radiant temperature (tr)? Please provide evidence.

\response{
            Here our response
            }

\item P7 L28-32. The authors address (also in methods and in the section 3.3) a highly relevant issue: the impact of microclimatic measurements on the accurate evaluation of thermal comfort. Unfortunately, thermal comfort databases suffer from significant biases due to several factors, including the absence of standardized measurement protocols (such as representative positions and height above the ground), incorrect or arbitrary assumptions (for instance, the belief in uniform and stable conditions), the use of inaccurate measurement devices, deficiencies in software, and the failure to assess clothing insulation. As a result, the PMV (Predicted Mean Vote) value can appear random, making it difficult to evaluate the category of indoor environmental quality. These issues have been previously discussed (e.g., in 10.1016/j.buildenv.2011.01.001) and should be emphasized more prominently in the current investigation. Such problems impact all PMV-related models, including adaptive models, where the operative temperature is often assumed to be the same as the air temperature.

\response{
            Here our response
            }

\item The database used for this investigation also includes high air velocity measurements (v>0.2 m/s). However, this does not rule out the possibility of local discomfort caused by draughts. What about the impact of these draughts on the overall thermal comfort as quantified by the Thermal Sensation Vote (TSV)? Are the authors confident that the TSV values collected at high air velocity are not influenced by local discomfort? Please provide a brief explanation.

\response{
            Here our response
            }

        \end{enumerate}

        % \textbf{Reviewer 2}
        % The paper introduces the authors' developed toolkit, CBE Clima Tool, a web application designed for analyzing and visualizing climate data. 
        % This tool has the potential to provide valuable practical support to AEC designers and engineers and has achieved significant success in terms of user adoption and applications. 
        % However, for publication as an academic journal paper, the authors should address the following aspects to more effectively showcase the originality and innovation of their work:

        % \begin{enumerate}
        %     \item It is important to clarify the key features that set the tool apart from existing weather tools. 
        %     Is the web-based or interactive feature the primary distinguishing factor? 
        %     To enhance the discussion, the authors should provide a detailed exploration of these features. For instance:
        %     What specific advantages does this web application offer over traditional software that performs local computation?
        %     Is there any empirical data available to quantify its interactive performance?

        %     \response{
        %     Thank you for the extremely valuable suggestions.
        %     Based on your feedback we have added a new sub-section within the Introduction section in which we describe in detail the reason why we believe our tool is extremely valuable to AEC designers and engineers.
        %     This sub-section is titled 2.4 Contribution.
        %     In this section, we also clarify the key features that set the tool apart from existing weather tools.
        %     Finally, we list the specific advantages that a web application tool offers over traditional software that performs local computation.
        %     }

        %     \item Which components of the tool are integral to highlighting its innovation? 
        %     While the paper introduces various models, such as UTCI and the adaptive thermal comfort model, it remains unclear which elements are at the core of the innovation. 
        %     In Section 5, the Sun Path Chart and Psychrometric Chart are presented as illustrative examples, but it should be acknowledged that other tools can also fulfil these functions.

        %     \response{
        %     We appreciate the reviewer's thoughtful feedback and would like to clarify the key features that distinguish CBE Clima Tool as an innovative solution for climate analysis tailored to sustainable building design.
        %     The core innovation of Clima extends beyond the calculation of thermal comfort indices, a capability shared by several cloud-based and desktop applications. 
        %     What sets Clima apart is its commitment to universal accessibility. 
        %     It empowers users worldwide with the ability to swiftly and reliably analyze and visualize intricate climate data.
        %     While other tools may offer similar graphical outputs, Clima stands out due to its exceptional ease of use, interactivity, and the superior quality of its graphical representations. 
        %     These features are invaluable to architects, engineers, students, and educators in effectively conveying climate-adapted building design concepts.
        %     Clima's web-based nature eliminates the need for local installations, ensuring cross-platform compatibility and seamless accessibility. 
        %     Furthermore, a fundamental innovation lies in its status as a free and open-source tool. 
        %     This transparency allows users to scrutinize and validate the underlying algorithms and calculations, fostering trust and confidence in the tool's capabilities.
        %     Our commitment to an open-source approach also encourages community-based development, facilitating collaborative enhancements and guaranteeing that Clima evolves through input from a diverse array of contributors. 
        %     Additionally, it grants users the opportunity to incorporate Clima's code into their own applications, further extending its impact within the research and engineering communities.
        %     In summary, the innovation of CBE Clima Tool lies in its global accessibility, user-friendliness, transparency, and its capacity to foster collaborative development, making it a valuable asset for sustainable building design professionals and researchers alike.
        %     We have summarised this answer in Sub-Section 2.4 of the revised manuscript.
        %     }

        %     \item Furthermore, the paper currently reads more like a report. To align with the style of an academic journal paper, it is recommended that the authors revise certain sections. For instance, the introduction could be enhanced by breaking down the lengthy one-page paragraph into several well-structured paragraphs.

        %     \response{
        %     We would like to express our gratitude to the reviewer for their constructive feedback and recommendations regarding the structure of our manuscript. 
        %     We acknowledge the importance of aligning the manuscript with the style and standards of an academic journal paper.
        %     Based on your suggestion, we have thoroughly revised the Introduction chapter. 
        %     Specifically, we have broken down the lengthy one-page paragraph into several well-structured paragraphs to enhance readability and clarity. 
        %     This restructuring allows for a more systematic presentation of the key concepts and objectives of our work.
        %     The revised Introduction now provides a more concise and focused overview of the background, motivation, and significance of CBE Clima Tool. 
        %     We have carefully organized the content to guide readers through the rationale behind our research, the challenges addressed by the tool, and its potential impact on the field of sustainable building design.
        %     We believe that these improvements will assist readers in comprehending the context and importance of our work.
        %     }

        % \end{enumerate}

% Closing statement
        Kind regards,

        \vspace*{5px}

        Federico Tartarini

    \end{letter}
\end{document}
%-----------------------------------------------------------------------------%

### 2025-01-10 Manuscript/src/sections/conclusions.tex
feat(acknowledgments): acknowledgments
paper: added limitations pmv ce
feat: started adding reviewers comments to response letter.

@@
 Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
+\todo{But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
 Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
+\todo{unclear what limits you are talking about, or who the audience is for your suggestions.
+The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.
+Your suggestions are actually just aimed at yourselves as keepers of the ComfortTool}
 The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
@@
 We then recommend that the \gls{55} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
+\todo{and why not the ISO Standard also?  After all, ASHRAE was copying ISO on that one}
 
@@
 This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.
+\todo{This is a major suggestion with very significant implications for energy etc that have not been discussed in the paper.
+There are surely reasons people have not used that scale on its own, for this purpose.}
 
-% todo Federico could you add an acknowledgement section? Paul Raftery
+\section*{Acknowledgements}\label{sec:acknowledgements}
+We would like to thank Prof. Edward Arens and Dr. Paul Raftery for their comments and suggestions on the manuscript.

### 2025-01-10 Manuscript/src/sections/introduction.tex
feat: started adding reviewers comments to response letter.

@@
 \section{Introduction}\label{sec:introduction}
+
+\todo[inline]{General notes:  
+The objective of the ASHRAE Std 55 approach is to substitute the Pierce two-node model for the PMV model for modeling heat loss under elevated air speeds in warm conditions.  
+The CE method is intended to provide a PMV sensation prediction from the more physiologically comprehensive two-node model.  
+The work on this paper has raised a number of interesting future research topics that we should discuss.}
+\todo[inline, color=blue!40]{
+FT: I agree that it would be ideal if Std 55 replaced the PMV with a more accurate model, but in its current form, Std 55 2023 still fundamentally uses the PMV model to calculate the thermal sensation.
+The PMV CE is a modified version of the PMV, which is not more physiologically comprehensive than the PMV model since as shown in the flow chart the PMV CE uses the same code as the PMV to calculate the thermal sensation.
+Only temperature is adjusted to account for elevated air movement but the other environmental or personal parameters are not adjusted.
+Moreover, as highlighted in Section 1.1 the use of the CE introduces several errors.
+}
+
 \mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} standard~\cite{iso7730} in its original form.
@@
 To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, Humphreys2002, doherty_evaluation_1988}.
+
+\todo[inline]{General notes:  
+The PMV model consists of a heat balance model and an empirical equation that predicts thermal sensation (PMV) from heat balance.
+(need to describe this further, especially the simplified way that sweating is treated by the heat balance model).
+In 2010 ASHRAE Std 55 modified the PMV heat balance calculation to make it more capable of predicting the cooling caused by elevated air movement under warm conditions.
+It uses the Gagge two-node model to compute the convective and evaporative heat loss from the body.
+This model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss.
+The model output can be expressed as Standard Effective Temperature (SET) (provide citations to Gagge/Foblets).
+SET is implemented in Std 55 as an index to quantify the equivalent temperature reduction caused by air movement cooling.
+The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.
+This approach is intended to minimize errors introduced by the Fanger heat balance model’s reduced evaporative capacity on the predicted PMV result.
+}
+\todo[inline, color=blue!40]{
+FT similar to my previous comment. Firstly, I would like to mention here that the PMV does not at all try to estimate sweating. This is a problem, but neither the PMV nor the PMV CE do that.
+The Std 55 did not modify the PMV heat balance equation, instead it simply and only reduced the air and mean radiant temperatures by the cooling effect, this is not equivalent to doing an accurate heat balance calculation.
+The PMV CE consequently does not: 1) uses the Gagge two-node model to compute the convective and evaporative heat loss from the body 2) model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss.
+In addition, another comment about the PMV CE is that the cooling effect is not only applied in warm conditions but in fact it is applied in any conditions.
+SET is also a rather confusing variable since it is calculated not for a reference environment with standardised clo and met.
+But rather it is the temperature of a hypothetical isothermal environment at 50\% rh, <0.1 m/s (20 fpm) average air speed Va, and in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level.
+Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} this however do not ensure that the person is not feeling hot in this condition or sweating, consequently does not solve the issues that the original PMV has.
+}
 
@@
 According to the standard, the measured value should be adjusted using this equation \ac{v} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
+\todo[size=\small]{you might explain that this equation is used in the ISO Standard as well}
+\todo[inline, color=blue!40]{Thank you for the feedback I will also create a flowchart for the PMV ISO and include it in the appendix.}
 The value of \ac{ce} is calculated using the \ac{set} equation.
@@
 We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
+\todo{It would be great to add 0.6m/s to this figure.   That would be the objective for ceiling fan designs.  There are plenty of lab studies that have measured results in that region.}
+\todo[inline, color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
 \begin{figure}[!htb]
@@
 
-The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment~\cite{huang_applicability_2014}.
+The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds ~\cite{huang_applicability_2014}.
 The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}. 
@@
 The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
-However, despite this limitation of the \ac{pmv} model, we have identified several limitations of the \ac{pmv-ce} model, which we list below.
+While the \ac{pmv} model has its limitations, our analysis has revealed additional limitations specific to the \ac{pmv-ce} model, as outlined below.
+\todo{This is a confusing dichotomy}
+\todo[color=blue!40]{FT. Is the new sentence better?}
+
 % \mycite{huang_applicability_2014}, using 19 averaged data points obtained from a set of experiments involving 30 subjects, showed the \ac{set} model could be used to estimate the thermal sensation with two different ways of estimating the airspeed.
@@
     \item We could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the ASHRAE standard over the original \ac{pmv} model.
+    \todo[inline]{true enough, but in the standard the objective is to have a rational physiology model. 
+    The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
+    \todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}
     \item Although the \ac{ce} is used to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model.
@@
     Therefore, in this scenario, the \ac{pmv-ce} model also cannot accurately predict the \ac{tsv}.
+    \todo[inline]{This is really confusing.  
+    You are evaluating the Cooling Effect in still air, when there is basically no Cooling Effect to evaluate.    
+    You are using as an example 29C in still air and 80\% RH which of course will cause sweating.  
+    You are citing skin wettedness values that have not been introduced in this paper (and are not part of any standard).
+Where did they come from?
+You allude to results from Huang 2014 that the reader is supposed to know about.
+The skin wettedness discussion is irrelevant unless you go into a lot more physiological detail here.}
     \item The previous example, highlights another limitation of the \ac{pmv-ce}.
@@
     \item It should be noted that over the years, the critical value of \ac{v} has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+    \todo[inline]{(where does the term ‘critical temperature’ come from?
+It is a threshold value and not critical to any predicted outcome.
+The changes over the years were due to converting from historical graphics, objections from diffuser manufacturers concerned about lawsuits, Peter Simmonds, etc.
+None of them constitute a ‘limitation’ to the model}
     \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
     However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
+    \todo[inline]{I think this argument is incorrect.
+SET is only used as a consistent method of obtaining a meaningful delta operative temperature.
+It does not change the environment.
+The environment is kept at the original RH.}
     \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
+    \todo[inline]{This may or may not be significant.
+If it is significant, the process could be simplified to a total of three calculations while remaining a rational method.
+What’s in there now was Tyler’s (CBE’s) choice.
+Charlie and I have used it in the simpler way.
+We could propose this to the Standards Committee if we wanted to.
+We would have to calculate two validation tables for the two approaches.
+}
 \end{enumerate}
@@
 Several other formulations of the \ac{pmv} model have been proposed.
-Among the most notable formulations there are the \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
+Among the most notable formulations there are the \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \todo{explain how this differs from the CE approach since it looks like the same thing} \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
 \mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
@@
 Some notable works that tried to determine the accuracy of the \ac{pmv} model include the work from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
+\todo{Doherty and I did not evaluate air movement which hardly anyone was considering back then.}
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.

### 2025-01-10 Manuscript/src/response-letter.tex
rebuttal: formatted rebuttal letter

@@
 
-            \item P4 L17. Icl is the basic clothing insulation. Total clothing insulation accounts for the air insulation and the clothing area factor (see ISO 9920).
+            \item P4 L17.
+            Icl is the basic clothing insulation.
+            Total clothing insulation accounts for the air insulation and the clothing area factor (see ISO 9920).
 
@@
 
-            \item P4 L34-38. "The rationale for the development of the PMVCE model is that the original PMV
+            \item P4 L34-38.
+            ``The rationale for the development of the PMVCE model is that the original PMV
             formulation does not accurately estimate convective and evaporative heat losses from
-the skin to the environment [10]." If this is the rationale, why CE is subtracted from both
-the dry-bulb air temperature (tdb) and mean radiant temperature (tr)? Please provide evidence.
+            the skin to the environment [10].`` If this is the rationale, why CE is subtracted from both
+            the dry-bulb air temperature (tdb) and mean radiant temperature (tr)?
+            Please provide evidence.
 
@@
 
-        % \textbf{Reviewer 2}
-        % The paper introduces the authors' developed toolkit, CBE Clima Tool, a web application designed for analyzing and visualizing climate data. 
-        % This tool has the potential to provide valuable practical support to AEC designers and engineers and has achieved significant success in terms of user adoption and applications. 
-        % However, for publication as an academic journal paper, the authors should address the following aspects to more effectively showcase the originality and innovation of their work:
+        \clearpage
 
-        % \begin{enumerate}
-        %     \item It is important to clarify the key features that set the tool apart from existing weather tools. 
-        %     Is the web-based or interactive feature the primary distinguishing factor? 
-        %     To enhance the discussion, the authors should provide a detailed exploration of these features. For instance:
-        %     What specific advantages does this web application offer over traditional software that performs local computation?
-        %     Is there any empirical data available to quantify its interactive performance?
+        \textbf{Reviewer 2}
+        This article conducts a comparative analysis of the accuracy of the PMV model implemented in ISO 7730:2005 and ASHRAE 55:2023 using data from the ASHRAE 2 database.
+        The research has good practical significance, but there are some issues that need further clarification:
 
-        %     \response{
-        %     Thank you for the extremely valuable suggestions.
-        %     Based on your feedback we have added a new sub-section within the Introduction section in which we describe in detail the reason why we believe our tool is extremely valuable to AEC designers and engineers.
-        %     This sub-section is titled 2.4 Contribution.
-        %     In this section, we also clarify the key features that set the tool apart from existing weather tools.
-        %     Finally, we list the specific advantages that a web application tool offers over traditional software that performs local computation.
-        %     }
+        \begin{enumerate}
+            \item The full text discusses the PMV model in ISO 7730:2005 and the PMVCE model in ASHRAE 55:2023, that is, the PMV models in the standards.
+            When Professor Fanger initially proposed the PMV model, there were certain applicable conditions.
+            Does the article aim to explore the accuracy of the PMV model, or the inappropriate application of this model in the two standards?
 
-        %     \item Which components of the tool are integral to highlighting its innovation? 
-        %     While the paper introduces various models, such as UTCI and the adaptive thermal comfort model, it remains unclear which elements are at the core of the innovation. 
-        %     In Section 5, the Sun Path Chart and Psychrometric Chart are presented as illustrative examples, but it should be acknowledged that other tools can also fulfil these functions.
+            \response{Here our reply}
+            \item It is stated in the abstract that `` The ISO7730:2005 and ASHRAE55:2023 are the most widely referenced thermal comfort standards worldwide, and their different PMV formulations are a source of confusion.`` However, it does not actually clarify what kind of confusion has been caused.
 
-        %     \response{
-        %     We appreciate the reviewer's thoughtful feedback and would like to clarify the key features that distinguish CBE Clima Tool as an innovative solution for climate analysis tailored to sustainable building design.
-        %     The core innovation of Clima extends beyond the calculation of thermal comfort indices, a capability shared by several cloud-based and desktop applications. 
-        %     What sets Clima apart is its commitment to universal accessibility. 
-        %     It empowers users worldwide with the ability to swiftly and reliably analyze and visualize intricate climate data.
-        %     While other tools may offer similar graphical outputs, Clima stands out due to its exceptional ease of use, interactivity, and the superior quality of its graphical representations. 
-        %     These features are invaluable to architects, engineers, students, and educators in effectively conveying climate-adapted building design concepts.
-        %     Clima's web-based nature eliminates the need for local installations, ensuring cross-platform compatibility and seamless accessibility. 
-        %     Furthermore, a fundamental innovation lies in its status as a free and open-source tool. 
-        %     This transparency allows users to scrutinize and validate the underlying algorithms and calculations, fostering trust and confidence in the tool's capabilities.
-        %     Our commitment to an open-source approach also encourages community-based development, facilitating collaborative enhancements and guaranteeing that Clima evolves through input from a diverse array of contributors.
-        %     Additionally, it grants users the opportunity to incorporate Clima's code into their own applications, further extending its impact within the research and engineering communities.
-        %     In summary, the innovation of CBE Clima Tool lies in its global accessibility, user-friendliness, transparency, and its capacity to foster collaborative development, making it a valuable asset for sustainable building design professionals and researchers alike.
-        %     We have summarised this answer in Sub-Section 2.4 of the revised manuscript.
-        %     }
+            \response{Here our reply}
 
-        %     \item Furthermore, the paper currently reads more like a report.
-To align with the style of an academic journal paper, it is recommended that the authors revise certain sections.
-For instance, the introduction could be enhanced by breaking down the lengthy one-page paragraph into several well-structured paragraphs.
+            \item In lines 32--34, ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.`` In the calculation of PMVCE, the metabolic rate and clothing thermal resistance will also be corrected.
+            So why does it claim that there are differences in the calculation results of the two models only when the ``V`` is greater than 0.1?
+Besides, does the ``V`` that the author wants to express refer to the original measured value or the adjusted V value?
 
-        %     \response{
-        %     We would like to express our gratitude to the reviewer for their constructive feedback and recommendations regarding the structure of our manuscript. 
-        %     We acknowledge the importance of aligning the manuscript with the style and standards of an academic journal paper.
-        %     Based on your suggestion, we have thoroughly revised the Introduction chapter. 
-        %     Specifically, we have broken down the lengthy one-page paragraph into several well-structured paragraphs to enhance readability and clarity. 
-        %     This restructuring allows for a more systematic presentation of the key concepts and objectives of our work.
-        %     The revised Introduction now provides a more concise and focused overview of the background, motivation, and significance of CBE Clima Tool. 
-        %     We have carefully organized the content to guide readers through the rationale behind our research, the challenges addressed by the tool, and its potential impact on the field of sustainable building design.
-        %     We believe that these improvements will assist readers in comprehending the context and importance of our work.
-        %     }
+            \response{Here our reply}
 
-        % \end{enumerate}
+            \item In the description of limitations on line 50, the first item does not seem to be a limitation of the model itself.
+            It is recommended to include it in the statement of the research purpose and significance.
 
-% Closing statement
+            \response{Here our reply}
+
+            \item In the data selection, there is no selection of environmental types.
+            If the aim is to test the accuracy of the PMV model, were data from naturally ventilated environments, outdoor environments, etc.\ also selected?
+
+            \response{Here our reply}
+
+            \item In the data selection in lines 168--171, it has been stated that ``the PMV should only be used when its absolute value is lower than 2``.
+            So why were data with |PMV| $\leq$ 3.5 retained?
+            Regarding the retention of TSV data, it is said that ``thermal sensation was measured with at least a seven-point scale``, which indicates that there are different scales.
+            When the scale sizes are different, the same number represents different meanings.
+            Is it reasonable to retain data in the form of |TSV| $\leq$ 3?
+
+            \response{Here our reply}
+
+            \item There are problems in the analysis of Figure 5i.
+            The figure shows that subjects with slightly warm and slightly cold thermal sensations have the intention to further change the environment, but this does not mean that they are in an uncomfortable state.
+            Even when the thermal sensation is neutral, people with different thermal preferences may still have expectations for environmental changes.
+            The expectation of changing the environment and the evaluation of the current environment should not be confused.
+            It is recommended to re-analyze after clarifying the meanings of thermal comfort, thermal sensation, and thermal preference.
+
+            \response{Here our reply}
+
+            \item It should be noted that the PMV model proposed by Fanger is applicable to steady-state indoor thermal environments that do not deviate too much from neutrality.
+            Therefore, it is unreasonable to use overheated or overcooled data to verify the accuracy of the PMV model, and the demonstrated accuracy will also be relatively low.
+
+            \response{Here our reply}
+
+            \item For highlights 4--5, whether in field studies or laboratory studies, PMV values generally ranging from -1 to 1 are defined as thermal comfort, which is appropriate for steady-state indoor environments close to neutrality.
+            However, if it is narrowed down to -0.5 to 0.5, the thermal comfort range is greatly reduced, which will increase the building operation energy consumption and is not recommended.
+
+            \response{Here our reply}
+
+            \item Some conclusion contents lack data and theoretical support.
+
+            \response{Here our reply}
+
+        \end{enumerate}
+
+        \clearpage
+
+        \textbf{Reviewer 3}
+        The manuscript presents an excellent analysis of the discrepancies between two versions of the PMV model and evaluates their prediction accuracy.
+        The analysis is informative and definitely relevant for researchers and professionals within thermal comfort.
+        The structure of the manuscript and the presentation of the topic are well laid out.
+
+        \begin{enumerate}
+            \item I have a only a specific few comments, which are listed below.
+            However, my main concern deals with the way the records representing individuals in the comfort DB are used to first calculate PMV for individuals and then compare with individual TSVs.
+            As the authors write ``The intended aim of the PMV model is not to accurately predict each individual thermal response from participants.
+            The PMV model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.`` But isn't this analysis relying on calculation of each individual TSV without considering that this is not what the model should be used for?
+            The approach is the same across a wealth of studies that examine the accuracy of the model, but that does not make it more correct.
+            I actually doubt that the DB can be reliably used to estimate the prediction accuracy, as there is no information of the mean TSV of groups of people exposed to the same thermal environment.
+
+            \response{Here our response}
+
+            \item Although less critical, another general comment is that it is not clear what is meant by low predicition accuracy, as is written already in the abstract - when is the prediction accurqacy low?
+            We aim to estimate what people perceive, which seriously suffers from individual differences and many others challenges.
+            What is the prediction accuracy of related models?
+            Or is the low prediction accuracy an outcome of using the prediction model wrongly (see first general comment).
+
+            \response{Here our response}
+
+            \item Ln. 103 - what is meant by formulation error?
+            Unsuitable equations or constants that are wrong?
+
+            \response{Here our response}
+
+            \item Sources of error.
+            Order of importance - is it really the heat transfer calculation that is the most important - not the attempt to calculate thermal sensation based on physical variables
+
+            \response{Here our response}
+
+            \item ``Additionally, the PMV model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
+            In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature [24].`` Indeeed, the thermoregulatory system maintains a stable body temperature, but the body still exchanges heat with the environment, so suggest to revise this sentence.
+
+            \response{Here our response}
+
+            \item P 23 ln 416.
+            Calculation of heat loss by sweating.
+            Actually, this equation applies to people in thermal comfort, which should probably be added to the statement, so heat lost due to sweating does not appear to be related only with met rate.
+
+            \response{Here our response}
+
+            \item ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent a TSV vote.`` This is wrong - the PMV was never meant to represent a TSV vote, but the average of TSV votes of a large group of people.
+
+            \response{Here our response}
+
+            \item Comment on measurement of air speed.
+            ISO7726, which is used as a guide for measurements in ISO7730, recommends measurement of air speed only at abdomen level in homogeneous environments in both class C (comfort) and S (stress).
+            The air speed measured in this height should be used to calculate the PMV\@.
+            When comparing 7730 and 55, especially with emphasis on air speed, this should probably be taken into account.
+
+            \response{Here our response}
+
+        \end{enumerate}
+
         Kind regards,

### 2025-01-10 Manuscript/src/response-letter.tex
rebuttal: answered some comments

@@
 
+\usepackage[textsize=scriptsize]{todonotes}
+
 % Name of sender
@@
 \newcommand{\response}[1]{\textcolor{blue}{\textbf{Author response:} #1}}
-\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO}: #1}}
 
@@
 
+            \todo[color=yellow]{Federico - check what the PMV model uses for clothing insulation as input and change the text accordingly}
             \response{
@@
 
+            \todo{Stefano - is the following answer true? Did not you pulish the paper which explains why the CE is subtracted from both the tdb and tr?}
             \response{
-                Here our response
+                Thank you for pointing this issue out, this is clearly a important issue about the PMVCE model.
+                Unfortunately we are not certain why the PMVCE subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
+                In Section 1.1 we have tried to summarise all the knowledge we have on the PMVCE model, in particular we stated explained that while the ASHRAE 55:2023 standard explains how the PMVCE works.
+                There is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
+                Arens et al. (2009) and Yang et al. (2015) provide a partial justification for the PMVCE model, they do not fully explain why these specific changes to the PMV inputs were made.
             }
@@
 
+            \todo[color=yellow]{Federico - add the following sentences and reference them in the comment below}
             \response{
-                Here our response
+                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data.
+                We have added a sentence in the introduction to highlight the importance of the quality of the input data.
+                We have also added a sentence in the limitations section to acknowledge the limitations of the PMV model.
             }
@@
 
+            \todo{Stefano shall we consider adding a sentence about the local discomfort in the paper?
+            Can we say that this is potentially a limitation of the PMV model?}
             \response{
-                Here our response
+                We agree with the reviewer that draughts can be the cause local discomfort and consequently influence the overall thermal comfort.
+                However, the PMV model does not estimate local discomfort and can be used to predict thermal sensation in a space with air speeds up to 0.2 m/s.
+                This, with other factors, can lead to a discrepancy between the PMV and the TSV.
             }
@@
 
-            \response{Here our reply}
-            \item It is stated in the abstract that `` The ISO7730:2005 and ASHRAE55:2023 are the most widely referenced thermal comfort standards worldwide, and their different PMV formulations are a source of confusion.`` However, it does not actually clarify what kind of confusion has been caused.
+            \todo{Stefano - I cannot fully understand the question, can you help me?}
+            \response{This manuscript aims to explore the accuracy of the PMV model implemented in the ISO 7730:2005 and ASHRAE 55:2023 standards.
+            We did this by comparing the results of the PMV model with the thermal sensation votes collected in the ASHRAE Global Thermal Comfort Database II v2.1.
+            In addition to comparing the accuracy of the two PMV models, we also discussed in the paper that the PMV model is inappropriately used by some authors to predict thermal sensation of individuals outside its range of applicability.
+            As a consequence, we recommend limiting the applicability of the PMV model to the range specified in the Table 2 of the manuscript.
+            }
 
-            \response{Here our reply}
+            \item It is stated in the abstract that `` The ISO7730:2005 and ASHRAE55:2023 are the most widely referenced thermal comfort standards worldwide, and their different PMV formulations are a source of confusion.``
+            However, it does not actually clarify what kind of confusion has been caused.
 
-            \item In lines 32--34, ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.`` In the calculation of PMVCE, the metabolic rate and clothing thermal resistance will also be corrected.
+            \todo[color=yellow]{Federico - add the following sentences and reference it in the comment below}
+            \response{Thank you for pointing this out.
+            We should have explained this better in the manuscript.
+            The confusion arises from the fact that the two standards use different PMV formulations and the results of the two models can differ.
+            Consequently, for a given thermal environment, a user may obtain different results depending on the standard used.
+            This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.
+            We have added a sentence in the introduction to clarify this point.
+            }
+
+            \item In lines 32--34, ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.``
+            In the calculation of PMVCE, the metabolic rate and clothing thermal resistance will also be corrected.
             So why does it claim that there are differences in the calculation results of the two models only when the ``V`` is greater than 0.1?
@@
 
-            \response{Here our reply}
+            \todo[color=yellow]{Federico - add the following sentences and reference it in the comment below}
+            \response{
+            Please let us clarify this point.
+            1) Neither the PMV nor the PMVCE model correct the metabolic rate.
+            2) Both the PMV and PMVCE models correct the clothing insulation based on the metabolic rate.
+            3) Both the PMV and PMVCE models correct the air speed based on the metabolic rate using the same equation.
+                The value of V refers to the adjusted air speed.
+                We have spedified this in the nomeclature section of the manuscript.
+                However, we strongly agree with the reviewer that this point is not clear and this issue arises from how the ASHRAE 55:2023 standard is written.
+                The authors of the ASHRAE 55:2023 standard voted not to have two variables one for the measured air speed and one for the adjusted air speed.
+                We have added a sentence in the manuscript to raise this issue about the wording of the ASHRAE 55:2023 standard.
+            }
 

### 2025-01-13 Manuscript/src/response-letter.tex
rebuttal: partially answered all comments

@@
 }
+
+% package to strike out text
+\usepackage[normalem]{ulem}
 
@@
                 Unfortunately we are not certain why the PMVCE subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
-                In Section 1.1 we have tried to summarise all the knowledge we have on the PMVCE model, in particular we stated explained that while the ASHRAE 55:2023 standard explains how the PMVCE works.
-                There is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
+                In Section 1.1 we have tried to summarise all the publicly available information on the PMVCE model.
+                While the ASHRAE 55:2023 standard explains how the PMVCE works, there is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
                 Arens et al. (2009) and Yang et al. (2015) provide a partial justification for the PMVCE model, they do not fully explain why these specific changes to the PMV inputs were made.
+                Moreover, we were not able to find any other peer-reviewed publication that explains in detail why the PMVCE model subtracts the cooling effect from both the air and mean radiant temperature.
             }
@@
             \response{
-                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data.
+                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data and that the ASHRAE Global Thermal Comfort Database II v2.1 has limitations.
                 We have added a sentence in the introduction to highlight the importance of the quality of the input data.
                 We have also added a sentence in the limitations section to acknowledge the limitations of the PMV model.
+                However, despite the fact that this database has some limitations, we decided to use it because, to the best of our knowledge, the ASHRAE Global Thermal Comfort Database II v2.1 is the most comprehensive database available for thermal comfort research.
+                The database is widely used in the field, and only includes data published in peer-reviewed studies.
+                Moreover we belive that errors in the inputs data, if not systematic, should be mitigated by the large number of data points in the database.
             }
 
-            \item The database used for this investigation also includes high air velocity measurements (v>0.2 m/s).
+            \item The database used for this investigation also includes high air velocity measurements (v$>$0.2 m/s).
             However, this does not rule out the possibility of local discomfort caused by draughts.
@@
             Can we say that this is potentially a limitation of the PMV model?}
+            \todo[color=yellow]{Federico - add the following sentences and reference them in the comment below}
             \response{
-                We agree with the reviewer that draughts can be the cause local discomfort and consequently influence the overall thermal comfort.
-                However, the PMV model does not estimate local discomfort and can be used to predict thermal sensation in a space with air speeds up to 0.2 m/s.
-                This, with other factors, can lead to a discrepancy between the PMV and the TSV.
+                We agree with the reviewer that draughts can be the cause local discomfort and consequently influence the overall thermal comfort sensation.
+                This is a limitation of the PMV model, as it does not account for local discomfort.
+                Dspite the fact that the PMV model does not estimate local discomfort, both standards state that it can be used to predict thermal sensation in a space with air speeds up to 1.0 m/s.
+                Hence, we decided to test the PMV model in the range of air speeds up to 1.0 m/s.
+                This, with other factors, can lead to a discrepancy observed between the PMV and the TSV.
+                Interestingly, we observed that both PMV models have a higher accuracy and lower bias in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s.
+                This would suggest that while local discomfort due to air movement may affect the overall predicion of the PMV model, it did not affect the results of our analysis.
+                We decided to include a sentence in the manuscipt to highlight the role that local discomfort can play in the overall thermal comfort sensation and the fact that the PMV model does not account for it.
             }
@@
             \todo{Stefano - I cannot fully understand the question, can you help me?}
-            \response{This manuscript aims to explore the accuracy of the PMV model implemented in the ISO 7730:2005 and ASHRAE 55:2023 standards.
+            \response{This manuscript aims to determine and compare the accuracy of the PMV models implemented in the ISO 7730:2005 and ASHRAE 55:2023 standards.
             We did this by comparing the results of the PMV model with the thermal sensation votes collected in the ASHRAE Global Thermal Comfort Database II v2.1.
@@
 
-            \item It is stated in the abstract that `` The ISO7730:2005 and ASHRAE55:2023 are the most widely referenced thermal comfort standards worldwide, and their different PMV formulations are a source of confusion.``
+            \item It is stated in the abstract that ``The ISO7730:2005 and ASHRAE55:2023 are the most widely referenced thermal comfort standards worldwide, and their different PMV formulations are a source of confusion.''
             However, it does not actually clarify what kind of confusion has been caused.
@@
             We should have explained this better in the manuscript.
-            The confusion arises from the fact that the two standards use different PMV formulations and the results of the two models can differ.
-            Consequently, for a given thermal environment, a user may obtain different results depending on the standard used.
+            The confusion arises from the fact that the two standards use different PMV formulations and the results of the two models are different for the same thermal environment when the air speed is higher than 0.1 m/s.
             This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.
@@
 
-            \item In lines 32--34, ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.``
+            \item In lines 32--34, ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.
             In the calculation of PMVCE, the metabolic rate and clothing thermal resistance will also be corrected.
-            So why does it claim that there are differences in the calculation results of the two models only when the ``V`` is greater than 0.1?
-            Besides, does the ``V`` that the author wants to express refer to the original measured value or the adjusted V value?
+            So why does it claim that there are differences in the calculation results of the two models only when the ``V'' is greater than 0.1?
+            Besides, does the ``V'' that the author wants to express refer to the original measured value or the adjusted V value?
 
             \todo[color=yellow]{Federico - add the following sentences and reference it in the comment below}
+            \todo{Stefano - Shall we consider changing how we refer to the air speed and relative air speed in the manuscript?}
+            \todo[color=red]{Federico - check point 2 since we may decide to change the dynamic clothing insulation for the PMV model}
+            \todo[color=red]{Federico - check how we refer to the air speed in the manuscript}
             \response{
-            Please let us clarify this point.
+                Please let us clarify this point by answering the questions raised by the reviewr separately:
                 1) Neither the PMV nor the PMVCE model correct the metabolic rate.
-            2) Both the PMV and PMVCE models correct the clothing insulation based on the metabolic rate.
+                Both models use as an input the metabolic rate of the participants which is wither measured or estimated by the researcher.
+                2) Both the PMV and PMVCE models correct the clothing insulation based on the metabolic rate using the same equation.
                 3) Both the PMV and PMVCE models correct the air speed based on the metabolic rate using the same equation.
-                The value of V refers to the adjusted air speed.
+                The value of V in the paper refers to the adjusted air speed.
                 We have spedified this in the nomeclature section of the manuscript.
@@
 
-            \response{Here our reply}
+            \response{
+                We completely agree with the reviewer that the first limitation is not a limitation of the model itself.
+                Hence, we have removed it from the limitations section and moved it to the beginning of Section 1.1.
+            }
 
@@
 
-            \response{Here our reply}
+            \response{
+                We made the decision to include all the data in the ASHRAE Global Thermal Comfort Database II v2.1 in the analysis regardless of the environmental type.
+                This is because the PMV model, being a heat balance model, can be used to predict thermal sensation in a wide range of environments.
+                Our aims was to determine the accuracy of the PMV model in predicting thermal sensation as a function of the input parameters.
+                Environmental type is not an input parameter of the PMV model, hence we did not consider it in the analysis.
+                We are aware that previous studies have shown that the PMV model may have limitations in predicting thermal sensation in non mechanically ventilated environments.
+                However, since both standards do not specify the type of environments for which the PMV model can be used, we decided to include all the data in the ASHRAE Global Thermal Comfort Database II v2.1 in the analysis.
+                It should be noted that we did not include data from outdoor environments in the analysis since the ASHRAE Global Thermal Comfort Database II v2.1 only includes data from ``real'' buildings occupied by ``real'' people doing their normal day-to-day activities.
+                The database does not include data from outdoor environments nor from climate chambers studies.
+            }
 
@@
             So why were data with |PMV| $\leq$ 3.5 retained?
-            Regarding the retention of TSV data, it is said that ``thermal sensation was measured with at least a seven-point scale``, which indicates that there are different scales.
+
+            \response{
+            We retained data with |PMV| $\leq$ 3.5 because the PMV model is scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
+            Moreover, as explained in the paper the thermal snesation votes in the ASHRAE Global Thermal Comfort Database II v2.1 are collected using a 7-point scale which ranges from -3 to +3.
+            Consequently, we decided to retain data with |PMV| $\leq$ 3.5 and |TSV| $\leq$ 3 in the analysis to compare the accuracy of the PMV model in predicting thermal sensation over the full range of the thermal sensation scale.
+            However, since the PMV model is not recommended to be used when its absolute value is higher than 2, we strived always to presented the accuracy of the model within its range of applicability.
+            For example, in Figure 6 we present the accuracy of the PMV model in predicting each thermal sensation category separately.
+            In Table 1, we present the F1 score of the PMV model in predicting thermal sensation when its absolute value is lower than 1.5.
+            We acknowledge that we did not report the bias of the PMV model in predicting thermal sensation when its absolute value is higher than 2.
+            We may include this figure if the reviewers believe it is necessary.
+            }
+
+            \item Regarding the retention of TSV data, it is said that ``thermal sensation was measured with at least a seven-point scale'', which indicates that there are different scales.
             When the scale sizes are different, the same number represents different meanings.
@@
 
-            \response{Here our reply}
+            \response{
+            Thank you for pointing this out.
+            We made a mistake in the body manuscript and we should have written that the ``thermal sensation was measured with \sout{at least} a seven-point scale''.
+            The peer-reviewed paper which describe the ASHRAE Global Thermal Comfort Database II v2.1 states that the thermal sensation votes were collected using a seven-point scale and coded as follows ``-3 cold, -2 cool, -1 slightly cool, 0 neutral, +1 slightly warm, +2 warm, +3 hot''.
+            We have corrected this mistake in the manuscript.
+            }
 
@@
 
+            \todo{Stefano - Let's discuss this point during the meeting}
             \response{Here our reply}
@@
 
-            \response{Here our reply}
+            \response{
+            We agree with the reviewer that the PMV model is applicable to steady-state indoor thermal environments that do not deviate too much from neutrality.
+            However, the PMV model is widely used in the scientific literature to predict thermal sensation in a wide range of thermal environments.
+            While the ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
+            Consequently, as preiously explained we decided to include all the data in the ASHRAE Global Thermal Comfort Database II v2.1 in the analysis.
+            We, however, tried to further analyze the accuracy of the PMV model in predicting thermal sensation within its range of applicability, for example in Table 1 where we present the F1 score of the PMV model in predicting thermal sensation when its absolute value is lower than 1.5.
+            }
 
@@
 
-            \response{Here our reply}
+            \response{
+            Thank you for pointing out this potential issue.
+            We agree that limiting the applicability of the PMV model to the range -0.5 to 0.5 may in some instances increase the building operation energy consumption.
+            However, we also show that if the building is correctly designed and operated, for example by providing individual control over the air speed, this may not be the case.
+            For example, in Figure 2 of the manuscript we show that the PMV predicts that occupants (met=1.2 and clo=0.5) are comfortable at t=27.7C when the air speed is 0.4 m/s.
+            An indoor air temperature of 27.7C is higher than the average indoor air temperature in offices in the US, Europe, AU and Singapore in which is generally in the range of 22-24C.
+            Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated and participants do not have individual control over the indoor environment.
+            This, however, is a discussion that goes beyond the scope of this paper.
+            It should also be noted that we recommend limiting the applicability of the PMV model to the range -0.5 to 0.5 only because our results show that the PMV model has a low accuracy in predicting thermal sensation outside this range.
+            }
 
@@
 
+            \todo{Stefano - Is the reviewer referring to the section about the assumption that people who are ‘slightly warm’ or
+            ‘slightly cool’ are thermally comfortable is incorrect?}
             \response{Here our reply}
@@
             As the authors write ``The intended aim of the PMV model is not to accurately predict each individual thermal response from participants.
-            The PMV model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.`` But isn't this analysis relying on calculation of each individual TSV without considering that this is not what the model should be used for?
+            The PMV model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.``
+            But isn't this analysis relying on calculation of each individual TSV without considering that this is not what the model should be used for?
             The approach is the same across a wealth of studies that examine the accuracy of the model, but that does not make it more correct.
@@
 
+            \todo{Stefano - Would you like to answer this question? I have the same concern as the reviewer but I know that you have a different opinion}
             \response{Here our response}
 
-            \item Although less critical, another general comment is that it is not clear what is meant by low predicition accuracy, as is written already in the abstract - when is the prediction accurqacy low?
+            \item Although less critical, another general comment is that it is not clear what is meant by low predicition accuracy, as is written already in the abstract - when is the prediction accuracy low?
             We aim to estimate what people perceive, which seriously suffers from individual differences and many others challenges.
@@
 
-            \response{Here our response}
+            \todo[color=yellow]{Federico - add a few more additional sentences about the bias in the manuscript}
+            \response{
+            As previously explained we acknowledge that the PMV model is not intended to predict the thermal sensation of each individual participant.
+            We have acknowledged this limitation in the manuscript and discussed it in Section 3.2.
+            To compensate for this limitation of our analysis, we calculated the bias of the model as previously done by Humphreys and Nicol (2002) to determine if the model is systematically over or under predicting the thermal sensation of the participants.
+            Individual differences in human subjects can be assumed to be random and distributed around the mean.
+            This is the same assumption that PMV model also does when trying to predict the average thermal sensation of a large group of occupants.
+            Consequently, we believe that calcualting the bias of the model is a valid approach to determine the overall accuracy of the model.
+            }
 
@@
 
-            \response{Here our response}
+            \response{
+            We apologize for the confusion.
+            We meant both the formulation of the PMV model and the constants used in the model.
+            We have added a sentence in the manuscript to clarify this point.
+            }
 
@@
 
+            \todo{Stefano - do you understand this comment?}
             \response{Here our response}
@@
             \item ``Additionally, the PMV model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
-            In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature [24].`` Indeeed, the thermoregulatory system maintains a stable body temperature, but the body still exchanges heat with the environment, so suggest to revise this sentence.
+            In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature [24].``
+            Indeeed, the thermoregulatory system maintains a stable body temperature, but the body still exchanges heat with the environment, so suggest to revise this sentence.
 
-            \response{Here our response}
+            \response{
+            We apologize for the confusion.
+            We have revised the sentence to clarify this point.
+            }
 
@@
 
+            \todo{Stefano - I do not understand this comment, can you help me?}
             \response{Here our response}
@@
 
-            \response{Here our response}
+            \response{
+            Thank you for pointing this out.
+            We changed the sentence as follows: ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent the average thermal sensation of a large group of occupants.''
+            }
 
@@
 
-            \response{Here our response}
+            \todo{Stefano - How shall we address this comment?}
+            \response{
+            Thank you for pointing this out.
+            Most of the studies in the ASHRAE Global Thermal Comfort Database II v2.1 measured air speed one single height.
+            When we specifically compared the the PMV and PMVCE models in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s and for those studies that measured air speed at three different heights, we used the same input data for both models since also the PMV ISO specifies that in non homogeneous environments the air speed should be measured at different heights.
+            }
 

### 2025-01-13 Manuscript/src/sections/introduction.tex
rebuttal: partially answered all comments

@@
 The \gls{55} standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
-The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented.
+The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented, nor for example, why the \ac{ce} is subtracted from the \ac{tdb} and \ac{tr}.
 A partial justification of the model is mainly described in \mycite{arens_moving_2009} and secondarily in \mycite{yang_cooling_2015}.
+However, we could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the \gls{55} over the original \ac{pmv} model.
+
 The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
@@
 \begin{enumerate}
-    \item We could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the ASHRAE standard over the original \ac{pmv} model.
     \todo[inline]{true enough, but in the standard the objective is to have a rational physiology model. 
@@
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
-Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
+Humphreys and Nicol estimated the effects of measurement and formulation error (i.e., incorrect equations or constants) on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
 They used the ASHRAE Global Thermal Comfort database~I and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.

### 2025-01-13 Manuscript/src/sections/methodology.tex
rebuttal: partially answered all comments

@@
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with at least a seven-point scale and the \ac{pmv} has no upper or lower boundary, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
+However, since the thermal sensation was measured with a seven-point scale and the \ac{pmv} has no upper or lower boundary, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
 

### 2025-01-13 Manuscript/src/sections/results.tex
feat: started adding reviewers comments to response letter.
rebuttal: partially answered all comments

@@
         \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
-        \caption{Figure a) shows the percentage of \ac{tpv} for each thermal sensation vote.
+        \caption{Figure a) shows the percentage of \ac{tpv} for each \ac{tsv}.
     The numbers on the right side of each bar show the number of points for each \ac{tsv}.
@@
     \caption{
-        Figure c) shows the percentage of participants who consider the space thermally acceptable for each thermal sensation vote.}
+        Figure c) shows the percentage of participants who consider the space thermally acceptable for each \ac{tsv}.}
         \label{fig:thermal_acceptability_by_ts}
@@
 The number of data points in each \ac{tsv} bin is different.
-\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
+\mycite{Cheung2019} grouped people based on the same heat loss or gains (same \ac{pmv} ranges), not the same \ac{tsv} as we did here, and then quantified the sensitivity of the model, i.e., how many times the \ac{pmv} model correctly predicted \ac{tsv}.
 They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.
@@
     \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
-    The x-axis shows the discrepancy (bias) between the predicted and self-reported thermal sensation vote.
+    The x-axis shows the discrepancy (bias) between the predicted and self-reported \ac{tsv}.
     We have binned the data into intervals of equal width.
@@
 The issues are reported in order of importance of what we believe may affect the accuracy of the model.
+\todo{Note that the section below does not mention the two-node model, the CE implementation of PMV, or the physiological differences between the ISO and ASHRAE models that could explain the observed differences or ‘error’.
+}
 
@@
 It also ignores vasoconstriction and vasodilation.
-Additionally, the \ac{pmv} model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
-In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature~\cite{romanovsky_thermoregulation_2018}.
+Additionally, the \ac{pmv} model erroneously assumes that the human body is not capable of maintaining a stable core temperature.
+A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicate that the hypothetical cylinder is either gaining or losing heat and consequently it is getting warmer or colder, respectively.
 Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer; however, this never precisely occurs.
-The human thermoregulatory system, in healthy humans, is constantly actively engaged to ensure a stable core temperature.
+The human thermoregulatory system, in healthy humans, is constantly actively engaged to ensure a stable core temperature~\cite{romanovsky_thermoregulation_2018}.
 The \ac{pmv} model also assumes that the human body is constantly losing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
@@
 
-The \ac{pmv} model uses the overall heat losses or gains to calculate a \ac{pmv} value which should represent a \ac{tsv} vote.
+The \ac{pmv} model uses the overall heat losses or gains to calculate a \ac{pmv} value which should represent the average \ac{tsv} of a large group of occupants.
 Heat losses or gains are only a proxy for the thermoregulation effort that the human body has to undergo to maintain a constant internal temperature.
 Fanger created this equation based on a limited dataset collected in a laboratory, and this equation was never updated or revisited~\cite{Fanger1970}.
-In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions~\cite{romanovsky_thermoregulation_2018}.
+In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a \ac{tsv}, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions~\cite{romanovsky_thermoregulation_2018}.
 

### 2025-01-13 main.py

@@ -18,7 +18,6 @@ from pythermalcomfort.psychrometrics import p_sat_torr
 from pythermalcomfort.utilities import check_standard_compliance_array
 from scipy import optimize
 from scipy import stats
-from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score
 from sklearn.metrics import f1_score
 from sklearn.metrics import r2_score, mean_absolute_error

### 2025-01-14 Manuscript/src/response-letter.tex
rebuttal: replied to first comment

@@
 
-            \todo[color=yellow]{Federico - check what the PMV model uses for clothing insulation as input and change the text accordingly}
             \response{
-                Here our response
+                Thank you for pointing this out.
+            While it is correct that the total clothing insulation accounts for the air insulation and the clothing area factor as per ISO 9920.
+            Both the PMV ISO and PMV CE models use the basic clothing insulation (Icl) as an input parameter.
+            ISO 7730:2005 -- please refer to page 3 and 4 where the standard states that Icl is the input to the PMV equation shown in Equation 1--4.
+            ASHRAE 55:2023 -- please refer to the computer code in appendix B and Section 5.2.2
+            Consequently we did not change the sentence in the manuscript.
             }

### 2025-01-16 Manuscript/src/response-letter.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

--- 2025-01-13 Manuscript/src/response-letter.tex
@@
 
-            \todo[color=yellow]{Federico - check what the PMV model uses for clothing insulation as input and change the text accordingly}
             \response{
-                Here our response
+                Thank you for pointing this out.
+                There was a typo in the manuscript in which we stated that $I_{cl}$ is the total clothing insulation.
+                This is incorrect and we now haved fixed this typo since both the PMV ISO and PMV CE models use the basic clothing insulation ($I_{cl}$) as an input parameter.
             }

### 2025-01-16 Manuscript/src/myacronyms.tex
feat: code to analyse data from studies with three air speed measurements
feat: updated the results using v_r instead of v
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

@@
     \acro{rh}[$RH$]{relative humidity\acroextra{, \%}}
-    \acro{v}[$V$]{average air speed\acroextra{, m/s}}
+    \acro{v}[$V$]{measured air speed\acroextra{, m/s}}
+    \acro{vr}[$V_{r}$]{relative air speed\acroextra{, m/s}}
     \acro{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, $^{\circ}$C}}
-    \acro{clo}[$I_{cl}$]{total clothing insulation\acroextra{, clo}}
-    \acro{i-cl}[$i_{cl}$]{permeation efficiency of water vapor through the clothing layer}
+    \acro{clo}[$I_{cl}$]{intrinsic clothing insulation from the skin to the outer surface under reference conditions\acroextra{, clo}}
+    \acro{clor}[$I_{cl,r}$]{actual intrinsic clothing insulation under given environmental and activities\acroextra{, clo}}
     \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
@@
 
-        \acro{i-cl}[$i_{cl}$]{permeation efficiency of water vapor through the clothing layer}
     \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
@@
 
-    \acro{7730}[ISO~7730:2005]{ISO~7730:2005 is a thermal comfort standard developed by ISO}
-    \acro{55}[ASHRAE~55:2023]{ASHRAE~55:2023 is a thermal comfort standard developed by ANSI and ASHRAE}
     \acro{pmv-ce}[PMV\textsubscript{CE}]{PMV calculated in accordance with the ASHRAE~55:2023}

### 2025-01-16 Manuscript/src/sections/appendix.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

@@
 In this section, we show the results of the analysis of the dataset with air speed measurements at three heights.
-Only \num{3063} entries in the \ac{db2} had air speed measurements at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+Only \num{3063} entries in the \ac{db2} had air speed measurements at three heights with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
 A summary of the papers that collected these data is provided in Table~\ref{tab:three_heights_papers}.
@@
 \end{figure}
-The adjustment of \ac{v} based on the \ac{met} significantly increased the base value of the air speed measured by all the studies.
-With the value of \ac{v} being \var{delta_v_mean} higher than the average of the other air speeds combined.
+The adjustment of \ac{vr} based on the \ac{met} significantly increased the base value of the air speed measured by all the studies.
+With the value of \ac{vr} being \var{delta_v_mean} higher than the average of the other air speeds combined.
 The discussion on whether this increase is justified is beyond the scope of this paper, since both standards require the user to adjust the air speed based on the \ac{met} value.
@@
     \caption{Comparison between the \ac{pmv} and \ac{pmv-ce} models for the dataset with air speed measured at three heights. 
-    The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data for the dataset with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and measured at three heights.
+    The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data for the dataset with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and measured at three heights.
     Raw data were then binned and a bubble chart (circle area is proportional to the number of votes in that bin) was superimposed over the regression curve to aid the visualization of a large dataset.
@@
 
-\subsection{Dataset with \ac{v} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}}\label{subsec:dataset-with-v-measured-at-three-heights-and-tdb-geq-24-celsius}
-We are showing here the results for the subset of data with \ac{v} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}.
+\subsection{Dataset with \ac{vr} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}}\label{subsec:dataset-with-v-measured-at-three-heights-and-tdb-geq-24-celsius}
+We are showing here the results for the subset of data with \ac{vr} measured at three heights and \ac{tdb}~$\geq$~\qty{24}{\celsius}.
 We are reporting these results since the \ac{pmv-ce} aims to more accurately predict the thermal sensation, compared to the \ac{pmv}, when air movement is used to cool the occupants.

### 2025-01-16 Manuscript/src/sections/conclusions.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

@@
 This means that the models are misleading the users.
-When only comparing the results of the two models for entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
+When only comparing the results of the two models for entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
 For example, the median bias of the \ac{pmv} model is \var{bias_median_pmv_0.2} while the median bias for the \ac{pmv-ce} is \var{bias_median_pmv_ce_0.2}.

### 2025-01-16 Manuscript/src/sections/introduction.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

@@
 \section{Introduction}\label{sec:introduction}
-
-\todo[inline]{General notes:  
-The objective of the ASHRAE Std 55 approach is to substitute the Pierce two-node model for the PMV model for modeling heat loss under elevated air speeds in warm conditions.  
-The CE method is intended to provide a PMV sensation prediction from the more physiologically comprehensive two-node model.  
-The work on this paper has raised a number of interesting future research topics that we should discuss.}
-\todo[inline, color=blue!40]{
-FT: I agree that it would be ideal if Std 55 replaced the PMV with a more accurate model, but in its current form, Std 55 2023 still fundamentally uses the PMV model to calculate the thermal sensation.
-The PMV CE is a modified version of the PMV, which is not more physiologically comprehensive than the PMV model since as shown in the flow chart the PMV CE uses the same code as the PMV to calculate the thermal sensation.
-Only temperature is adjusted to account for elevated air movement but the other environmental or personal parameters are not adjusted.
-Moreover, as highlighted in Section 1.1 the use of the CE introduces several errors.
-}
-
 \mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} standard~\cite{iso7730} in its original form.
@@
 
-\todo[inline]{General notes:  
+\todo[inline]{Ed's comment:
 The PMV model consists of a heat balance model and an empirical equation that predicts thermal sensation (PMV) from heat balance.
@@
 
-The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
+Figure~\ref{fig:flowchart_pmv_calculation} shows side by side the calculation routines for the \ac{pmv} and \ac{pmv-ce} models.
 \begin{figure}[!htb]
+    \begin{subfigure}[b]{\textwidth}
         \centering
@@
             \node (dec1) [decision, below of=start, yshift=-.25cm] {\acs{met} $> \qty{1}{met}$ };
-        \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\acs{v} = \acs{v}$+ 0.3 ($\acs{met}$-1)$};
+            \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\acs{vr} = \acs{v}$+ 0.3 ($\acs{met}$-1)$};
             \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\acs{met} $> \qty{1.2}{met}$ };
-        \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\acs{clo} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$};
-        \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\acs{v} $> \qty{0.1}{\m\per\s}$ };
-        \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\acs{pmv}(\acs{tdb}, \acs{tr}, \acs{rh}, \acs{v}, \acs{met}, \acs{clo})};
+            \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\acs{clor} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$};
+            \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\acs{vr} $> \qty{0.1}{\m\per\s}$ };
+            \node (end1) [startstop, below of=dec3, yshift=-1.75cm, xshift=-4cm] {\acs{pmv}(\acs{tdb}, \acs{tr}, \acs{rh}, \acs{vr}, \acs{met}, \acs{clor})};
             \node (pro3) [process, below of=dec3, yshift=-.25cm, xshift=3cm] {Calculate \acs{ce}};
-        \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\acs{pmv}(\acs{tdb} - \acs{ce}, \acs{tr} - \acs{ce}, \acs{rh}, \acs{v} = \qty{0.1}{\m\per\s}, \acs{met}, \acs{clo})};
+            \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\acs{pmv}(\acs{tdb} - \acs{ce}, \acs{tr} - \acs{ce}, \acs{rh}, \acs{vr} = \qty{0.1}{\m\per\s}, \acs{met}, \acs{clor})};
 
@@
         \label{fig:flowchart_pmv_ce}
+    \end{subfigure}
+    \par\bigskip % force a bit of vertical whitespace
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \begin{tikzpicture}[node distance=1cm and 3cm, transform shape]
+            \node (start) [startstop] {\acs{pmv} calculation as per \gls{7730}};
+            \node (dec1) [decision, below of=start, yshift=-.25cm] {\acs{met} $> \qty{1}{met}$ };
+            \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\acs{vr} = \acs{v}$+ 0.3 ($\acs{met}$-1)$};
+            \node (dec2) [process, below of=dec1, yshift=-0.75cm] {\acs{clor} = $\alpha \times$\acs{clo} };
+            \node (dec3) [process, below of=dec2, yshift=-0.75cm] {\acs{pmv}(\acs{tdb}, \acs{tr}, \acs{rh}, \acs{vr}, \acs{met}, \acs{clor})};
+
+            \draw [arrow] (start) -- (dec1);
+            \draw [arrow] (dec1) -- node[above, pos=0.3] {Yes} (pro1a);
+            \draw [arrow] (dec1) -- node[right, pos=0.3] {No} (dec2);
+            \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
+            \draw [arrow] (dec2) -- (dec3);
+        \end{tikzpicture}
+        \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{7730} standard.
+            $\alpha$ is the correction coefficient for \acs{clo}.
+            Since the \gls{7730} does contain some errors and it does not provide all the necessary information to calculate $\alpha$ we used the equations provided in the \gls{9920} to derive the value of $\alpha$.
+        In accordance with the \gls{7730} and \gls{9920} standards, since we did not have information about the walking speed of the participants, we estimated it using the provided equation.
+        }
+        \label{fig:flowchart_pmv_iso}
+    \end{subfigure}
+    \caption{Flowcharts depicting the steps for the calculation of the \ac{pmv} following the \gls{55} and \gls{7730} standards.}
+    \label{fig:flowchart_pmv_calculation}
 \end{figure}
-In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
-\ac{v} is not the measured value by air speed sensors.
-According to the standard, the measured value should be adjusted using this equation \ac{v} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
-\todo[size=\small]{you might explain that this equation is used in the ISO Standard as well}
-\todo[inline, color=blue!40]{Thank you for the feedback I will also create a flowchart for the PMV ISO and include it in the appendix.}
+In summary, as shown in Figure~\ref{fig:flowchart_pmv_ce} when the \ac{vr} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
+\ac{vr} is different from the \ac{v}.
+According to the standard, the measured value should be adjusted using this equation \ac{vr} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
+The standard also prescribes the use of \ac{clor} and not of \ac{clo} as an input.
+For \ac{met} $> \qty{1.2}{met}$, \acs{clor} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$.
 The value of \ac{ce} is calculated using the \ac{set} equation.
@@
 The resulting values become the new inputs in the \ac{pmv} model.
-Since the \ac{ce} accounts for convective and evaporative heat losses from the person to its environment, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
-The other three input parameters (\ac{clo}, \ac{rh}, \ac{met}) remain unchanged.
-Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
-To illustrate the extent of the differences between the outputs of the two models, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clo}~=~\qty{0.5}{clo}.
+Since the \ac{ce} accounts for convective and evaporative heat losses from the person to its environment, the value of \ac{vr}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
+The other three input parameters (\ac{clor}, \ac{rh}, \ac{met}) remain unchanged.
+% todo this is no longer true and it should be removed or rephrased
+Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
+To illustrate the extent of the differences between the outputs of the two models, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{vr} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
 We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
 \todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. There are plenty of lab studies that have measured results in that region.}
-\todo[inline, color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
+\todo[color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
 \begin{figure}[!htb]
@@
     \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
-    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{v}.
-    We set \ac{met}~=~\qty{1.2}{met}, \ac{clo}~=~\qty{0.5}{clo}.
+    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{vr}.
+    We set \ac{met}~=~\qty{1.2}{met}, \ac{clor}~=~\qty{0.5}{clo}.
     \label{fig:comfort_regios_pmv_pmvce}}
 \end{figure}
-The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
-For \ac{v}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.
+The results show that for \ac{vr}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
+For \ac{vr}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.
 
-The rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds ~\cite{huang_applicability_2014}.
+The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds ~\cite{huang_applicability_2014}.
 The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}.
@@
 The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
-While the \ac{pmv} model has its limitations, our analysis has revealed additional limitations specific to the \ac{pmv-ce} model, as outlined below.
-\todo{This is a confusing dichotomy}
-\todo[color=blue!40]{FT. Is the new sentence better?}
+\todo[inline]{true enough, but in the standard the objective is to have a rational physiology model.
+The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
+\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}
+
+Our analysis has, however, revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.
 
@@
 \begin{enumerate}
-    \todo[inline]{true enough, but in the standard the objective is to have a rational physiology model. 
-    The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
-    \todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}
-    \item Although the \ac{ce} is used to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model.
-    For example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{v}~=~\qty{0.15}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clo} of \qty{0.5}{clo}.
+    \item Although the \ac{ce} claims to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model.
+    For example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.15}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
     According to the Gagge's two-node model, which is used to calculate the \ac{set}, this hypothetical person would have \qty{30}{\percent} of their skin covered with sweat.
@@
     Even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating.
-    Therefore, in this scenario, the \ac{pmv-ce} model also cannot accurately predict the \ac{tsv}.
+    Therefore, in this scenario, where the \ac{pmv-ce} should excell compared to the \ac{pmv}, the \ac{pmv-ce} model still fails to accurately predict the \ac{tsv}.
     \todo[inline]{This is really confusing.
@@
     The \ac{rh} should be re-calculated, potentially assuming the humidity ratio to be constant, to reflect the new \ac{tdb} value.
-    \item It should be noted that over the years, the critical value of \ac{v} has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-    \todo[inline]{(where does the term ‘critical temperature’ come from?
-It is a threshold value and not critical to any predicted outcome.
-The changes over the years were due to converting from historical graphics, objections from diffuser manufacturers concerned about lawsuits, Peter Simmonds, etc.
-None of them constitute a ‘limitation’ to the model}
-    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-    However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
+    \item It should be noted that over the years, the threshold value of \ac{vr} after which the \ac{ce} is calculated, has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
+    However, these changes were not based on any scientific evidence and significantly affect the output of the model.
+    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{vr}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
+    However, while \ac{vr}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
     \todo[inline]{I think this argument is incorrect.
@@
 Several other formulations of the \ac{pmv} model have been proposed.
-Among the most notable formulations there are the \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \todo{explain how this differs from the CE approach since it looks like the same thing} \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
+Among the most notable formulations there are the \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET},  \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
 \mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
+It should be noted that, while both the \ac{pmvs} and \ac{pmvg} models are based on the \ac{set} model they significantly differ from the \ac{pmv-ce} model.
+Both the \ac{pmvs} and \ac{pmvg} are directly calculated within the \ac{set} model using the estimated heat gains and losses from the human body.
+
 Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models since they are incorporated in the most widely referenced thermal comfort standards worldwide (\gls{7730} and \gls{55}).
@@
 Some notable works that tried to determine the accuracy of the \ac{pmv} model include the work from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
-\todo{Doherty and I did not evaluate air movement which hardly anyone was considering back then.}
 They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
@@
 Reporting the classification accuracy of the \ac{pmv} formulations when people are grouped by their \ac{tsv} is particularly important for an unbalanced dataset like the \ac{db2} where most of the participants reported to be `neutral'.
-Finally, \mycite{Yao2022} only reported the overall bias for the whole dataset, even though the two formulations only differ when the \ac{v} exceeds \qty{0.1}{\m\per\s}.
+Finally, \mycite{Yao2022} only reported the overall bias for the whole dataset, even though the two formulations mainly differ when the \ac{vr} exceeds \qty{0.1}{\m\per\s}.
 

### 2025-01-16 Manuscript/src/sections/methodology.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

@@
 This filtered out approximately \var{entries_db_valid} of the data.
-In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
-This is because both \ac{pmv} formulations use the \ac{v} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
+In our analysis, we used the value \ac{vr} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
+This is because both \ac{pmv} formulations use the \ac{vr} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
 
@@
 \num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
-\num{0}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s},
+\num{0}~$\leq$~\ac{vr}~$\leq$~\qty{1}{\m\per\s},
 \num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
@@
 This could introduce a bias in the measurements and, consequently, in the results of the models.
-Moreover, since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
+Moreover, since the two \ac{pmv} formulations only differ when \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
 \begin{enumerate}[ {}1{)} ]
     \item all the data points with \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv};
-    \item the data points in 1) with \ac{v}~$\geq$~\qty{0.2}{\m\per\s};
-    \item the data points in 1) with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
+    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s};
+    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
     These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
@@
 The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
-In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
+In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
 We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.
@@
 However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
-These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{v}).
+These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
 If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.

### 2025-01-16 Manuscript/src/sections/results.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

@@
 Mainly because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
-This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
-The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v}~$\geq$~\qty{0.1}{\m\per\s}.
+This lack of data in the \ac{db2} is particularly relevant for \ac{vr}.
+The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}.
 The median value in the \ac{db2} is \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
-We also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v}~$\geq$~\var{v_95_perc_max}.
+We also could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{vr}~$\geq$~\var{v_95_perc_max}.
 
@@
 The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we compared the predicted value from the two models with measured \ac{tsv}.
-However, since the two models only differ when \ac{v}~$>$~\qty{0.1}{\m\per\s};
-we also present the results for a subset of the data where \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+However, since the two models only differ when \ac{vr}~$>$~\qty{0.1}{\m\per\s};
+we also present the results for a subset of the data where \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
 In this subset of data, the differences in the outputs of the two models are more pronounced.
-We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
+We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{vr} of \var{v_95_perc_max}.
 
@@
     \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants.
-    In a) and b) for the entire dataset, while in c and d for the dataset when \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+    In a) and b) for the entire dataset, while in c and d for the dataset when \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
     In a) and c) are shown the results for the \ac{pmv} model, and in b) and d) for the \ac{pmv-ce} model.
@@
 The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
-The overall accuracy of both \ac{pmv} models increase to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+The overall accuracy of both \ac{pmv} models increase to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
 The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.
@@
 
-In Table~\ref{tab:f1}, we report the F1 scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{v} measured at three heights with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}, and iv) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
+In Table~\ref{tab:f1}, we report the F1 scores for both models, i) using the whole dataset, ii) only keeping entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}, iii) only keeping data points that have \ac{vr} measured at three heights with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}, and iv) for those with a $\lvert \textrm{PMV}\lvert \leq 1.5$ and $\lvert \textrm{TSV}\lvert \leq 1.5$.
 \begin{table}[htb!]
@@
 The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
-For the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models only slightly improved.
-We then analyzed the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{v} measured at three heights.
+For the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the F1-macro scores of the \ac{pmv} and \ac{pmv-ce} models only slightly improved.
+We then analyzed the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{vr} measured at three heights.
 Both F1-macro scores decrease and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
@@
 Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
-The results highlight that the \ac{pmv-ce} has a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+The results highlight that the \ac{pmv-ce} has a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
 We report the median and inter-quantile ranges since data are not normally distributed.
@@
     \includegraphics[width=\textwidth]{figures/hist_discrepancies}
-    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s}.
+    \caption{Overall bias of the \ac{pmv} and \ac{pmv-ce} models for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
     The x-axis shows the discrepancy (bias) between the predicted and self-reported \ac{tsv}.
@@
 \end{figure*}
-When including only the data with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
+When including only the data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the median bias of the \ac{pmv} model reduces to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improves.
 The median bias for the \ac{pmv-ce} also decreases to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0} but it is higher than the one for the \ac{pmv}.
@@
 \subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
-Here we present the model bias as a function of the five input variables \ac{tdb}, \ac{v}, \ac{rh}, \ac{met}, and \ac{clo}.
+Here we present the model bias as a function of the five input variables \ac{tdb}, \ac{vr}, \ac{rh}, \ac{met}, and \ac{clo}.
 Results are shown in Figure~\ref{fig:bias_models}.
@@
 The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
-On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{v}~$\leq$~\qty{0.46}{\m\per\s}.
+On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.
 Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.

### 2025-01-16 main.py

@@ -565,6 +565,123 @@ def pmv(tdb, tr, vr, rh, met, clo, wme=0, standard="ISO", **kwargs):
     return pmv_ppd(tdb, tr, vr, rh, met, clo, wme, standard, **kwargs)["pmv"]
 
 
+def clothing_correction(clo, met, v):
+    i_cl_st = clo * 0.155
+    fcl = 1 + 0.3 * clo
+    # Static boundary layer thermal insulation in quiet air
+    i_a_st = 0.111
+    # Total static insulation
+    i_tot_st = i_cl_st + i_a_st / fcl
+    walk_sp = 0.0052 * (met - 58)
+    if walk_sp > 0.7:
+        walk_sp = 0.7
+    v_r = v
+    v_ux = v_r
+    if v_r > 3:
+        v_ux = 3
+    w_a_ux = walk_sp
+    if walk_sp > 1.5:
+        w_a_ux = 1.5
+    # correction for the dynamic total dry thermal insulation at or above 0.6 clo
+    corr_cl = 1.044 * math.exp(
+        (0.066 * v_ux - 0.398) * v_ux + (0.094 * w_a_ux - 0.378) * w_a_ux
+    )
+    if corr_cl > 1:
+        corr_cl = 1
+    # correction for the dynamic total dry thermal insulation at 0 clo
+    corr_ia = math.exp((0.047 * v_r - 0.472) * v_r + (0.117 * w_a_ux - 0.342) * w_a_ux)
+    if corr_ia > 1:
+        corr_ia = 1
+    corr_tot = corr_cl
+    if clo <= 0.6:
+        corr_tot = ((0.6 - clo) * corr_ia + clo * corr_cl) / 0.6
+    # total dynamic clothing insulation
+    i_tot_dyn = i_tot_st * corr_tot
+    # dynamic boundary layer thermal insulation
+    i_a_dyn = corr_ia * i_a_st
+    return i_tot_dyn - i_a_dyn / fcl
+
+
+def _correction_nude(_vr, _vw):
+    return np.exp(
+        -0.533 * (_vr - 0.15) + 0.069 * (_vr - 0.15) ** 2 - 0.462 * _vw + 0.201 * _vw**2
+    )
+
+
+def _correction_normal_clothing(_vr, _vw):
+    return np.exp(
[…]

### 2025-01-17 Manuscript/src/response-letter.tex
review: updated the whole paper with the new results I_cl_r ISO.

@@
 
+\usepackage{siunitx}
+
 % Name of sender
@@
 
+\usepackage[printonlyused]{acronym}
+\usepackage[nonumberlist,nogroupskip]{glossaries}
+\input{myglossary.tex}
+
 %--------------------------------------------------------------------------%
@@
 \begin{document}
+
+    \renewcommand{\baselinestretch}{0.75}\normalsize
+    \renewcommand{\aclabelfont}[1]{\textsc{\acsfont{#1}}}
+
+    \acrodef{tdb}[$t_{db}$]{dry-bulb air temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{twb}[$t_{wb}$]{wet-bulb air temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{ti}[$t_{i}$]{indoor air temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{tout}[$t_{out}$]{outdoor air temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{to}[$t_{o}$]{operative air temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{tcl}[$t_{cl}$]{clothing temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{tg}[$t_{g}$]{globe temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{rh}[$RH$]{relative humidity\acroextra{, \%}}
+    \acrodef{v}[$V$]{measured air speed\acroextra{, m/s}}
+    \acrodef{vr}[$V_{r}$]{relative air speed\acroextra{, m/s}}
+    \acrodef{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{clo}[$I_{cl}$]{intrinsic clothing insulation from the skin to the outer surface under reference conditions\acroextra{, clo}}
+    \acrodef{clor}[$I_{cl,r}$]{actual intrinsic clothing insulation under given environmental and activities\acroextra{, clo}}
+    \acrodef{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
+
+    \acrodef{athb}[ATHB]{Adaptive Thermal Heat Balance}
+    \acrodef{pmv}[PMV]{Predicted Mean Vote}
+    \acrodef{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
+    \acrodef{set}[SET]{Standard Effective Temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{ce}[CE]{Cooling Effect\acroextra{, $^{\circ}$C}}
+    \acrodef{tsv}[TSV]{Thermal Sensation Vote}
+    \acrodef{tpv}[TPV]{Thermal Preference Vote}
+    \acrodef{ef}[$e$]{expectancy factor}
+    \acrodef{af}[a]{adaptive coefficient}
+
+    \acrodef{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{work}[$W$]{rate of mechanical work accomplished\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{t-sk}[$t_{sk}$]{skin mean temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{t-sk-n}[$t_{sk,n}$]{neutral skin mean temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{t-cr}[$t_{cr}$]{core mean temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{t-re}[$t_{re}$]{rectal temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{t-cr-n}[$t_{cr,n}$]{neutral core mean temperature\acroextra{, $^{\circ}$C}}
+    \acrodef{r-cl}[$R_{cl}$]{thermal resistance of clothing\acroextra{, m\textsuperscript{2}K/W}}
+    \acrodef{r-e-cl}[$R_{e,cl}$]{evaporative heat transfer resistance of clothing layer\acroextra{, m\textsuperscript{2}kPa/W}}
+    \acrodef{f-cl}[$f_{cl}$]{clothing area factor $A_{cl}/A_{body}$\acroextra{, m\textsuperscript{2}K/W}}
+    \acrodef{h}[$h$]{sum of convective $h_{c}$ and radiative $h_{r}$ heat transfer coefficients\acroextra{, W/(m\textsuperscript{2}K)}}
+    \acrodef{h-r}[$h_{r}$]{linear radiative heat transfer coefficient\acroextra{, W/(m\textsuperscript{2}K)}}
+    \acrodef{h-c}[$h_{c}$]{convective heat transfer coefficient\acroextra{, W/(m\textsuperscript{2}K)}}
+    \acrodef{h-e}[$h_{e}$]{evaporative heat transfer coefficient\acroextra{, W/(m\textsuperscript{2}kPa)}}
+    \acrodef{a}[$\alpha$]{fraction of the total body mass considered
+    to be thermally in the skin compartment}
+    \acrodef{s-cr}[$S_{cr}$]{rate of heat storage in the core compartment\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{s-sk}[$S_{sk}$]{rate of heat storage in the skin compartment\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{s}[$S$]{rate of heat storage in the human body\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{e-res}[$E_{res}$]{rate of evaporative heat loss from respiration\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{e-dif}[$E_{dif}$]{rate of evaporative heat loss from moisture diffused through the skin\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{e-rsw}[$E_{rsw}$]{rate of evaporative heat loss from sweat evaporation\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{e-sk}[$E_{sk}$]{total rate of evaporative heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{e-max}[$E_{max}$]{maximum rate of evaporative heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{c-res}[$C_{res}$]{rate of convective heat loss from respiration\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{c-r}[$C + R$]{sensible heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{q-res}[$q_{res}$]{total rate of heat loss through respiration\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{q-sk}[$q_{sk}$]{total rate of heat loss from skin\acroextra{, W/m\textsuperscript{2}}}
+    \acrodef{w}[$w$]{skin wettedness}
+    \acrodef{w-max}[$w_{max}$]{skin wettedness practical upper limit}
+    \acrodef{m-sweat}[$m_{rsw}$]{rate at which regulatory sweat is generated\acroextra{, mL/h\textsuperscript{2}}}
+    \acrodef{m-bl}[$m_{bl}$]{skin blood flow\acroextra{, L/(hm\textsuperscript{2})}}
+    \acrodef{c-sw}[$c_{sw}$]{driving coefficient for regulatory sweating\acroextra{, g/(hKm\textsuperscript{2})}}
+
+    \acrodef{pmv-ce}[PMV\textsubscript{CE}]{PMV calculated in accordance with the ASHRAE~55:2023}
+    \acrodef{epmv}[ePMV]{Adjusted Predicted Mean Votes with Expectancy Factor}
+    \acrodef{apmv}[aPMV]{Adaptive Predicted Mean Vote}
+    \acrodef{pmvs}[PMV$_{SET}$]{PMV calculated using the SET temperature in place of $t_{db}$}
+    \acrodef{pmvg}[PMV$_{Gagge}$]{PMV calculated using the two-node heat balance model}
+    \acrodef{db2}[DB$_{comfort}$]{ASHRAE Global Thermal Comfort Database II}
+    \acrodef{lowess}[LOWESS]{LOcally WEighted Scatterplot Smoothing}
+
+    \renewcommand{\baselinestretch}{1}\normalsize
 
@@
 
-            \todo[color=yellow]{Federico - add the following sentences and reference them in the comment below}
             \response{
                 We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data and that the ASHRAE Global Thermal Comfort Database II v2.1 has limitations.
-                We have added a sentence in the introduction to highlight the importance of the quality of the input data.
-                We have also added a sentence in the limitations section to acknowledge the limitations of the PMV model.
-                However, despite the fact that this database has some limitations, we decided to use it because, to the best of our knowledge, the ASHRAE Global Thermal Comfort Database II v2.1 is the most comprehensive database available for thermal comfort research.
-                The database is widely used in the field, and only includes data published in peer-reviewed studies.
-                Moreover we belive that errors in the inputs data, if not systematic, should be mitigated by the large number of data points in the database.
+                While we previously mentioned in the manuscript that not all authors used the same measurement protocols.
+                Based on your valuable suggestion we have further improved Section 2.1. and better discussed the limitations of the dataset.
+                We have also added a few sentences in which we reference the work of d’Ambrosio Alfano et al. (2011) which have already investigated this issue.
+                We have referenced the work of d’Ambrosio Alfano et al.
+(2011) in the manuscript when we justify that the model can be considered accurate if the error between the PMV and the TSV is within $\pm$ 0.5; and we have referenced the work in Section 3.3 measurements errors.\\
+                Finally, while we recognised the limitations of the DB2 in section 2.1., we have also highlighted the measures we have taken to mitigate these limitations.
+                For example, we clarify that if we belive that errors in the inputs data are not systematic across all authors, the use of a large dataset should be able to mitigate the errors caused by the lack of standardised measurement protocols.
             }
@@
         \end{enumerate}
-
-        \clearpage
 
@@
                 We retained data with |PMV| $\leq$ 3.5 because the PMV model is scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
+                In addition, while the ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
                 Moreover, as explained in the paper the thermal snesation votes in the ASHRAE Global Thermal Comfort Database II v2.1 are collected using a 7-point scale which ranges from -3 to +3.
@@
                 We may include this figure if the reviewers believe it is necessary.
+                We have changed the text in the manuscript as follows: \textit{``Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2.
+                However, since the thermal sensation was measured with a seven-point scale, the \ac{pmv} has no upper or lower boundary, and the \gls{55} does not specify the range of applicability of the \ac{pmv-ce} model, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.''}
             }
@@
             \response{
-            Thank you for pointing out this potential issue.
+                Thank you for this very valuable comment.
                 We agree that limiting the applicability of the PMV model to the range -0.5 to 0.5 may in some instances increase the building operation energy consumption.
@@
                 Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated and participants do not have individual control over the indoor environment.
-            This, however, is a discussion that goes beyond the scope of this paper.
-            It should also be noted that we recommend limiting the applicability of the PMV model to the range -0.5 to 0.5 only because our results show that the PMV model has a low accuracy in predicting thermal sensation outside this range.
+                We have now included the following paragraph in the conclusion of the manuscript to address this issue: \textit{``We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
+            However, this is not the case.
+            As shown in Figure 2 increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office settiong to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
+            This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
+            Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy, but instead is that in buildings we cool the air rather than moving it.
+            A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}.''}
             }
@@
 
-            \todo{Stefano - Would you like to answer this question? I have the same concern as the reviewer but I know that you have a different opinion}
-            \response{Here our response}
+            \response{
+                Thank you for your valuable comment.
+                We completely agree with the reviewer that originally the PMV model was not intended to predict the thermal sensation of each individual participant and we have mentioned this as the opening sentence of Section 2.2.
+                However, the PMV model is widely used in the scientific literature \textit{``in both singly occupied spaces and areas accommodating several hundred people''}.
+                In Section 2.2 of the manuscript we have also added the following sentence \textit{``Partially because neither of the two standards specifies the minimum number of people needed to apply the model.''}
+                We are, however, completely aware of this issue and we have not neglected it.
+                In the manuscript we stated that \textit{``we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
+                These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
+                However, on their own, are a low-precision estimate of the overall accuracy of the model''}
+                We have also added the following sentence: \textit{``since the model is not expected to predict the exact \ac{tsv} of each participant}''.
+                We believe that \textit{``If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
+                The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)''}
+                Finally we would want to raise a concern we have about the vague definition of the PMV model in the ISO 7730 and ASHRAE 55 standards.
+                Stating that a model is intended to predict the average thermal sensation of an undefined large group of occupants is not a clear definition and in principle makes it impossible to determine the accuracy of the model.
+                At the same time it does not help the user to understand when the model can be used and when it cannot.
+                For example, can it be used to design a single office or a small appartment for a single person?
+                What about when the space is occupied by two or three people?
+                We believe that the Standards should be more clear about the range of applicability of the PMV model.
+                We have, therefore, added a sentence about this in the conclusion of the manuscript which reads as follows: \textit{``Finally, we would like to highlight that the current definition of the \ac{pmv} model ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
+                Both the \gls{55} and \gls{7730} standards should clarify the minimum number of occupants required to use the \ac{pmv} model.
+                Can it be used for a space with only one occupant?''}
+            }
 
-            \item Although less critical, another general comment is that it is not clear what is meant by low predicition accuracy, as is written already in the abstract - when is the prediction accuracy low?
+            \item Although less critical, another general comment is that it is not clear what is meant by low prediction accuracy, as is written already in the abstract - when is the prediction accuracy low?
             We aim to estimate what people perceive, which seriously suffers from individual differences and many others challenges.
@@
 
-            \todo[color=yellow]{Federico - add a few more additional sentences about the bias in the manuscript}
             \response{
-            As previously explained we acknowledge that the PMV model is not intended to predict the thermal sensation of each individual participant.
-            We have acknowledged this limitation in the manuscript and discussed it in Section 3.2.
+                We acknowledge that the PMV model is not intended to predict the thermal sensation of each individual participant.
+                We have acknowledged this limitation in the manuscript and discussed it in Section 2.2 and 3.2.
                 To compensate for this limitation of our analysis, we calculated the bias of the model as previously done by Humphreys and Nicol (2002) to determine if the model is systematically over or under predicting the thermal sensation of the participants.
                 Individual differences in human subjects can be assumed to be random and distributed around the mean.
-            This is the same assumption that PMV model also does when trying to predict the average thermal sensation of a large group of occupants.
-            Consequently, we believe that calcualting the bias of the model is a valid approach to determine the overall accuracy of the model.
+                We have then added the following sentence to the manuscript: \textit{``his is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.''}
+                Consequently, we believe that calcualting the bias of the \ac{pmv} models is a valid approach to determine the overall accuracy of the model.
             }

### 2025-01-17 Manuscript/src/cover-letter.tex
submission: first submission energy and buildings
review: updated the whole paper with the new results I_cl_r ISO.

@@
 We found that both models have low and similar prediction accuracy ($ \approx$\qty{34}{\percent}) and the PMV$_{CE}$ has higher bias than PMV in predicting thermal sensation in the subset of data with air speed greater than \qty{0.2}{\m\per\s}.
-Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm,' `hot,' or `cold.'
+Both models have an accuracy lower or equal to \qty{5}{\percent} when predicting either `warm', `hot', or `cold'.
 This value is lower than random guessing (\qty{14}{\percent}), meaning they are misleading the users.
@@
 Moreover, we recommend using the PMV model over the PMV$_{CE}$ model, as the latter is more complex and computationally intensive to run, while not being more accurate than the original PMV model.
-Moreover, we challenge the assumption that individuals who feel `slightly warm' or `slightly cool' should be considered thermally comfortable.
+Finally, we challenge the assumption that individuals who feel `slightly warm' or `slightly cool' should be considered thermally comfortable.
 

### 2025-01-17 Manuscript/src/sections/appendix.tex
review: updated the whole paper with the new results I_cl_r ISO.

@@
 In this section, we show the results of the analysis of the dataset with air speed measurements at three heights.
-Only \num{3063} entries in the \ac{db2} had air speed measurements at three heights with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+% todo check manually
+Only \num{3059} entries in the \ac{db2} had air speed measurements at three heights with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
 A summary of the papers that collected these data is provided in Table~\ref{tab:three_heights_papers}.
@@
 With the value of \ac{vr} being \var{delta_v_mean} higher than the average of the other air speeds combined.
-The discussion on whether this increase is justified is beyond the scope of this paper, since both standards require the user to adjust the air speed based on the \ac{met} value.
+The discussion on whether this increase is justified is beyond the scope of this paper, since both standards require the user to adjust the air speed based on the \ac{met} value using the same equation.
 
@@
 The \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
-The intercept of the \ac{pmv} model is \num{-.31} while the intercept of the \ac{pmv-ce} model is \num{-.41}.
+The intercept of the \ac{pmv} model is \var{intercept_v_02_three_pmv} while the intercept of the \ac{pmv-ce} model is \var{intercept_v_02_three_pmv_ce}.
 Moreover, the \ac{pmv-ce} \ac{lowess} curve is only marginally higher than \num{0} despite participants reporting a \ac{tsv} of \num{3}.
@@
 \end{figure}
-The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.39}.
-The interquartile range for the \ac{pmv} model is more centered around \num{0} than the \ac{pmv-ce} model.
-These results allow us to conclude that the \ac{pmv} model is more accurate than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
+% todo check manually
+The \ac{pmv} model has a bias of \num{-.39} while the \ac{pmv-ce} model has a bias of \num{-.38}.
+These results allow us to conclude that the \ac{pmv} has a similar accuracy than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
 
@@
 
-The \ac{pmv} model has a bias of \num{-.08} while the \ac{pmv-ce} model has a bias of \num{-.31}.
-Even for this subset of data, the \ac{pmv} model is more accurate than the \ac{pmv-ce} model; this result is consistent with the previous analysis and confirms that the \ac{pmv} model is more accurate than the \ac{pmv-ce}.
+The \ac{pmv} model has a bias of \num{-.28} while the \ac{pmv-ce} model has a bias of \num{-.31}.
+Even for this subset of data, the \ac{pmv} model is more accurate than the \ac{pmv-ce} model.

### 2025-01-17 Manuscript/src/sections/conclusions.tex
review: updated the whole paper with the new results I_cl_r ISO.

@@
 
-The overall accuracy and F1-macro score of the \ac{pmv} is \qty{33}{\percent} and \num{.16}, respectively.
-While the overall accuracy and F1-macro score of the \ac{pmv-ce} is \qty{34}{\percent} and \num{.16}, respectively.
-Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm,' `hot,' or `cold.'
+% todo check values manually
+The overall accuracy and F1-macro score of the \ac{pmv} is \var{overall_acc_pmv} and \num{.16}, respectively.
+While the overall accuracy and F1-macro score of the \ac{pmv-ce} is \var{overall_acc_pmv_ce} and \num{.15}, respectively.
+% todo check values manually
+Both models have an accuracy lower or equal to \qty{5}{\percent} when predicting either `warm', `hot', or `cold'.
 This value is lower than random guessing (\qty{14}{\percent}).
 This means that the models are misleading the users.
+% todo check values manually
 When only comparing the results of the two models for entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
@@
 Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988} provide corroborating evidence that supports our results.
+We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
+However, this is not the case.
+As shown in Figure~\ref{fig:comfort_regios_pmv_pmvce} increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office settiong to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
+This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
+Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy, but instead is that in buildings we cool the air rather than moving it.
+A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}~\cite{kent_energy_2023}.
 
@@
 The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
-We then recommend that the \gls{55} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
-\todo{and why not the ISO Standard also?  After all, ASHRAE was copying ISO on that one}
+We then recommend that both the \gls{55} and \gls{7730} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
 
@@
 
+Finally, we would like to highlight that the current definition of the \ac{pmv} model ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
+Both the \gls{55} and \gls{7730} standards should clarify the minimum number of occupants required to use the \ac{pmv} model.
+Can it be used for a space with only one occupant?
+
 \section*{Acknowledgements}\label{sec:acknowledgements}

### 2025-01-17 Manuscript/src/sections/introduction.tex
review: updated the whole paper with the new results I_cl_r ISO.

@@
             Since the \gls{7730} does contain some errors and it does not provide all the necessary information to calculate $\alpha$ we used the equations provided in the \gls{9920} to derive the value of $\alpha$.
-        In accordance with the \gls{7730} and \gls{9920} standards, since we did not have information about the walking speed of the participants, we estimated it using the provided equation.
+        The \gls{7730} standard, also provides an equation to calculat the walking speed if this information is not available.
         }
@@
 \end{figure}
-In summary, as shown in Figure~\ref{fig:flowchart_pmv_ce} when the \ac{vr} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
+In summary, as shown in Figure~\ref{fig:flowchart_pmv_ce}, when the \ac{vr} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
 \ac{vr} is different from the \ac{v}.
 According to the standard, the measured value should be adjusted using this equation \ac{vr} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
-The standard also prescribes the use of \ac{clor} and not of \ac{clo} as an input.
+The standard also prescribes the use of the \ac{clor} as an input rather than the \ac{clo}.
 For \ac{met} $> \qty{1.2}{met}$, \acs{clor} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$.
@@
 The other three input parameters (\ac{clor}, \ac{rh}, \ac{met}) remain unchanged.
-% todo this is no longer true and it should be removed or rephrased
-Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
-To illustrate the extent of the differences between the outputs of the two models, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{vr} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
+
+Figure~\ref{fig:flowchart_pmv_iso} shows the calculation routine for the \ac{pmv} model as per the \gls{7730} standard.
+The same equation as in the \gls{55} is used to adjust the \ac{vr} value.
+However, the \ac{clor} is calculated differently.
+In the \gls{7730}, the \ac{clor} varies as a function of \ac{clo}, \ac{met}, walking speed, and \ac{vr}.
+
+Consequently, while the two \ac{pmv} formulations mainly differ when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
+It should be noted that the two models also use different equations to calculate the \ac{clor}.
+
+To illustrate the extent of the differences between the outputs of the two models, only as a function of \ac{vr}, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) assuming the same \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
 We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
@@
 
-The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds ~\cite{huang_applicability_2014}.
+The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds~\cite{huang_applicability_2014}.
 The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}.
-Hence, if sweating is required to maintain the body in thermal equilibrium, the \ac{pmv} model assumes that the body would gain heat.
+Hence, if sweating is required to maintain the body in thermal equilibrium, the \ac{pmv} model assumes that the body would get hotter.
 The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
@@
 
-Our analysis has, however, revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.
-
-% \mycite{huang_applicability_2014}, using 19 averaged data points obtained from a set of experiments involving 30 subjects, showed the \ac{set} model could be used to estimate the thermal sensation with two different ways of estimating the airspeed.
-% They developed a model to relate the SET and the TSV. 
-% They also plotted the \ac{pmv} versus the TSV and provided a visual analysis of its performance.
-They concluded that the SET is better than the \ac{pmv} model in predicting thermal sensation votes when subjects are exposed to air movements, but they did not report the performance of the \ac{pmv} model.
-% Moreover, they did not compare the accuracy of the \ac{pmv-ce} with the \ac{pmv} model.
+Despite this claim from the authors of the \ac{pmv-ce}, our analysis has, however, revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.
 
 \begin{enumerate}
-    \item Although the \ac{ce} claims to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model.
+    \item Although the \ac{ce} claims to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model, hence it has the same limitations.
     For example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.15}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
@@
     \item It should be noted that over the years, the threshold value of \ac{vr} after which the \ac{ce} is calculated, has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-    However, these changes were not based on any scientific evidence and significantly affect the output of the model.
+    However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.
     \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{vr}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
@@
 We could propose this to the Standards Committee if we wanted to.
-We would have to calculate two validation tables for the two approaches.
-}
+We would have to calculate two validation tables for the two approaches.}
 \end{enumerate}

### 2025-01-17 Manuscript/src/sections/methodology.tex
review: updated the whole paper with the new results I_cl_r ISO.

@@
 \Ac{tsv} were collected using a right-here-right-now survey and environmental parameters were measured and logged in the proximity of participants when they completed the survey.
-The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
+The \ac{db2} only contains data collected indoors by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
 
@@
 This filtered out approximately \var{entries_db_valid} of the data.
-In our analysis, we used the value \ac{vr} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
-This is because both \ac{pmv} formulations use the \ac{vr} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
 
@@
 Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
-However, since the thermal sensation was measured with a seven-point scale and the \ac{pmv} has no upper or lower boundary, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
+However, since the thermal sensation was measured with a seven-point scale, the \ac{pmv} has no upper or lower boundary, and the \gls{55} does not specify the range of applicability of the \ac{pmv-ce} model, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
 
-In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same methodology.
+In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same standardised methodology.
 For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
-This could introduce a bias in the measurements and, consequently, in the results of the models.
-Moreover, since the two \ac{pmv} formulations only differ when \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
+The absence of standardized measurement protocols, the use of inaccurate or un-calibrated sensors, and the use of different reference tables to estimate \ac{clo} and \ac{met} may all introduce an error in the measurements and, consequently, in the results of both models.
+This issue has been already evidenced in the literature by \mycite{dambrosio_alfano_role_2011}.
+They showed that the reliability of the \ac{pmv} model is strongly affected by the accuracy of the input data.
+In our analysis, we compensate this issue by:
+\begin{enumerate}[ {}1{)} ]
+    \item comparing the results of the two models using the same dataset; hence, the different in the output is mainly due to the different formulations of the models;
+    \item using a large dataset to reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards since it can be assumed that measurement errors random across different studies.
+    \item using only datasets that have been published in peer-reviewed journals.
+%    \item excluding all those studies that did not measure all the environmental parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
+\end{enumerate}
+
+In addition, since the two \ac{pmv} formulations mainly differ when \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
 \begin{enumerate}[ {}1{)} ]
@@
     \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s};
-    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
+    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
     These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
@@
 The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
-The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
 In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
@@
 \subsection{Model Validation}\label{subsec:model-validation}
-The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, but it is commonly used both in singly occupied zones and in areas with several hundred people.
+The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, but it is often applied in both singly occupied spaces and areas accommodating several hundred people.
+Partially because neither of the two standards specifies the minimum number of people needed to apply the model.
 We first grouped participants' responses by their \ac{tsv}, and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
@@
 These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
-However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
+However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002} since the model is not expected to predict the exact \ac{tsv} of each participant.
 These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
@@
 The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
-According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
-However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers.
-Consequently, we expanded the range to \num{-.5} and \num{0.5}.
+This is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.
+According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{\pm 0.25}.
+However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers and even when environmental parameters are measured within the accuracy of the standards, the output of the model may vary by more than \num{\pm 0.25}~\cite{dambrosio_alfano_role_2011}.
+Consequently, we expanded the range to \num{\pm 0.5}.
 This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
-We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
+We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range.
+Ideally, we would want to have the \ac{pmv} to have higher precision.
+however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale and by the accuracy of the inputs.
 When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.

### 2025-01-17 Manuscript/src/sections/results.tex
review: updated the whole paper with the new results I_cl_r ISO.

@@
 The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
+We are also presenting the values of \ac{v} and \ac{clo} for comparison.
 \begin{figure*}[htb!]
@@
 This lack of data in the \ac{db2} is particularly relevant for \ac{vr}.
-The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}.
+The \ac{pmv-ce} only calculates the \ac{ce} for \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}.
 The median value in the \ac{db2} is \var{v_median} suggesting that most of the data points were collected in buildings with low air movement.
@@
 
-The percentage of \ac{tpv} grouped by each \ac{tsv} is shown in Figure~\ref{fig:bar_plot_tp_by_ts} a).
+The percentage of \ac{tpv} grouped by each \ac{tsv} is shown in Figure~\ref{fig:bar_plot_tp_by_ts}~a).
 \begin{figure*}[htb!]
@@
 This finding challenges the above assumption.
-Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to under-counting thermal discomfort.
+Assuming that all people who are `slightly warm' or `slightly cool' are comfortable may lead to under-counting thermal discomfort.
 On the other hand, some participants who reported to be `cool', `cold', `warm', or `hot' wanted `no change'.
@@
 The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we compared the predicted value from the two models with measured \ac{tsv}.
-However, since the two models only differ when \ac{vr}~$>$~\qty{0.1}{\m\per\s};
-we also present the results for a subset of the data where \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
-In this subset of data, the differences in the outputs of the two models are more pronounced.
+In the subset of data where \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}, the differences in the outputs of the two models are more pronounced.
 We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{vr} of \var{v_95_perc_max}.
@@
 Both \ac{pmv} formulations underestimate the conditions at which participants are thermally dissatisfied with their environment.
-The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \qty{33}{\percent} and \qty{34}{\percent}, respectively, when using the whole dataset.
-The overall accuracy of both \ac{pmv} models increase to \qty{34}{\percent} and \qty{35}{\percent}, respectively when only considering entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+The overall accuracies of the \ac{pmv} and \ac{pmv-ce} are \var{overall_acc_pmv} and \var{overall_acc_pmv_ce}, respectively, when using the whole dataset.
+The overall accuracy of both \ac{pmv} models changed to \var{overall_acc_pmv_0.2} and \var{overall_acc_pmv_ce_0.2}, respectively when only considering entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+% todo check data manually
 The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.
@@
 \end{table}
+% todo check data manually
 The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
@@
 We then analyzed the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and with \ac{vr} measured at three heights.
-Both F1-macro scores decrease and the \ac{pmv-ce} (\num{0.16}) is lower than the value for the \ac{pmv}~ (\num{0.17}).
+The score of the \ac{pmv-ce} decreased (F1-macro~=~\num{0.16}) and it is lower than the value for the \ac{pmv} (F1-macro~=~\num{0.17}).
 Our results show that the assumption that the \ac{pmv-ce} better takes into account the effect of elevated air speed is false.
@@
 This ideal prediction is referred to as the identity line, shown as the brown dashed line in the figure.
-The intercepts for the \ac{pmv} is \qty{-0.22} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
-This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{54}{\percent} and \qty{58}{\percent} of the times, respectively.
+The intercepts for the \ac{pmv} is \var{intercept_all_pmv} and for the \ac{pmv-ce} is \var{intercept_all_pmv_ce} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
+% todo check data manually
+This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{51}{\percent} and \qty{58}{\percent} of the times, respectively.
 Both \ac{pmv} formulations always under-predict the thermal sensation of people who reported to be on the extremes of the \ac{tsv};
-the slope of the curve was significantly lower than 1.
+the slope of the curve is significantly lower than 1.
 
@@
 Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
+% todo check data manually
 Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
-The results highlight that the \ac{pmv-ce} has a higher bias than the \ac{pmv} both when calculating the bias for all the data and for the subset of data with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
-We report the median and inter-quantile ranges since data are not normally distributed.
-%  We cannot change into a letter value plot since in that case, it would be hard to highlight the number of responses that are within the range from -.5 to 5.
 \begin{figure*}[htb!]
@@
     We colored the bars in the range from \num{-.5} to \num{.5} to highlight the number of responses that are within this range.
-    The chart also shows the median and inter-quantile ranges of the data.
+    We report the median and inter-quantile ranges since data are not normally distributed.
     }
@@
 \subsubsection{Model Bias as Function of Five Input Variables}\label{subsubsec:model-bias-variable}
-Here we present the model bias as a function of the five input variables \ac{tdb}, \ac{vr}, \ac{rh}, \ac{met}, and \ac{clo}.
+Here we present the model bias as a function of \ac{tdb}, \ac{vr}, \ac{rh}, and \ac{met}.
 Results are shown in Figure~\ref{fig:bias_models}.
@@
 The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
-On the contrary, the bias of the \ac{pmv-ce} diverges more from 0 and is significantly worse (\num{-.33} versus \num{-0.13}) than the one for the \ac{pmv} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.
-Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius} and \ac{clo}~$\leq$~\qty{.5}{clo}.
+% todo check data manually
+On the contrary, the bias of the \ac{pmv-ce}~=~\num{-.33} is higher than \ac{pmv}~=~\num{-0.23} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.
+Both models have a median bias higher or equal to $\lvert0.5\lvert$ for \ac{tdb}~$\leq$~\qty{21}{\celsius}.
 
-These findings, along with those shown in Table~\ref{tab:f1} and earlier figures, demonstrate that the \ac{pmv-ce} model has lower accuracy and higher bias compared to the \ac{pmv} at elevated air speeds, even though the \ac{pmv-ce} model is more complex and requires more computational resources.
+These findings, along with those shown in Table~\ref{tab:f1} and earlier figures, demonstrate that the \ac{pmv-ce} model at elevated air speeds is not an improvement compared to the \ac{pmv}, even though the \ac{pmv-ce} model is more complex and requires more computational resources.
 
@@
 The issues are reported in order of importance of what we believe may affect the accuracy of the model.
-\todo{Note that the section below does not mention the two-node model, the CE implementation of PMV, or the physiological differences between the ISO and ASHRAE models that could explain the observed differences or ‘error’.
+\todo{Ed's comment - Note that the section below does not mention the two-node model, the CE implementation of PMV, or the physiological differences between the ISO and ASHRAE models that could explain the observed differences or ‘error’.
 }
@@
 Additionally, the \ac{pmv} model erroneously assumes that the human body is not capable of maintaining a stable core temperature.
-A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicate that the hypothetical cylinder is either gaining or losing heat and consequently it is getting warmer or colder, respectively.
+A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicate that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently it is getting warmer or colder, respectively.
+
 Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer; however, this never precisely occurs.
 The human thermoregulatory system, in healthy humans, is constantly actively engaged to ensure a stable core temperature~\cite{romanovsky_thermoregulation_2018}.
+
 The \ac{pmv} model also assumes that the human body is constantly losing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
@@
 It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy only in determining when people are `neutral'.
-This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis~\cite{ romanovsky_thermoregulation_2018}.
+This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, evaporation, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis~\cite{ romanovsky_thermoregulation_2018}.
 On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the thermal neutrality zone.
@@
 Instrumentation is not always calibrated and the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the participant's proximity, and estimating clothing and metabolic rates from tables is a complex and prone-to-error task.
+Even when variables are measured within the standards' accuracy requirements, the output of the \ac{pmv} value may vary enough to wrongly estimate the correct \ac{tsv}~\cite{dambrosio_alfano_role_2011}.
 In an ideal scenario, errors would be randomly distributed and cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.

### 2025-01-17 Manuscript/src/sections/titleAbstract.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r
review: updated the whole paper with the new results I_cl_r ISO.

@@
         We compared the accuracy of the two \ac{pmv} models, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that both models have low and similar prediction accuracy ($\approx$\qty{34}{\percent}).
-        The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{v}~$\geq$~\qty{0.2}{\m\per\s}.
-        Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm', `hot', or `cold'.
+        We found that overall both models only predict \ac{tsv} correctly only one third of the times.
+        The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+%        todo manually check values
+        Both models have an accuracy lower or equal to \qty{5}{\percent} when predicting either `warm', `hot', or `cold'.
         This value is lower than random guessing (\qty{14}{\percent}), meaning they are misleading users.
@@
         Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort.
+%        todo manually check values
         In the \acl{db2}, \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
@@
 
+%    todo change the graphical abstract
     \begin{graphicalabstract}
@@
         \item We compared the accuracy of the \ac{pmv} models in \gls{7730} and \gls{55}.
-        \item Both models have low and similar prediction accuracy ($\approx$\qty{34}{\percent}).
+        \item Both models have low and similar prediction accuracy (\ac{pmv}~=~\var{overall_acc_pmv} -- \ac{pmv-ce}~=~\var{overall_acc_pmv_ce}).
         \item The \ac{pmv-ce} has a higher bias than \ac{pmv} in predicting thermal sensation.

### 2025-01-17 main.py

@@ -131,9 +131,12 @@ var_names = {
     "ta": r"$t_{db}$",
     "tr": r"$\overline{t_{r}}$",
     "top": r"$t_{o}$",
-    "vel_r": r"$V$",
+    "vel_r": r"$V_r$",
+    "vel": r"$V$",
     "rh": r"RH",
     "clo": r"$I_{cl}$",
+    "clo_d": "ASHRAE\n" + r"$I_{cl,r}$",
+    "i_cl_r": "ISO\n" + r"$I_{cl,r}$",
     "met": r"$M$",
     "thermal_sensation": "Thermal Sensation Vote (TSV)",
     "thermal_sensation_round": "Thermal Sensation Vote (TSV)",
@@ -164,8 +167,11 @@ var_units = {
     "ta": r"$^{\circ}$C",
     "tr": r"$^{\circ}$C",
     "vel_r": r"m/s",
+    "vel": r"m/s",
     "rh": r"%",
     "clo": r"clo",
+    "clo_d": r"clo",
+    "i_cl_r": r"clo",
     "met": r"met",
 }
 
@@ -1314,11 +1320,11 @@ def importing_filtering_processing(load_preprocessed=False):
     # ax.set(ylabel="Clothing insulation [clo]")
 
     # calculation in accordance with the ISO 9920
-    i_a = 0.7  # todo 9920 default value
+    i_a = 0.7
     f_cl = 1 + 0.28 * df_.clo
     i_t = df_.clo + i_a / f_cl
     v_walk = 0.3 * (df_.met - 1)
-    # v_walk = 0  # todo check what to use
+    # v_walk = 0
     # v_walk[v_walk > 0.7] = 0.7
     v_r = df_.vel + 0.3 * (df_.met - 1)
     # v_r[v_r < df_.vel] = df_.vel
@@ -1800,9 +1806,11 @@ def plot_error_prediction(data):
 
 
 def plot_distribution_variable():
-    f, axs = plt.subplots(1, 6, constrained_layout=True, figsize=(7, 3))
+    f, axs = plt.subplots(1, 9, constrained_layout=True, figsize=(7, 3))
 
-    for ix, var in enumerate(["ta", "tr", "rh", "vel_r", "clo", "met"]):
+    for ix, var in enumerate(
+        ["ta", "tr", "rh", "met", "vel", "vel_r", "clo", "clo_d", "i_cl_r"]
+    ):
         sns.boxenplot(
             y=var,
             data=df,
[…]

### 2025-01-20 Manuscript/src/sections/introduction.tex
review(federico): finalised review

@@
 Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
-For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards.
+For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards, despite the fact that the two models have different outputs under the same environmental and personal conditions as shown in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
+This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.
+
 In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} standards, respectively.

### 2025-01-20 Manuscript/src/sections/titleAbstract.tex
review(federico): finalised review

@@
         One is the \ac{pmv-ce} standard.
-        The \gls{7730} and \gls{55} are the most widely referenced thermal comfort standards worldwide, and their different \ac{pmv} formulations are a source of confusion.
 
-        We compared the accuracy of the two \ac{pmv} models, against \var{entries_db_used} thermal sensation votes collected in buildings.
+        The \gls{7730} and \gls{55} are the most widely referenced thermal comfort standards worldwide, hence, we compared the accuracy of the two \ac{pmv} models, against \var{entries_db_used} thermal sensation votes collected in buildings.
         We found that overall both models only predict \ac{tsv} correctly only one third of the times.

### 2025-01-20 test_f1_score.py

@@ -1,266 +1,0 @@
-import pandas as pd
-from sklearn.model_selection import train_test_split
-from sklearn.linear_model import LogisticRegression
-from sklearn.metrics import confusion_matrix
-import matplotlib.pyplot as plt
-from sklearn.metrics import accuracy_score, recall_score, precision_score, f1_score
-from sklearn.metrics import classification_report
-import numpy as np
-from pythermalcomfort.models import two_nodes, pmv
-from pprint import pprint
-
-pprint(two_nodes(tdb=35, tr=35, v=0.1, rh=50, met=1.1, clo=0.5))
-pprint(pmv(tdb=35, tr=35, vr=0.1, rh=50, met=1.1, clo=0.5, standard="ASHRAE"))
-pprint(
-    pmv(
-        tdb=35,
-        tr=35,
-        vr=0.1,
-        rh=50,
-        met=1.1,
-        clo=0.5,
-        standard="ISO",
-        limit_inputs=False,
-    )
-)
-
-
-f, ax = plt.subplots(2, 2, constrained_layout=True)
-ax = ax.flatten()
-for clo in [0.5, 1]:
-    for met in [1.1, 2]:
-        results = []
-        for t in np.arange(20, 50, 0.1):
-            r = two_nodes(tdb=t, tr=t, v=0.1, rh=50, met=met, clo=clo)
-            r["tdb"] = t
-            r["clo"] = clo
-            results.append(r)
-
-        df = pd.DataFrame(results)
-        ax[0].plot(df["_set"], df["w"], label=f"{clo=}-{met=}")
-        ax[2].plot(df["tdb"], df["w"], label=f"{clo=}-{met=}")
-        ax[1].plot(df["_set"], df["pmv_gagge"], label=f"{clo=}-{met=}")
-        ax[3].plot(df["tdb"], df["pmv_gagge"], label=f"{clo=}-{met=}")
-        ax[0].legend()
-        ax[1].legend()
-        ax[3].legend()
-        t_pmv_3 = df[df["pmv_gagge"] < 3]["_set"].max()
-        ax[1].axvline(x=t_pmv_3)
-        ax[0].axvline(x=t_pmv_3)
-        t_pmv_3 = df[df["pmv_gagge"] < 3]["tdb"].max()
-        ax[3].axvline(x=t_pmv_3)
-        ax[1].axhline(y=3)
-        ax[3].axhline(y=3)
[…]

### 2025-01-20 Manuscript/src/sections/titleAbstract.tex
Updates from Overleaf

--- 2025-01-17 Manuscript/src/sections/titleAbstract.tex
@@
         We compared the accuracy of the two \ac{pmv} models, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that overall both models only predict \ac{tsv} correctly only one third of the times.
-        The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+        We found that overall both models only predict \ac{tsv} correctly only one-third of the time.
+        The \ac{pmv-ce} has a higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{vr}~$\geq$~\qty{0.2}{\m\per\s}.
 %        todo manually check values

### 2025-01-20 Manuscript/src/response-letter.tex

@@
     \acrodef{pmv}[PMV]{Predicted Mean Vote}
+    \acrodef{pmvce}[PMV$_{CE}$]{ASHRAE Predicted Mean Vote}
     \acrodef{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
@@
 
+        \response{
+            Thank you for the very kind feedback and for providing very detailed comments.
+            We also wish you a Merry Christmas and a Happy New Year.
+        }
+
         \begin{enumerate}
@@
                 Thank you for pointing this out.
-                There was a typo in the manuscript in which we stated that $I_{cl}$ is the total clothing insulation.
-                This is incorrect and we now haved fixed this typo since both the PMV ISO and PMV CE models use the basic clothing insulation ($I_{cl}$) as an input parameter.
+                There was a typo in the manuscript in which we stated that \ac{clo} is the total clothing insulation.
+                This is incorrect and we now haved fixed the error since both the PMV ISO and PMV CE models use the basic clothing insulation (\ac{clo}) as an input parameter.
             }
@@
             \item P4 L34-38.
-            ``The rationale for the development of the PMVCE model is that the original PMV
+            ``The rationale for the development of the \acs{pmvce} model is that the original PMV
             formulation does not accurately estimate convective and evaporative heat losses from
@@
             \response{
-                Thank you for pointing this issue out, this is clearly a important issue about the PMVCE model.
-                Unfortunately we are not certain why the PMVCE subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
-                In Section 1.1 we have tried to summarise all the publicly available information on the PMVCE model.
-                While the ASHRAE 55:2023 standard explains how the PMVCE works, there is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
-                Arens et al. (2009) and Yang et al. (2015) provide a partial justification for the PMVCE model, they do not fully explain why these specific changes to the PMV inputs were made.
-                Moreover, we were not able to find any other peer-reviewed publication that explains in detail why the PMVCE model subtracts the cooling effect from both the air and mean radiant temperature.
+                Thank you for pointing this issue out, this is clearly a important issue about the \ac{pmvce} model.
+                Unfortunately we are not certain why the \ac{pmvce} subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
+                In Section 1.1 we have tried to summarise all the publicly available information on the \ac{pmvce} model.
+                While the ASHRAE 55:2023 standard explains how the \ac{pmvce} works, there is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
+                Arens et al. (2009) and Yang et al. (2015) provide a partial justification for the \ac{pmvce} model, they do not fully explain why these specific changes to the PMV inputs were made.
             }
@@
             \response{
-                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data and that the ASHRAE Global Thermal Comfort Database II v2.1 has limitations.
+                We agree with the reviewer that: 1) the accuracy of the PMV model is influenced by the quality of the input data and that the \ac{db2} has limitations; 2) the PMV model has limitations in evaluating narrow categories of indoor environmental quality.\\
+                \textbf{Measurements errors}\\
                 While we previously mentioned in the manuscript that not all authors used the same measurement protocols.
@@
                 We have also added a few sentences in which we reference the work of d’Ambrosio Alfano et al. (2011) which have already investigated this issue.
-                We have referenced the work of d’Ambrosio Alfano et al.
+                We have also referenced the work of d’Ambrosio Alfano et al.
 (2011) in the manuscript when we justify that the model can be considered accurate if the error between the PMV and the TSV is within $\pm$ 0.5; and we have referenced the work in Section 3.3 measurements errors.\\
-                Finally, while we recognised the limitations of the DB2 in section 2.1., we have also highlighted the measures we have taken to mitigate these limitations.
-                For example, we clarify that if we belive that errors in the inputs data are not systematic across all authors, the use of a large dataset should be able to mitigate the errors caused by the lack of standardised measurement protocols.
+                While we recognised the limitations of the DB2 in section 2.1., we have also highlighted the measures we have taken to mitigate these limitations.
+                For example, we clarify that if we belive that errors in the inputs data are not systematic across all authors, the use of a large dataset should be able to mitigate the errors caused by the lack of standardised measurement protocols.\\
+                \textbf{Categorise the indoor environment}\\
+                In the paper we also explain the issues related to categorising the indoor environment and we explain the issues related with grouping the data as we have done in Fig. 6.
+                Consequently, in the manuscript, we avoided focusing on the classification issue but instead we decided to present the results in terms of difference between the calculated \ac{pmv} value and the reported \ac{tsv}.
             }
@@
             Can we say that this is potentially a limitation of the PMV model?}
-            \todo[color=yellow]{Federico - add the following sentences and reference them in the comment below}
+            \todo[color=yellow]{Federico - check if you need to add the local discomfort sentence in the paper}
             \response{
-                We agree with the reviewer that draughts can be the cause local discomfort and consequently influence the overall thermal comfort sensation.
-                This is a limitation of the PMV model, as it does not account for local discomfort.
-                Dspite the fact that the PMV model does not estimate local discomfort, both standards state that it can be used to predict thermal sensation in a space with air speeds up to 1.0 m/s.
+                We agree with the reviewer that draughts can be the cause local discomfort and consequently influence the overall thermal comfort sensation even when the person is in estimated to be thermally neutral.
+                This is a limitation of the PMV model, which it does not account for local discomfort.
+                In addition, despite the fact that the PMV model does not estimate local discomfort, both standards state that it can be used to predict thermal sensation in a space with air speeds up to 1.0 m/s.
                 Hence, we decided to test the PMV model in the range of air speeds up to 1.0 m/s.
                 This, with other factors, can lead to a discrepancy observed between the PMV and the TSV.
-                Interestingly, we observed that both PMV models have a higher accuracy and lower bias in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s.
+                Interestingly, we observed that both \ac{pmv} models have a higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.
                 This would suggest that while local discomfort due to air movement may affect the overall predicion of the PMV model, it did not affect the results of our analysis.
                 We decided to include a sentence in the manuscipt to highlight the role that local discomfort can play in the overall thermal comfort sensation and the fact that the PMV model does not account for it.
+                We are aware that both Stadards provide equations to calculate thermal discomfort but these equations are not included within the \ac{pmv} model, hence, we did not include them in this analysis.
+                The scope of this paper was to compare only the two \ac{pmv} formulations and not the discomfort equations.
             }
@@
 
-            \todo{Stefano - I cannot fully understand the question, can you help me?}
-            \response{This manuscript aims to determine and compare the accuracy of the PMV models implemented in the ISO 7730:2005 and ASHRAE 55:2023 standards.
-            We did this by comparing the results of the PMV model with the thermal sensation votes collected in the ASHRAE Global Thermal Comfort Database II v2.1.
-            In addition to comparing the accuracy of the two PMV models, we also discussed in the paper that the PMV model is inappropriately used by some authors to predict thermal sensation of individuals outside its range of applicability.
-            As a consequence, we recommend limiting the applicability of the PMV model to the range specified in the Table 2 of the manuscript.
+            \todo{Stefano - I cannot fully understand the question, please check my answer.}
+            \response{This manuscript aims to determine and compare the accuracy of the \ac{pmv} models included in the ISO 7730:2005 and ASHRAE 55:2023 standards.
+            We did this by comparing the results of the \ac{pmv} model with the thermal sensation votes collected in the \ac{db2}.
+            In addition to comparing the accuracy of the two \ac{pmv} models, we also discussed in the paper that the \ac{pmv} model is inappropriately used by some authors to predict thermal sensation of individuals outside its range of applicability.
+            As a consequence, we recommend limiting the applicability of the \ac{pmv} model to the range specified in the Table 2 of the manuscript.
             }
@@
 
-            \todo[color=yellow]{Federico - add the following sentences and reference it in the comment below}
             \response{Thank you for pointing this out.
             We should have explained this better in the manuscript.
-            The confusion arises from the fact that the two standards use different PMV formulations and the results of the two models are different for the same thermal environment when the air speed is higher than 0.1 m/s.
+            The confusion arises from the fact that the two standards use different PMV formulations and the results of the two models are different for the same environmental conditions.
+            The output is significantly different when \ac{vr}~$>$~\qty{0.1}{\m\per\s} as shown in Fig. 2.
             This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.
-            We have added a sentence in the introduction to clarify this point.
+            We have modified the abstract and removed the following text: ``, and their different PMV formulations are a source of confusion.'' since it was not clear.
+            We have also change the text in Section 1.4 which now reads: ``Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
+            For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards, despite the fact that the two models have different outputs under the same environmental and personal conditions as shown in Figure~2.
+This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.''
             }
@@
 
-            \todo[color=yellow]{Federico - add the following sentences and reference it in the comment below}
-            \todo{Stefano - Shall we consider changing how we refer to the air speed and relative air speed in the manuscript?}
-            \todo[color=red]{Federico - check point 2 since we may decide to change the dynamic clothing insulation for the PMV model}
-            \todo[color=red]{Federico - check how we refer to the air speed in the manuscript}
             \response{
-                Please let us clarify this point by answering the questions raised by the reviewr separately:
-                1) Neither the PMV nor the PMVCE model correct the metabolic rate.
-                Both models use as an input the metabolic rate of the participants which is wither measured or estimated by the researcher.
-                2) Both the PMV and PMVCE models correct the clothing insulation based on the metabolic rate using the same equation.
-                3) Both the PMV and PMVCE models correct the air speed based on the metabolic rate using the same equation.
-                The value of V in the paper refers to the adjusted air speed.
+                Thank you for asking for clarifications on this points.
+                We acknowledge that the text in the manuscript was not clear and we have modified it to better describe the differences between the PMV and \ac{pmvce} models.
+                Please let us answer all your points separately below:\\
+                1) Neither the \ac{pmv} nor the \ac{pmvce} model correct the metabolic rate.
+                Both models use as an input the metabolic rate of the participants which is wither measured or estimated by the researcher.\\
+                2) The PMV and \ac{pmvce} models calculate \ac{clor} using different equations.
+                In ordere, to clearly explain this point we have added a new flowchart in Figure 1 which shows the differences between the \ac{pmv} and \ac{pmvce} assumptions.\\
+                3) Both the PMV and \ac{pmvce} models calculate the \acf{vr} using the same equation.
+                In the paper we now clearly differenciate between \acf{v} and \acf{vr}.
                 We have spedified this in the nomeclature section of the manuscript.
-                However, we strongly agree with the reviewer that this point is not clear and this issue arises from how the ASHRAE 55:2023 standard is written.
-                The authors of the ASHRAE 55:2023 standard voted not to have two variables one for the measured air speed and one for the adjusted air speed.
-                We have added a sentence in the manuscript to raise this issue about the wording of the ASHRAE 55:2023 standard.
+                4) we have reworded the sentence ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.'' in the manuscript to better explain when the two models differ.
             }
@@
             \response{
-                We made the decision to include all the data in the ASHRAE Global Thermal Comfort Database II v2.1 in the analysis regardless of the environmental type.
+                We made the decision to include all the data from the \ac{db2} in the analysis regardless of the environmental type.
                 This is because the PMV model, being a heat balance model, can be used to predict thermal sensation in a wide range of environments.
@@
                 We are aware that previous studies have shown that the PMV model may have limitations in predicting thermal sensation in non mechanically ventilated environments.
-                However, since both standards do not specify the type of environments for which the PMV model can be used, we decided to include all the data in the ASHRAE Global Thermal Comfort Database II v2.1 in the analysis.
-                It should be noted that we did not include data from outdoor environments in the analysis since the ASHRAE Global Thermal Comfort Database II v2.1 only includes data from ``real'' buildings occupied by ``real'' people doing their normal day-to-day activities.
-                The database does not include data from outdoor environments nor from climate chambers studies.
+                However, since both standards do not specify the type of environments for which the PMV model can and should be used, we decided to include all the data in the \ac{db2} in the analysis.
+                It should be noted that we did not include data from outdoor environments in the analysis since the \ac{db2} only includes data from \textit{```real' buildings occupied by `real' people doing their normal day-to-day activities.''}
+                The database, hence, does not include data from outdoor environments nor from climate chambers studies.
             }
 
-            \item In the data selection in lines 168--171, it has been stated that ``the PMV should only be used when its absolute value is lower than 2``.
-            So why were data with |PMV| $\leq$ 3.5 retained?
+            \item In the data selection in lines 168--171, it has been stated that ``the PMV should only be used when its absolute value is lower than 2''.
+            So why were data with $|$PMV$|$ $\leq$ 3.5 retained?
 
             \response{
-                We retained data with |PMV| $\leq$ 3.5 because the PMV model is scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
+                We retained data with $|$PMV$|$ $\leq$ 3.5 because the PMV model is scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
+                Hence, we wanted to inform our readers of the accuracy of the PMV model in predicting thermal sensation over the full range of \ac{tsv}.
                 In addition, while the ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
-                Moreover, as explained in the paper the thermal snesation votes in the ASHRAE Global Thermal Comfort Database II v2.1 are collected using a 7-point scale which ranges from -3 to +3.
-                Consequently, we decided to retain data with |PMV| $\leq$ 3.5 and |TSV| $\leq$ 3 in the analysis to compare the accuracy of the PMV model in predicting thermal sensation over the full range of the thermal sensation scale.
+                Moreover, as explained in the paper, the thermal sensation votes in the \ac{db2} are collected using a 7-point scale which ranges from -3 to +3.
+                Consequently, we decided to retain data with $|$PMV$|$~$\leq$~3.5 and $|$TSV$|$~$\leq$~3 in the analysis to compare the accuracy of the PMV model in predicting thermal sensation over the full range of \ac{tsv}.\\
                 However, since the PMV model is not recommended to be used when its absolute value is higher than 2, we strived always to presented the accuracy of the model within its range of applicability.
@@
             When the scale sizes are different, the same number represents different meanings.
-            Is it reasonable to retain data in the form of |TSV| $\leq$ 3?
+            Is it reasonable to retain data in the form of $|$TSV$|$ $\leq$ 3?
 
@@
                 We made a mistake in the body manuscript and we should have written that the ``thermal sensation was measured with \sout{at least} a seven-point scale''.
-                The peer-reviewed paper which describe the ASHRAE Global Thermal Comfort Database II v2.1 states that the thermal sensation votes were collected using a seven-point scale and coded as follows ``-3 cold, -2 cool, -1 slightly cool, 0 neutral, +1 slightly warm, +2 warm, +3 hot''.
+                The peer-reviewed paper which describe the \ac{db2} states that the thermal sensation votes were collected using a seven-point scale and coded as follows ``-3 cold, -2 cool, -1 slightly cool, 0 neutral, +1 slightly warm, +2 warm, +3 hot''.
                 We have corrected this mistake in the manuscript.
@@
 
-            \todo{Stefano - Let's discuss this point during the meeting}
+            \todo{Stefano - Could you please address this point?}
             \response{Here our reply}
@@
                 While the ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
-                Consequently, as preiously explained we decided to include all the data in the ASHRAE Global Thermal Comfort Database II v2.1 in the analysis.
+                Consequently, as previously explained in asnswer 6, we decided to include all the data in the \ac{db2} in the analysis.
                 We, however, tried to further analyze the accuracy of the PMV model in predicting thermal sensation within its range of applicability, for example in Table 1 where we present the F1 score of the PMV model in predicting thermal sensation when its absolute value is lower than 1.5.
@@
                 We agree that limiting the applicability of the PMV model to the range -0.5 to 0.5 may in some instances increase the building operation energy consumption.
-                However, we also show that if the building is correctly designed and operated, for example by providing individual control over the air speed, this may not be the case.
-                For example, in Figure 2 of the manuscript we show that the PMV predicts that occupants (met=1.2 and clo=0.5) are comfortable at t=27.7C when the air speed is 0.4 m/s.
-                An indoor air temperature of 27.7C is higher than the average indoor air temperature in offices in the US, Europe, AU and Singapore in which is generally in the range of 22-24C.
-                Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated and participants do not have individual control over the indoor environment.
+                However, if the building is correctly designed and operated, for example by providing individual control over the air speed, this may not be the case.
+                For example, in Figure 2 of the manuscript we show that the PMV predicts that occupants (met=1.2 and \ac{clor}=0.5) are comfortable at \ac{tdb}=27.7C when \ac{vr}~=~0.4 m/s.
+                A \ac{tdb}=27.7C is higher than the average indoor air temperature in offices in the US, Europe, AU and Singapore in which is generally in the range of \qtyrange{22}{24}{\celsius}.
+                Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated, participants do not have individual control over the indoor environment, or are not allowed to adapt their clothing to the indoor environment.
                 We have now included the following paragraph in the conclusion of the manuscript to address this issue: \textit{``We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
-            However, this is not the case.
-            As shown in Figure 2 increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office settiong to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
-            This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
+                However, this is not always the case.
+                As shown in Figure 2 increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office setting to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
+                This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks with \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
                 Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy, but instead is that in buildings we cool the air rather than moving it.
@@
 
-            \todo{Stefano - Is the reviewer referring to the section about the assumption that people who are ‘slightly warm’ or
-            ‘slightly cool’ are thermally comfortable is incorrect?}
+            \todo{Stefano - Is the reviewer referring to the section about the assumption that people who are ‘slightly warm’ or ‘slightly cool’ are thermally comfortable is incorrect?
+            If so could you please review it and answer the reviewer?
+            Feel free to reference the previous answer about the same topic.}
             \response{Here our reply}
@@
 
+        \response{
+            Thank you for your positive feedback.
+        }
+
         \begin{enumerate}
-            \item I have a only a specific few comments, which are listed below.
+            \item I have only a specific few comments, which are listed below.
             However, my main concern deals with the way the records representing individuals in the comfort DB are used to first calculate PMV for individuals and then compare with individual TSVs.
@@
                 Thank you for your valuable comment.
-                We completely agree with the reviewer that originally the PMV model was not intended to predict the thermal sensation of each individual participant and we have mentioned this as the opening sentence of Section 2.2.
+                We completely agree with the reviewer that originally the \ac{pmv} was not intended to predict the thermal sensation of each individual participant and we have mentioned this in several sections on the paper, for example in Section 2.2.
                 However, the PMV model is widely used in the scientific literature \textit{``in both singly occupied spaces and areas accommodating several hundred people''}.
-                In Section 2.2 of the manuscript we have also added the following sentence \textit{``Partially because neither of the two standards specifies the minimum number of people needed to apply the model.''}
+                In Section 2.2 of the manuscript we have also added the following sentence \textit{``Partially because neither of the two standards specifies the minimum number of people needed to apply the model.''}.\\
                 We are, however, completely aware of this issue and we have not neglected it.
@@
                 These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
-                However, on their own, are a low-precision estimate of the overall accuracy of the model''}
+                However, on their own, are a low-precision estimate of the overall accuracy of the model''.}
                 We have also added the following sentence: \textit{``since the model is not expected to predict the exact \ac{tsv} of each participant}''.
                 We believe that \textit{``If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
-                The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)''}
-                Finally we would want to raise a concern we have about the vague definition of the PMV model in the ISO 7730 and ASHRAE 55 standards.
+                The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)''}.\\
+                Finally we would like to raise a concern we have about the vague definition of the PMV model in both the ISO 7730 and ASHRAE 55 standards.
                 Stating that a model is intended to predict the average thermal sensation of an undefined large group of occupants is not a clear definition and in principle makes it impossible to determine the accuracy of the model.
@@
                 What about when the space is occupied by two or three people?
-                We believe that the Standards should be more clear about the range of applicability of the PMV model.
-                We have, therefore, added a sentence about this in the conclusion of the manuscript which reads as follows: \textit{``Finally, we would like to highlight that the current definition of the \ac{pmv} model ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
+                We believe that the Standards should provide a clearer definition of the range of applicability of the PMV model.
+                We have, therefore, added a sentence about this in the conclusion of the manuscript which reads as follows: \textit{``Finally, we would like to highlight that the current definition of the \ac{pmv} model which states that the model aims ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
                 Both the \gls{55} and \gls{7730} standards should clarify the minimum number of occupants required to use the \ac{pmv} model.
-                Can it be used for a space with only one occupant?''}
+                For example, can it be used for a space with only one occupant?''}
             }
@@
             \response{
-                We acknowledge that the PMV model is not intended to predict the thermal sensation of each individual participant.
+                As per our previous answer, we acknowledge that the PMV model is not intended to predict the thermal sensation of each individual participant.
                 We have acknowledged this limitation in the manuscript and discussed it in Section 2.2 and 3.2.
                 To compensate for this limitation of our analysis, we calculated the bias of the model as previously done by Humphreys and Nicol (2002) to determine if the model is systematically over or under predicting the thermal sensation of the participants.
-                Individual differences in human subjects can be assumed to be random and distributed around the mean.
-                We have then added the following sentence to the manuscript: \textit{``his is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.''}
-                Consequently, we believe that calcualting the bias of the \ac{pmv} models is a valid approach to determine the overall accuracy of the model.
+                Individual differences in human subjects can be assumed to be random and distributed around the mean of a `typical person'.
+                We have then added the following sentence to the manuscript: \textit{``this is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.''}
+                Consequently, we believe that calcualting the bias of the \ac{pmv} models is a valid approach to determine the overall accuracy of the model and that the low prediction accuracy is not an outcome of using the model wrongly.
             }
@@
 
-            \todo{Stefano - do you understand this comment?}
+            \todo{Stefano - how shall we reply to this comment?}
             \response{Here our response}
@@
                 We apologize for the confusion.
-                We have revised the sentence to clarify this point.
+                We have revised the sentence to clarify this point and now reads as follows: \textit{``Additionally, the PMV model erroneously assumes that the human body is not capable of maintaining a stable core temperature.
+                A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicates that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently it is getting warmer or colder, respectively.
+                }
             }
@@
 
-            \item ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent a TSV vote.`` This is wrong - the PMV was never meant to represent a TSV vote, but the average of TSV votes of a large group of people.
+            \item ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent a TSV vote.'' This is wrong - the PMV was never meant to represent a TSV vote, but the average of TSV votes of a large group of people.
 
@@
 
-            \todo{Stefano - How shall we address this comment?}
             \response{
                 Thank you for pointing this out.
-                Most of the studies in the ASHRAE Global Thermal Comfort Database II v2.1 measured air speed one single height.
-                When we specifically compared the the PMV and PMVCE models in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s and for those studies that measured air speed at three different heights, we used the same input data for both models since also the PMV ISO specifies that in non homogeneous environments the air speed should be measured at different heights.
+                Most of the studies in the \ac{db2} measured air speed one single height.
+                Only \num{3059} entries out of \num{49245} entries in the \ac{db2} measured air speed at three different heights.
+                Hence, this should not be significant issue in the analysis.
+%                When we specifically compared the the \ac{pmv} and \ac{pmvce} models in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s and for those studies that measured air speed at three different heights, we used the same input data for both models since also the PMV ISO specifies that in non homogeneous environments the air speed should be measured at different heights.
             }

### 2025-01-20 Manuscript/src/sections/conclusions.tex

@@
 
-Finally, we would like to highlight that the current definition of the \ac{pmv} model ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
+Finally, we would like to highlight that the current definition of the \ac{pmv} model which states that the model aims ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
 Both the \gls{55} and \gls{7730} standards should clarify the minimum number of occupants required to use the \ac{pmv} model.
-Can it be used for a space with only one occupant?
+For example, can it be used for a space with only one occupant?
 

### 2025-01-20 Manuscript/src/sections/titleAbstract.tex

--- 2025-01-20 Manuscript/src/sections/titleAbstract.tex
@@
         The \gls{7730} and \gls{55} are the most widely referenced thermal comfort standards worldwide, hence, we compared the accuracy of the two \ac{pmv} models, against \var{entries_db_used} thermal sensation votes collected in buildings.
-        We found that overall both models only predict \ac{tsv} correctly only one third of the times.
-        The \ac{pmv-ce} has higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+        We found that overall both models only predict \ac{tsv} correctly only one-third of the time.
+        The \ac{pmv-ce} has a higher bias than \ac{pmv} in predicting thermal sensation in the subset of data with \acl{vr}~$\geq$~\qty{0.2}{\m\per\s}.
+
 %        todo manually check values

### 2025-01-20 Manuscript/src/response-letter.tex
review: fixed typos

@@
 
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \hypersetup{
@@
         This manuscript is well-written and showcases your strong capabilities in addressing thermal comfort issues.
-        I have just a few comments to share with you Wishing you a Merry Christmas and a Happy New Year.
+        I have just a few comments to share with you, wishing you a Merry Christmas and a Happy New Year.
 
@@
                 There was a typo in the manuscript in which we stated that \ac{clo} is the total clothing insulation.
-                This is incorrect and we now haved fixed the error since both the PMV ISO and PMV CE models use the basic clothing insulation (\ac{clo}) as an input parameter.
+                This is incorrect and we now have fixed the error since both the PMV ISO and PMV CE models use the basic clothing insulation (\ac{clo}) as an input parameter.
             }
@@
 
-            \todo{Stefano - is the following answer true? Did not you pulish the paper which explains why the CE is subtracted from both the tdb and tr?}
+            \todo{Stefano - is the following answer true? Did not you publish the paper that explains why the CE is subtracted from both the tdb and tr?}
             \response{
-                Thank you for pointing this issue out, this is clearly a important issue about the \ac{pmvce} model.
-                Unfortunately we are not certain why the \ac{pmvce} subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
+                Thank you for pointing this issue out, this is an important issue about the \ac{pmvce} model.
+                Unfortunately, we are not certain why the \ac{pmvce} subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
                 In Section 1.1 we have tried to summarise all the publicly available information on the \ac{pmvce} model.
                 While the ASHRAE 55:2023 standard explains how the \ac{pmvce} works, there is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
-                Arens et al. (2009) and Yang et al. (2015) provide a partial justification for the \ac{pmvce} model, they do not fully explain why these specific changes to the PMV inputs were made.
+                Arens et al.
+(2009) and Yang et al.
+(2015) provide a partial justification for the \ac{pmvce} model, but they do not fully explain why these specific changes to the PMV inputs were made.
             }
@@
             \item P7 L28-32.
-            The authors address (also in methods and in the section 3.3) a highly relevant issue: the impact of microclimatic measurements on the accurate evaluation of thermal comfort.
+            The authors address (also in methods and section 3.3) a highly relevant issue: the impact of microclimatic measurements on the accurate evaluation of thermal comfort.
             Unfortunately, thermal comfort databases suffer from significant biases due to several factors, including the absence of standardized measurement protocols (such as representative positions and height above the ground), incorrect or arbitrary assumptions (for instance, the belief in uniform and stable conditions), the use of inaccurate measurement devices, deficiencies in software, and the failure to assess clothing insulation.
@@
                 \textbf{Measurements errors}\\
-                While we previously mentioned in the manuscript that not all authors used the same measurement protocols.
-                Based on your valuable suggestion we have further improved Section 2.1. and better discussed the limitations of the dataset.
-                We have also added a few sentences in which we reference the work of d’Ambrosio Alfano et al. (2011) which have already investigated this issue.
+                We previously mentioned in the manuscript that not all authors used the same measurement protocols.
+                Based on your valuable suggestion we have further improved Section 2.1. and discussed the limitations of the dataset.
+                We have also added a few sentences in which we reference the work of d’Ambrosio Alfano et al. (2011) who have already investigated this issue.
                 We have also referenced the work of d’Ambrosio Alfano et al.
-(2011) in the manuscript when we justify that the model can be considered accurate if the error between the PMV and the TSV is within $\pm$ 0.5; and we have referenced the work in Section 3.3 measurements errors.\\
+(2011) in the manuscript when we justify that the model can be considered accurate if the error between the PMV and the TSV is within $\pm$ 0.5, and we have referenced the work in Section 3.3 measurements errors.\\
                 While we recognised the limitations of the DB2 in section 2.1., we have also highlighted the measures we have taken to mitigate these limitations.
-                For example, we clarify that if we belive that errors in the inputs data are not systematic across all authors, the use of a large dataset should be able to mitigate the errors caused by the lack of standardised measurement protocols.\\
+                For example, we clarify that if we believe that errors in the input data are not systematic across all authors, the use of a large dataset should be able to mitigate the errors caused by the lack of standardised measurement protocols.\\
                 \textbf{Categorise the indoor environment}\\
-                In the paper we also explain the issues related to categorising the indoor environment and we explain the issues related with grouping the data as we have done in Fig. 6.
-                Consequently, in the manuscript, we avoided focusing on the classification issue but instead we decided to present the results in terms of difference between the calculated \ac{pmv} value and the reported \ac{tsv}.
+                In the paper, we also explain the issues related to categorising the indoor environment and we explain the issues related to grouping the data as we have done in Fig. 6.
+                Consequently, in the manuscript, we avoided focusing on the classification issue but instead, we decided to present the results in terms of the difference between the calculated \ac{pmv} value and the reported \ac{tsv}.
             }
@@
             \item The database used for this investigation also includes high air velocity measurements (v$>$0.2 m/s).
-            However, this does not rule out the possibility of local discomfort caused by draughts.
-            What about the impact of these draughts on the overall thermal comfort as quantified by the Thermal Sensation Vote (TSV)?
+            However, this does not rule out the possibility of local discomfort caused by drafts.
+            What about the impact of these drafts on the overall thermal comfort as quantified by the Thermal Sensation Vote (TSV)?
             Are the authors confident that the TSV values collected at high air velocity are not influenced by local discomfort?
@@
             \response{
-                We agree with the reviewer that draughts can be the cause local discomfort and consequently influence the overall thermal comfort sensation even when the person is in estimated to be thermally neutral.
-                This is a limitation of the PMV model, which it does not account for local discomfort.
-                In addition, despite the fact that the PMV model does not estimate local discomfort, both standards state that it can be used to predict thermal sensation in a space with air speeds up to 1.0 m/s.
+                We agree with the reviewer that drafts can cause local discomfort and consequently influence the overall thermal comfort sensation even when the person is estimated to be thermally neutral.
+                This is a limitation of the PMV model, which does not account for local discomfort.
+                In addition, even though the PMV model does not estimate local discomfort, both standards state that it can be used to predict thermal sensation in a space with air speeds up to 1.0 m/s.
                 Hence, we decided to test the PMV model in the range of air speeds up to 1.0 m/s.
                 This, with other factors, can lead to a discrepancy observed between the PMV and the TSV.
-                Interestingly, we observed that both \ac{pmv} models have a higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.
-                This would suggest that while local discomfort due to air movement may affect the overall predicion of the PMV model, it did not affect the results of our analysis.
-                We decided to include a sentence in the manuscipt to highlight the role that local discomfort can play in the overall thermal comfort sensation and the fact that the PMV model does not account for it.
+                Interestingly, we observed that both \ac{pmv} models have higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.
+                This would suggest that while local discomfort due to air movement may affect the overall prediction of the PMV model, it did not affect the results of our analysis.
+                We decided to include a sentence in the manuscript to highlight the role that local discomfort can play in the overall thermal comfort sensation and the fact that the PMV model does not account for it.
                 We are aware that both Stadards provide equations to calculate thermal discomfort but these equations are not included within the \ac{pmv} model, hence, we did not include them in this analysis.
@@
         This article conducts a comparative analysis of the accuracy of the PMV model implemented in ISO 7730:2005 and ASHRAE 55:2023 using data from the ASHRAE 2 database.
-        The research has good practical significance, but there are some issues that need further clarification:
+        The research has good practical significance, but some issues need further clarification:
 
@@
             We did this by comparing the results of the \ac{pmv} model with the thermal sensation votes collected in the \ac{db2}.
-            In addition to comparing the accuracy of the two \ac{pmv} models, we also discussed in the paper that the \ac{pmv} model is inappropriately used by some authors to predict thermal sensation of individuals outside its range of applicability.
-            As a consequence, we recommend limiting the applicability of the \ac{pmv} model to the range specified in the Table 2 of the manuscript.
+            In addition to comparing the accuracy of the two \ac{pmv} models, we also discussed in the paper that the \ac{pmv} model is inappropriately used by some authors to predict the thermal sensation of individuals outside its range of applicability.
+            As a consequence, we recommend limiting the applicability of the \ac{pmv} model to the range specified in Table 2 of the manuscript.
             }
@@
             \item It is stated in the abstract that ``The ISO7730:2005 and ASHRAE55:2023 are the most widely referenced thermal comfort standards worldwide, and their different PMV formulations are a source of confusion.''
-            However, it does not actually clarify what kind of confusion has been caused.
+            However, it does not clarify what kind of confusion has been caused.
 
@@
             We have modified the abstract and removed the following text: ``, and their different PMV formulations are a source of confusion.'' since it was not clear.
-            We have also change the text in Section 1.4 which now reads: ``Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
-            For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards, despite the fact that the two models have different outputs under the same environmental and personal conditions as shown in Figure~2.
+            We have also changed the text in Section 1.4 which now reads: ``Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
+            For example, the WELL certification allows compliance with \gls{7730} and \gls{55} standards, even though the two models have different outputs under the same environmental and personal conditions as shown in Figure~2.
 This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.''
@@
             \response{
-                Thank you for asking for clarifications on this points.
+                Thank you for asking for clarifications on this point.
                 We acknowledge that the text in the manuscript was not clear and we have modified it to better describe the differences between the PMV and \ac{pmvce} models.
@@
                 1) Neither the \ac{pmv} nor the \ac{pmvce} model correct the metabolic rate.
-                Both models use as an input the metabolic rate of the participants which is wither measured or estimated by the researcher.\\
+                Both models use as an input the metabolic rate of the participants which is either measured or estimated by the researcher.\\
                 2) The PMV and \ac{pmvce} models calculate \ac{clor} using different equations.
-                In ordere, to clearly explain this point we have added a new flowchart in Figure 1 which shows the differences between the \ac{pmv} and \ac{pmvce} assumptions.\\
+                In order, to clearly explain this point we have added a new flowchart in Figure 1 which shows the differences between the \ac{pmv} and \ac{pmvce} assumptions.\\
                 3) Both the PMV and \ac{pmvce} models calculate the \acf{vr} using the same equation.
-                In the paper we now clearly differenciate between \acf{v} and \acf{vr}.
-                We have spedified this in the nomeclature section of the manuscript.
+                In the paper, we now clearly differentiate between \acf{v} and \acf{vr}.
+                We have specified this in the nomenclature section of the manuscript.
                 4) we have reworded the sentence ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.'' in the manuscript to better explain when the two models differ.
@@
             \response{
-                We made the decision to include all the data from the \ac{db2} in the analysis regardless of the environmental type.
+                We decided to include all the data from the \ac{db2} in the analysis regardless of the environmental type.
                 This is because the PMV model, being a heat balance model, can be used to predict thermal sensation in a wide range of environments.
-                Our aims was to determine the accuracy of the PMV model in predicting thermal sensation as a function of the input parameters.
+                We aimed to determine the accuracy of the PMV model in predicting thermal sensation as a function of the input parameters.
                 Environmental type is not an input parameter of the PMV model, hence we did not consider it in the analysis.
-                We are aware that previous studies have shown that the PMV model may have limitations in predicting thermal sensation in non mechanically ventilated environments.
+                We are aware that previous studies have shown that the PMV model may have limitations in predicting thermal sensation in non-mechanically ventilated environments.
                 However, since both standards do not specify the type of environments for which the PMV model can and should be used, we decided to include all the data in the \ac{db2} in the analysis.
                 It should be noted that we did not include data from outdoor environments in the analysis since the \ac{db2} only includes data from \textit{```real' buildings occupied by `real' people doing their normal day-to-day activities.''}
-                The database, hence, does not include data from outdoor environments nor from climate chambers studies.
+                The database, hence, does not include data from outdoor environments or climate chamber studies.
             }
@@
             \response{
-                We retained data with $|$PMV$|$ $\leq$ 3.5 because the PMV model is scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
+                We retained data with $|$PMV$|$ $\leq$ 3.5 because the PMV model in scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
                 Hence, we wanted to inform our readers of the accuracy of the PMV model in predicting thermal sensation over the full range of \ac{tsv}.
-                In addition, while the ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
+                In addition, while ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
                 Moreover, as explained in the paper, the thermal sensation votes in the \ac{db2} are collected using a 7-point scale which ranges from -3 to +3.
                 Consequently, we decided to retain data with $|$PMV$|$~$\leq$~3.5 and $|$TSV$|$~$\leq$~3 in the analysis to compare the accuracy of the PMV model in predicting thermal sensation over the full range of \ac{tsv}.\\
-                However, since the PMV model is not recommended to be used when its absolute value is higher than 2, we strived always to presented the accuracy of the model within its range of applicability.
+                However, since the PMV model is not recommended to be used when its absolute value is higher than 2, we strived always to present the accuracy of the model within its range of applicability.
                 For example, in Figure 6 we present the accuracy of the PMV model in predicting each thermal sensation category separately.
@@
                 We made a mistake in the body manuscript and we should have written that the ``thermal sensation was measured with \sout{at least} a seven-point scale''.
-                The peer-reviewed paper which describe the \ac{db2} states that the thermal sensation votes were collected using a seven-point scale and coded as follows ``-3 cold, -2 cool, -1 slightly cool, 0 neutral, +1 slightly warm, +2 warm, +3 hot''.
+                The peer-reviewed paper which describes the \ac{db2} states that the thermal sensation votes were collected using a seven-point scale and coded as follows ``-3 cold, -2 cool, -1 slightly cool, 0 neutral, +1 slightly warm, +2 warm, +3 hot''.
                 We have corrected this mistake in the manuscript.
@@
 
-            \item It should be noted that the PMV model proposed by Fanger is applicable to steady-state indoor thermal environments that do not deviate too much from neutrality.
+            \item It should be noted that the PMV model proposed by Fanger applies to steady-state indoor thermal environments that do not deviate too much from neutrality.
             Therefore, it is unreasonable to use overheated or overcooled data to verify the accuracy of the PMV model, and the demonstrated accuracy will also be relatively low.
@@
             \response{
-                We agree with the reviewer that the PMV model is applicable to steady-state indoor thermal environments that do not deviate too much from neutrality.
+                We agree with the reviewer that the PMV model applies to steady-state indoor thermal environments that do not deviate too much from neutrality.
                 However, the PMV model is widely used in the scientific literature to predict thermal sensation in a wide range of thermal environments.
-                While the ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
-                Consequently, as previously explained in asnswer 6, we decided to include all the data in the \ac{db2} in the analysis.
+                While ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
+                Consequently, as previously explained in answer 6, we decided to include all the data in the \ac{db2} in the analysis.
                 We, however, tried to further analyze the accuracy of the PMV model in predicting thermal sensation within its range of applicability, for example in Table 1 where we present the F1 score of the PMV model in predicting thermal sensation when its absolute value is lower than 1.5.
@@
 
-            \item For highlights 4--5, whether in field studies or laboratory studies, PMV values generally ranging from -1 to 1 are defined as thermal comfort, which is appropriate for steady-state indoor environments close to neutrality.
+            \item For highlights 4--5, whether in field studies or laboratory studies, PMV values generally ranging from -1 to 1 is defined as thermal comfort, which is appropriate for steady-state indoor environments close to neutrality.
             However, if it is narrowed down to -0.5 to 0.5, the thermal comfort range is greatly reduced, which will increase the building operation energy consumption and is not recommended.
@@
                 Thank you for this very valuable comment.
-                We agree that limiting the applicability of the PMV model to the range -0.5 to 0.5 may in some instances increase the building operation energy consumption.
+                We agree that limiting the applicability of the PMV model to the range from \numrange{-.5}{.5} may in some instances increase the building operation energy consumption.
                 However, if the building is correctly designed and operated, for example by providing individual control over the air speed, this may not be the case.
-                For example, in Figure 2 of the manuscript we show that the PMV predicts that occupants (met=1.2 and \ac{clor}=0.5) are comfortable at \ac{tdb}=27.7C when \ac{vr}~=~0.4 m/s.
-                A \ac{tdb}=27.7C is higher than the average indoor air temperature in offices in the US, Europe, AU and Singapore in which is generally in the range of \qtyrange{22}{24}{\celsius}.
-                Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated, participants do not have individual control over the indoor environment, or are not allowed to adapt their clothing to the indoor environment.
+                For example, in Figure 2 of the manuscript, we show that the PMV predicts that occupants (met=1.2 and \ac{clor}=0.5) are comfortable at \ac{tdb}=27.7C when \ac{vr}~=~0.4 m/s.
+                A \ac{tdb}=27.7C is higher than the average indoor air temperature in offices in the US, Europe, AU and Singapore which is generally in the range of \qtyrange{22}{24}{\celsius}.
+                Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated, participants do not have individual control over the indoor environment or are not allowed to adapt their clothing to the indoor environment.
                 We have now included the following paragraph in the conclusion of the manuscript to address this issue: \textit{``We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
@@
                 This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks with \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
-                Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy, but instead is that in buildings we cool the air rather than moving it.
+                Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy but instead is that in buildings we cool the air rather than moving it.
                 A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}.''}
@@
         \textbf{Reviewer 3}
-        The manuscript presents an excellent analysis of the discrepancies between two versions of the PMV model and evaluates their prediction accuracy.
-        The analysis is informative and definitely relevant for researchers and professionals within thermal comfort.
+        The manuscript presents an excellent analysis of the discrepancies between the two versions of the PMV model and evaluates their prediction accuracy.
+        The analysis is informative and relevant for researchers and professionals within thermal comfort.
         The structure of the manuscript and the presentation of the topic are well laid out.
@@
             \item I have only a specific few comments, which are listed below.
-            However, my main concern deals with the way the records representing individuals in the comfort DB are used to first calculate PMV for individuals and then compare with individual TSVs.
-            As the authors write ``The intended aim of the PMV model is not to accurately predict each individual thermal response from participants.
+            However, my main concern deals with the way the records representing individuals in the comfort DB are used to first calculate PMV for individuals and then compare them with individual TSVs.
+            As the authors write: ``The intended aim of the PMV model is not to accurately predict individual thermal response from participants.
             The PMV model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.``
-            But isn't this analysis relying on calculation of each individual TSV without considering that this is not what the model should be used for?
+            But isn't this analysis relying on the calculation of individual TSV without considering that this is not what the model should be used for?
             The approach is the same across a wealth of studies that examine the accuracy of the model, but that does not make it more correct.
-            I actually doubt that the DB can be reliably used to estimate the prediction accuracy, as there is no information of the mean TSV of groups of people exposed to the same thermal environment.
+            I doubt that the DB can be reliably used to estimate the prediction accuracy, as there is no information on the mean TSV of groups of people exposed to the same thermal environment.
 
@@
                 Thank you for your valuable comment.
-                We completely agree with the reviewer that originally the \ac{pmv} was not intended to predict the thermal sensation of each individual participant and we have mentioned this in several sections on the paper, for example in Section 2.2.
+                We completely agree with the reviewer that originally the \ac{pmv} was not intended to predict the thermal sensation of each participant and we have mentioned this in several sections of the paper, for example in Section 2.2.
                 However, the PMV model is widely used in the scientific literature \textit{``in both singly occupied spaces and areas accommodating several hundred people''}.
@@
                 We are, however, completely aware of this issue and we have not neglected it.
-                In the manuscript we stated that \textit{``we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
+                In the manuscript, we stated that \textit{``we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
                 These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
@@
                 We believe that \textit{``If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
-                The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)''}.\\
-                Finally we would like to raise a concern we have about the vague definition of the PMV model in both the ISO 7730 and ASHRAE 55 standards.
+                The standard deviation would reflect the combined effect of the people's differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)''}.\\
+                Finally, we would like to raise a concern we have about the vague definition of the PMV model in both the ISO 7730 and ASHRAE 55 standards.
                 Stating that a model is intended to predict the average thermal sensation of an undefined large group of occupants is not a clear definition and in principle makes it impossible to determine the accuracy of the model.
-                At the same time it does not help the user to understand when the model can be used and when it cannot.
-                For example, can it be used to design a single office or a small appartment for a single person?
+                At the same time, it does not help the user to understand when the model can be used and when it cannot.
+                For example, can it be used to design a single office or a small apartment for a single person?
                 What about when the space is occupied by two or three people?
@@
 
-            \item Although less critical, another general comment is that it is not clear what is meant by low prediction accuracy, as is written already in the abstract - when is the prediction accuracy low?
-            We aim to estimate what people perceive, which seriously suffers from individual differences and many others challenges.
+            \item Although less critical, another general comment is that it is not clear what is meant by low prediction accuracy, as is written already in the abstract -- when is the prediction accuracy low?
+            We aim to estimate what people perceive, which seriously suffers from individual differences and many other challenges.
             What is the prediction accuracy of related models?
-            Or is the low prediction accuracy an outcome of using the prediction model wrongly (see first general comment).
+            Or is the low prediction accuracy an outcome of using the prediction model wrongly (see first general comment)?
 
             \response{
-                As per our previous answer, we acknowledge that the PMV model is not intended to predict the thermal sensation of each individual participant.
-                We have acknowledged this limitation in the manuscript and discussed it in Section 2.2 and 3.2.
-                To compensate for this limitation of our analysis, we calculated the bias of the model as previously done by Humphreys and Nicol (2002) to determine if the model is systematically over or under predicting the thermal sensation of the participants.
+                As per our previous answer, we acknowledge that the PMV model is not intended to predict the thermal sensation of each participant.
+                We have acknowledged this limitation in the manuscript and discussed it in Sections 2.2 and 3.2.
+                To compensate for this limitation of our analysis, we calculated the bias of the model as previously done by Humphreys and Nicol (2002) to determine if the model is systematically over or under-predicting the thermal sensation of the participants.
                 Individual differences in human subjects can be assumed to be random and distributed around the mean of a `typical person'.
-                We have then added the following sentence to the manuscript: \textit{``this is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.''}
-                Consequently, we believe that calcualting the bias of the \ac{pmv} models is a valid approach to determine the overall accuracy of the model and that the low prediction accuracy is not an outcome of using the model wrongly.
+                We have then added the following sentence to the manuscript: \textit{``This is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.''}
+                Consequently, we believe that calculating the bias of the \ac{pmv} models is a valid approach to determine the overall accuracy of the model and that the low prediction accuracy is not an outcome of using the model wrongly.
             }
@@
             \item Sources of error.
-            Order of importance - is it really the heat transfer calculation that is the most important - not the attempt to calculate thermal sensation based on physical variables
+            Order of importance -- is it the heat transfer calculation that is the most important -- not the attempt to calculate thermal sensation based on physical variables
 
@@
             In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature [24].``
-            Indeeed, the thermoregulatory system maintains a stable body temperature, but the body still exchanges heat with the environment, so suggest to revise this sentence.
+            Indeed, the thermoregulatory system maintains a stable body temperature, but the body still exchanges heat with the environment, so suggest revising this sentence.
 
@@
             Calculation of heat loss by sweating.
-            Actually, this equation applies to people in thermal comfort, which should probably be added to the statement, so heat lost due to sweating does not appear to be related only with met rate.
+            This equation applies to people in thermal comfort, which should probably be added to the statement, so heat lost due to sweating does not appear to be related only to met rate.
 
@@
 
-            \item ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent a TSV vote.'' This is wrong - the PMV was never meant to represent a TSV vote, but the average of TSV votes of a large group of people.
+            \item ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent a TSV vote.'' This is wrong -- the PMV was never meant to represent a TSV vote, but the average of TSV votes of a large group of people.
 
@@
             \item Comment on measurement of air speed.
-            ISO7726, which is used as a guide for measurements in ISO7730, recommends measurement of air speed only at abdomen level in homogeneous environments in both class C (comfort) and S (stress).
-            The air speed measured in this height should be used to calculate the PMV\@.
-            When comparing 7730 and 55, especially with emphasis on air speed, this should probably be taken into account.
+            ISO7726, which is used as a guide for measurements in ISO7730, recommends the measurement of air speed only at the abdomen level in homogeneous environments in both class C (comfort) and S (stress).
+            The air speed measured at this height should be used to calculate the PMV\@.
+            When comparing 7730 and 55, especially with an emphasis on air speed, this should probably be taken into account.
 
@@
                 Thank you for pointing this out.
-                Most of the studies in the \ac{db2} measured air speed one single height.
+                Most of the studies in the \ac{db2} measured air speed at one single height.
                 Only \num{3059} entries out of \num{49245} entries in the \ac{db2} measured air speed at three different heights.
-                Hence, this should not be significant issue in the analysis.
-%                When we specifically compared the the \ac{pmv} and \ac{pmvce} models in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s and for those studies that measured air speed at three different heights, we used the same input data for both models since also the PMV ISO specifies that in non homogeneous environments the air speed should be measured at different heights.
+                Hence, this should not be a significant issue in the analysis.
+%                When we specifically compared the \ac{pmv} and \ac{pmvce} models in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s and for those studies that measured air speed at three different heights, we used the same input data for both models since also the PMV ISO specifies that in non-homogeneous environments the air speed should be measured at different heights.
             }

### 2025-01-20 Manuscript/src/sections/appendix.tex
review: fixed typos

@@
     \input{tables/publications_with_three_heights.tex}
-    \caption{The number of entries used in this analysis grouped by publication.
+    \caption{The number of entries used in this analysis, grouped by publication.
     We did not include papers with less than \num{5} entries.}
@@
 The \ac{pmv} model has a bias of \num{-.39} while the \ac{pmv-ce} model has a bias of \num{-.38}.
-These results allow us to conclude that the \ac{pmv} has a similar accuracy than the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
+These results allow us to conclude that the \ac{pmv} has a similar accuracy to the \ac{pmv-ce} model for the dataset with air speed measured at three heights.
 

### 2025-01-20 Manuscript/src/sections/conclusions.tex
review: fixed typos

@@
 The \ac{pmv-ce} was intentionally developed to more accurately estimate the effect of elevated air movement~\cite{arens_moving_2009}.
-We determined the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models by comparing their results to the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1.
+We determined the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models by comparing their results to the \ac{tsv} available in the \acl{db2}.
 
@@
 \subsection{Implications for the \gls{55} and \gls{7730} Standards and \ac{pmv} Users}\label{subsec:implications-for-the-ashrae-55-and-iso-7730-standards}
-Researchers, educators and practitioners worldwide may wonder which model they should use.
+Researchers, educators, and practitioners worldwide may wonder which model they should use.
 The \gls{55} standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that proves it has superior prediction performances.
@@
 Based on our results, we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
-In other words, we only recommend using it for determining whether an environment may be deemed thermally neutral by a large group of occupants.
+In other words, we only recommend using it to determine whether an environment may be deemed thermally neutral by a large group of occupants.
 The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classify if individuals are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
@@
 However, this is not the case.
-As shown in Figure~\ref{fig:comfort_regios_pmv_pmvce} increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office settiong to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
+As shown in Figure~\ref{fig:comfort_regios_pmv_pmvce} increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office setting to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
 This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
@@
 Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
-\todo{unclear what limits you are talking about, or who the audience is for your suggestions.
-The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.
-Your suggestions are actually just aimed at yourselves as keepers of the ComfortTool}
+\todo{unclear what limits you are talking about, or who the audience is for your suggestions.  The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.  
+Your suggestions are just aimed at you as keepers of the ComfortTool}
 The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
@@
         \ac{met} & \qtyrange{1}{2}{met} \\
-        \ac{pmv} & \numrange{-.5}{.5}{} \\
+        \ac{pmv} & \numrange{-.5}{.5} \\
         \bottomrule

### 2025-01-20 Manuscript/src/sections/introduction.tex
review: fixed typos

--- 2025-01-17 Manuscript/src/sections/introduction.tex
@@
 \mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} standard~\cite{iso7730} in its original form.
-The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories `cold,' `cool,' `slightly cool,' `neutral,' `slightly warm,' `warm,' and `hot.'~\cite{iso7730, ashrae552023}.
+The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \numrange{-3}{3} corresponding to the categories `cold,' `cool,' `slightly cool,' `neutral,' `slightly warm,' `warm,' and `hot.'~\cite{iso7730, ashrae552023}.
 The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
-It is the most extensively used thermal comfort index.
-A search within the article title, abstract, and keyword on Scopus with the search term `predicted mean vote' returns approximately \num{2050} documents.
-Out of those \num{1200} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
-This highlights this model's extensive adoption and use among the scientific community.
-To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, Humphreys2002, doherty_evaluation_1988}.
+It is the most widely used thermal comfort index.
+Searching within the article title, abstract, and keyword on Scopus with the search term `predicted mean vote' returns approximately \num{2050} documents.
+Of these \num{1200} are peer-reviewed scientific articles published in the research areas of engineering, environmental science, social sciences, and energy.
+This highlights the extensive adoption and use of this model among the scientific community.
+To this date, the \ac{pmv} remains to be the most utilized thermal comfort model, even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, Humphreys2002, doherty_evaluation_1988}.
 
@@
 \todo[inline, color=blue!40]{
-    FT similar to my previous comment. Firstly, I would like to mention here that the PMV does not at all try to estimate sweating. This is a problem, but neither the PMV nor the PMV CE do that.
+    FT. Similar to my previous comment. Firstly, I would like to mention here that the PMV does not at all try to estimate sweating. This is a problem, but neither the PMV nor the PMV CE do that.
     The Std 55 did not modify the PMV heat balance equation, instead it simply and only reduced the air and mean radiant temperatures by the cooling effect, this is not equivalent to doing an accurate heat balance calculation.
-    The PMV CE consequently does not: 1) uses the Gagge two-node model to compute the convective and evaporative heat loss from the body 2) model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss.
-    In addition, another comment about the PMV CE is that the cooling effect is not only applied in warm conditions but in fact it is applied in any conditions.
+    The PMV CE consequently does not: 1) use the Gagge two-node model to compute the convective and evaporative heat loss from the body and 2) model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss.
+    In addition, another comment about the PMV CE is that the cooling effect is not only applied in warm conditions but is applied in any conditions.
     SET is also a rather confusing variable since it is calculated not for a reference environment with standardised clo and met.
 But rather it is the temperature of a hypothetical isothermal environment at 50\% rh, <0.1 m/s (20 fpm) average air speed Va, and in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level.
-    Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} this however do not ensure that the person is not feeling hot in this condition or sweating, consequently does not solve the issues that the original PMV has.
+    Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} this however does not ensure that the person is not feeling hot in this condition or sweating, consequently does not solve the issues that the original PMV has.
 }
@@
 However, we could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the \gls{55} over the original \ac{pmv} model.
+\todo[inline]{true enough, but in the standard, the objective is to have a rational physiology model.
+The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
+\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}
 
-Figure~\ref{fig:flowchart_pmv_calculation} shows side by side the calculation routines for the \ac{pmv} and \ac{pmv-ce} models.
+Figure~\ref{fig:flowchart_pmv_calculation} shows side-by-side the calculation routines for the \ac{pmv} and \ac{pmv-ce} models.
 \begin{figure}[!htb]
@@
             $\alpha$ is the correction coefficient for \acs{clo}.
-            Since the \gls{7730} does contain some errors and it does not provide all the necessary information to calculate $\alpha$ we used the equations provided in the \gls{9920} to derive the value of $\alpha$.
-        The \gls{7730} standard, also provides an equation to calculat the walking speed if this information is not available.
+            Since the \gls{7730} does contain some errors and does not provide all the necessary information to calculate $\alpha$ we used the equations provided in the \gls{9920} to derive the value of $\alpha$.
+        The \gls{7730} standard, also provides an equation to calculate the walking speed if this information is not available.
         }
@@
 
-Consequently, while the two \ac{pmv} formulations mainly differ when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
+Consequently, the two \ac{pmv} formulations mainly differ when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
 It should be noted that the two models also use different equations to calculate the \ac{clor}.
@@
 We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
-\todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. There are plenty of lab studies that have measured results in that region.}
+\todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. Many lab studies have measured results in that region.}
 \todo[color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
@@
 
-The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds~\cite{huang_applicability_2014}.
-The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}.
+The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate the convective and evaporative heat losses from the skin to the environment at elevated air speeds~\cite{huang_applicability_2014}.
+The \ac{pmv} does not accurately estimate evaporative heat losses through sweating because it assumes that sweating is constant and only varies as a function of \ac{met}.
 Hence, if sweating is required to maintain the body in thermal equilibrium, the \ac{pmv} model assumes that the body would get hotter.
 The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
-\todo[inline]{true enough, but in the standard the objective is to have a rational physiology model.
-The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
-\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}
 
-Despite this claim from the authors of the \ac{pmv-ce}, our analysis has, however, revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.
+However, despite this claim of the authors of \ac{pmv-ce}, our analysis has revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.
 
@@
     For example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.15}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
-    According to the Gagge's two-node model, which is used to calculate the \ac{set}, this hypothetical person would have \qty{30}{\percent} of their skin covered with sweat.
+    According to Gagge's two-node model, which is used to calculate the \ac{set}, this hypothetical person would have \qty{30}{\percent} of their skin covered with sweat.
     Based on the results from \mycite{huang_applicability_2014}, the \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating.
@@
     Even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating.
-    Therefore, in this scenario, where the \ac{pmv-ce} should excell compared to the \ac{pmv}, the \ac{pmv-ce} model still fails to accurately predict the \ac{tsv}.
+    Therefore, in this scenario, where the \ac{pmv-ce} should excel compared to the \ac{pmv}, the \ac{pmv-ce} model still fails to accurately predict the \ac{tsv}.
     \todo[inline]{This is really confusing.
-    You are evaluating the Cooling Effect in still air, when there is basically no Cooling Effect to evaluate.
+    You are evaluating the Cooling Effect in still air, when there is no Cooling Effect to evaluate.
     You are using as an example 29C in still air and 80\% RH which of course will cause sweating.
@@
 You allude to results from Huang 2014 that the reader is supposed to know about.
-The skin wettedness discussion is irrelevant unless you go into a lot more physiological detail here.}
+The skin-wettedness discussion is irrelevant unless you go into a lot more physiological detail here.}
     \item The previous example, highlights another limitation of the \ac{pmv-ce}.
@@
     However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.
-    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{vr}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-    However, while \ac{vr}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh} to \qty{50}{\percent} and standardized the clothing to the activity.
+    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{vr}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
+    However, while \ac{vr}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh}~=~\qty{50}{\percent} and standardized the clothing to the activity.
     \todo[inline]{I think this argument is incorrect.
@@
 What’s in there now was Tyler’s (CBE’s) choice.
-Charlie and I have used it in the simpler way.
+Charlie and I have used it more simply.
 We could propose this to the Standards Committee if we wanted to.
@@
 Among the most notable formulations there are the \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET},  \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
-\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
-It should be noted that, while both the \ac{pmvs} and \ac{pmvg} models are based on the \ac{set} model they significantly differ from the \ac{pmv-ce} model.
-Both the \ac{pmvs} and \ac{pmvg} are directly calculated within the \ac{set} model using the estimated heat gains and losses from the human body.
+\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some models mentioned above.
+It should be noted that, while both the \ac{pmvs} and \ac{pmvg} models are based on the \ac{set} model, they significantly differ from the \ac{pmv-ce} model.
+Both the \ac{pmvs} and \ac{pmvg} are directly calculated within the \ac{set} model using the estimated heat gains and losses of the human body.
 
-Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models since they are incorporated in the most widely referenced thermal comfort standards worldwide (\gls{7730} and \gls{55}).
+Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models, since they are incorporated in the most widely referenced thermal comfort standards worldwide (\gls{7730} and \gls{55}).
 
@@
 The \ac{pmv} was free from bias only when it was used to predict thermal neutrality~\cite{Humphreys2002}.
-\mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
+\mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the \ac{db2}.
 They found that the thermal sensation predicted by the PMV model, on average, is one full thermal sensation scale unit away from the subject’s responses, confirming the results of~\mycite{Humphreys2002}.
@@
 \subsection{Objectives}\label{subsec:aim-and-objectives}
-Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
-For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards.
+Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators, and practitioners worldwide, since both models are widely used in building codes, guidelines, and certification programs.
+For example, the WELL certification allows both compliance with \gls{7730} and \gls{55} standards, even though the two models have different outputs under the same environmental and personal conditions, as shown in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
+This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.
+
 In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} standards, respectively.

### 2025-01-20 Manuscript/src/sections/methodology.tex
review: fixed typos

@@
 and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
-These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} datapoints.
+These criteria removed an additional \var{entries_db_filtered_by_limit_inputs} data points.
 Consequently, the final dataset used for the analysis contains \var{entries_db_used} entries.
@@
 
-In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same standardised methodology.
+In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same standardized methodology.
 For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
-The absence of standardized measurement protocols, the use of inaccurate or un-calibrated sensors, and the use of different reference tables to estimate \ac{clo} and \ac{met} may all introduce an error in the measurements and, consequently, in the results of both models.
-This issue has been already evidenced in the literature by \mycite{dambrosio_alfano_role_2011}.
+The absence of standardized measurement protocols, the use of inaccurate or uncalibrated sensors, and the use of different reference tables to estimate \ac{clo} and \ac{met} may all introduce an error in the measurements and, consequently, in the results of both models.
+This issue has already been evidenced in the literature by \mycite{dambrosio_alfano_role_2011}.
 They showed that the reliability of the \ac{pmv} model is strongly affected by the accuracy of the input data.
@@
 \begin{enumerate}[ {}1{)} ]
-    \item comparing the results of the two models using the same dataset; hence, the different in the output is mainly due to the different formulations of the models;
-    \item using a large dataset to reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards since it can be assumed that measurement errors random across different studies.
+    \item comparing the results of the two models using the same dataset; hence, the difference in the output is mainly due to the different formulations of the models;
+    \item using a large dataset to reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards, since it can be assumed that measurement errors are random across different studies.
     \item using only datasets that have been published in peer-reviewed journals.
@@
 If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
-The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
-This is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.
+The standard deviation would reflect the combined effect of the people's differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
+This is a similar assumption to the one used by the \ac{pmv} model, which ignores individual differences and calculates the average thermal sensation of a `typical' average person.
 According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{\pm 0.25}.
@@
 Ideally, we would want to have the \ac{pmv} to have higher precision.
-however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale and by the accuracy of the inputs.
+However, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale and by the accuracy of the inputs.
 When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.

### 2025-01-20 Manuscript/src/sections/results.tex
review: fixed typos

@@
 \end{figure*}
-Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex, and are almost equally distributed among males (\qty{52}{\percent}) and females.
+Less than half (\var{entries_with_sex_info}) of the total entries have information about the participant's sex and are almost equally distributed among males (\qty{52}{\percent}) and females.
 Ages are not normally distributed, the same is true for the running mean outdoor temperature.
@@
         Figure b) shows the total number of data points grouped by \ac{tsv}.
-    Each bin in the right figure has more data points than in the left figure, since less data on \ac{tpv} were available.}
+    Each bin in the right figure has more data points than in the left figure since less data on \ac{tpv} were available.}
     \label{fig:bar_plot_tp_by_ts}
@@
 In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable~\cite{Fanger1970, schweiker2020evaluating}.
-Fanger, in his seminal book on thermal comfort wrote that ``The dissatisfied are defined here as those who vote \num{-2} (cool) or \num{-3} (cold), \num{2} (warm) or \num{3} (hot).
+Fanger, in his seminal book on thermal comfort, wrote that ``The dissatisfied are defined here as those who vote \num{-2} (cool) or \num{-3} (cold), \num{2} (warm) or \num{3} (hot).
 One could perhaps object that those voting \num{-1} or \num{1} were not included also, but as evidenced by Gagge et al., real discomfort is first expressed by those voting higher than \num{2} or lower than \num{-2}.''~\cite{Fanger1970}.
@@
 \subsection{Model Bias}\label{subsec:model-bias}
-The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
+The intended aim of the \ac{pmv} model is not to accurately predict individual thermal responses from participants.
 The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
@@
 Moreover, grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
-To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
+To compensate for this, we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
 \begin{figure*}[htb!]
@@
 \subsubsection{Model Overall Bias}\label{subsubsec:model-overall-bias}
-We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
+We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and \ac{pmv-ce} predictions from the self-reported \ac{tsv}.
 Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
 % todo check data manually
-Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
+Overall, both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
 \begin{figure*}[htb!]
@@
     We have binned the data into intervals of equal width.
-    We colored the bars in the range from \num{-.5} to \num{.5} to highlight the number of responses that are within this range.
+    We colored the bars in the range from \numrange{-.5}{.5} to highlight the number of responses that are within this range.
     We report the median and inter-quantile ranges since data are not normally distributed.
@@
 The x-axis tick label shows the bin range.
-All the boxen-plots have the same area to aid visualization, even though the number of points in each bin is not constant.
+All the boxen plots have the same area to aid visualization, even though the number of points in each bin is not constant.
 The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
@@
     \item if the previous assumption holds, research is needed to validate the correlation coefficients used by Fanger to map the heat losses/gains and thermal stress;
-    \item currently a single correlation coefficient is used for the hot and cold side but this may not be a valid assumption and two different sets of equations may be used;
+    \item currently a single correlation coefficient is used for the hot and cold side, but this may not be a valid assumption and two different sets of equations may be used;
     \item the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
@@
 The latter hypothesis -- 4) anchoring -- has gained significant momentum in the literature, which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equations~\cite{Yao2022, Toftum2002}.
-Supporting evidence was provided by \mycite{schweiker2020evaluating}, who showed that the distances between the anchors are not constant and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
+Supporting evidence was provided by \mycite{schweiker2020evaluating}, who showed that the distances between the anchors are not constant, and a layperson may not perceive the \ac{tsv} scale as researchers believe they do~\cite{schweiker2019scales, schweiker2020evaluating}.
 
@@
 This is a problem since the \ac{pmv} model assumes that the air speed is uniform across the human body, and this is not always the case.
-Measuring the air speed at a single height may lead to record higher values than the averaged velocities that are required for the standards’ comfort models.
-This is because the anemometer is almost always placed in relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently the associated sensation prediction will be cooler than it should be.
+Measuring the air speed at a single height may lead to recording higher values than the averaged velocities that are required for the standards’ comfort models.
+This is because the anemometer is almost always placed in a relatively open space between mid-body and head level, avoiding measuring the usually sheltered lower body level, consequently, the associated sensation prediction will be cooler than it should be.
 To account for this error, we only included the studies where the air speed was measured at three different heights in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
@@
 Using a continuous scale, may in principle be a more robust approach since the distances between the anchors are not constant~\cite{schweiker2019scales, schweiker2020evaluating}.
-They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
+They also showed that the \ac{tsv} assumption of its independence from contextual factors such as climate, season, and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
 An alternative solution would be to transform the \ac{pmv} output to an ordinal seven-point scale.

### 2025-01-28 Manuscript/src/response-letter.tex
review: addressed all reviewer's comments (with Ed's comments)

@@
 % Opening statement
-        \opening{Dear Athanasios Tzempelikos and Reviewers,}
+        \opening{Dear Professor Tzempelikos and Reviewers,}
 
@@
 
-        We would like to thank you for your time and your valuable feedback.
-        Please find below our answers to all your comments.
+        We thank you for your time and your valuable feedback. 
+        We addressed all your comments.
+        You can find our answers to all your comments below.
 
@@
             Thank you for the very kind feedback and for providing very detailed comments.
-            We also wish you a Merry Christmas and a Happy New Year.
+            Thank you for the wishes, we hope you had a joyful and restful break.
         }
@@
 
-            \todo{Stefano - is the following answer true? Did not you publish the paper that explains why the CE is subtracted from both the tdb and tr?}
             \response{
                 Thank you for pointing this issue out, this is an important issue about the \ac{pmvce} model.
-                Unfortunately, we are not certain why the \ac{pmvce} subtracts the cooling effect from both the air and mean radiant temperature, as we were not involved in the model's development.
+                Unfortunately, we are unsure why the \ac{pmvce} subtracts the cooling effect from the air and mean radiant temperature. 
                 In Section 1.1 we have tried to summarise all the publicly available information on the \ac{pmvce} model.
-                While the ASHRAE 55:2023 standard explains how the \ac{pmvce} works, there is no peer-reviewed publication that quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
-                Arens et al.
-(2009) and Yang et al.
-(2015) provide a partial justification for the \ac{pmvce} model, but they do not fully explain why these specific changes to the PMV inputs were made.
+                While the ASHRAE 55:2023 standard explains how the \ac{pmvce} works, no peer-reviewed publication quantifies the accuracy improvements of the model as implemented in the ASHRAE standard over the original PMV model.
+                We informally discussed this with some of the authors.
+                Arens et al. (2009) partially justify the \ac{pmvce} model, but they do not fully explain why these specific changes to the PMV inputs were made. 
+                The original model (ASHRAE Standard 55-2013) accounted for the entire cooling effect of air movement on SET by applying it solely to the dry bulb temperature. 
+                However, Yang et al.
+(2015) highlighted substantial differences between the cooling effect measured with a thermal manikin (dry losses only) and the effect calculated using the approach from ASHRAE Standard 55-2013.
+                Specifically, they found that the cooling effect estimated by the standard was nearly double that of the manikin’s measurement. 
+                This discrepancy led to situations where the difference between air temperature and mean radiant temperature (MRT) appeared unreasonably large. 
+                To address this issue, Yang et al. demonstrated that applying the cooling effect to both air temperature and MRT produced much closer estimations. 
+                This revised approach was subsequently adopted in the updated version of the standard. 
+                It’s important to note that the thermal manikin used in their study measured only dry heat losses and the assessment was done in conditions in which sweating was low. 
+                Yang et al. concluded: "To assess which method for SET* implementation and which air speed measurements should be used, experiments with human subjects are needed". 
+                To our knowledge experiments have not been done yet.  
             }
@@
             \response{
-                We agree with the reviewer that: 1) the accuracy of the PMV model is influenced by the quality of the input data and that the \ac{db2} has limitations; 2) the PMV model has limitations in evaluating narrow categories of indoor environmental quality.\\
+                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data and that the \ac{db2} has substantial limitations, as mentioned by the reviewer.
                 \textbf{Measurements errors}\\
-                We previously mentioned in the manuscript that not all authors used the same measurement protocols.
-                Based on your valuable suggestion we have further improved Section 2.1. and discussed the limitations of the dataset.
+                We previously mentioned in the manuscript that not all studies included in the \ac{db2} used the same measurement protocols. 
+                Based on your valuable suggestion, we have further improved Section 2.1. and discussed the limitations of the dataset.
                 We have also added a few sentences in which we reference the work of d’Ambrosio Alfano et al. (2011) who have already investigated this issue.
@@
                 While we recognised the limitations of the DB2 in section 2.1., we have also highlighted the measures we have taken to mitigate these limitations.
-                For example, we clarify that if we believe that errors in the input data are not systematic across all authors, the use of a large dataset should be able to mitigate the errors caused by the lack of standardised measurement protocols.\\
-                \textbf{Categorise the indoor environment}\\
-                In the paper, we also explain the issues related to categorising the indoor environment and we explain the issues related to grouping the data as we have done in Fig. 6.
+                For example, we clarify that if we believe that errors in the input data are not systematic across all authors, the use of a large dataset could, at least in part, counterbalance the errors caused by the lack of standardised measurement protocols.
+                In science and technology, researchers often face the challenge of balancing sample size and data quality within a fixed financial or time budget. 
+                Both extremes offer valuable insights: small, laboratory, high-quality datasets provide precise information, while large, less accurate datasets obtained in field studies enable more "ecologically valid" and broader trends and patterns to emerge.
+                In recent years, the value of large, less precise datasets has profoundly transformed the way science is conducted, driving new discoveries and methodologies.\\
+                \textbf{Categorise the thermal environment}\\
+                In the paper, we also explain the issues related to categorising the thermal environment and we explain the issues related to grouping the data as we have done in Fig. 6.
                 Consequently, in the manuscript, we avoided focusing on the classification issue but instead, we decided to present the results in terms of the difference between the calculated \ac{pmv} value and the reported \ac{tsv}.
@@
 
-            \todo{Stefano shall we consider adding a sentence about the local discomfort in the paper?
-            Can we say that this is potentially a limitation of the PMV model?}
-            \todo[color=yellow]{Federico - check if you need to add the local discomfort sentence in the paper}
-            \response{
-                We agree with the reviewer that drafts can cause local discomfort and consequently influence the overall thermal comfort sensation even when the person is estimated to be thermally neutral.
-                This is a limitation of the PMV model, which does not account for local discomfort.
-                In addition, even though the PMV model does not estimate local discomfort, both standards state that it can be used to predict thermal sensation in a space with air speeds up to 1.0 m/s.
-                Hence, we decided to test the PMV model in the range of air speeds up to 1.0 m/s.
-                This, with other factors, can lead to a discrepancy observed between the PMV and the TSV.
-                Interestingly, we observed that both \ac{pmv} models have higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.
-                This would suggest that while local discomfort due to air movement may affect the overall prediction of the PMV model, it did not affect the results of our analysis.
-                We decided to include a sentence in the manuscript to highlight the role that local discomfort can play in the overall thermal comfort sensation and the fact that the PMV model does not account for it.
-                We are aware that both Stadards provide equations to calculate thermal discomfort but these equations are not included within the \ac{pmv} model, hence, we did not include them in this analysis.
-                The scope of this paper was to compare only the two \ac{pmv} formulations and not the discomfort equations.
-            }
-
+            \response{In a cool environment, high air speed can cause local discomfort and consequently influence the overall thermal sensation.
+            In a warm environment, high air speed can be perceived as beneficial and also influences the overall thermal sensation.
+Prof.
+Jørn Toftum summarized the dual nature of air movement in the following article: Toftum, Jørn.
+“Air Movement – Good or Bad?” Indoor Air 14 (August 2, 2004): 40–45.
+https://doi.org/10.1111/j.1600-0668.2004.00271.x.
+            In this study, we did not attempt to evaluate local discomfort models, such as radiant asymmetry, thermal stratification, floor temperature, draft, ankle draft, etc. 
+            Some of these models differ significantly between the ISO and ASHRAE standards. 
+            Besides airspeed, we anticipate that other environmental factors, such as air temperature stratification, could influence overall thermal sensation. 
+            However, since neither standard mathematically integrates local discomfort models with the overall PMV, they are treated independently, therefore we chose not to include them in our analysis.
+            Minor note: Interestingly, we observed that both \ac{pmv} models have higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.}
         \end{enumerate}
@@
             When Professor Fanger initially proposed the PMV model, there were certain applicable conditions.
-            Does the article aim to explore the accuracy of the PMV model, or the inappropriate application of this model in the two standards?
+            Does the article aim to explore the accuracy of the PMV model or the inappropriate application of this model in the two standards?
 
-            \todo{Stefano - I cannot fully understand the question, please check my answer.}
-            \response{This manuscript aims to determine and compare the accuracy of the \ac{pmv} models included in the ISO 7730:2005 and ASHRAE 55:2023 standards.
-            We did this by comparing the results of the \ac{pmv} model with the thermal sensation votes collected in the \ac{db2}.
-            In addition to comparing the accuracy of the two \ac{pmv} models, we also discussed in the paper that the \ac{pmv} model is inappropriately used by some authors to predict the thermal sensation of individuals outside its range of applicability.
-            As a consequence, we recommend limiting the applicability of the \ac{pmv} model to the range specified in Table 2 of the manuscript.
+            \response{We aimed to determine and compare the accuracy of the \ac{pmv} models implemented in the ISO 7730:2005 and ASHRAE 55:2023 standards. 
+            We did not plan to assess the accuracy of the original Professor Fanger model. 
+            As far as we know, the ISO 7730:2005 implementation is closer to the original model than the ASHRAE 55:2023.
             }
@@
             We should have explained this better in the manuscript.
-            The confusion arises from the fact that the two standards use different PMV formulations and the results of the two models are different for the same environmental conditions.
-            The output is significantly different when \ac{vr}~$>$~\qty{0.1}{\m\per\s} as shown in Fig. 2.
+            The confusion arises from the fact that the two standards use different PMV algorithms and the results of the two models differ for the same environmental conditions, the difference is most pronounced when \ac{vr}~$>$~\qty{0.1}{\m\per\s} as shown in Fig.
+2.
             This can lead to confusion when comparing results from different studies or when trying to apply the results in practice. 
+            Some researchers may want guidance to decide which model they should use to report their data. 
+            Some practitioners may have the option to select the model to use, for example, some certification programs, including LEED, let the designer decide which model to use. 
             We have modified the abstract and removed the following text: ``, and their different PMV formulations are a source of confusion.'' since it was not clear.
@@
                 1) Neither the \ac{pmv} nor the \ac{pmvce} model correct the metabolic rate.
-                Both models use as an input the metabolic rate of the participants which is either measured or estimated by the researcher.\\
+                Both models use as an input the metabolic rate of the participants which is either measured or estimated.\\
                 2) The PMV and \ac{pmvce} models calculate \ac{clor} using different equations.
@@
             \response{
-                We completely agree with the reviewer that the first limitation is not a limitation of the model itself.
+                We agree with the reviewer that the first limitation is not a limitation of the model itself.
                 Hence, we have removed it from the limitations section and moved it to the beginning of Section 1.1.
@@
                 We decided to include all the data from the \ac{db2} in the analysis regardless of the environmental type.
-                This is because the PMV model, being a heat balance model, can be used to predict thermal sensation in a wide range of environments.
-                We aimed to determine the accuracy of the PMV model in predicting thermal sensation as a function of the input parameters.
-                Environmental type is not an input parameter of the PMV model, hence we did not consider it in the analysis.
-                We are aware that previous studies have shown that the PMV model may have limitations in predicting thermal sensation in non-mechanically ventilated environments.
-                However, since both standards do not specify the type of environments for which the PMV model can and should be used, we decided to include all the data in the \ac{db2} in the analysis.
-                It should be noted that we did not include data from outdoor environments in the analysis since the \ac{db2} only includes data from \textit{```real' buildings occupied by `real' people doing their normal day-to-day activities.''}
+                This is because the PMV model can be applied to predict thermal sensation across a wide range of environments, including naturally conditioned spaces. 
+                It is worth noting that standards, such as ASHRAE 55:2023, allow the use of “adaptive comfort” in “occupant-controlled naturally conditioned spaces” when all the conditions outlined in Section 5.4.1 are met.
+                However, they do not prohibit the use of the PMV model in naturally ventilated spaces. 
+                In fact, the PMV model can be applied in "all occupied spaces within the scope of this standard" (Section 5.3.1). 
+                It should be noted that we did not include data from outdoor environments in the analysis since the \ac{db2} only includes data from \textit{```real' buildings occupied by `real' people doing their normal day-to-day activities.''}.
                 The database, hence, does not include data from outdoor environments or climate chamber studies.
@@
             \response{
-                We retained data with $|$PMV$|$ $\leq$ 3.5 because the PMV model in scientific peer-reviewed papers is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
+                We retained data with $|$PMV$|$ $\leq$ 3.5 because the PMV model in scientific peer-reviewed papers and in practice is widely used to predict the thermal sensation of people even when its absolute value is higher than 2.
                 Hence, we wanted to inform our readers of the accuracy of the PMV model in predicting thermal sensation over the full range of \ac{tsv}.
@@
                 Consequently, we decided to retain data with $|$PMV$|$~$\leq$~3.5 and $|$TSV$|$~$\leq$~3 in the analysis to compare the accuracy of the PMV model in predicting thermal sensation over the full range of \ac{tsv}.\\
-                However, since the PMV model is not recommended to be used when its absolute value is higher than 2, we strived always to present the accuracy of the model within its range of applicability.
+                However, since the PMV model is not recommended to be used when its absolute value is higher than 2, we strived to present the model's accuracy within its range of applicability.
                 For example, in Figure 6 we present the accuracy of the PMV model in predicting each thermal sensation category separately.
                 In Table 1, we present the F1 score of the PMV model in predicting thermal sensation when its absolute value is lower than 1.5.
-                We acknowledge that we did not report the bias of the PMV model in predicting thermal sensation when its absolute value is higher than 2.
-                We may include this figure if the reviewers believe it is necessary.
                 We have changed the text in the manuscript as follows: \textit{``Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2.
@@
 
-            \todo{Stefano - Could you please address this point?}
-            \response{Here our reply}
+            \response{Thermal comfort is defined as the state of mind that expresses \textbf{satisfaction} with the thermal environment. 
+            Your comment rightly highlights distinctions related to thermal preference, thermal sensation, and thermal comfort. 
+            However, it is important to note that thermal comfort can only be assessed directly through a thermal satisfaction scale (or a thermal comfort scale, though this risks tautology (circular definition) in its definition).
+            We believe thermal preference is a more practical and actionable scale than thermal sensation for building control, as it explicitly indicates what the system should do. 
+            For instance, if someone states they “prefer cooler,” it is unequivocal that the system should lower the temperature in an air-conditioned building or recommend opening a window in a naturally ventilated space.
+            In contrast, if someone reports feeling “slightly warm,” it remains unclear how the system should respond; some may decide to classify it as comfortable, others as uncomfortable. 
+            While both scales are valid and have their merits, we favor the use of thermal preference and have reported its implications in this study. 
+            It is worth noting that this distinction is a secondary objective of our work, but we believe it is an important point to address. 
+            We modified the text as reported in answer 11 and made our assumption explicit, so, it is not presented as a universal fact, but conditional to our assumption).}
 
@@
             \response{
-                We agree with the reviewer that the PMV model applies to steady-state indoor thermal environments that do not deviate too much from neutrality.
-                However, the PMV model is widely used in the scientific literature to predict thermal sensation in a wide range of thermal environments.
                 While ISO 7730 specifies that the PMV model should only be used when its absolute value is lower than 2, the ASHRAE 55 does not specify the range of applicability of the PMV model.
@@
             }
-
             \item For highlights 4--5, whether in field studies or laboratory studies, PMV values generally ranging from -1 to 1 is defined as thermal comfort, which is appropriate for steady-state indoor environments close to neutrality.
@@
                 Thank you for this very valuable comment.
-                We agree that limiting the applicability of the PMV model to the range from \numrange{-.5}{.5} may in some instances increase the building operation energy consumption.
-                However, if the building is correctly designed and operated, for example by providing individual control over the air speed, this may not be the case.
+                The ASHRAE Standard defines comfort as PMV between -0.5 and 0.5. 
+                The European standard proposes four classes. 
+                We agree that limiting the applicability of the PMV model to the range from \numrange{-.5}{.5} may, in some instances, increase the building operation energy consumption.
+                However, in some conditions that is not necessarily the case, for example, by providing individual control over the air speed.
                 For example, in Figure 2 of the manuscript, we show that the PMV predicts that occupants (met=1.2 and \ac{clor}=0.5) are comfortable at \ac{tdb}=27.7C when \ac{vr}~=~0.4 m/s.
-                A \ac{tdb}=27.7C is higher than the average indoor air temperature in offices in the US, Europe, AU and Singapore which is generally in the range of \qtyrange{22}{24}{\celsius}.
-                Consequently, we believe that mechanically ventilated buildings are not efficient not because the PMV mode is too restrictive, but because the buildings are not correctly designed and operated, participants do not have individual control over the indoor environment or are not allowed to adapt their clothing to the indoor environment.
+                A \ac{tdb}=27.7C is higher than the average indoor air temperature in offices in the US, Europe, Australia and Singapore which is generally in the range of \qtyrange{22}{24}{\celsius}.
                 We have now included the following paragraph in the conclusion of the manuscript to address this issue: \textit{``We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
@@
                 This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks with \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
-                Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy but instead is that in buildings we cool the air rather than moving it.
+                Hence, we would like to point out that the \ac{pmv} model should be used in a range where it is accurate, and then the designer should look for solutions that provide a more sustainable built environment.
                 A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}.''}
@@
             \item Some conclusion contents lack data and theoretical support.
-
-            \todo{Stefano - Is the reviewer referring to the section about the assumption that people who are ‘slightly warm’ or ‘slightly cool’ are thermally comfortable is incorrect?
-            If so could you please review it and answer the reviewer?
-            Feel free to reference the previous answer about the same topic.}
-            \response{Here our reply}
+            \response{
+            We assume that the reviewer referred to the thermal preference conclusions. 
+            We modified it by making our assumption, as described in the answer above, explicit and clear. 
+            We modified the text as follows: "If we assume that selecting “warmer” or “cooler” on the thermal preference scale indicates thermal discomfort, then our results suggest that individuals who report being slightly warm' or slightly cool’ on the thermal sensation scale are thermally comfortable is incorrect.
+            This misinterpretation can lead to an underestimation of thermal discomfort. 
+            In the \ac{db2} \qty{68}{\percent} of participants who reported to be `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who reported to be `slightly cool' wanted to be `warmer'.}
 
@@
                 Thank you for your valuable comment.
-                We completely agree with the reviewer that originally the \ac{pmv} was not intended to predict the thermal sensation of each participant and we have mentioned this in several sections of the paper, for example in Section 2.2.
+                We agree with the reviewer that originally the \ac{pmv} was not intended to predict the thermal sensation of each participant and we have mentioned this in several sections of the paper, for example in Section 2.2.
                 However, the PMV model is widely used in the scientific literature \textit{``in both singly occupied spaces and areas accommodating several hundred people''}.
                 In Section 2.2 of the manuscript we have also added the following sentence \textit{``Partially because neither of the two standards specifies the minimum number of people needed to apply the model.''}.\\
-                We are, however, completely aware of this issue and we have not neglected it.
+                We are, however, aware of this issue and we have not neglected it.
                 In the manuscript, we stated that \textit{``we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
@@
                 Finally, we would like to raise a concern we have about the vague definition of the PMV model in both the ISO 7730 and ASHRAE 55 standards.
-                Stating that a model is intended to predict the average thermal sensation of an undefined large group of occupants is not a clear definition and in principle makes it impossible to determine the accuracy of the model.
+                Stating that a model is intended to predict the average thermal sensation of an undefined large group of occupants is not a clear definition and, in principle, it makes it impossible to determine the accuracy of the model.
                 At the same time, it does not help the user to understand when the model can be used and when it cannot.
-                For example, can it be used to design a single office or a small apartment for a single person?
+                For example, can it be used to design a single office?
                 What about when the space is occupied by two or three people?
@@
             \response{
-                As per our previous answer, we acknowledge that the PMV model is not intended to predict the thermal sensation of each participant.
-                We have acknowledged this limitation in the manuscript and discussed it in Sections 2.2 and 3.2.
-                To compensate for this limitation of our analysis, we calculated the bias of the model as previously done by Humphreys and Nicol (2002) to determine if the model is systematically over or under-predicting the thermal sensation of the participants.
+                Estimating perceived thermal comfort is inherently challenging due to individual differences and contextual factors, both of which significantly influence prediction accuracy. 
+                There is no universal definition of high, medium, or low accuracy, as accuracy is context-dependent and should be evaluated based on the potential consequences of an error. 
+                For example, in critical applications such as detecting cancer or making autonomous vehicle braking decisions, extremely high accuracy is essential. 
+                In less consequential scenarios, such as determining whether an environment feels warm or cool, more flexibility is acceptable. 
+                Negligible accuracy in any context can be considered equivalent to random guessing. 
+                In our case, with seven thermal sensation categories, random guessing would result in an accuracy of approximately 14 percent. 
+                Achieving a correct prediction one out of three times (around 33 percent) can be considered low accuracy because it is equivalent to a very simple model based solely on temperature achieved a similar level of accuracy (see Section 6 and Figure 7 of Cheung, Toby, Stefano Schiavon, Thomas Parkinson, Peixian Li, and Gail Brager, “Analysis of the Accuracy on PMV – PPD Model Using the ASHRAE Global Thermal Comfort Database II,” Building and Environment, 153, no.
+15 (April 15, 2019): 205–17.
+https://doi.org/10.1016/j.buildenv.2019.01.055).
+                One way to evaluate accuracy is to compare it with what could be achieved by other models. 
+                For example, personal comfort models, now applied in smart thermostats like Nest or in car HVAC systems such as Tesla’s, have demonstrated accuracy levels exceeding 95 percent. 
+                However, these models are not yet widely applicable to design scenarios. 
+                Nevertheless, they tell us where we should aim. 
+                We also drew on previous studies, such as the work by Humphreys and Nicol (2002), which provide a benchmark for evaluating accuracy. 
+                In line with their approach, we calculated the bias of the model to determine whether it systematically over- or under-predicts participants’ thermal sensations.
                 Individual differences in human subjects can be assumed to be random and distributed around the mean of a `typical person'.
@@
 
-            \todo{Stefano - how shall we reply to this comment?}
-            \response{Here our response}
+            \response{We have combined the discussion of heat transfer issues and the intrinsic complexity of predicting people’s responses into a single paragraph in the paper, we did not rank them. 
+            The title of this paragraph has been updated to: \textit{``Heat Balance Equations and Thermal Sensation Predictions.} 
+            In the text, we emphasize: \textit{``The issues are reported in order of importance of what we believe may affect the accuracy of the model.} 
+            This statement clarifies that we do not have quantifiable metrics to rank these factors; instead, the order reflects our expert judgment regarding their potential impact on model accuracy.}
 
@@
             \response{
-                We apologize for the confusion.
+                We apologize for the mistake.
                 We have revised the sentence to clarify this point and now reads as follows: \textit{``Additionally, the PMV model erroneously assumes that the human body is not capable of maintaining a stable core temperature.
-                A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicates that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently it is getting warmer or colder, respectively.
+                A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicates that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently, it is getting warmer or colder, respectively.
                 }
@@
 
-            \todo{Stefano - I do not understand this comment, can you help me?}
-            \response{Here our response}
+            \response{Agree, we modified the text as follows: \textit{``This amount of heat loss is included in the equation even if a person is in thermal neutrality (feeling comfortable) or has a negative heat gain (feeling cooler than neutral).}}
 
@@
             \response{
-                Thank you for pointing this out.
-                We changed the sentence as follows: ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent the average thermal sensation of a large group of occupants.''
+                Please see our answer above about the weak definition of what is "a large group of people" and the fact that standards allow the PMV model to be used in places occupied by only one person.
+                Nevertheless, we changed the sentence as follows: \textit{``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent the average thermal sensation of a large group of occupants.}
             }
@@
                 Hence, this should not be a significant issue in the analysis.
-%                When we specifically compared the \ac{pmv} and \ac{pmvce} models in predicting thermal sensation in the subset of data with air speed greater than 0.2 m/s and for those studies that measured air speed at three different heights, we used the same input data for both models since also the PMV ISO specifies that in non-homogeneous environments the air speed should be measured at different heights.
             }
@@
 
-        Federico Tartarini
+        Federico Tartarini and Stefano Schiavon
 

### 2025-01-28 Manuscript/src/sections/conclusions.tex
review: addressed all reviewer's comments (with Ed's comments)

@@
 
-Our results also show that the assumption that people who are `slightly warm' or `slightly cool' are thermally comfortable is incorrect and leads to under-counting thermal discomfort.
-In the \ac{db2} \qty{68}{\percent} of participants who were `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who were `slightly cool' wanted to be `warmer'.
+If we assume that selecting `warmer' or `cooler' on the thermal preference scale indicates thermal discomfort, then our results suggest that individuals who report being `slightly warm' or `slightly cool’ on the thermal sensation scale are thermally comfortable is incorrect.
+This misinterpretation can lead to an underestimation of thermal discomfort. 
+In the \ac{db2} \qty{68}{\percent} of participants who reported to be `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who reported to be `slightly cool' wanted to be `warmer'.
 
@@
 The \gls{55} standard explains how the \ac{pmv-ce} works, however, we could not find a peer-reviewed publication that proves it has superior prediction performances.
-The \ac{pmv-ce} is more complex and computationally intensive to run, while not being more accurate than the original \ac{pmv} model.
+The \ac{pmv-ce} is more complex and computationally intensive to run, while not being more accurate than the original \ac{pmv} model, moreover, it does not explicitly solve the lack of proper sweat rate modeling of the \ac{pmv} model.
 Hence, we suggest researchers should calculate the \ac{pmv} values according to the \gls{7730} standards.
 
-The very low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits should be reconsidered.
+The low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits should be reconsidered.
 Based on our results, we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
@@
 The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classify if individuals are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
-Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988} provide corroborating evidence that supports our results.
+The \gls{55} standard specifies only the conditions that are considered comfortable (within the 0.5 range) but does not provide any explicit interpretation for values above or below that range, other than stating that they are not compliant with the standard.
+Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988} provided corroborating evidence that supports our results.
 We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption. 
-However, this is not the case.
+However, this is not necessarily the case.
 As shown in Figure~\ref{fig:comfort_regios_pmv_pmvce} increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office setting to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
 This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
-Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy, but instead is that in buildings we cool the air rather than moving it.
+Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy but instead is that we cool the air rather than move it.
 A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}~\cite{kent_energy_2023}.
@@
 Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
-\todo{But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
+\todo{EA. But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
+\todo[inline, color=blue!40]{FT. We clarified this issue above.}
 Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
@@
 Your suggestions are just aimed at you as keepers of the ComfortTool}
+\todo[inline, color=blue!40]{FT. The limits are the one described in Table 2, for example, we suggest to use the model only from 0.3 and 1.5 clo and met between 1 and 2.}
 The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
@@
 This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.
-\todo{This is a major suggestion with very significant implications for energy etc that have not been discussed in the paper.
+\todo{EA.
+This is a major suggestion with very significant implications for energy etc that have not been discussed in the paper.
 There are surely reasons people have not used that scale on its own, for this purpose.}
+\todo[inline, color=blue!40]{FT.
+We discussed this in Figure 5a and in its related text.
+The PMV model assume that people are comfortable when they claim to be slightly warm or slightly cool, this is explicit in Fanger's book.}
 
@@
 \section*{Acknowledgements}\label{sec:acknowledgements}
-We would like to thank Prof. Edward Arens and Dr. Paul Raftery for their comments and suggestions on the manuscript.
+We would like to thank Prof. Edward Arens, Dr. Charlie Huizenga and Dr. Paul Raftery for their comments and suggestions on the manuscript.

### 2025-01-28 Manuscript/src/sections/introduction.tex
review: addressed all reviewer's comments (with Ed's comments)

@@
 \todo[inline, color=blue!40]{
-    FT. Similar to my previous comment. Firstly, I would like to mention here that the PMV does not at all try to estimate sweating. This is a problem, but neither the PMV nor the PMV CE do that.
-    The Std 55 did not modify the PMV heat balance equation, instead it simply and only reduced the air and mean radiant temperatures by the cooling effect, this is not equivalent to doing an accurate heat balance calculation.
-    The PMV CE consequently does not: 1) use the Gagge two-node model to compute the convective and evaporative heat loss from the body and 2) model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss.
-    In addition, another comment about the PMV CE is that the cooling effect is not only applied in warm conditions but is applied in any conditions.
-    SET is also a rather confusing variable since it is calculated not for a reference environment with standardised clo and met.
-But rather it is the temperature of a hypothetical isothermal environment at 50\% rh, <0.1 m/s (20 fpm) average air speed Va, and in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level.
-    Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} this however does not ensure that the person is not feeling hot in this condition or sweating, consequently does not solve the issues that the original PMV has.
+    FT. Firstly, it is important to note that the PMV does not aim to estimate sweating. 
+    While this is a limitation, neither the PMV nor the PMV CE models include this capability.  
+    The Std 55 did not modify the PMV heat balance equation but instead reduced the air and mean radiant temperatures by the cooling effect, assuming this adequately accounts for sweating. 
+    However, this approach does not equate to an accurate heat balance calculation where sweat rate is explicitly modeled, as seen in the Gagge two-node model.  
+    Consequently, the PMV CE does not: 1) use the Gagge two-node model to compute convective and evaporative heat loss from the body, nor 2) explicitly model physiological cooling in warm conditions where sweat evaporation becomes a key source of heat loss.
+    Secondly, in the PMV CE model, the cooling effect is applied in all conditions, not just in warm ones.  
+    Thirdly, the definition of SET can also be somewhat unclear. 
+    It is not calculated for a reference environment with standardised clothing and metabolic rates. 
+    Instead, it represents the temperature of a hypothetical isothermal environment at 50\% RH, $<$0.1 m/s (20 fpm) average air velocity, where the total heat loss from the skin of an imaginary occupant (wearing clothing standardised for the activity) equals the heat loss from a person in the actual environment with their actual clothing and activity level.
+    Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} However, this does not guarantee that the person will not feel hot or sweat in this condition and, therefore, does not address the inherent issues of the original PMV model.
 }
@@
 However, we could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the \gls{55} over the original \ac{pmv} model.
-\todo[inline]{true enough, but in the standard, the objective is to have a rational physiology model.
+\todo[inline]{EA. true enough, but in the standard, the objective is to have a rational physiology model.
 The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
-\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}
+\todo[inline, color=blue!40]{FT.
+The PMV CE and the Gagge model are different, hence, publications about the latter do not sufficiently describe the rationale of the new model, moreover, the accuracy of the PMV CE was not tested}
 
@@
 \todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. Many lab studies have measured results in that region.}
-\todo[color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
+\todo[color=blue!40]{FT. In the manuscript, we decided not to include a Figure for 0.6m/s since in the ASHRAE DB 2 we hardly reach that value.
+I am happy to generate a figure for you if need be}
 \begin{figure}[!htb]
@@
 \begin{enumerate}
-    \item Although the \ac{ce} claims to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model, hence it has the same limitations.
-    For example, let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.15}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
-    According to Gagge's two-node model, which is used to calculate the \ac{set}, this hypothetical person would have \qty{30}{\percent} of their skin covered with sweat.
-    Based on the results from \mycite{huang_applicability_2014}, the \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating.
-    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this scenario, is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement (delta between \qty{0.15}{\m\per\s} and \qty{0.1}{\m\per\s}) which is equal to \qty{.17}{\celsius}.
+    \item Although the \ac{ce} claims to better account for the heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same \ac{pmv} model equations, hence it has the same limitations.
+    This can be shown with a thought-example.
+    Let's assume that a person is exposed to the following environmental conditions \ac{tdb}~=~\ac{tr}~=~\qty{29}{\celsius}, \ac{rh}~=~\qty{80}{\percent}, \ac{vr}~=~\qty{0.25}{\m\per\s}, they have a \ac{met} of \qty{1}{met} and a \ac{clor} of \qty{0.5}{clo}.
+    In these conditions, a person will be sweating; in fact, according to Gagge's two-node model, which is used to calculate the \ac{set}, a person would have approximately \qty{30}{\percent} of their skin covered with sweat.
+    The \ac{pmv} model, in this scenario, cannot accurately predict the \ac{tsv} since the person is sweating \mycite{huang_applicability_2014}.
+    However, the only difference between the \ac{pmv} and \ac{pmv-ce} in this thought example is that the \ac{pmv-ce} accounts for the cooling effect due to the air movement (delta between \qty{0.25}{\m\per\s} and \qty{0.1}{\m\per\s}) which is equal to \qty{1.1}{\celsius}.
     Even after reducing both \ac{tdb} and \ac{tr} by the \ac{ce}, the person is still estimated to be sweating.
@@
     You are citing skin wettedness values that have not been introduced in this paper (and are not part of any standard). 
-Where did they come from?
-You allude to results from Huang 2014 that the reader is supposed to know about.
+    Where did they come from? You allude to results from Huang 2014 that the reader is supposed to know about. 
     The skin-wettedness discussion is irrelevant unless you go into a lot more physiological detail here.}
+    \todo[inline, color=blue!40]{FT.
+We developed this thought-example to demonstrate that while the PMVCE seeks to improve the modeling of sweat rates, it does not alter the underlying PMV equations.
+    As a result, there are situations where the PMVCE may still fail to provide accurate predictions.
+    We have modified the example and changed the air speed from 0.15 to 0.25 m/s}
     \item The previous example, highlights another limitation of the \ac{pmv-ce}.
@@
     However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.
-    \item The \ac{set} temperature which is used to calculate the \ac{ce}, is the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{vr}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
-    However, while \ac{vr}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} fixed the \ac{rh}~=~\qty{50}{\percent} and standardized the clothing to the activity.
-    \todo[inline]{I think this argument is incorrect.
-SET is only used as a consistent method of obtaining a meaningful delta operative temperature.
-It does not change the environment.
-The environment is kept at the original RH.}
     \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
+    In some contexts, like in yearly building performance simulations, this could make a difference. 
     \todo[inline]{This may or may not be significant.
@@
 We would have to calculate two validation tables for the two approaches.}
+    \todo[inline, color=blue!40]{FT. I agree that this may or may not be significant, depending on the context and the computation speed could be improved.}
 \end{enumerate}

### 2025-01-28 Manuscript/src/sections/results.tex
review: addressed all reviewer's comments (with Ed's comments)

@@
 In this section, we try to identify some of the main reasons that could explain the difference in the results from the two \ac{pmv} models when compared against the \ac{tsv} available in the \ac{db2}.
-Moreover, we tried to explain why we observed that the \ac{pmv-ce} model has a lower accuracy than the \ac{pmv} in predicting thermal sensation.
 The issues are reported in order of importance of what we believe may affect the accuracy of the model. 
-\todo{Ed's comment - Note that the section below does not mention the two-node model, the CE implementation of PMV, or the physiological differences between the ISO and ASHRAE models that could explain the observed differences or ‘error’.
-}
+In this paragraph, we refer to the \ac{pmv} model, but the consideration can also be applied to the \ac{pmv-ce} model given its formulation.
 
-\paragraph{Heat Balance Equation and Calculation of the PMV Value}
+\paragraph{Heat Balance Equations and Thermal Sensation Predictions}
 The \ac{pmv} model uses simplified heat balance equations to estimate the heat losses and gains from the human body to its surrounding environment.
-This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
+This is a possible source of error since the model considers the human body to be a cylinder with constant width, all uniformly covered with clothing, where the heat is all generated in its core.
 This is a significant simplification, as it overlooks several important factors, such as the uneven coverage of clothing across different parts of the body, the non-uniformity of metabolic heat generation, and the variations of mass-to-surface ratios.
@@
 Additionally, the \ac{pmv} model erroneously assumes that the human body is not capable of maintaining a stable core temperature.
-A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicate that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently it is getting warmer or colder, respectively.
+A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicate that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently, it is getting warmer or colder, respectively.
 
@@
 The \ac{pmv} model also assumes that the human body is constantly losing a certain amount of heat through sweating and this heat loss (W/m\textsuperscript{2}) only varies as a function of \ac{met} and it is calculated using the following equation $0.42\times($\ac{met}$\times58.15 - 58.15)$~\cite{Fanger1970}.
-This amount of heat loss is included in the equation even if the cylinder has a negative heat gain, or in other words, the person is estimated to be feeling cold.
+This amount of heat loss is included in the equation even if a person is in thermal neutrality (feeling comfortable) or has a negative heat gain (feeling cooler than neutral).
 
@@
 Heat losses or gains are only a proxy for the thermoregulation effort that the human body has to undergo to maintain a constant internal temperature.
-Fanger created this equation based on a limited dataset collected in a laboratory, and this equation was never updated or revisited~\cite{Fanger1970}.
-In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a \ac{tsv}, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction to cope with warm and cold conditions~\cite{romanovsky_thermoregulation_2018}.
+Fanger created this equation based on a limited dataset collected in a laboratory, which was never updated or revisited~\cite{Fanger1970}.
+In the \ac{pmv} model, the heat losses and gains are mapped using constant coefficients to a \ac{tsv}, disregarding the fact that the human body employs different strategies, i.e., vasodilation and vasoconstriction, to cope with warm and cold conditions~\cite{romanovsky_thermoregulation_2018}.
 

### 2025-01-28 Manuscript/src/sections/conclusions.tex
review: removed ed's comments

@@
 Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
-\todo{EA. But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
-\todo[inline, color=blue!40]{FT. We clarified this issue above.}
 Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
-\todo{unclear what limits you are talking about, or who the audience is for your suggestions.  The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.  
-Your suggestions are just aimed at you as keepers of the ComfortTool}
-\todo[inline, color=blue!40]{FT. The limits are the one described in Table 2, for example, we suggest to use the model only from 0.3 and 1.5 clo and met between 1 and 2.}
 The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
@@
 This scale more clearly allows an HVAC system to decide whether to change or not the thermal environment.
-\todo{EA.
-This is a major suggestion with very significant implications for energy etc that have not been discussed in the paper.
-There are surely reasons people have not used that scale on its own, for this purpose.}
-\todo[inline, color=blue!40]{FT.
-We discussed this in Figure 5a and in its related text.
-The PMV model assume that people are comfortable when they claim to be slightly warm or slightly cool, this is explicit in Fanger's book.}
 

### 2025-01-28 Manuscript/src/sections/introduction.tex
review: removed ed's comments

@@
 
-\todo[inline]{Ed's comment:
-The PMV model consists of a heat balance model and an empirical equation that predicts thermal sensation (PMV) from heat balance.
-(need to describe this further, especially the simplified way that sweating is treated by the heat balance model).
-In 2010 ASHRAE Std 55 modified the PMV heat balance calculation to make it more capable of predicting the cooling caused by elevated air movement under warm conditions.
-It uses the Gagge two-node model to compute the convective and evaporative heat loss from the body.
-This model explicitly models physiological cooling in warm conditions where evaporation of sweat becomes a significant source of heat loss.
-The model output can be expressed as Standard Effective Temperature (SET) (provide citations to Gagge/Foblets).
-SET is implemented in Std 55 as an index to quantify the equivalent temperature reduction caused by air movement cooling.
-The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.
-This approach is intended to minimize errors introduced by the Fanger heat balance model’s reduced evaporative capacity on the predicted PMV result.
-}
-\todo[inline, color=blue!40]{
-    FT. Firstly, it is important to note that the PMV does not aim to estimate sweating. 
-    While this is a limitation, neither the PMV nor the PMV CE models include this capability.  
-    The Std 55 did not modify the PMV heat balance equation but instead reduced the air and mean radiant temperatures by the cooling effect, assuming this adequately accounts for sweating. 
-    However, this approach does not equate to an accurate heat balance calculation where sweat rate is explicitly modeled, as seen in the Gagge two-node model.  
-    Consequently, the PMV CE does not: 1) use the Gagge two-node model to compute convective and evaporative heat loss from the body, nor 2) explicitly model physiological cooling in warm conditions where sweat evaporation becomes a key source of heat loss.
-    Secondly, in the PMV CE model, the cooling effect is applied in all conditions, not just in warm ones.  
-    Thirdly, the definition of SET can also be somewhat unclear. 
-    It is not calculated for a reference environment with standardised clothing and metabolic rates. 
-    Instead, it represents the temperature of a hypothetical isothermal environment at 50\% RH, $<$0.1 m/s (20 fpm) average air velocity, where the total heat loss from the skin of an imaginary occupant (wearing clothing standardised for the activity) equals the heat loss from a person in the actual environment with their actual clothing and activity level.
-    Finally, as you mentioned, \textit{The reduced temperature is entered into the PMV heat balance model in a still air condition, from which the PMV index is calculated.} However, this does not guarantee that the person will not feel hot or sweat in this condition and, therefore, does not address the inherent issues of the original PMV model.
-}
-
 \subsection{\ac{pmv-ce} model}\label{subsec:pmv-ce-limitations}
@@
 However, we could not find a peer-reviewed scientific publication that quantified the accuracy improvements of the model as implemented in the \gls{55} over the original \ac{pmv} model.
-\todo[inline]{EA. true enough, but in the standard, the objective is to have a rational physiology model.
-The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
-\todo[inline, color=blue!40]{FT.
-The PMV CE and the Gagge model are different, hence, publications about the latter do not sufficiently describe the rationale of the new model, moreover, the accuracy of the PMV CE was not tested}
 
@@
 We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
-\todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. Many lab studies have measured results in that region.}
-\todo[color=blue!40]{FT. In the manuscript, we decided not to include a Figure for 0.6m/s since in the ASHRAE DB 2 we hardly reach that value.
-I am happy to generate a figure for you if need be}
 \begin{figure}[!htb]
@@
     Therefore, in this scenario, where the \ac{pmv-ce} should excel compared to the \ac{pmv}, the \ac{pmv-ce} model still fails to accurately predict the \ac{tsv}.
-    \todo[inline]{This is really confusing.
-    You are evaluating the Cooling Effect in still air, when there is no Cooling Effect to evaluate.
-    You are using as an example 29C in still air and 80\% RH which of course will cause sweating.
-    You are citing skin wettedness values that have not been introduced in this paper (and are not part of any standard). 
-    Where did they come from? You allude to results from Huang 2014 that the reader is supposed to know about. 
-    The skin-wettedness discussion is irrelevant unless you go into a lot more physiological detail here.}
-    \todo[inline, color=blue!40]{FT.
-We developed this thought-example to demonstrate that while the PMVCE seeks to improve the modeling of sweat rates, it does not alter the underlying PMV equations.
-    As a result, there are situations where the PMVCE may still fail to provide accurate predictions.
-    We have modified the example and changed the air speed from 0.15 to 0.25 m/s}
     \item The previous example, highlights another limitation of the \ac{pmv-ce}.
@@
     In some contexts, like in yearly building performance simulations, this could make a difference. 
-    \todo[inline]{This may or may not be significant.
-If it is significant, the process could be simplified to a total of three calculations while remaining a rational method.
-What’s in there now was Tyler’s (CBE’s) choice.
-Charlie and I have used it more simply.
-We could propose this to the Standards Committee if we wanted to.
-We would have to calculate two validation tables for the two approaches.}
-    \todo[inline, color=blue!40]{FT. I agree that this may or may not be significant, depending on the context and the computation speed could be improved.}
 \end{enumerate}

### 2025-01-28 main.py

@@ -30,10 +30,12 @@ warnings.filterwarnings("ignore")
 psychrolib.SetUnitSystem(psychrolib.SI)
 
 
-results = two_nodes(29, 29, 0.15, 80, 1, 0.5, 0)
-results["w"]
-results = cooling_effect(29, 29, 0.15, 80, 1, 0.5, 0)
-print(results)
+results = two_nodes(29, 29, 0.25, 80, 1, 0.5, 0)
+print(results["w"])
+ce = cooling_effect(29, 29, 0.25, 80, 1, 0.5, 0)
+print(ce)
+results = two_nodes(29 - ce, 29 - ce, 0.25, 80, 1, 0.5, 0)
+print(results["w"])
 
 palette_tp = [
     "#06A6EE",

### 2025-01-28 Manuscript/src/response-letter.tex
submission: second submission Building and Environment

@@
                 It’s important to note that the thermal manikin used in their study measured only dry heat losses and the assessment was done in conditions in which sweating was low. 
-                Yang et al. concluded: "To assess which method for SET* implementation and which air speed measurements should be used, experiments with human subjects are needed". 
+                Yang et al. concluded: \textit{``To assess which method for SET* implementation and which air speed measurements should be used, experiments with human subjects are needed''.}
                 To our knowledge experiments have not been done yet.  
@@
             \response{
-                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data and that the \ac{db2} has substantial limitations, as mentioned by the reviewer.
-                \textbf{Measurements errors}\\
+                We agree with the reviewer that the accuracy of the PMV model is influenced by the quality of the input data and that the \ac{db2} has substantial limitations, as mentioned by the reviewer.\\
+                \textbf{Measurements errors}:
                 We previously mentioned in the manuscript that not all studies included in the \ac{db2} used the same measurement protocols. 
@@
                 In science and technology, researchers often face the challenge of balancing sample size and data quality within a fixed financial or time budget. 
-                Both extremes offer valuable insights: small, laboratory, high-quality datasets provide precise information, while large, less accurate datasets obtained in field studies enable more "ecologically valid" and broader trends and patterns to emerge.
+                Both extremes offer valuable insights: small, laboratory, high-quality datasets provide precise information, while large, less accurate datasets obtained in field studies enable more ``ecologically valid'' and broader trends and patterns to emerge.
                 In recent years, the value of large, less precise datasets has profoundly transformed the way science is conducted, driving new discoveries and methodologies.\\
@@
 “Air Movement – Good or Bad?” Indoor Air 14 (August 2, 2004): 40–45.
-https://doi.org/10.1111/j.1600-0668.2004.00271.x.
+https://doi.org/10.1111/j.1600-0668.2004.00271.x.\\
             In this study, we did not attempt to evaluate local discomfort models, such as radiant asymmetry, thermal stratification, floor temperature, draft, ankle draft, etc. 
@@
             Besides airspeed, we anticipate that other environmental factors, such as air temperature stratification, could influence overall thermal sensation. 
-            However, since neither standard mathematically integrates local discomfort models with the overall PMV, they are treated independently, therefore we chose not to include them in our analysis.
-            Minor note: Interestingly, we observed that both \ac{pmv} models have higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.}
+            However, since neither standard mathematically integrates local discomfort models with the overall PMV, they are treated independently, therefore we chose not to include them in our analysis.\\
+            \textbf{Minor note:} Interestingly, we observed that both \ac{pmv} models have higher accuracy and lower bias in predicting \ac{tsv} in the subset of data with air speed greater than 0.2 m/s.}
         \end{enumerate}
@@
             Some practitioners may have the option to select the model to use, for example, some certification programs, including LEED, let the designer decide which model to use. 
-            We have modified the abstract and removed the following text: ``, and their different PMV formulations are a source of confusion.'' since it was not clear.
-            We have also changed the text in Section 1.4 which now reads: ``Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
+            We have modified the abstract and removed the following text: \textit{``, and their different PMV formulations are a source of confusion.''} since it was not clear.
+            We have also changed the text in Section 1.4 which now reads: \textit{``Choosing between the \ac{pmv} and \ac{pmv-ce} is a source of confusion for researchers, educators and practitioners worldwide since both models are widely used in building codes, guidelines and certification programs.
             For example, the WELL certification allows compliance with \gls{7730} and \gls{55} standards, even though the two models have different outputs under the same environmental and personal conditions as shown in Figure~2.
-This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.''
+This can lead to confusion when comparing results from different studies or when trying to apply the results in practice.''}
             }
@@
                 In the paper, we now clearly differentiate between \acf{v} and \acf{vr}.
-                We have specified this in the nomenclature section of the manuscript.
+                We have specified this in the nomenclature section of the manuscript.\\
                 4) we have reworded the sentence ``Consequently, for a given thermal environment, results of the two PMV formulations differ only when the value of V is higher than 0.1m/s.'' in the manuscript to better explain when the two models differ.
@@
                 This is because the PMV model can be applied to predict thermal sensation across a wide range of environments, including naturally conditioned spaces. 
-                It is worth noting that standards, such as ASHRAE 55:2023, allow the use of “adaptive comfort” in “occupant-controlled naturally conditioned spaces” when all the conditions outlined in Section 5.4.1 are met.
+                It is worth noting that standards, such as ASHRAE 55:2023, allow the use of ``adaptive comfort” in ``occupant-controlled naturally conditioned spaces” when all the conditions outlined in Section 5.4.1 are met.
                 However, they do not prohibit the use of the PMV model in naturally ventilated spaces. 
-                In fact, the PMV model can be applied in "all occupied spaces within the scope of this standard" (Section 5.3.1). 
-                It should be noted that we did not include data from outdoor environments in the analysis since the \ac{db2} only includes data from \textit{```real' buildings occupied by `real' people doing their normal day-to-day activities.''}.
+                In fact, the PMV model can be applied in ``all occupied spaces within the scope of this standard" (Section 5.3.1).
+                It should be noted that we did not include data from outdoor environments in the analysis since the \ac{db2} only includes data from \textit{`` `real' buildings occupied by `real' people doing their normal day-to-day activities.''}.
                 The database, hence, does not include data from outdoor environments or climate chamber studies.
@@
             It is worth noting that this distinction is a secondary objective of our work, but we believe it is an important point to address. 
-            We modified the text as reported in answer 11 and made our assumption explicit, so, it is not presented as a universal fact, but conditional to our assumption).}
+            We modified the text as reported in answer 11 and made our assumption explicit, so, it is not presented as a universal fact, but conditional to our assumption.}
 
@@
             \item Some conclusion contents lack data and theoretical support.
+
             \response{
@@
             \response{
-                Please see our answer above about the weak definition of what is "a large group of people" and the fact that standards allow the PMV model to be used in places occupied by only one person.
+                Please see our answer above about the weak definition of what is ``a large group of people" and the fact that standards allow the PMV model to be used in places occupied by only one person.
                 Nevertheless, we changed the sentence as follows: \textit{``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent the average thermal sensation of a large group of occupants.}

### 2025-03-07 Manuscript/src/sections/conclusions.tex
feat: changes from Overleaf before publication

@@
 The \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.
-Assuming that people who are `slightly warm' or `slightly cool' are thermally neutral leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
+Assuming that people who are `slightly warm' or `slightly cool' are comfortable leads to undercounting thermal discomfort in our dataset, which is not a valid assumption.
 We recommend the use of the thermal preference scale to determine how people perceive their thermal environment. 

### 2025-03-07 Manuscript/src/sections/introduction.tex
feat: changes from Overleaf before publication

@@
 The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some changes to the \ac{pmv} inputs were implemented, nor for example, why the \ac{ce} is subtracted from the \ac{tdb} and \ac{tr}.
+It is a common practice of normative standards not to provide explanations.
 A partial justification of the model is mainly described in \mycite{arens_moving_2009} and secondarily in \mycite{yang_cooling_2015}.
@@
     \item It should be noted that over the years, the threshold value of \ac{vr} after which the \ac{ce} is calculated, has been changed from \qty{0.15}{\m\per\s} in the ASHRAE~55:2013~\cite{ASHRAE552013} to \qty{0.2}{\m\per\s} in the ASHRAE~55:2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
-    However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.
+    These changes significantly affected the output and applicability of the model.
     \item Calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the \ac{set} model.
@@
 We used the \ac{tsv} recorded in the \acf{db2}.
-We aim to determine which \ac{pmv} model is most accurate and the models' applicability limits.
+We aim to determine which \ac{pmv} model is more accurate and the models' applicability limits.

### 2025-03-07 Manuscript/src/sections/methodology.tex
feat: changes from Overleaf before publication

@@
 However, since the thermal sensation was measured with a seven-point scale, the \ac{pmv} has no upper or lower boundary, and the \gls{55} does not specify the range of applicability of the \ac{pmv-ce} model, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.
+The \gls{55} classifies “combinations of indoor thermal environmental factors and personal factors that produce satisfactory thermal environmental conditions” into only two categories: compliant or non-compliant with the standard.
+It defines compliance as $\mid$\ac{pmv-ce}$\mid \leq 0.5$.
+Consequently, its accuracy could have been evaluated by relating the standard-compliant (or not) conditions and the thermal satisfaction (dissatisfaction) votes.
+However, this metric is rarely collected and is absent from the \ac{db2}.
 
@@
 They showed that the reliability of the \ac{pmv} model is strongly affected by the accuracy of the input data.
+Moreover, it is unclear if the clothing insulation of the chair has been consistently included or excluded in \ac{clo}.
 In our analysis, we compensate this issue by:

### 2025-03-07 Manuscript/src/sections/titleAbstract.tex
feat: changes from Overleaf before publication

@@
         Hence, we suggest reducing the applicability limit of both models within the range $\mid$\ac{pmv}$\mid \leq 0.5$.
-        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally neutral leads to under-counting thermal discomfort.
+        Finally, the assumption that individuals who feel `slightly warm' or `slightly cool' can be considered thermally comfortable leads to under-counting thermal discomfort.
 %        todo manually check values

### 2025-03-07 Manuscript/src/sections/appendix.tex
submission: published version

@@
 In this section, we show the results of the analysis of the dataset with air speed measurements at three heights.
-% todo check manually
 Only \num{3059} entries in the \ac{db2} had air speed measurements at three heights with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
@@
 The \ac{pmv} model is more accurate than the \ac{pmv-ce} model.
-The intercept of the \ac{pmv} model is \var{intercept_v_02_three_pmv} while the intercept of the \ac{pmv-ce} model is \var{intercept_v_02_three_pmv_ce}.
+The intercept of both \ac{pmv} models is \var{intercept_v_02_three_pmv_ce}.
 Moreover, the \ac{pmv-ce} \ac{lowess} curve is only marginally higher than \num{0} despite participants reporting a \ac{tsv} of \num{3}.
@@
 \end{figure}
-% todo check manually
 The \ac{pmv} model has a bias of \num{-.39} while the \ac{pmv-ce} model has a bias of \num{-.38}.

### 2025-03-07 Manuscript/src/sections/results.tex
submission: published version

@@
 The overall accuracy of both \ac{pmv} models changed to \var{overall_acc_pmv_0.2} and \var{overall_acc_pmv_ce_0.2}, respectively when only considering entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s}.
-% todo check data manually
 The \ac{pmv-ce} model outperforms the \ac{pmv} model for \ac{tsv}~=~0 and -1.
@@
 \end{table}
-% todo check data manually
 The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}) for all data.
@@
 The intercepts for the \ac{pmv} is \var{intercept_all_pmv} and for the \ac{pmv-ce} is \var{intercept_all_pmv_ce} which means that the model has a bias of less than one-quarter of a thermal sensation interval.
-% todo check data manually
 This result is aligned with those presented in Figure~\ref{fig:bar_stacked_model_accuracy} where the \ac{pmv} and \ac{pmv-ce} models can predict people who reported to be neutral \qty{51}{\percent} and \qty{58}{\percent} of the times, respectively.
@@
 Figure~\ref{fig:hist_discrepancies} contains the summary statistics.
-% todo check data manually
 Overall, both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
@@
 The results depict that the use of the \ac{pmv-ce} model does not improve the bias of the \ac{pmv} results across all input variables.
-% todo check data manually
 On the contrary, the bias of the \ac{pmv-ce}~=~\num{-.33} is higher than \ac{pmv}~=~\num{-0.23} for \qty{0.3}{\m\per\s}~$<$~\ac{vr}~$\leq$~\qty{0.46}{\m\per\s}.

### 2025-03-07 main.py

@@ -3621,7 +3621,7 @@ def compare_pmv_pmv_ce_comfort_region():
         axs[ix].set(
             ylabel="Relative Humidity (%)",
             xlabel="Temperature (°C)",
-            title=r"$V$" + f" = {v} m/s",
+            title=r"$V_r$" + f" = {v} m/s",
             ylim=(0, 100),
             yticks=np.arange(0, 110, 25),
         )
